---
title: "Early Prediction of Breast Cancer Progression-Free Survival"
format: 
  revealjs:
    smaller: true
---

```{r}
#| label: setup
#| include: false

# library(conflicted)
library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(gtsummary)
library(targets)

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

knitr::opts_chunk$set(cache.path = file.path("/wscratch", Sys.getenv("USER"), "adc-early-predict/cache/"))

options(mc.cores = 4)

ggplot2::theme_set(theme_minimal())

source(here("src", "util.R"))
```

```{r}
#| label: priors
#| cache: true

tumor_priors <- lst(
  pop_tumor_gp_rho_alpha = 7.3,
  pop_tumor_gp_rho_beta = 7.5, 
)

pfs_priors <- lst(
  log_lambda_gp_intercept_mean = -1,
  log_lambda_gp_intercept_sd = 0.25,
  log_lambda_gp_alpha_sd = 0.25,
  log_lambda_gp_rho_alpha = 7.3,
  log_lambda_gp_rho_beta = 7.5, 
  tumor_stim_intercept_sd = 0.125,
  log_lambda_gp_trial_alpha_sd = 0.0625,
  log_lambda_gp_trial_intercept_sd_sd = 0.25,
  tumor_stim_coef_sd = c(0.125, 0.125, 0.0625) / 2,
  tumor_stim_trial_coef_sd_sd = 0.0625
)
```

```{r}
#| label: models
#| code-fold: false   

tumor_model <- cmdstan_model(here("src", "tumor.stan"))
pfs_model <- cmdstan_model(here("src", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)
```

```{r}
#| label: load-trials-data
#| cache: true

data_details <- tribble(
  ~ idap,          ~ trial, 
  "idap_20231204", "Breast01",
  "idap_20231129", "Breast02",  
  "idap_20231124", "Breast04"
) |> 
  mutate(across(everything(), factor))

# Demographics 
sdtm_dm <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_dm", "sdtm")))

# Tumor analysis_data
adam_adtr <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adtr", "adam")))

# Subject-level data
adam_adsl <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adsl", "adam")))

adam_adcm <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adcm", "adam")))

# Time to event
adam_adtte <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adtte", "adam")) |> 
           map(\(d) mutate(d, cnsr = cnsr == 1)))
```

```{r}
#| label: prepare-analysis-data
#| cache: true
#| dependson: load-trials-data
#| code-fold: false 

tumor_analysis_data <- adam_adtr |> 
  unnest(data) |> 
  filter(
    fct_match(paramcd, "LDIAM"),
    # There are a number of assessments that show up in the data but in SDTM they are marked as "NOT DONE"
    !is.na(trdy), !is.na(aval),
    tracptfl,
    aval > 0, # Drop all zero diameter assessments; these appear to be missing measurements. 
  ) %>% 
  select(studyid, idap, trial, 
         usubjid, trlnkid, tuloc, tulat, trtsdt, trtedt, adt, ends_with("fl"), trdy, trtp, trta, visit, visitnum, mmdiam = aval) |> 
  mutate(
    week = if_else(trdy > 0, trdy, trdy + 1) %/% 7, 
    treated_week = adt >= trtsdt,
  ) |> 
  arrange(usubjid, trlnkid, visitnum) |> 
  group_by(usubjid, trlnkid) |> 
  mutate(tuloc = first(tuloc), tulat = first(tulat) |> na_if("")) |> # Only the first assessment has the location of the tumor. Fix that.
  ungroup() |>
  # Turns out that some assessments are not listed in order of the days they were done. Look at subject DS8201-A-U303-10033028.
  arrange(studyid, usubjid, trlnkid, trdy) |> 
  nest(tumor_history = c(adt, trdy, week, visit, visitnum, tsplitfl, tmergefl, aocc01fl, mmdiam, treated_week)) |> 
  mutate(# For some weeks, there are multiple (unscheduled) visits that might happen. I'm just going to use the average for tumor sizes within those
         # weeks.
    tumor_history = map(
      tumor_history, 
      \(h) nest(h, visits = c(adt, trdy, visit, visitnum, mmdiam)) |> 
        mutate(
          n_week_visits = map_int(visits, nrow),
          mmdiam = map_dbl(visits, \(v) mean(v$mmdiam))
        )
    ),
    n_tumor_measures = map_int(tumor_history, nrow),
    
    min_t = map_int(tumor_history, \(h) min(h$week)),
    max_t = map_int(tumor_history, \(h) max(h$week)),
  ) |> 
  pack(flags = ends_with("fl")) |> 
  nest(patient_tumors = !c(studyid, trial, idap, usubjid, trtsdt, trtedt, trtp, trta)) |> 
  rowwise() |> 
  mutate(
    patient_min_t = min(patient_tumors$min_t),
    patient_max_t = max(patient_tumors$max_t),
  ) |>
  ungroup() |> 
  mutate(
    patient_t_width = patient_max_t - patient_min_t + 1,
    n_tumors = map_int(patient_tumors, nrow),
  ) |> 
  left_join(
    adam_adtte |> 
      unnest(data) |> 
      filter(
        (fct_match(paramcd, "PFS1") & fct_match(studyid, c("DS8201-A-U301", "DS8201-A-U303"))) | 
          (fct_match(paramcd, "PFS") & fct_match(studyid, "DS8201-A-U201")), 
        fct_match(parqual, "CENTRAL")
      ) |> 
        select(studyid, usubjid, aval, death = dthfl, dthdt, fpddt, cnsr),
    by = c("studyid", "usubjid")
  ) |> 
  mutate(
    right_censored = cnsr,
    death_week = floor(time_length(dthdt - trtsdt, unit = "weeks")),
    progress_week = floor(time_length(fpddt - trtsdt, unit = "weeks")),
    progress_week = if_else(death, coalesce(progress_week, death_week), 
                                   if_else(right_censored & is.na(progress_week), max(progress_week, na.rm = TRUE), progress_week)),
    aval_week = aval %/% 7,
    
    treated = 
      (fct_match(trial, "Breast01") & fct_match(trtp, "5.4 mg/kg")) | 
      (fct_match(trial, c("Breast02", "Breast04")) & fct_match(trtp, "T-DXd")),
    
    # The model expects "pfs" to record the number of weeks that were progress free, so I need to find the last assessment week before
    # disease progress was detected.
    pfs = pmap_int(
      lst(progress_week, death_week, right_censored, patient_tumors), function(progress_week, death_week, right_censored, patient_tumors) {
        pre_progress_weeks <- unnest(patient_tumors, tumor_history) |>
          distinct(week) |>
          filter(right_censored | week < min(progress_week, death_week, na.rm = TRUE)) |>
          pull(week)

        # There are a few patients who just have a single post treatment visit
        if (length(pre_progress_weeks) > 0) max(pre_progress_weeks) else NA_integer_
      }
    ),

    death_week = coalesce(death_week, 0L),
    
    n_measures = map(patient_tumors, \(tu) tu$n_tumor_measures),
    t_measure = map(patient_tumors, \(tu) map(tu$tumor_history, \(tr) pull(tr, week))),
  ) |> 
  left_join(
    unnest(adam_adsl, data) |> 
      transmute(
        studyid, usubjid, country, age, sex, race, ecogbl, hormonr, her2ic1, nreg, 
        region = coalesce(region1, region2)
      ), 
    by = c("studyid", "usubjid")
  ) %>%
  mutate(
    interval_censored = with(., pfs_model$functions$identify_censoring(pfs, death_week, n_tumors, unlist(n_measures), unlist(t_measure))[[1]]) |> 
      pmax(0) # Issue https://github.com/azu-oncology-rd/ods-adc-early-predict-2023/issues/36
  ) |> 
  nest_join(unnest(adam_adcm, data) |> select(usubjid, cmcat, response, cmdecod), by = c("usubjid"), name = "med_data") |> 
  mutate(
    med_data = map(med_data, \(cm) mutate(cm, response = str_extract(response, "CR|NA|PR|PD|SD|UNK")))
  )
  
pfs_analysis_data <- tumor_analysis_data |> 
  filter(
    !is.na(pfs), pfs >= 0, # No missing pfs and progress must be assured to have happened after treatment
    patient_min_t <= 0, # At least one screening assessment needed,
    patient_max_t > 0, # and at least one post treatment assessment.
  )

treated_pfs_analysis_data <- pfs_analysis_data |> 
  filter(treated) 

treated_pfs_stan_data <- prepare_pfs_stan_data(treated_pfs_analysis_data, tumor_priors, pfs_priors)

treated_trial_pfs_stan_data <- treated_pfs_analysis_data |> 
  nest(.by = trial) |> 
  deframe() |> 
  imap(\(d, tr) prepare_pfs_stan_data(mutate(d, trial = tr), tumor_priors, pfs_priors))

treated_obs_km <- map_dfr(treated_trial_pfs_stan_data, function(sdata) {
  with(sdata, {
    list(lb = pfs, ub = pfs + interval_censored) |> 
      map_dfr(function(s) {
        pfs_model$functions$estimate_kaplan_meier(s, right_censored, max(s)) |>
          set_names(c("s", "n", "c", "e")) |>
          as_tibble() |> 
          mutate(t = seq(0, n() - 1))
      }, .id = "btype")
    }
  )
}, .id = "trial")

hist_breaks <- seq(10, 150, by = 10)
```

# Overiew

## Main results

::: {.r-fit-text}

Can early data on lesion sizes lead to a _meaningful_ improvement in predicting progression-free survival? 

:::

* We are interested in predicting progression-free survival (PFS) from early assessments of target lesion sizes.
* Predictors: diameters of target lesions assessed in 
  - Last pre-screening assessment
  - First assessment after treatment begins (~ 6 weeks after)
* Prediction accuracy improves compared to a baseline no predictors model. This was evaluated using
  - Qualitative evaluation of the model's sensitivity to lesion sizes.
  - Out-of-sample accuracy using various measures 

## Implications of these results

How is this useful? 

Improved prediction of survival conditional on early assessments has the potential to

* Provide early decisions on study progression.
* Provide information about later stage studies:
  - Predicted outcomes.
  - Sample size requirements.
* Identify inclusion/exclusions criteria for later stage studies.
* Personalize patient-level treatment decisions.

## Limitations of this exploratory analysis

* Only two assessments are used for training and prediction.
* Prediction improvement is mostly driven by the _number_ of target lesions.
* Prediction is not causal; it is prognostic within the treatment arm. 
* No other covariates' interaction with lesion sizes are evaluated.
* Results are restricted to prediction within, not across, trials.
* Still unclear if the magnitude of prediction accuracy improvement is _meaningful_.

## Next steps 

* More time needs to be committed to address the caveats described.
* A better understanding of the causal prediction identification strategy.
* Exploit the full lesion size history available. 

## Data used in analysis

* We used data from the following trials:
    1. Breast01: phase II dose finding study (HER2-positive metastatic breast cancer).
    2. Breast02: phase III study comparing to physician choice of drug (HER2-positive metastatic breast cancer). 
    2. Breast03^[Pending]: phase III study comparing to trastuzumab emtansine (HER2-positive metastatic breast cancer).
    3. Breast04: This is a phase III study on patients with HER2-low metastatic breast cancer (HER2-low metastatic breast cancer).
* We restricted our analysis to the primary arms of these trials, receiving the same dosage of trastuzumab deruxtecan.

## Subject-level characteristics at baseline patients included in analysis {.scrollable} 

```{r}
#| label: tbl-subject-level-summary

over_64 <- function(x) sum(x >= 65)

pfs_analysis_data |>
  filter(treated) |> 
  mutate(
    sex = fct_recode(sex, "Female" = "F", "Male" = "M"),
    race = str_to_title(race) |> fct_recode("Missing data" = "") |> fct_relevel("Other", "Missing data", after = Inf),
    region = fct_collapse(region, "Europe + Israel" = c("EU", "Europe", "Europe + Israel")),
    hormonr = str_to_title(hormonr) |> fct_collapse(Unknown = c("Unknown", "Indeterminate")),
    her2ic1 = str_to_title(her2ic1) |> fct_collapse("Not Examined" = "Ne"),
    
    sum_tumor_size = map_dbl(
      patient_tumors, \(tr) map(tr$tumor_history, \(th) filter(th, week <= 0) |> filter(min_rank(week) == n()) |> pull(mmdiam)) |> 
        compact() |> 
        unlist() |> 
        sum()
    ) |> 
      divide_by(10) # cm
  ) |> 
  tbl_summary(
    by = trial,
    include = c(age, sex, race, region, ecogbl, hormonr, her2ic1, sum_tumor_size, nreg),
    label = list(
      age = "Age (years)", sex = "Sex", race = "Race", region = "Region", ecogbl = "ECOG performance-status score", hormonr = "Hormone-receptor status",
      her2ic1 = "HER2 expression", sum_tumor_size = "Sum of diameters of target lesions (cm)", nreg = "Number of previous regimens"
    ),
    type = list(age = "continuous2"),
    statistic = list(
      age = c("{median} ({min} - {max})", "{over_64}"), sum_tumor_size = "{median} ({min} - {max})",
      nreg = "{median} ({min} - {max})"
    ),
  ) |> 
  add_stat_label(label = list(age = c("Median (Range)", ">64 yr"))) 
```

## Previous systemic cancer therapy for patients included in analysis 

```{r}
#| label: tbl-prior-med-summary

pfs_analysis_data |>
  filter(treated) |> 
  unnest(med_data) |> 
  filter(fct_match(cmcat, "PRIOR CANCER SYSTEMIC THERAPY")) |> 
  mutate(
    med = fct_collapse(
      cmdecod, 
      "Trastuzumab" = "TRASTUZUMAB", 
      "Trastuzumab emtansine" = "TRASTUZUMAB EMTANSINE", 
      "Pertuzumab" = "PERTUZUMAB",
      "Taxane" = "TAXANES",
      other_level = "Other"
    )
  ) |> 
  tbl_summary(
    by = trial,
    include = c(med),
    label = list(
      med = "Previous systemic cancer therapy"
    ),
    statistic = list(
    ),
  ) 
```

# Analysis

```{r}
#| label: survival-fit
#| cache: true
#| cache-lazy: false

# prior_no_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^prior_no_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
no_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^no_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
tumor_count_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^tumor_count_pfs-", full.names = TRUE), check_diagnostics = FALSE)
# prior_two_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^prior_two_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
two_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^two_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
tumor_change_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^tumor_change_pfs-", full.names = TRUE), check_diagnostics = FALSE)

```

```{r}
#| label: predict-error 
#| cache: true

linear_loss <- function(m, c, a, ac, k0 = 1, k1 = 1) if_else(c & (ac | a >= m), 0, if_else(m >= a, k0, k1) * abs(m - a))
squared_error_loss <- function(m, c, a, ac) if_else(c & (ac | a >= m), 0, (m - a)^2)

asym_linear_loss <- partial(linear_loss, k1 = 3)

no_tumor_loo <- no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
tumor_count_loo <- tumor_count_pfs_res$loo(save_psis = TRUE, cores = 12)
two_tumor_loo <- two_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
tumor_change_loo <- tumor_change_pfs_res$loo(save_psis = TRUE, cores = 12)

no_tumor_asym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, treated_pfs_stan_data, asym_linear_loss)
two_tumor_asym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, treated_pfs_stan_data, asym_linear_loss)
no_tumor_sym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, treated_pfs_stan_data, linear_loss)
two_tumor_sym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, treated_pfs_stan_data, linear_loss)
no_tumor_sqerr_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, treated_pfs_stan_data, squared_error_loss)
two_tumor_sqerr_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, treated_pfs_stan_data, squared_error_loss)

loss_data <- bind_rows(
  lst(no_tumor_sym_loss, 
      two_tumor_sym_loss, 
      no_tumor_asym_loss, 
      two_tumor_asym_loss,
      no_tumor_sqerr_loss,
      two_tumor_sqerr_loss), 
  .id = "model"
) |> 
  mutate(loss_type = str_extract(model, "(a?sym)|sqerr"), model = str_extract(model, "(ic_ignored_)?(no|two)_tumor"))

```

```{r}
#| label: predict-error-diff
#| cache: true

loss_diff_data <- loss_data |> 
  filter(fct_match(loss_type, "sym")) %>% {
    left_join(
      filter(., fct_match(model, "two_tumor")), 
      filter(., fct_match(model, "no_tumor")), 
      by = c("id", "trial"), relationship = "one-to-one")
  } |> 
  mutate(loss_diff = value.x - value.y) |> 
  arrange(loss_diff) |> 
  left_join(two_tumor_pfs_res |> spread_rvars(rep_pfs[i], rep_right_censored[i]), by = c("id" = "i")) |> 
  left_join(no_tumor_pfs_res |> spread_rvars(rep_pfs[i], rep_right_censored[i]), by = c("id" = "i"), suffix = c("_two_tumor", "_no_tumor")) |> 
  bind_cols(select(treated_pfs_analysis_data, true_pfs = pfs)) |> 
  mutate(across(starts_with("rep_pfs"), \(p) p - true_pfs, .names = "centered_{.col}")) 
```


## What are the candidate models we are comparing?

* *Baseline Model*: No covariates for any of the patient or tumor characteristics.
* *Linear Model*: Using a linear model of the lesion diameters (early assessments^[The last pre-screening assessment and the first assessment during treatment.]).
* *Proportional Change Model*: Using a model of proportional change in lesion diameters (early assessments).

## Estimated survival 

* No visual difference in estimated survival between candidate models.

::: {.panel-tabset}

### Survival Curves

```{r}
#| label: fig-survival-curves
#| fig-cap: Analysis is restricted to subjects who survived up to the first assessment first after treatment started. Treatment epoch starts at t > 0. Ribbon shows the 80% credible interval. 
#| cache: true

lst(no_tumor = no_tumor_pfs_res, two_tumor = two_tumor_pfs_res, tumor_change = tumor_change_pfs_res) |> 
  plot_km(treated_pfs_analysis_data, facet_arm = TRUE) +
  geom_step(aes(x = t, y = s, group = btype, linetype = "Kaplan-Meier"), data = treated_obs_km |> filter(fct_match(btype, "lb"))) +
  scale_linetype_manual("", values = c("Kaplan-Meier" = "dotted")) + 
  labs(title = "", subtitle = "") 
```

### Survival Histogram 

```{r}
#| label: fig-pfs-hist-post
#| cache: true

lst(no_tumor = no_tumor_pfs_res, two_tumor = two_tumor_pfs_res) |> 
  compact() |> 
  plot_pfs_hist_posterior(treated_pfs_stan_data) +
  stat_bin(aes(pfs, color = "All"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data) +
  stat_bin(aes(pfs, color = "Uncensored"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data |> filter(!right_censored)) +
  scale_color_discrete("Observed Survival") +
  theme(legend.position = "bottom") +
  NULL
```

:::

## Hazard ratio

::: {.panel-tabset}

### Overall Hazard Ratio

* Captures the variation in risk conditional on early assessments.
* Histogram captures variations due to both (a) the effect of lesions and (b) uncertainty.

```{r}
#| label: fig-predictor-variation
#| fig-cap: Hazard ratio, for the linear model. Line intervals identify the 50%, 80%, and 90% credible intervals.
#| cache: true
#| cache-lazy: false 

median_coef <- lst(post = two_tumor_pfs_res) |> 
  map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |> 
  map_dfr(\(f) spread_rvars(f, tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, k]), .id = "fit") |> 
  mutate(across(starts_with("tumor_stim"), median)) |> 
  pivot_wider(values_from = tumor_stim_trial_coef, names_prefix = "tumor_stim_trial_coef", names_from = k)

tumor_design_matrix <- with(
  treated_pfs_stan_data, {
    n_screening_t <- pfs_model$functions$calc_n_screening_t(n_patient_tumors, n_measures, t_measure)
    
    pfs_model$functions$standardize_nonzero_tumor_sizes(tumor_size) |>
      pluck(3) |> 
      pfs_model$functions$prepare_early_tumors_design_matrix(n_patient_tumors, n_measures, n_screening_t)
  }
)

tumor_design_matrix <- cbind(1, tumor_design_matrix, tumor_design_matrix[,1] * tumor_design_matrix[,2])

patient_kappa_data <- tibble(kappa = tumor_design_matrix, trial = with(treated_pfs_stan_data, rep(patient_trial, n_patient_tumors))) |>  
  left_join(median_coef, by = "trial") %>% 
  mutate(
    kappa = rowSums(kappa * select(., starts_with("tumor_stim"))),
    patient_index = rep(seq(treated_pfs_stan_data$n_patients), treated_pfs_stan_data$n_patient_tumors)
  ) |> 
  group_by(patient_index) |> 
  summarize(kappa = exp(-first(tumor_stim_trial_intercept) * mean(treated_pfs_stan_data$n_patient_tumors) + sum(kappa)))

patient_kappa_full_dist_data <- two_tumor_pfs_res |> 
  spread_rvars(total_time_invar_tumor_stim[patient_index]) |> 
  mutate(kappa = exp(total_time_invar_tumor_stim))

patient_kappa_data |> 
  mutate(var = "median") |> 
  bind_rows(
    patient_kappa_full_dist_data |>
      unnest_rvars() |> 
      mutate(var = "post")
  ) |> 
  filter(fct_match(var, "post")) |> 
  ggplot(aes(kappa)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  stat_histinterval(alpha = 0.5, breaks = ggdist::breaks_fixed(width = 0.1), point_size = 0, .width = c(0.5, 0.8, 0.9)) +
  #geom_histogram(aes(y = after_stat(density)), fill = "black", position = "identity", alpha = 0.25, binwidth = 0.1) +
  # scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(x = latex2exp::TeX(r"{$\kappa$}"), y = "Density") +
  coord_cartesian(xlim = c(0, 5)) +
  NULL
```

### Sample of Patient Hazard Ratios

```{r, fig.height=6, fig.width=8}
#| label: fig-patient-hazard-ratios
#| fig-cap: Hazard ratios for the 100 patients with the greatest increase. Line intervals identify the 50% and 80% credible intervals.
#| fig-height: 4 

loss_diff_data |> 
  left_join(patient_kappa_full_dist_data, by = c("id" = "patient_index"), relationship = "one-to-one") |> 
  mutate(
    median_log_kappa = median(total_time_invar_tumor_stim),
    diff_type = loss_diff <= 0.0,
  ) |> 
  # arrange(diff_type, median_log_kappa) |> 
  arrange(median(kappa)) |> 
  # head(200) |> 
  group_by(trial) |> 
  mutate(id = seq(n())) |> 
  ungroup() |> 
  ggplot() +
  stat_pointinterval(aes(xdist = kappa, y = id), point_size = 0.5, alpha = 0.5, linewidth = 0.125, .width = c(0.8)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_continuous(latex2exp::TeX(r"{Hazard Ratio ($\kappa$)}"), breaks = seq(0, 7)) +
  labs(y = "Patients") +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), axis.text.y = element_blank()) +
  facet_grid(vars(trial), space = "free_y", scales = "free_y") +
  # coord_cartesian(xlim = c(0, 6.5)) +
  coord_trans(x = "log") +
  NULL
```

:::

## Challenges in assessing predictive accuracy

* Qualitative visual inspection of survival curves is only helpful in the case of very strong signals.
* Posterior plots of hazard ratios confound model sensitivity to lesion sizes with uncertainty.
* Evaluating model accuracy in predicting outcomes used to fit model bias results.
* It is not always clear what level of improved accuracy is sufficient. 

## Out-of-sample prediction accuracy

* In order to assess prediction accuracy we need assess prediction of _future data_.
* In the absence of such true _out-of-sample_ data, we utilize _leave-one-out cross-validation (LOO-CV)_: 
    1. Iteratively, fit the model holding out a single subject,
    2. Predict survival of the held-out subject, and
    3. Calculate the prediction error.
    4. Take an average of the prediction errors.
    
## How do we calculate prediction error?

* **Root mean squared error (RMSE)**: Places an increasingly high penalty on larger errors.
* **Log likelihood**: Use the probability of data conditional on estimated model parameters.
* **Linear loss function**: Use an explicit function capturing the seriousness of errors in our context.

While the first two accuracy measures are well known, we need to spend more time on the loss functions.

## Root mean squared error loss function

* We see that in terms of RMSE the linear model outperforms the baseline model for all trials. 

```{r, fig.height=2, fig.width=8}
#| label: fig-rmse
#| fig-cap: Comparing RMSE for the baseline and linear models, separated by trials.

loss_data |> 
  filter(fct_match(loss_type, "sqerr")) |> 
  group_by(model, trial) |> 
  summarize(rmse = sqrt(mean(value)), .groups = "drop") |> 
  ggplot() +
  geom_col(aes(y = trial, x = rmse, fill = model), position = "dodge", alpha = 0.5) +
  scale_fill_discrete("Model", labels = c(no_tumor = "Baseline", two_tumor = "Linear Lesion Size")) +
  labs(x = "RMSE", y = "Trial")
```

## Log likelihood scoring function

* We see that the linear model out performs the baseline model.
* We include a version of the linear model that does not include lesion sizes, just their number.
* Accuracy improvement appears to be driven by the number of lesions and not really by their size. 

```{r, fig.height=2, fig.width=8}
#| label: fig-loo-cv
#| fig-cap: Expected log likelihood for the various model. Error bars characterize the standard errors since we rely on approximation to calculate LOO-CV.
#| fig-height: 2
#| fig-width: 8

loo_compare(lst("Baseline" = no_tumor_loo, "Tumor Count Only" = tumor_count_loo, "Linear Lesion Size" = two_tumor_loo))[, c("elpd_loo", "elpd_diff", "se_diff")] |> 
  as_tibble(rownames = "model") |>
  mutate(se_diff = if_else(se_diff > 0, se_diff, NA)) |>  
  ggplot(aes(y = model)) +
  geom_col(aes(elpd_loo), alpha = 0.5) +
  geom_errorbar(aes(xmin = elpd_loo - se_diff, xmax = elpd_loo + se_diff), width = 0.25) +
  labs(x = "Expected log likelihood", y = "Model") +
  coord_cartesian(xlim = c(-1125, -1000)) 
```


## Linear loss function

:::: {.columns}

::: {.column width="50%"}

* This is simple measure of the difference between the predicted and actual PFS, accounting for censoring.
* Unlike RMSE we do not increasingly penalize high prediction error.
* We consider 
  - **Symmetric loss function**: under- and over-estimation are not treated differently.
  - **Asymmetric loss function**: Over-estimation of survival is pnealized higher than under-estimation -- 3 times higher.

:::

::: {.column width="50%"}

```{r}
#| label: fig-linear-loss
#| fig-cap: Linear loss functions, one symmetric and the other asymmetric, placing greater loss on overestimation of survival.
#| fig-width: 5
#| fig-height: 3

tibble(delta = seq(-10, 10), sym_loss = abs(delta), asym_loss = if_else(delta < 0, 3, 1) * abs(delta)) |> 
  pivot_longer(sym_loss:asym_loss) |> 
  ggplot(aes(delta, value)) +
  geom_line(aes(y = value, color = name)) +
  scale_color_discrete("Loss Type", labels = c("sym_loss" = "Symmetric", "asym_loss" = "Asymmetric")) +
  labs(x = latex2exp::TeX(r"{Prediction Error}"), y = "Loss") +
  NULL
```

:::

::::

## Expected Loss 

```{r}
#| label: loss-table

loss_data |> 
  group_by(loss_type, model, trial) |> 
  summarize(expected_loss = mean(value), .groups = "drop") |> 
  mutate(expected_loss = if_else(fct_match(loss_type, "sqerr"), sqrt(expected_loss), expected_loss)) |> 
  pivot_wider(names_from = trial, values_from = expected_loss) |> 
  mutate(
    loss_type = fct_recode(loss_type, "Asymmetric" = "asym", "Symmetric" = "sym", "RMSE" = "sqerr") |> 
      fct_relevel("RMSE", after = Inf),
    model = fct_recode(model, "No Tumor Response" = "no_tumor", "Linear Response" = "two_tumor"),
  ) |> 
  arrange(loss_type) |> 
  rename("Loss Type" = loss_type, "Model" = model) |>
  kableExtra::kable(digits = 2) |> 
  kableExtra::kable_styling(font_size = 15)
```


<!-- ```{r} -->
<!-- #| label: pfs-loo -->
<!-- #| cache: true -->

<!-- loo_compare(lst("No Tumor Response" = no_tumor_loo, "Tumor Count Only" = tumor_count_loo, "Linear Response" = two_tumor_loo, "Proportional Change Response" = tumor_change_loo))[, c("elpd_loo", "elpd_diff", "se_diff")] |>  -->
<!--   kableExtra::kable( -->
<!--     col.names = c("Expected Log Pointwise Predictive Density", "Difference", "SE"),  -->
<!--     digits = 2,  -->
<!--     label = c("Linear Model", "Proportional Change", "No Tumor Response") -->
<!--   ) |>  -->
<!--   kableExtra::kable_styling() -->
<!-- ``` -->


<!-- # Introduction -->

<!-- ## Question -->

<!-- This is an exploratory analysis aimed at answering the questions: -->

<!-- * In general: Are there observable strong and deep early responses to antibody-drug conjugates (ADCs) that are predictive of long term progression free survival (PFS)? -->
<!-- * Specifically: Are observable target tumor sizes measured at the pre-screening and the first post-treatment visits predictive of PFS? -->

<!-- ## Why? -->

<!-- * Predict patients' response to ADCs based on early proxy outcomes to inform treatment decisions. -->
<!-- * Inform decisions on if and when parallel/overlapping studies should be launched. -->
<!-- * Inform exclusion/inclusion decisions on follow-up studies. -->

<!-- ## How? {.smaller} -->

<!-- * Conducted a Bayesian hierarchical survival meta-analysis of four clinical trials evaluating ENHERTU (Trastuzumab Deruxtecan). -->
<!-- * Allowed patient hazard to respond to tumor sizes in a non-aggregated manner.  -->
<!-- * Conducted an out-of-sample evaluation of prediction accuracy and compared to a model with no response to tumor sizes. -->
<!-- * Conducted qualitative assessment of model sensitivity to tumor sizes. -->

<!-- ## Data {.smaller}  -->

<!-- 1. Breast01: phase II dose finding study (HER2-positive metastatic breast cancer). -->
<!-- 2. Breast02: phase III study comparing to physician choice of drug (HER2-positive metastatic breast cancer).  -->
<!-- 2. Breast03^[Pending]: phase III study comparing to trastuzumab emtansine (HER2-positive metastatic breast cancer). -->
<!-- 3. Breast04: This is a phase III study on patients with HER2-low metastatic breast cancer (HER2-low metastatic breast cancer). -->

<!-- * We evaluated survival with and without response to tumor sizes  -->

<!-- ## Observed Survival (Kaplan-Meier) -->

<!-- ```{r} -->
<!-- #| label: fig-survival-dynamics -->
<!-- #| warning: false -->

<!-- km_res_calculated <- pfs_analysis_data |> -->
<!--   group_by(trial, treated) |>   -->
<!--   group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "pfs"), .keep = TRUE) |>   -->
<!--   bind_rows()   -->

<!-- km_res_from_data <- pfs_analysis_data |> -->
<!--   group_by(trial, treated) |>   -->
<!--   group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "progress_week"), .keep = TRUE) |>   -->
<!--   bind_rows()   -->

<!-- km_res <- bind_rows(calculated = km_res_calculated, from_data = km_res_from_data, .id = "pfs_type") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| labels: kaplan-meier-upper-and-lower-bounds -->

<!-- km_res |>  -->
<!--   filter(fct_match(pfs_type, "calculated")) |> -->
<!--   ggplot() + -->
<!--   geom_step(aes(x = t, y = s, color = btype)) + -->
<!--   scale_color_discrete("Survival up to", labels = c(lb = "First possible week", ub = "Last possible week")) + -->
<!--   labs(y = "S(t)") + -->
<!--   facet_grid(vars(treated), vars(trial), scales = "free_x", labeller = labeller(treated = label_both)) + -->
<!--   guides(color = "none") + -->
<!--   NULL -->
<!-- ``` -->

<!-- ::: aside -->

<!-- We show both the lower and upper bounds for PFS, given interval censoring.  -->

<!-- ::: -->

<!-- # Analysis  -->

<!-- ```{r} -->
<!-- #| label: survival-fit -->
<!-- #| cache: true -->
<!-- #| cache-lazy: false -->

<!-- # prior_no_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^prior_no_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE) -->
<!-- no_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^no_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE) -->
<!-- tumor_count_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^tumor_count_pfs-", full.names = TRUE), check_diagnostics = FALSE) -->
<!-- # prior_two_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^prior_two_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE) -->
<!-- two_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^two_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE) -->
<!-- tumor_change_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^tumor_change_pfs-", full.names = TRUE), check_diagnostics = FALSE) -->
<!-- ``` -->


<!-- ## Survival Model {.smaller} -->

<!-- * We use a piecewise-constant proportional hazard survival model. -->
<!-- * Survival was grouped over weekly intervals. -->
<!-- * Base hazard is modeled as a Gaussian process smoothed over intervals. -->
<!-- * Patient time-invariant hazard is modeled as a linear model of  -->
<!--     i) of the two first tumor size assessments, plus their interaction -->
<!--     ii) of the proportional change in each tumor's size -->
<!-- * Each tumor individually to the a patient's hazard -- tumor sizes are not aggregated. -->
<!-- * We adjust for both right and interval censoring. -->
<!-- * All parameters are modeled hierarchically to share information across trials. -->

<!-- ## Survival Plot -->

<!-- ```{r} -->
<!-- #| label: fig-no-tumor-kaplan-meier -->
<!-- #| cache: true -->

<!-- lst(no_tumor = no_tumor_pfs_res, two_tumor = two_tumor_pfs_res, tumor_change = tumor_change_pfs_res) |>  -->
<!--   plot_km(treated_pfs_analysis_data, facet_arm = TRUE) + -->
<!--   geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) + -->
<!--   labs(title = "", subtitle = "")  -->
<!-- ``` -->

<!-- ::: aside -->

<!-- Restricted to the treated arms with the same dosage. -->

<!-- ::: -->

<!-- ## Two Tumor Linear Model Parameters  -->

<!-- ```{r} -->
<!-- #| label: two-tumor-coef -->
<!-- #| cache: true -->

<!-- lst(post = two_tumor_pfs_res) |>  -->
<!--   map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |>  -->
<!--   map_dfr( -->
<!--     \(f) gather_rvars(f, tumor_stim_intercept, tumor_stim_coef[t], tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, t]),  -->
<!--     .id = "fit" -->
<!--   ) |>  -->
<!--   mutate( -->
<!--     .variable = if_else(is.na(trial), .variable, str_glue("{.variable}[{trial}]")), -->
<!--     .variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]")), -->
<!--   ) |>  -->
<!--   ggplot() + -->
<!--   stat_halfeye(aes(xdist = .value, y = .variable), size = 0.5, linewidth = 0.9, fill = alpha("black", 0.25), color = "black", .width = c(0.8)) + -->
<!--   geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   labs(x = "", y = "Tumor Model Parameters") + -->
<!--   coord_cartesian(xlim = c(0, 0.5)) -->
<!-- ``` -->

<!-- # Model Comparison -->

<!-- ## Approximate Leave-One-Out Cross Validation {#sec-loo-cv} -->

<!-- ```{r} -->
<!-- #| label: pfs-loo -->
<!-- #| cache: true -->

<!-- no_tumor_loo <- no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12) -->
<!-- tumor_count_loo <- tumor_count_pfs_res$loo(save_psis = TRUE, cores = 12) -->
<!-- two_tumor_loo <- two_tumor_pfs_res$loo(save_psis = TRUE, cores = 12) -->
<!-- tumor_change_loo <- tumor_change_pfs_res$loo(save_psis = TRUE, cores = 12) -->

<!-- loo_compare(lst("No Tumor Response" = no_tumor_loo, "Tumor Count Only" = tumor_count_loo, "Linear Response" = two_tumor_loo, "Proportional Change Response" = tumor_change_loo))[, c("elpd_loo", "elpd_diff", "se_diff")] |>  -->
<!--   kableExtra::kable( -->
<!--     col.names = c("Expected Log Pointwise Predictive Density", "Difference", "SE"),  -->
<!--     digits = 2,  -->
<!--     label = c("Linear Model", "Proportional Change", "No Tumor Response") -->
<!--   ) |>  -->
<!--   kableExtra::kable_styling() -->
<!-- ``` -->


<!-- ## Comparison of Out-of-Sample Loss {.smaller} -->


<!-- :::: {.columns} -->

<!-- ::: {.column width="40%"} -->

<!-- * From this analysis we see that including tumors does reduce expected loss. -->
<!-- * At the highest end, expected loss is reduced by ~7 weeks asymmetric loss (~2 week symmetric loss). -->
<!-- * However, as we saw in the [cross validation](#sec-loo-cv) results, this reduction in loss appears to be driven by the number of tumors rather than their sizes. -->

<!-- ::: -->

<!-- ::: {.column width="60%"} -->

<!-- ::: {.panel-tabset} -->


<!-- ### CDF  -->

<!-- ```{r} -->
<!-- #| label: fig-loss-cdf -->
<!-- #| fig-width: 5 -->
<!-- #| fig-cap: Empirical CDF for patients' mean loss. -->

<!-- loss_data |>  -->
<!--   mutate(loss_type = fct_recode(loss_type, "Asymmetric" = "asym", "Symmetric" = "sym")) |>  -->
<!--   ggplot(aes(value)) + -->
<!--   stat_ecdf(aes(color = model)) + -->
<!--   scale_color_discrete("Model", labels = c(no_tumor = "No Tumor Response", two_tumor = "Linear Response"), aesthetics = c("color", "fill")) + -->
<!--   labs(x = "Loss", y = "CDF") + -->
<!--   facet_grid(vars(trial), vars(loss_type), scales = "free", margins = "trial") + -->
<!--   theme(legend.position = "top", strip.text.y = element_text(angle = 0)) + -->
<!--   NULL -->
<!-- ``` -->

<!-- ### Histogram  -->

<!-- ```{r} -->
<!-- #| label: fig-loss-hist -->
<!-- #| fig-width: 5 -->
<!-- #| fig-cap: Histogram of patients' mean loss. -->

<!-- loss_data |>  -->
<!--   mutate(loss_type = fct_recode(loss_type, "Asymmetric" = "asym", "Symmetric" = "sym")) |>  -->
<!--   plot_loss(binwidth = 6) + -->
<!--   scale_color_discrete("Model", labels = c(no_tumor = "No Tumor Response", two_tumor = "Linear Response"), aesthetics = c("color", "fill")) -->
<!-- ``` -->

<!-- ::: -->

<!-- ::: -->

<!-- :::: -->

<!-- # Tumor Size Sensitivity -->

<!-- ## Variation in Proportional Hazard in Response to Tumor Size -->

<!-- How much does the patient specific component of the proportional hazard respond to tumor sizes in the data? -->

<!-- ```{r} -->
<!-- #| label: predictor-variation -->
<!-- #| cache: true -->
<!-- #| cache-lazy: false  -->

<!-- median_coef <- lst(post = two_tumor_pfs_res) |>  -->
<!--   map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |>  -->
<!--   map_dfr(\(f) spread_rvars(f, tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, k]), .id = "fit") |>  -->
<!--   mutate(across(starts_with("tumor_stim"), median)) |>  -->
<!--   pivot_wider(values_from = tumor_stim_trial_coef, names_prefix = "tumor_stim_trial_coef", names_from = k) -->

<!-- tumor_design_matrix <- with( -->
<!--   treated_pfs_stan_data, { -->
<!--     n_screening_t <- pfs_model$functions$calc_n_screening_t(n_patient_tumors, n_measures, t_measure) -->

<!--     pfs_model$functions$standardize_nonzero_tumor_sizes(tumor_size) |> -->
<!--       pluck(3) |>  -->
<!--       pfs_model$functions$prepare_early_tumors_design_matrix(n_patient_tumors, n_measures, n_screening_t) -->
<!--   } -->
<!-- ) -->

<!-- tumor_design_matrix <- cbind(1, tumor_design_matrix, tumor_design_matrix[,1] * tumor_design_matrix[,2]) -->

<!-- patient_kappa_data <- tibble(kappa = tumor_design_matrix, trial = with(treated_pfs_stan_data, rep(patient_trial, n_patient_tumors))) |>   -->
<!--   left_join(median_coef, by = "trial") %>%  -->
<!--   mutate( -->
<!--     kappa = rowSums(kappa * select(., starts_with("tumor_stim"))), -->
<!--     patient_index = rep(seq(treated_pfs_stan_data$n_patients), treated_pfs_stan_data$n_patient_tumors) -->
<!--   ) |>  -->
<!--   group_by(patient_index) |>  -->
<!--   summarize(kappa = exp(-first(tumor_stim_trial_intercept) * mean(treated_pfs_stan_data$n_patient_tumors) + sum(kappa))) -->

<!-- patient_kappa_full_dist_data <- two_tumor_pfs_res |>  -->
<!--   spread_rvars(total_time_invar_tumor_stim[patient_index]) |>  -->
<!--   mutate(kappa = exp(total_time_invar_tumor_stim)) -->

<!-- patient_kappa_data |>  -->
<!--   mutate(var = "median") |>  -->
<!--   bind_rows( -->
<!--     patient_kappa_full_dist_data |> -->
<!--       unnest_rvars() |>  -->
<!--       mutate(var = "post") -->
<!--   ) |>  -->
<!--   filter(fct_match(var, "post")) |>  -->
<!--   ggplot(aes(kappa)) + -->
<!--   geom_vline(xintercept = 1, linetype = "dotted") + -->
<!--   geom_histogram(aes(y = after_stat(density)), fill = "black", position = "identity", alpha = 0.25, binwidth = 0.1) + -->
<!--   # scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) + -->
<!--   labs(x = latex2exp::TeX(r"{$\kappa$}"), y = "Density") + -->
<!--   coord_cartesian(xlim = c(0, 5)) + -->
<!--   NULL -->
<!-- ``` -->

<!-- ## Variation in Probability of Disease Progression -->

<!-- ```{r} -->
<!-- #| label: dp-prob-variation -->
<!-- #| cache: true -->

<!-- two_tumor_pfs_res |>  -->
<!--   spread_rvars(log_lambda[t]) |>  -->
<!--   mutate( -->
<!--     lambda = exp(log_lambda), -->
<!--     median_baseline_dp_prob = 1 - exp(- median(lambda)), -->
<!--     patient_dp_prob = map(lambda, \(l) 1 - exp(- median(l) * patient_kappa_data$kappa)) %>%  -->
<!--       { array(unlist(.), dim = c(nrow(patient_kappa_data), length(.))) } |>  -->
<!--       rvar(), -->

<!--     patient_full_dist_dp_prob = map(lambda, \(l) draws_of(1 - exp(- median(l) * patient_kappa_full_dist_data$kappa))) %>% -->
<!--       { array(unlist(.), dim = c(prod(dim(first(.))), length(.))) } |>  -->
<!--       rvar() -->
<!--   ) |>  -->
<!--   ggplot(aes(t)) + -->
<!--   # geom_line(aes(y = median_baseline_dp_prob)) + -->
<!--   # Why didn't I use pivot? We can't have different sized rvars in the same column. -->
<!--   # stat_lineribbon(aes(ydist = patient_dp_prob, fill = "median"), alpha = 0.25, .width = 0.8) + -->
<!--   stat_lineribbon(aes(ydist = patient_full_dist_dp_prob), fill = "black", alpha = 0.25, .width = 0.8) + -->
<!--   scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) + -->
<!--   labs(y = latex2exp::TeX(r"{1 - \alpha}")) + -->
<!--   NULL -->
<!-- ``` -->

<!-- # Discussion -->

<!-- ## Limitations  -->

<!-- * We are currently using only the first two tumor assessments to make predictions about survival at every interval. -->
<!--     - Consider using time-variant tumor sizes; capturing both the dynamics of tumor change and hazard. -->
<!-- * The main questions are ultimately causal while we are not treating them as in this analysis.  -->

<!-- ## Conclusion -->

<!-- * Cross validation shows that  -->
<!--     - A tumor response model is more predictive than a no tumor response model. -->
<!--     - Tumor response intercepts seem to be the main predictive driver of this model. -->
<!-- * Expected loss is the lowest for models that respond to tumor sizes. -->