---
title: "Early Prediction of Breast Cancer Progress Free Survival"
format: revealjs
---

```{r}
#| label: setup
#| include: false

# library(conflicted)
library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(gtsummary)
library(targets)

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

knitr::opts_chunk$set(cache.path = file.path("/wscratch", Sys.getenv("USER"), "adc-early-predict/cache/"))

options(mc.cores = 4)

ggplot2::theme_set(theme_minimal())

source(here("src", "util.R"))
```

```{r}
#| label: priors
#| cache: true

tumor_priors <- lst(
  pop_tumor_gp_rho_alpha = 7.3,
  pop_tumor_gp_rho_beta = 7.5, 
)

pfs_priors <- lst(
  log_lambda_gp_intercept_mean = -1,
  log_lambda_gp_intercept_sd = 0.25,
  log_lambda_gp_alpha_sd = 0.25,
  log_lambda_gp_rho_alpha = 7.3,
  log_lambda_gp_rho_beta = 7.5, 
  tumor_stim_intercept_sd = 0.125,
  log_lambda_gp_trial_alpha_sd = 0.0625,
  log_lambda_gp_trial_intercept_sd_sd = 0.25,
  tumor_stim_coef_sd = c(0.125, 0.125, 0.0625) / 2,
  tumor_stim_trial_coef_sd_sd = 0.0625
)
```

```{r}
#| label: models
#| code-fold: false   

tumor_model <- cmdstan_model(here("src", "tumor.stan"))
pfs_model <- cmdstan_model(here("src", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)
```

```{r}
#| label: load-trials-data
#| cache: true

data_details <- tribble(
  ~ idap,          ~ trial, 
  "idap_20231204", "Breast01",
  "idap_20231129", "Breast02",  
  "idap_20231124", "Breast04"
) |> 
  mutate(across(everything(), factor))

# Trial arms
sdtm_ta <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_ta", "sdtm")))

# Trial elements 
sdtm_te <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_te", "sdtm")))

# Protocol deviations
# sdtm_dv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dv.parquet") |> 
#   rename_with(str_to_lower)

# Inclusion/exclusion
# sdtm_ie <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_ie.parquet") |> 
#   rename_with(str_to_lower)

# Demographics 
sdtm_dm <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_dm", "sdtm")))

# Death details 
# sdtm_dd <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dd.parquet") |> 
#   rename_with(str_to_lower)

# Exposure 
sdtm_ex <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_ex", "sdtm")))

# Tumor identification
# sdtm_tu <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tu.parquet") |> 
#   rename_with(str_to_lower)

# Trial visits 
# sdtm_tv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tv.parquet") |> 
#   rename_with(str_to_lower)

# Subject visits
# sdtm_sv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_sv.parquet") |> 
#   rename_with(str_to_lower)

# Tumor measurements/assessments
# sdtm_tr <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tr.parquet") |>
# sdtm_tr <- read_rds("/scratch/ewfteams/dpo0083/sdtm/prod/data/deid_tr.rds") |> 
  # rename_with(str_to_lower) |> 
  # mutate(tracptfl = fct_match(tracptfl, "Y"))

# Disease response 
# sdtm_rs <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_rs.parquet") |> 
#   rename_with(str_to_lower)

# Tumor analysis_data
adam_adtr <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adtr", "adam")))
  
# Tumor analysis_data
# adam_adef <- map_dfr(idaps, \(idap) read_rds(file.path(team_dir, idap, "adam/prod/data/deid_adef.rds")), .id = "idap") |> 
#   rename_with(str_to_lower) |> 
#   mutate(across(ends_with("fl"), \(fl) fct_match(fl, "Y")))

# Time to event
adam_adtte <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adtte", "adam")) |> 
           map(\(d) mutate(d, cnsr = cnsr == 1)))
```

```{r}
#| label: prepare-analysis-data
#| cache: true
#| dependson: load-trials-data
#| code-fold: false 

tumor_analysis_data <- adam_adtr |> 
  unnest(data) |> 
  filter(
    # fct_match(paramcd, c("DIAMETER", "DIAM")),
    fct_match(paramcd, "LDIAM"),
    # There are a number of assessments that show up in the data but in SDTM they are marked as "NOT DONE"
    !is.na(trdy), !is.na(aval),
    tracptfl,
    # adt <= trtedt, # Drop assessments occurring after treatment ended.
    aval > 0, # Drop all zero diameter assessments; these appear to be missing measurements. 
  ) %>% 
  select(studyid, idap, trial, 
         usubjid, trlnkid, tuloc, tulat, trtsdt, trtedt, adt, ends_with("fl"), trdy, trtp, trta, visit, visitnum, mmdiam = aval) |> 
  mutate(
    week = if_else(trdy > 0, trdy, trdy + 1) %/% 7, 
    treated_week = adt >= trtsdt,
  ) |> 
  arrange(usubjid, trlnkid, visitnum) |> 
  group_by(usubjid, trlnkid) |> 
  mutate(tuloc = first(tuloc), tulat = first(tulat) |> na_if("")) |> # Only the first assessment has the location of the tumor. Fix that.
  ungroup() |>
  # Turns out that some assessments are not listed in order of the days they were done. Look at subject DS8201-A-U303-10033028.
  arrange(studyid, usubjid, trlnkid, trdy) |> 
  nest(tumor_history = c(adt, trdy, week, visit, visitnum, tsplitfl, tmergefl, aocc01fl, mmdiam, treated_week)) |> 
  mutate(# For some weeks, there are multiple (unscheduled) visits that might happen. I'm just going to use the average for tumor sizes within those
         # weeks.
    tumor_history = map(
      tumor_history, 
      \(h) nest(h, visits = c(adt, trdy, visit, visitnum, mmdiam)) |> 
        mutate(
          n_week_visits = map_int(visits, nrow),
          mmdiam = map_dbl(visits, \(v) mean(v$mmdiam))
        )
    ),
    n_tumor_measures = map_int(tumor_history, nrow),
    
    min_t = map_int(tumor_history, \(h) min(h$week)),
    max_t = map_int(tumor_history, \(h) max(h$week)),
  ) |> 
  pack(flags = ends_with("fl")) |> 
  nest(patient_tumors = !c(studyid, trial, idap, usubjid, trtsdt, trtedt, trtp, trta)) |> 
  rowwise() |> 
  mutate(
    patient_min_t = min(patient_tumors$min_t),
    patient_max_t = max(patient_tumors$max_t),
  ) |>
  ungroup() |> 
  mutate(
    patient_t_width = patient_max_t - patient_min_t + 1,
    n_tumors = map_int(patient_tumors, nrow),
  ) |> 
  left_join(
    adam_adtte |> 
      unnest(data) |> 
      filter(
        (fct_match(paramcd, "PFS1") & fct_match(studyid, c("DS8201-A-U301", "DS8201-A-U303"))) | 
          (fct_match(paramcd, "PFS") & fct_match(studyid, "DS8201-A-U201")), 
        fct_match(parqual, "CENTRAL")
      ) |> 
        select(studyid, usubjid, aval, death = dthfl, dthdt, fpddt, cnsr),
    by = c("studyid", "usubjid")
  ) |> 
  mutate(
    right_censored = cnsr,
    death_week = floor(time_length(dthdt - trtsdt, unit = "weeks")),
    progress_week = floor(time_length(fpddt - trtsdt, unit = "weeks")),
    progress_week = if_else(death, coalesce(progress_week, death_week), 
                                   if_else(right_censored & is.na(progress_week), max(progress_week, na.rm = TRUE), progress_week)),
    aval_week = aval %/% 7,
    
    treated = 
      (fct_match(trial, "Breast01") & fct_match(trtp, "5.4 mg/kg")) | 
      (fct_match(trial, c("Breast02", "Breast04")) & fct_match(trtp, "T-DXd")),
    
    # The model expects "pfs" to record the number of weeks that were progress free, so I need to find the last assessment week before
    # disease progress was detected.
    pfs = pmap_int(
      lst(progress_week, death_week, right_censored, patient_tumors), function(progress_week, death_week, right_censored, patient_tumors) {
        pre_progress_weeks <- unnest(patient_tumors, tumor_history) |>
          distinct(week) |>
          filter(right_censored | week < min(progress_week, death_week, na.rm = TRUE)) |>
          pull(week)

        # There are a few patients who just have a single post treatment visit
        if (length(pre_progress_weeks) > 0) max(pre_progress_weeks) else NA_integer_
      }
    ),

    death_week = coalesce(death_week, 0L),
    
    n_measures = map(patient_tumors, \(tu) tu$n_tumor_measures),
    t_measure = map(patient_tumors, \(tu) map(tu$tumor_history, \(tr) pull(tr, week))),
  ) |> 
  left_join(select(unnest(sdtm_dm, data), studyid, usubjid, country), by = c("studyid", "usubjid")) %>%
  mutate(
    interval_censored = with(., pfs_model$functions$identify_censoring(pfs, death_week, n_tumors, unlist(n_measures), unlist(t_measure))[[1]]) |> 
      pmax(0) # Issue https://github.com/azu-oncology-rd/ods-adc-early-predict-2023/issues/36
  )
  
pfs_analysis_data <- tumor_analysis_data |> 
  filter(
    !is.na(pfs), pfs >= 0, # No missing pfs and progress must be assured to have happened after treatment
    patient_min_t <= 0, # At least one screening assessment needed,
    patient_max_t > 0, # and at least one post treatment assessment.
  )

treated_pfs_analysis_data <- pfs_analysis_data |> 
  filter(treated) 

treated_pfs_stan_data <- prepare_pfs_stan_data(treated_pfs_analysis_data, tumor_priors, pfs_priors)

treated_trial_pfs_stan_data <- treated_pfs_analysis_data |> 
  nest(.by = trial) |> 
  deframe() |> 
  imap(\(d, tr) prepare_pfs_stan_data(mutate(d, trial = tr), tumor_priors, pfs_priors))

treated_obs_km <- map_dfr(treated_trial_pfs_stan_data, function(sdata) {
  with(sdata, {
    list(lb = pfs, ub = pfs + interval_censored) |> 
      map_dfr(function(s) {
        pfs_model$functions$estimate_kaplan_meier(s, right_censored, max(s)) |>
          set_names(c("s", "n", "c", "e")) |>
          as_tibble() |> 
          mutate(t = seq(0, n() - 1))
      }, .id = "btype")
    }
  )
}, .id = "trial")
```

# Introduction

## Question

This is an exploratory analysis aimed at answering the questions:

* In general: Are there observable strong and deep early responses to antibody-drug conjugates (ADCs) that are predictive of long term progression free survival (PFS)?
* Specifically: Are observable target tumor sizes measured at the pre-screening and the first post-treatment visits predictive of PFS?

## Why?

* Predict patients' response to ADCs based on early proxy outcomes to inform treatment decisions.
* Inform decisions on if and when parallel/overlapping studies should be launched.
* Inform exclusion/inclusion decisions on follow-up studies.

## How? {.smaller}

* Conducted a Bayesian hierarchical survival meta-analysis of four clinical trials evaluating ENHERTU (Trastuzumab Deruxtecan).
* Allowed patient hazard to respond to tumor sizes in a non-aggregated manner. 
* Conducted an out-of-sample evaluation of prediction accuracy and compared to a model with no response to tumor sizes.
* Conducted qualitative assessment of model sensitivity to tumor sizes.

## Data {.smaller} 

1. Breast01: phase II dose finding study (HER2-positive metastatic breast cancer).
2. Breast02: phase III study comparing to physician choice of drug (HER2-positive metastatic breast cancer). 
2. Breast03^[Pending]: phase III study comparing to trastuzumab emtansine (HER2-positive metastatic breast cancer).
3. Breast04: This is a phase III study on patients with HER2-low metastatic breast cancer (HER2-low metastatic breast cancer).

* We evaluated survival with and without response to tumor sizes 

## Observed Survival (Kaplan-Meier)

```{r}
#| label: fig-survival-dynamics
#| warning: false

km_res_calculated <- pfs_analysis_data |>
  group_by(trial, treated) |>  
  group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "pfs"), .keep = TRUE) |>  
  bind_rows()  

km_res_from_data <- pfs_analysis_data |>
  group_by(trial, treated) |>  
  group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "progress_week"), .keep = TRUE) |>  
  bind_rows()  

km_res <- bind_rows(calculated = km_res_calculated, from_data = km_res_from_data, .id = "pfs_type")
```

```{r}
#| labels: kaplan-meier-upper-and-lower-bounds

km_res |> 
  filter(fct_match(pfs_type, "calculated")) |>
  ggplot() +
  geom_step(aes(x = t, y = s, color = btype)) +
  scale_color_discrete("Survival up to", labels = c(lb = "First possible week", ub = "Last possible week")) +
  labs(y = "S(t)") +
  facet_grid(vars(treated), vars(trial), scales = "free_x", labeller = labeller(treated = label_both)) +
  guides(color = "none") +
  NULL
```

::: aside

We show both the lower and upper bounds for PFS, given interval censoring. 

:::

# Analysis 

```{r}
#| label: survival-fit
#| cache: true
#| cache-lazy: false

# prior_no_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^prior_no_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
no_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^no_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
tumor_count_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^tumor_count_pfs-", full.names = TRUE), check_diagnostics = FALSE)
# prior_two_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^prior_two_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
two_tumor_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^two_tumor_pfs-", full.names = TRUE), check_diagnostics = FALSE)
tumor_change_pfs_res <- as_cmdstan_fit(dir(file.path(tmp_dir, "fit"), pattern = "^tumor_change_pfs-", full.names = TRUE), check_diagnostics = FALSE)
```


## Survival Model {.smaller}

* We use a piecewise-constant proportional hazard survival model.
* Survival was grouped over weekly intervals.
* Base hazard is modeled as a Gaussian process smoothed over intervals.
* Patient time-invariant hazard is modeled as a linear model of 
    i) of the two first tumor size assessments, plus their interaction
    ii) of the proportional change in each tumor's size
* Each tumor individually to the a patient's hazard -- tumor sizes are not aggregated.
* We adjust for both right and interval censoring.
* All parameters are modeled hierarchically to share information across trials.

## Survival Plot

```{r}
#| label: fig-no-tumor-kaplan-meier
#| cache: true

lst(no_tumor = no_tumor_pfs_res, two_tumor = two_tumor_pfs_res, tumor_change = tumor_change_pfs_res) |> 
  plot_km(treated_pfs_analysis_data, facet_arm = TRUE) +
  geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) +
  labs(title = "", subtitle = "") 
```

::: aside

Restricted to the treated arms with the same dosage.

:::

## Mean and Median PFS

```{r}
#| label: two-tumor-mean-median
#| cache: true

lst(no_tumor = no_tumor_pfs_res, two_tumor = two_tumor_pfs_res, tumor_change = tumor_change_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "model") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = trial, color = model), position = "dodge", .width = 0.8) +
  scale_color_discrete("Model") +
  facet_wrap(vars(stat), ncol = 1) +
  labs(subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```


::: aside

Estimates conditional on being uncensored.

:::

## Two Tumor Linear Model Parameters 

```{r}
#| label: two-tumor-coef
#| cache: true

lst(post = two_tumor_pfs_res) |> 
  map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |> 
  map_dfr(
    \(f) gather_rvars(f, tumor_stim_intercept, tumor_stim_coef[t], tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, t]), 
    .id = "fit"
  ) |> 
  mutate(
    .variable = if_else(is.na(trial), .variable, str_glue("{.variable}[{trial}]")),
    .variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]")),
  ) |> 
  ggplot() +
  stat_halfeye(aes(xdist = .value, y = .variable), size = 0.5, linewidth = 0.9, fill = alpha("black", 0.25), color = "black", .width = c(0.8)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  labs(x = "", y = "Tumor Model Parameters") +
  coord_cartesian(xlim = c(0, 0.5))
```

# Model Comparison

## Approximate Leave-One-Out Cross Validation {.smaller}

```{r}
#| label: pfs-loo
#| cache: true

no_tumor_loo <- no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
tumor_count_loo <- tumor_count_pfs_res$loo(save_psis = TRUE, cores = 12)
two_tumor_loo <- two_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
tumor_change_loo <- tumor_change_pfs_res$loo(save_psis = TRUE, cores = 12)

loo_compare(lst("No Tumor Response" = no_tumor_loo, "Tumor Count Only" = tumor_count_loo, "Linear Response" = two_tumor_loo, "Proportional Change Response" = tumor_change_loo))[, c("elpd_loo", "elpd_diff", "se_diff")] |> 
  kableExtra::kable(
    col.names = c("Expected Log Pointwise Predictive Density", "Difference", "SE"), 
    digits = 2, 
    label = c("Linear Model", "Proportional Change", "No Tumor Response")
  ) |> 
  kableExtra::kable_styling()
```

## Loss Function

```{r}
#| label: loss-functions

linear_loss <- function(m, c, a, ac, k0 = 1, k1 = 1) if_else(c & (ac | a >= m), 0, if_else(m >= a, k0, k1) * abs(m - a))
```

```{r}
#| label: fig-linear-loss

tibble(delta = seq(-10, 10), sym_loss = abs(delta), asym_loss = if_else(delta < 0, 3, 1) * abs(delta)) |> 
  pivot_longer(sym_loss:asym_loss) |> 
  ggplot(aes(delta, value)) +
  geom_line(aes(y = value, color = name), linewidth = 2, alpha = 0.25) +
  scale_color_discrete("Loss Type", labels = c("sym_loss" = "Symmetric", "asym_loss" = "Asymmetric")) +
  labs(x = latex2exp::TeX(r"{$y - \tilde{y}$}"), y = "Loss")
```

## Comparison of Out-of-sample Loss {.smaller}

```{r}
#| label: loo-predict-other-utility 
#| cache: true

asym_linear_loss <- partial(linear_loss, k1 = 3)

no_tumor_asym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, treated_pfs_stan_data, asym_linear_loss)
two_tumor_asym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, treated_pfs_stan_data, asym_linear_loss)
no_tumor_sym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, treated_pfs_stan_data, linear_loss)
two_tumor_sym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, treated_pfs_stan_data, linear_loss)

loss_data <- bind_rows(
  lst(no_tumor_sym_loss, 
      two_tumor_sym_loss, 
      no_tumor_asym_loss, 
      two_tumor_asym_loss), 
  .id = "model"
) |> 
  mutate(loss_type = str_extract(model, "(a?sym)|sqerr"), model = str_extract(model, "(ic_ignored_)?(no|two)_tumor"))
```

::: {.panel-tabset}

### Expected Loss 

```{r}
loss_data |> 
  group_by(loss_type, model, trial) |> 
  summarize(expected_loss = mean(value), .groups = "drop") |> 
  pivot_wider(names_from = trial, values_from = expected_loss) |> 
  mutate(
    loss_type = fct_recode(loss_type, "Asymmetric" = "asym", "Symmetric" = "sym"),
    model = fct_recode(model, "No Tumor Response" = "no_tumor", "Linear Response" = "two_tumor")
  ) |> 
  kableExtra::kable(digits = 2) |> 
  kableExtra::kable_styling()
```

### CDF 

```{r}
#| label: fig-loss-cdf

loss_data |> 
  mutate(loss_type = fct_recode(loss_type, "Asymmetric" = "asym", "Symmetric" = "sym")) |> 
  ggplot(aes(value)) +
  stat_ecdf(aes(color = model)) +
  scale_color_discrete("Model", labels = c(no_tumor = "No Tumor Response", two_tumor = "Linear Response"), aesthetics = c("color", "fill")) +
  labs(x = "Loss", y = "CDF") +
  facet_grid(vars(trial), vars(loss_type), scales = "free", margins = "trial") +
  theme(legend.position = "top", strip.text.y = element_text(angle = 0)) +
  NULL
```

### Histogram 

```{r}
#| label: fig-loss-hist

loss_data |> 
  mutate(loss_type = fct_recode(loss_type, "Asymmetric" = "asym", "Symmetric" = "sym")) |> 
  plot_loss(binwidth = 6) +
  scale_color_discrete("Model", labels = c(no_tumor = "No Tumor Response", two_tumor = "Linear Response"), aesthetics = c("color", "fill"))
```

:::

# Tumor Size Sensitivity

## Variation in Proportional Hazard in Response to Tumor Size

How much does the patient specific component of the proportional hazard respond to tumor sizes in the data?

```{r}
#| label: predictor-variation
#| cache: true
#| cache-lazy: false 

median_coef <- lst(post = two_tumor_pfs_res) |> 
  map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |> 
  map_dfr(\(f) spread_rvars(f, tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, k]), .id = "fit") |> 
  mutate(across(starts_with("tumor_stim"), median)) |> 
  pivot_wider(values_from = tumor_stim_trial_coef, names_prefix = "tumor_stim_trial_coef", names_from = k)

tumor_design_matrix <- with(
  treated_pfs_stan_data, {
    n_screening_t <- pfs_model$functions$calc_n_screening_t(n_patient_tumors, n_measures, t_measure)
    
    pfs_model$functions$standardize_nonzero_tumor_sizes(tumor_size) |>
      pluck(3) |> 
      pfs_model$functions$prepare_early_tumors_design_matrix(n_patient_tumors, n_measures, n_screening_t)
  }
)

tumor_design_matrix <- cbind(1, tumor_design_matrix, tumor_design_matrix[,1] * tumor_design_matrix[,2])

patient_kappa_data <- tibble(kappa = tumor_design_matrix, trial = with(treated_pfs_stan_data, rep(patient_trial, n_patient_tumors))) |>  
  left_join(median_coef, by = "trial") %>% 
  mutate(
    kappa = rowSums(kappa * select(., starts_with("tumor_stim"))),
    patient_index = rep(seq(treated_pfs_stan_data$n_patients), treated_pfs_stan_data$n_patient_tumors)
  ) |> 
  group_by(patient_index) |> 
  summarize(kappa = exp(-first(tumor_stim_trial_intercept) * mean(treated_pfs_stan_data$n_patient_tumors) + sum(kappa)))

patient_kappa_full_dist_data <- two_tumor_pfs_res |> 
  spread_rvars(total_time_invar_tumor_stim[patient_index]) |> 
  mutate(kappa = exp(total_time_invar_tumor_stim))

patient_kappa_data |> 
  mutate(var = "median") |> 
  bind_rows(
    patient_kappa_full_dist_data |>
      unnest_rvars() |> 
      mutate(var = "post")
  ) |> 
  filter(fct_match(var, "post")) |> 
  ggplot(aes(kappa)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  geom_histogram(aes(y = after_stat(density)), fill = "black", position = "identity", alpha = 0.25, binwidth = 0.1) +
  # scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(x = latex2exp::TeX(r"{$\kappa$}"), y = "Density") +
  coord_cartesian(xlim = c(0, 5)) +
  NULL
```

## Variation in Probability of Disease Progression

```{r}
#| label: dp-prob-variation
#| cache: true

two_tumor_pfs_res |> 
  spread_rvars(log_lambda[t]) |> 
  mutate(
    lambda = exp(log_lambda),
    median_baseline_dp_prob = 1 - exp(- median(lambda)),
    patient_dp_prob = map(lambda, \(l) 1 - exp(- median(l) * patient_kappa_data$kappa)) %>% 
      { array(unlist(.), dim = c(nrow(patient_kappa_data), length(.))) } |> 
      rvar(),
    
    patient_full_dist_dp_prob = map(lambda, \(l) draws_of(1 - exp(- median(l) * patient_kappa_full_dist_data$kappa))) %>%
      { array(unlist(.), dim = c(prod(dim(first(.))), length(.))) } |> 
      rvar()
  ) |> 
  ggplot(aes(t)) +
  # geom_line(aes(y = median_baseline_dp_prob)) +
  # Why didn't I use pivot? We can't have different sized rvars in the same column.
  # stat_lineribbon(aes(ydist = patient_dp_prob, fill = "median"), alpha = 0.25, .width = 0.8) +
  stat_lineribbon(aes(ydist = patient_full_dist_dp_prob), fill = "black", alpha = 0.25, .width = 0.8) +
  scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(y = latex2exp::TeX(r"{1 - \alpha}")) +
  NULL
```

# Discussion

## Limitations 

* We are currently using only the first two tumor assessments to make predictions about survival at every interval.
    - Consider using time-variant tumor sizes; capturing both the dynamics of tumor change and hazard.
* The main questions are ultimately causal while we are not treating them as in this analysis. 

## Conclusion

* Cross validation shows that 
    - A tumor response model is more predictive than a no tumor response model.
    - Tumor response intercepts seem to be the main predictive driver of this model.
* Expected loss is the lowest for models that respond to tumor sizes.