---
title: "Data Exploration"
subtitle: "Overview of the publication analysis dataset"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
all_analysis_data <- pub_tar_read("all_analysis_data")
```

This page provides an overview of the combined publication dataset (497 patients across two trials).

## Tumor Size Dynamics {#sec-tumor-dynamics}

```{r}
#| label: fig-sld-trajectories
#| fig-cap: Sum of target lesion diameters (SLD) over time by trial. Last observation for right-censored patients highlighted.
#| fig-width: 10
#| fig-height: 5

all_analysis_data |>
  unnest(visit_data) |>
  ggplot(aes(ady, mmsumdiam)) +
  geom_line(aes(group = usubjid), alpha = 0.07) +
  geom_point(size = 0.4, alpha = 0.3) +
  geom_point(
    aes(color = "Right Censored"),
    size = 0.8,
    data = \(d) filter(d, right_censored) |>
      group_by(usubjid) |>
      filter(min_rank(week) == n())
  ) +
  scale_color_discrete("", type = AZ_palette[-1]) +
  scale_x_continuous("Analysis Day") +
  scale_y_continuous("SLD [mm]") +
  facet_wrap(vars(trial), scales = "free", labeller = labeller(trial = str_to_upper)) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

## PFS Distribution {#sec-pfs}

```{r}
#| label: fig-pfs-distribution
#| fig-cap: PFS distribution split by event type (right-censored, progressive disease, or death without progression).

all_analysis_data |>
  mutate(event_type = case_when(
    right_censored         ~ "censored",
    progression_before_death ~ "pd",
    !progression_before_death & death ~ "death"
  )) |>
  ggplot(aes(pfs)) +
  geom_histogram(
    aes(fill = event_type),
    binwidth = 4, alpha = 0.5, position = "identity"
  ) +
  scale_fill_discrete(
    "Event Type",
    label = c(censored = "Right Censored", death = "Death without PD", pd = "Progressive Disease"),
    type = AZ_palette[-c(1:3)]
  ) +
  scale_x_continuous(
    "PFS [Months]",
    breaks = months_to_weeks(seq(0, 36, 6)),
    label  = label_weeks_to_months
  ) +
  labs(y = "Count") +
  facet_wrap(vars(trial), labeller = labeller(trial = str_to_upper), scales = "free") +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

## RECIST Classification {#sec-recist}

Comparison between recorded clinical RECIST classifications and those calculated from target lesion SLD alone. Discordance arises because clinical RECIST includes non-target lesions and new lesions.

```{r}
#| label: fig-recist-confusion
#| fig-cap: Recorded vs calculated RECIST response at each visit (target lesion SLD only).
#| fig-width: 10
#| fig-height: 5

all_analysis_data |>
  select(trial, visit_data) |>
  unnest(visit_data) |>
  filter(!is.na(response)) |>
  count(trial, response, det_response) |>
  group_by(trial, response) |>
  mutate(p = n / sum(n)) |>
  ungroup() |>
  plot_confusion_matrix(response, det_response, p, n, label_size = 3) +
  facet_wrap(vars(trial), labeller = labeller(.default = str_to_upper)) +
  labs(x = "Recorded Response", y = "Calculated Response")
```

## Target vs Recorded PFS {#sec-target-pfs}

```{r}
#| label: fig-target-vs-recorded-pfs
#| fig-cap: "Deterministic target lesion PFS vs recorded clinical PFS, coloured by event type. Points well below the diagonal: clinical event was death or non-target/new-lesion PD while target SLD never crossed the PD threshold."
#| fig-width: 10
#| fig-height: 5

all_analysis_data |>
  mutate(event_type = case_when(
    right_censored                          ~ "Censored",
    !det_right_censored                     ~ "Target PD",
    progression_before_death                ~ "Non-target / New lesion PD",
    death                                   ~ "Death without PD"
  )) |>
  mutate(recorded_pfs_months = (pfs + interval_censored + 1) / (52 / 12),
         det_pfs_months = det_pfs / (52 / 12)) |>
  ggplot(aes(recorded_pfs_months, det_pfs_months, color = event_type)) +
  geom_abline(slope = 1, linetype = "dashed", alpha = 0.4) +
  geom_point(alpha = 0.6, size = 1.2) +
  scale_color_manual(
    "Event Type",
    values = c(
      "Censored"                   = AZ_platinum,
      "Target PD"                  = AZ_navy,
      "Non-target / New lesion PD" = AZ_gold,
      "Death without PD"           = AZ_pink
    )
  ) +
  scale_x_continuous("Recorded PFS [Months]", breaks = seq(0, 36, 6)) +
  scale_y_continuous("Target Tumour PFS [Months]", breaks = seq(0, 36, 6)) +
  facet_wrap(vars(trial), labeller = labeller(trial = str_to_upper)) +
  theme(legend.position = "bottom")
```

## Baseline Characteristics {#sec-baseline}

```{r}
#| label: fig-baseline
#| fig-cap: Baseline covariate distributions by trial.
#| fig-width: 10
#| fig-height: 6
#| message: false

all_analysis_data |>
  select(usubjid, trial, age, male, ecog, hgb, ldh_log, albumin) |>
  pivot_longer(!c(usubjid, trial)) |>
  ggplot() +
  geom_histogram(
    aes(value, after_stat(density), fill = trial),
    position = "identity", alpha = 0.5, bins = 30
  ) +
  scale_fill_discrete("", type = AZ_palette, label = str_to_upper) +
  facet_wrap(
    vars(name), scales = "free",
    labeller = labeller(name = c(
      age = "Age", male = "Male", ecog = "ECOG",
      hgb = "Haemoglobin", ldh_log = "log(LDH)", albumin = "Albumin"
    ))
  ) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```
