---
title: "Model Validation"
subtitle: "Leave-future-out cross-validation and OOS accuracy assessment"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
lfo_store <- file.path(output_path, "publication", "lfo", "_targets")
```

This page presents out-of-sample predictive accuracy evaluated via leave-future-out (LFO)
cross-validation. LFO respects the temporal structure of the data: at each cutoff, all future
observations are held out and predicted using only data up to that point.

::: {.callout-note}
## Methodology
At each LFO fold the model is refit on data up to a cutoff date, then RECIST responses are
predicted for all patients with future visits. Pareto-smoothed importance sampling (PSIS) is
used to approximate refits where the LOO approximation is adequate, with exact MCMC refits
for high-influence observations.
:::

## Accuracy Assessment {#sec-accuracy}

### RECIST Prediction Accuracy

Out-of-sample confusion matrix showing model sensitivity (recall) for each RECIST category,
aggregated across all LFO folds.

::: {layout-ncol=2}

```{r}
#| label: fig-oos-confusion-matrix
#| fig-cap: "Out-of-sample (leave-future-out) confusion matrix showing sensitivity for each RECIST category. Cell values represent P(Predicted response | Observed response)."
#| fig-width: 5

oos_cm <- lfo_tar_read("oos_confusion_matrix")
plot_oos_confusion_matrix(oos_cm)
```

```{r}
#| label: fig-oos-confusion-sankey
#| fig-cap: "Sankey diagram showing the flow from observed to predicted RECIST responses. Flow width represents prediction proportions; percentages indicate correct prediction rates."
#| fig-width: 6
#| fig-height: 5

plot_oos_confusion_sankey(oos_cm)
```

:::

### RECIST Specificity Analysis

For each category X, the specificity matrix shows the distribution of predictions when the
true response is **NOT** X.

```{r}
#| label: fig-oos-specificity-matrix
#| fig-cap: "Out-of-sample specificity matrix. Each column shows P(Predicted response | Observed ≠ category). Higher proportions in off-diagonal cells indicate better specificity."
#| fig-width: 6

plot_oos_specificity_matrix(oos_cm)
```

::: {.callout-note}
## Understanding Sensitivity vs Specificity
- **Confusion matrix**: column-wise proportions — P(Predicted X | Observed X), i.e., sensitivity/recall
- **Specificity matrix**: P(Predicted not X | Observed not X) for each category X
- Higher off-diagonal values in the specificity matrix indicate correct rejection of false positives
:::

### Summary Metrics

```{r}
#| label: tbl-accuracy-metrics
#| tbl-cap: "Out-of-sample accuracy, sensitivity, and specificity with 90% credible intervals."

compute_oos_accuracy_metrics(oos_cm) |>
  gt() |>
  tab_header(
    title = "RECIST Prediction Performance"
  ) |>
  cols_label(
    metric = "Metric",
    class = "Class",
    median = "Median",
    q5 = "5%",
    q95 = "95%"
  ) |>
  tab_spanner(
    label = "Percentiles",
    columns = c("q5", "q95")
  ) |>
  fmt_percent(
    columns = c(median, q5, q95),
    decimals = 1
  )
```

### RECIST Prediction Over Time

Tracking per-category sensitivity across LFO folds, showing how prediction accuracy evolves
as more data accumulates.

```{r}
#| label: fig-recist-sensitivity-over-time
#| fig-cap: "Running sensitivity for each RECIST category across LFO folds. Ribbons show ±1 SE computed within each posterior draw across folds."
#| fig-width: 10
#| fig-height: 6

lfo_tar_read("full_oos_confusion_matrix") |>
  plot_recist_sensitivity_over_time()
```

## KM Forecast Evolution {#sec-km-evolution}

PFS Kaplan-Meier forecasts at three evenly-distributed LFO cutoffs, showing how the model's
survival prediction develops as more patient data accumulates over the trial.

::: {.callout-note}
## Cutoff selection
Three cutoffs are drawn at roughly the 25th, 50th, and 75th percentile of the LFO run,
spanning early, mid, and late enrollment. Each panel shows the posterior predictive KM at
that information cutoff — earlier panels use fewer patients and carry wider uncertainty.
:::

::: {.panel-tabset}

### Sample-Based PFS

```{r}
#| label: fig-km-evolution-sample
#| fig-cap: "Sample-based PFS KM forecast at three LFO cutoffs. Each panel re-simulates enrolled patients forward conditional on their last observed visit; shaded band is 95% credible interval."
#| fig-width: 12
#| fig-height: 5

lfo_cutoff_km <- lfo_tar_read_pattern("lfo_cutoff_km_est")
target_arms <- grep("lilly_cxcr4", levels(lfo_cutoff_km$trial), value = TRUE)
plot_km_evolution(lfo_cutoff_km, km_est_var = sample_pfs_km_est, trials = target_arms)
```

### Multistate PFS

```{r}
#| label: fig-km-evolution-ms
#| fig-cap: "Multistate-hazard PFS KM forecast at three LFO cutoffs. Uses the illness-death model's 0→1 (progression detection) endpoint."
#| fig-width: 12
#| fig-height: 5

lfo_tar_read_pattern("lfo_cutoff_km_est") |>
  plot_km_evolution(km_est_var = sample_ms_pfs_km_est, trials = target_arms)
```

### Superpopulation PFS

```{r}
#| label: fig-km-evolution-spop
#| fig-cap: "Superpopulation PFS KM forecast at three LFO cutoffs. Patients re-simulated unconditionally from week 1 (population-forecast estimand); systematically left of sample-based by design."
#| fig-width: 12
#| fig-height: 5

lfo_tar_read_pattern("lfo_cutoff_km_est") |>
  plot_km_evolution(km_est_var = spop_pfs_km_est, trials = target_arms)
```

:::
