---
title: "Model Validation"
subtitle: "Leave-future-out cross-validation and OOS accuracy assessment"
execute:
  eval: true
  echo: false
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
lfo_store <- file.path(output_path, "publication", "lfo", "_targets")
```

This page presents out-of-sample predictive accuracy evaluated via leave-future-out (LFO)
cross-validation. LFO respects the temporal structure of the data: at each cutoff, all future
observations are held out and predicted using only data up to that point.

::: {.callout-note}
## Methodology
At each LFO fold the model is refit on data up to a cutoff date, then RECIST responses are
predicted for all patients with future visits. Pareto-smoothed importance sampling (PSIS) is
used to approximate refits where the LOO approximation is adequate, with exact MCMC refits
for high-influence observations.
:::

## Accuracy Assessment {#sec-accuracy}

### RECIST Prediction Accuracy

Out-of-sample confusion matrix showing model sensitivity (recall) for each RECIST category,
aggregated across all LFO folds.

::: {layout-ncol=2}

```{r}
#| label: fig-oos-confusion-matrix
#| fig-cap: "Out-of-sample (leave-future-out) confusion matrix showing sensitivity for each RECIST category. Cell values represent P(Predicted response | Observed response)."
#| fig-width: 5

oos_cm <- lfo_tar_read("oos_confusion_matrix")
plot_oos_confusion_matrix(oos_cm)
```

```{r}
#| label: fig-oos-confusion-sankey
#| fig-cap: "Sankey diagram showing the flow from observed to predicted RECIST responses. Flow width represents prediction proportions; percentages indicate correct prediction rates."
#| fig-width: 6
#| fig-height: 5

plot_oos_confusion_sankey(oos_cm)
```

:::

### RECIST Specificity Analysis

For each category X, the specificity matrix shows the distribution of predictions when the
true response is **NOT** X.

```{r}
#| label: fig-oos-specificity-matrix
#| fig-cap: "Out-of-sample specificity matrix. Each column shows P(Predicted response | Observed ≠ category). Higher proportions in off-diagonal cells indicate better specificity."
#| fig-width: 6

plot_oos_specificity_matrix(oos_cm)
```

::: {.callout-note}
## Understanding Sensitivity vs Specificity
- **Confusion matrix**: column-wise proportions — P(Predicted X | Observed X), i.e., sensitivity/recall
- **Specificity matrix**: P(Predicted not X | Observed not X) for each category X
- Higher off-diagonal values in the specificity matrix indicate correct rejection of false positives
:::

### Summary Metrics

```{r}
#| label: tbl-accuracy-metrics
#| tbl-cap: "Out-of-sample accuracy, sensitivity, and specificity with 90% credible intervals."

compute_oos_accuracy_metrics(oos_cm) |>
  gt() |>
  tab_header(
    title = "RECIST Prediction Performance"
  ) |>
  cols_label(
    metric = "Metric",
    class = "Class",
    median = "Median",
    q5 = "5%",
    q95 = "95%"
  ) |>
  tab_spanner(
    label = "Percentiles",
    columns = c("q5", "q95")
  ) |>
  fmt_percent(
    columns = c(median, q5, q95),
    decimals = 1
  )
```

### RECIST Prediction Over Time

Tracking per-category sensitivity across LFO folds, showing how prediction accuracy evolves
as more data accumulates.

```{r}
#| label: fig-recist-sensitivity-over-time
#| fig-cap: "Running sensitivity for each RECIST category across LFO folds. Ribbons show ±1 SE computed within each posterior draw across folds."
#| fig-width: 10
#| fig-height: 6

lfo_tar_read("full_oos_confusion_matrix") |>
  plot_recist_sensitivity_over_time(dco_dates = c())
```

## KM Forecast Evolution {#sec-km-evolution}

PFS Kaplan-Meier forecasts at three evenly-distributed LFO cutoffs, showing how the model's
survival prediction develops as more patient data accumulates over the trial.

::: {.callout-note}
## Cutoff selection
Three cutoffs are drawn at roughly the 25th, 50th, and 75th percentile of the LFO run,
spanning early, mid, and late enrollment. Each panel shows the posterior predictive KM at
that information cutoff — earlier panels use fewer patients and carry wider uncertainty.
:::

::: {.panel-tabset}

### Sample-Based PFS

```{r}
#| label: fig-km-evolution-sample
#| fig-cap: "Sample-based PFS KM forecast at three LFO cutoffs. Each panel re-simulates enrolled patients forward conditional on their last observed visit; shaded band is 80% credible interval. The \"At cutoff\" series is the observed KM censored at each cutoff; the \"Overall\" series is the final full-data observed KM (mature endpoint, identical across panels) for reference."
#| fig-width: 12
#| fig-height: 5

lfo_cutoff_km <- lfo_tar_read_pattern("lfo_cutoff_km_est")
target_arms <- grep("lilly_cxcr4", levels(lfo_cutoff_km$trial), value = TRUE)

all_analysis_data <- pub_tar_read("all_analysis_data")

cutoff_meta <- lfo_cutoff_km |>
  distinct(cutoff_calendar_day, n_target_observed)

censor_at_cutoff <- function(data, cutoff_day) {
  data |>
    filter(calendar_day <= cutoff_day) |>
    mutate(
      followup_weeks = (cutoff_day - calendar_day) / 7,
      pfs_cut        = pmin(pfs, followup_weeks),
      censored_cut   = right_censored | (pfs > followup_weeks)
    )
}

obs_km_cutoffs <- purrr::map(seq_len(nrow(cutoff_meta)), \(i) {
  cd  <- cutoff_meta$cutoff_calendar_day[[i]]
  n_t <- cutoff_meta$n_target_observed[[i]]
  all_analysis_data |>
    filter(forcats::fct_match(trial, "lilly_cxcr4")) |>
    mutate(group = forcats::fct_drop(interaction(trial, arm, sep = "_"))) |>
    censor_at_cutoff(cd) |>
    get_km_res(pfs_cut, censored_cut, by = group) |>
    tidyr::unnest(km_data) |>
    transmute(
      trial,
      btype,
      cutoff_calendar_day = cd,
      n_target_observed   = n_t,
      t, s, c
    )
}) |>
  dplyr::bind_rows()

# Final full-data PFS KM (no cutoff censoring) — the mature observed endpoint,
# drawn as a dashed reference across every cutoff panel.
final_pfs_km <- all_analysis_data |>
  filter(forcats::fct_match(trial, "lilly_cxcr4")) |>
  mutate(group = forcats::fct_drop(interaction(trial, arm, sep = "_"))) |>
  get_km_res(pfs, right_censored, by = group) |>
  tidyr::unnest(km_data) |>
  filter(btype == "ub") |>
  transmute(trial, t, s, c)

plot_km_evolution(
  lfo_cutoff_km,
  km_est_var    = sample_pfs_km_est,
  trials        = target_arms,
  obs_km_data   = obs_km_cutoffs,
  final_km_data = final_pfs_km
)
```

### Multistate PFS

```{r}
#| label: fig-km-evolution-ms
#| fig-cap: "Multistate-hazard PFS KM forecast at three LFO cutoffs. Uses the illness-death model's 0→1 (progression detection) endpoint. The \"At cutoff\" series is the observed KM censored at each cutoff; the \"Overall\" series is the final full-data observed KM (mature endpoint, identical across panels) for reference."
#| fig-width: 12
#| fig-height: 5

plot_km_evolution(lfo_cutoff_km, km_est_var = sample_ms_pfs_km_est, trials = target_arms, obs_km_data = obs_km_cutoffs, final_km_data = final_pfs_km)
```

### Superpopulation PFS

```{r}
#| label: fig-km-evolution-spop
#| fig-cap: "Superpopulation PFS KM forecast at three LFO cutoffs. Patients re-simulated unconditionally from week 1 (population-forecast estimand); systematically left of sample-based by design. The \"At cutoff\" series is the observed KM censored at each cutoff; the \"Overall\" series is the final full-data observed KM (mature endpoint, identical across panels) for reference."
#| fig-width: 12
#| fig-height: 5

plot_km_evolution(lfo_cutoff_km, km_est_var = spop_pfs_km_est, trials = target_arms, obs_km_data = obs_km_cutoffs, final_km_data = final_pfs_km)
```

:::

## OS Forecast Evolution {#sec-os-km-evolution}

Overall survival Kaplan-Meier forecasts at the same three LFO cutoffs, showing how the model's
OS prediction develops as trial data accumulates.

::: {.callout-note}
## Interpretation
Earlier cutoffs have wider uncertainty and fewer observed deaths. Convergence toward the final
observed OS KM as the cutoff advances confirms that the multistate hazard model is calibrated
for OS, not just PFS.
:::

::: {.panel-tabset}

### Sample-Based OS

```{r}
#| label: fig-os-km-evolution-sample
#| fig-cap: "Sample-based OS KM forecast at three LFO cutoffs. Each panel re-simulates enrolled patients forward; shaded band is 80% credible interval. The \"At cutoff\" series is the observed OS KM censored at each cutoff; the \"Overall\" series is the full-data observed OS KM (identical across panels) for reference."
#| fig-width: 12
#| fig-height: 5

lfo_cutoff_os_km <- lfo_tar_read_pattern("lfo_cutoff_os_km_est")

obs_os_km_cutoffs <- purrr::map(seq_len(nrow(cutoff_meta)), \(i) {
  cd  <- cutoff_meta$cutoff_calendar_day[[i]]
  n_t <- cutoff_meta$n_target_observed[[i]]
  all_analysis_data |>
    filter(forcats::fct_match(trial, "lilly_cxcr4")) |>
    mutate(
      group   = forcats::fct_drop(interaction(trial, arm, sep = "_")),
      os_time = if_else(death, death_week, patient_max_t),
      os_cens = !death
    ) |>
    censor_at_cutoff(cd) |>
    mutate(
      os_cut      = pmin(os_time, followup_weeks),
      os_cens_cut = os_cens | (os_time > followup_weeks)
    ) |>
    get_km_res(os_cut, os_cens_cut, by = group) |>
    tidyr::unnest(km_data) |>
    transmute(
      trial,
      btype,
      cutoff_calendar_day = cd,
      n_target_observed   = n_t,
      t, s, c
    )
}) |>
  dplyr::bind_rows()

# Final full-data OS KM (no cutoff censoring) — mature observed endpoint.
final_os_km <- all_analysis_data |>
  filter(forcats::fct_match(trial, "lilly_cxcr4")) |>
  mutate(
    group   = forcats::fct_drop(interaction(trial, arm, sep = "_")),
    os_time = if_else(death, death_week, patient_max_t),
    os_cens = !death
  ) |>
  get_km_res(os_time, os_cens, by = group) |>
  tidyr::unnest(km_data) |>
  filter(btype == "ub") |>
  transmute(trial, t, s, c)

plot_km_evolution(
  lfo_cutoff_os_km,
  km_est_var    = sample_os_km_est,
  trials        = target_arms,
  obs_km_data   = obs_os_km_cutoffs,
  final_km_data = final_os_km
)
```

### Superpopulation OS

```{r}
#| label: fig-os-km-evolution-spop
#| fig-cap: "Superpopulation OS KM forecast at three LFO cutoffs. Patients re-simulated unconditionally from week 1 (population-forecast estimand). The \"At cutoff\" series is the observed OS KM censored at each cutoff; the \"Overall\" series is the full-data observed OS KM (identical across panels) for reference."
#| fig-width: 12
#| fig-height: 5

plot_km_evolution(lfo_cutoff_os_km, km_est_var = spop_os_km_est, trials = target_arms, obs_km_data = obs_os_km_cutoffs, final_km_data = final_os_km)
```

:::
