---
title: "Overall Survival"
subtitle: "OS Kaplan-Meier forecasts and observed data"
execute:
  freeze: true
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
os_km_data  <- pub_tar_read("all_tumor_ssls_km_os_rvar")
obs_os_data <- pub_tar_read("km_trial_os") |>
  filter(fct_match(btype, "ub")) |>
  select(!quantiles) |>
  unnest(km_data)
```

## OS Kaplan-Meier Curves {#sec-os-km}

::: {.callout-note collapse="false"}
## Two estimands: sample-based vs superpopulation

As with PFS, the two tabs answer different questions and are **not** expected to
coincide:

- **Sample-based** is the posterior-predictive check — each enrolled patient
  re-simulated *conditional on their observed survival to last follow-up*. This
  is the curve to compare against the observed OS KM.

- **Superpopulation** re-simulates the same cohort *unconditionally, from week 1*.
  Discarding already-survived follow-up shifts the curve left of the
  sample-based and observed curves. This is the population-forecast estimand —
  predicted OS for the cohort before observing its survival — so its
  displacement is the expected signature of unconditional re-simulation, not a
  lack of fit.
:::

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-os-km-sample
#| fig-cap: Sample-based OS KM (posterior) by trial.
#| fig-width: 10
#| fig-height: 5

os_km_data |>
  filter(fit_type == "posterior") |>
  plot_km(
    km_est_var  = sample_os_km_est,
    facet_col   = trial,
    obs_km_data = obs_os_data,
    endpoint    = "os"
  )
```

### Superpopulation

```{r}
#| label: fig-os-km-spop
#| fig-cap: "Superpopulation OS KM (posterior) by trial: the enrolled cohort re-simulated unconditionally from week 1 (the population-forecast estimand). It sits left of the sample-based and observed curves by design — see the note above."
#| fig-width: 10
#| fig-height: 5

os_km_data |>
  filter(fit_type == "posterior") |>
  plot_km(
    km_est_var  = spop_os_km_est,
    facet_col   = trial,
    obs_km_data = obs_os_data,
    endpoint    = "os"
  )
```

### Prior vs Posterior

```{r}
#| label: fig-os-km-prior-post
#| fig-cap: Prior vs posterior OS KM (sample-based) by trial.
#| fig-width: 10
#| fig-height: 5

os_km_data |>
  plot_km(
    km_est_var  = sample_os_km_est,
    facet_col   = trial,
    group       = fit_type,
    alpha_group = fit_type,
    endpoint    = "os"
  )
```

:::
