---
title: "Patient Dynamics"
subtitle: "Individual tumor trajectories and RECIST predictions"
execute:
  freeze: true
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
sld_rc    <- pub_tar_read("tumor_ssls_staged_sld_rvar_right_censored_posterior")
sld_unc   <- pub_tar_read("tumor_ssls_staged_sld_rvar_uncensored_posterior")
recist_rc <- pub_tar_read("tumor_ssls_staged_recist_rvar_right_censored_posterior")
```

## SLD Forecasts — Right-Censored Patients {#sec-sld-censored}

Individual SLD trajectories for patients who are right-censored (progression-free at data cutoff), with model forecasts extended beyond the last observed measurement.

```{r}
#| label: fig-sld-censored
#| fig-cap: SLD trajectories and forecasts for right-censored patients (posterior).
#| fig-width: 10
#| fig-height: 12

sld_rc |> plot_sld_forecast_censored()
```

## SLD Posterior Predictive Check — Uncensored Patients {#sec-sld-ppc}

Replicated vs observed SLD for patients who had a PFS event. Only the observed window is shown — no forecast beyond the event.

```{r}
#| label: fig-sld-ppc
#| fig-cap: Observed vs posterior-predicted SLD for uncensored patients (observation window only).
#| fig-width: 10
#| fig-height: 12

sld_unc |>
  filter(stage == "obs") |>
  plot_sld_forecast_censored()
```

## RECIST Predictions — Right-Censored Patients {#sec-recist-censored}

```{r}
#| label: fig-recist-censored
#| fig-cap: RECIST response predictions for right-censored patients (posterior).
#| fig-width: 10
#| fig-height: 12

recist_rc |> plot_recist_censored()
```
