---
title: "Survival Analysis"
subtitle: "Kaplan-Meier forecasts and observed PFS"
execute:
  freeze: true
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
km_data     <- pub_tar_read("all_tumor_ssls_km_rvar")
obs_km_data <- pub_tar_read("km_trial_pfs") |>
  filter(fct_match(btype, "ub")) |>
  select(!quantiles) |>
  unnest(km_data)
```

## PFS Kaplan-Meier Curves {#sec-pfs-km}

::: {.callout-note collapse="false"}
## Two estimands: sample-based vs superpopulation

The two tabs answer different questions and are **not** expected to coincide:

- **Sample-based** is a posterior-predictive check. Each enrolled patient is
  re-simulated forward *conditional on having survived event-free up to their
  last observed visit*. This is the curve to compare against the observed KM —
  it tracks the data to within roughly ±0.02–0.04 across the follow-up window.

- **Superpopulation** re-simulates the same cohort *unconditionally, from week 1*.
  By discarding each patient's already-survived early follow-up, the cohort is
  re-exposed to the high early-progression hazard window, so the curve sits
  roughly 4–5 weeks to the left of the sample-based curve (largest separation,
  ≈0.16–0.19, near week 24, closing to near-zero by week 60+). This is the
  population-forecast estimand — the predicted PFS for the cohort *before*
  observing any of its survival, not a fit to the realised data. Its
  displacement from the observed KM is the mechanical signature of unconditional
  re-simulation, not a misspecification.
:::

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-pfs-km-sample
#| fig-cap: Sample-based PFS KM (posterior) by trial. Shaded band shows 80% credible interval.
#| fig-width: 10
#| fig-height: 5

km_data |>
  filter(fit_type == "posterior") |>
  plot_km(
    km_est_var  = sample_pfs_km_est,
    facet_col   = trial,
    obs_km_data = obs_km_data
  )
```

### Superpopulation

```{r}
#| label: fig-pfs-km-spop
#| fig-cap: "Superpopulation PFS KM (posterior) by trial: the enrolled cohort re-simulated unconditionally from week 1 (the population-forecast estimand). It sits left of both the sample-based and observed curves by design — see the note above. Wider uncertainty reflects the unconditional re-simulation."
#| fig-width: 10
#| fig-height: 5

km_data |>
  filter(fit_type == "posterior") |>
  plot_km(
    km_est_var  = spop_pfs_km_est,
    facet_col   = trial,
    obs_km_data = obs_km_data
  )
```

### Prior vs Posterior

```{r}
#| label: fig-pfs-km-prior-post
#| fig-cap: Prior vs posterior PFS KM (sample-based) by trial.
#| fig-width: 10
#| fig-height: 5

km_data |>
  plot_km(
    km_est_var  = sample_pfs_km_est,
    facet_col   = trial,
    group       = fit_type,
    alpha_group = fit_type
  )
```

:::
