---
title: "Leave-Future-Out Cross-Validation"
subtitle: "Temporal cross-validation for longitudinal survival data"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

**Created:** January 2026
**Status:** Production-ready
**Module:** `stan/ssls/sf-ssls-lfo.stan`

::: {.callout-tip}
## Related Pages
- See [Model Validation](../analysis/model-validation.qmd) for LFO results and ELPD comparisons
- See [Model Specification](model-specification.qmd) for the underlying tumor dynamics model
- See [Architecture](architecture.qmd) for code organization
:::

---

## Overview {#sec-lfo-overview}

### What is Leave-Future-Out Cross-Validation?

Leave-future-out (LFO) cross-validation is a method for evaluating predictive performance of time-series and longitudinal models. Unlike traditional cross-validation methods, LFO respects the **temporal structure** of the data by only using past observations to predict future ones.

**Key Principle:** At each time point $t$, the model is trained on data up to $t$ and evaluated on its ability to predict observations after $t$.

### Why LFO for Oncology Trials?

In clinical trials with longitudinal tumor measurements:

1. **Staggered patient enrollment**: Patients enter the study at different calendar times
2. **Variable follow-up**: Patients have different observation periods
3. **Temporal dependence**: Future tumor measurements depend on past dynamics
4. **Clinical relevance**: We want to know how well we can predict future outcomes given current data

LFO provides a realistic assessment of predictive accuracy because it mimics the actual clinical decision-making scenario: using available data at a data cutoff to forecast future outcomes.

---

## LFO vs LOO: Key Differences {#sec-lfo-vs-loo}

### Leave-One-Out (LOO) Cross-Validation

**Approach:** Hold out one observation (or patient) at a time, train on all remaining data, and evaluate on the held-out observation.

**Problem for longitudinal data:**

- LOO ignores temporal structure
- A model trained on "future" observations can predict "past" ones
- This creates **information leakage** that inflates predictive accuracy estimates
- Does not reflect real-world forecasting scenarios

### Leave-Future-Out (LFO) Cross-Validation

**Approach:** Define a calendar time cutoff, train only on data before the cutoff, and evaluate on data after the cutoff.

**Benefits:**

- **No information leakage**: Model never sees future data during training
- **Realistic evaluation**: Mimics actual clinical data cutoff scenarios
- **Temporal integrity**: Respects causal structure of longitudinal data
- **Clinically relevant**: Answers "how well can we predict future outcomes?"

### Visual Comparison

```{r}
#| label: fig-lfo-vs-loo-comparison
#| fig-cap: Comparison between LFO and LOO cross-validation approaches for longitudinal clinical trial data.
#| fig-width: 12
#| fig-height: 5

source(here::here("r/plot_lfo_diagram.R"))
```

::: {.callout-note}
## Key Observations from the Diagram

**LFO (Left Panel):**

- Multiple time-based cutoffs divide the study into training (navy) and test (gold) periods
- All patients contribute to both training and testing at different cutoffs
- Test window starts at cutoff and extends forward in time

**LOO (Right Panel):**

- Patient-based folds: each patient (pink) held out once
- Training data includes observations from all time points of non-held-out patients
- Can use "future" patient data to predict "past" patient outcomes
:::

---

## Mathematical Framework {#sec-lfo-math}

### Expected Log Predictive Density (ELPD)

The primary metric for comparing models using LFO is the **expected log predictive density (ELPD)**:

$$
\text{ELPD}_{\text{LFO}} = \sum_{c=1}^{C} \sum_{i \in \mathcal{T}_c} \sum_{t > t_c} \log p(y_{i,t} \mid y_{1:t_c}, \theta)
$$

where:

- $C$: number of cutoffs
- $\mathcal{T}_c$: set of patients with observations after cutoff $c$
- $t_c$: time index of cutoff $c$
- $y_{i,t}$: observation for patient $i$ at time $t$
- $y_{1:t_c}$: all observations up to cutoff time $t_c$
- $\theta$: model parameters

### Posterior Predictive Distribution

For each cutoff, we compute the posterior predictive distribution:

$$
p(y_{i,t} \mid y_{1:t_c}) = \int p(y_{i,t} \mid \theta) \, p(\theta \mid y_{1:t_c}) \, d\theta
$$

This integral is approximated using MCMC samples from the posterior $p(\theta \mid y_{1:t_c})$.

### Log-Likelihood Components

In our joint tumor-survival model, the log-likelihood decomposes into:

$$
\log p(y \mid \theta) = \underbrace{\log p(\text{SLD} \mid \theta)}_{\text{Tumor component}} + \underbrace{\log p(\text{PFS}_{\text{other}} \mid \theta)}_{\text{Other events component}}
$$

This allows us to separately evaluate:

1. **Tumor model accuracy**: How well do we predict future SLD measurements?
2. **Other events accuracy**: How well do we predict non-target progression and death?
3. **Joint accuracy**: Combined predictive performance

### Comparing Models via ELPD Difference

To compare two models, we compute:

$$
\Delta \text{ELPD} = \text{ELPD}_{\text{model A}} - \text{ELPD}_{\text{model B}}
$$

**Interpretation:**

- $\Delta \text{ELPD} > 0$: Model A has better predictive performance
- $\Delta \text{ELPD} < 0$: Model B has better predictive performance
- Standard error quantifies uncertainty in the comparison

::: {.callout-note}
## ELPD Scale

ELPD values are on the log probability scale. A difference of 1 ELPD unit corresponds to a factor of $e \approx 2.7$ in probability ratio.
:::

---

## Implementation Details {#sec-lfo-implementation}

### Cutoff Definition

LFO cutoffs are defined by **calendar time**, not visit number:

```stan
// From _sf-ssls-lfo-data.stan
int cutoff_calendar_day;  // The calendar day defining the cutoff
```

For each patient $i$, we determine:

1. **Last observed visit before cutoff**: `cutoff_last_visit_idx[i]`
2. **Test visits after cutoff**: All visits with `visit_calendar_day > cutoff_calendar_day`
3. **Whether patient was enrolled**: `cutoff_observed_mask[i] == 1` if patient had visits before cutoff

### Training Phase

The model is fit using only pre-cutoff data:

```stan
// Tumor likelihood (only up to cutoff)
for (i in 1:n_patients) {
  if (cutoff_last_visit_idx[i] > 0) {
    int cutoff_idx = cutoff_last_visit_idx[i];
    normalized_sld[visit_start:cutoff_idx] ~
      sf_log_space_obs(states[visit_start:cutoff_idx], measure_sd, ...);
  }
}

// Other events likelihood (cutoff-aware)
patient_response_lp[, 1] = calc_pch_loglik(
  cutoff_ic_other_events_pfs,      // Censored at cutoff
  cutoff_other_events_right_censored,
  ...,
  cutoff_last_visit_week           // Bounded by cutoff time
);
```

### Out-of-Sample Prediction

For each patient with post-cutoff observations, we generate predictions:

1. **State forecasting**: Propagate latent tumor states forward from last observed state

```stan
// Compute forecast states using constant rates
(full_forecast_expected, full_forecast) = sf_log_space_trajectory_ncp(
  patient_states[visit_size],  // Last observed state as initial
  forecast_time,
  exp(patient_log_decrease_rate[i, 1]),
  exp(patient_log_growth_rate[i, 1]),
  ...
);
```

2. **SLD prediction**: Generate predicted SLD with measurement noise

```stan
forecast_patient_log_sld = to_vector(normal_rng(
  forecast_mean_patient_log_sld,
  rep_vector(measure_sd, n_oos_visits)
));
```

3. **RECIST classification**: Apply RECIST criteria to predicted trajectories

```stan
array[treat_visit_size + n_oos_visits] int full_predict_overall_recist =
  calculate_target_recist(
    exp(append_row(rep_mean_patient_log_sld, forecast_mean_patient_log_sld)) * 10,
    n_patient_screening_visits[i]
  );
```

### Log-Likelihood Computation

For each test observation, we compute the log-likelihood:

```stan
// Tumor component
real tumor_ll = sf_log_space_obs_lpdf(
  normalized_sld[start_idx:end_idx] | states[start_idx:end_idx],
  measure_sd, log_lod - log_baseline_sld[i]
);

// Other events component
real oe_ll = calc_pch_loglik(
  obs_oe_pfs, obs_oe_censored, obs_oe_ic, 0,
  patient_log_surv, test_start, test_end
)[1];

// Joint log-likelihood
patient_log_lik[n, m_rel, patient_idx] = tumor_ll + oe_ll;
```

### Confusion Matrix Generation

We track RECIST prediction accuracy via a confusion matrix:

```stan
// rows = observed RECIST, cols = predicted RECIST
for (t_idx in 1:n_curr_testing_visits) {
  int obs_val = recist[start_idx + t_idx - 1];
  int pred_val = oos_recist[oos_recist_start + test_start_offset + t_idx - 1];

  if (obs_val <= PD) {
    oos_recist_confusion_matrix[n, m_rel][obs_val, pred_val] += 1;
  }
}
```

---

## Special Considerations {#sec-lfo-considerations}

### Handling Newly Enrolled Patients

Patients who enrolled **after** the cutoff are excluded from LFO evaluation:

```stan
// Only evaluate patients who were observed at cutoff
if (start_idx > 0 && end_idx >= start_idx && cutoff_observed_mask[i] == 1) {
  // ... compute log-likelihood
}
```

**Rationale:** We cannot evaluate predictive performance for patients the model has never seen. These patients would require a different evaluation approach (e.g., leave-one-trial-out).

### Progressive Disease Handling

Once a patient reaches PD (progressive disease), they remain PD:

```stan
// Check if there was already a PD in the in-sample period
int had_insample_pd = treat_visit_size > 0 &&
  full_predict_overall_recist[treat_visit_size] == PD;

if (had_insample_pd == 1) {
  // All forecast visits are PD since patient already had PD before cutoff
  full_predict_overall_recist[(treat_visit_size + 1):] = rep_array(PD, n_oos_visits);
}
```

### Other Events Integration

The model integrates other events (non-target progression, death) with target RECIST:

```stan
// If other events cause PD before target lesion PD
if (!forecast_other_events_censored) {
  // Find first forecast visit at or after other events PFS
  // Mark all subsequent visits as PD
  if (forecast_other_events_visit_idx <= n_oos_visits) {
    full_predict_overall_recist[(treat_visit_size + forecast_other_events_visit_idx):] =
      rep_array(PD, n_oos_visits - forecast_other_events_visit_idx + 1);
  }
}
```

---

## Practical Usage {#sec-lfo-usage}

### Running LFO Cross-Validation

LFO is implemented as a separate Stan model (`sf-ssls-lfo.stan`) that takes additional data:

```r
# Prepare LFO-specific data
lfo_data <- prepare_lfo_data(
  stan_data,
  cutoff_calendar_day = as.Date("2024-06-01")
)

# Compile LFO model
lfo_model <- cmdstan_model("stan/ssls/sf-ssls-lfo.stan")

# Run LFO
lfo_fit <- lfo_model$sample(
  data = lfo_data,
  iter_warmup = 1000,
  iter_sampling = 1000,
  chains = 4
)
```

### Extracting Results

```r
# Extract ELPD components
tumor_elpd <- lfo_fit$draws("patient_log_lik_tumor")
oe_elpd <- lfo_fit$draws("patient_log_lik_oe")
joint_elpd <- lfo_fit$draws("patient_log_lik")

# Extract confusion matrix
confusion <- lfo_fit$draws("oos_recist_confusion_matrix")
```

### Multiple Cutoffs

For robust evaluation, run LFO at multiple cutoffs:

```r
cutoff_dates <- seq(
  as.Date("2024-01-01"),
  as.Date("2024-12-01"),
  by = "2 months"
)

lfo_results <- map(cutoff_dates, function(cutoff) {
  lfo_data <- prepare_lfo_data(stan_data, cutoff_calendar_day = cutoff)
  lfo_model$sample(data = lfo_data, ...)
})
```

---

## Interpretation Guidelines {#sec-lfo-interpretation}

### ELPD Differences

When comparing models:

| ELPD Difference (SE) | Interpretation |
|---------------------|----------------|
| $> 4$ SE | Strong evidence for better model |
| $2-4$ SE | Moderate evidence |
| $< 2$ SE | Inconclusive / models similar |

### Confusion Matrix Metrics

From the RECIST confusion matrix, compute:

- **Accuracy**: Overall correct predictions / total predictions
- **Sensitivity (PD)**: True PD / (True PD + False non-PD)
- **Specificity (PD)**: True non-PD / (True non-PD + False PD)

::: {.callout-tip}
## Clinical Relevance

For treatment decisions, **sensitivity for PD** is particularly important - we want to correctly identify patients who will progress so appropriate interventions can be planned.
:::

### Time-Dependent Performance

Prediction accuracy typically:

- **Decreases** with forecast horizon (predicting further into the future is harder)
- **Improves** with more training data (later cutoffs have more historical observations)
- **Varies** by patient characteristics (some trajectories are more predictable)

---

## Comparison with Alternative Methods {#sec-lfo-alternatives}

### LOO (Leave-One-Out)

| Aspect | LFO | LOO |
|--------|-----|-----|
| Temporal integrity | Preserved | Violated |
| Information leakage | None | Possible |
| Clinical relevance | High | Low |
| Computational cost | Multiple fits | Single fit + PSIS |
| Implementation | More complex | Standard |

### K-Fold CV

| Aspect | LFO | K-Fold |
|--------|-----|--------|
| Fold definition | Time-based | Random |
| Within-patient correlation | Respected | Broken |
| Staggered enrollment | Handled naturally | Problematic |

### Prequential (Rolling Origin)

LFO is a form of prequential validation:

- **Prequential**: Expand training set one observation at a time
- **LFO**: Expand training set by calendar time cutoffs

LFO is more practical for clinical data where observations are naturally grouped by calendar time (data cutoffs).

---

## References {#sec-lfo-references}

1. Bürkner, P.-C., Gabry, J., & Vehtari, A. (2020). Approximate leave-future-out cross-validation for Bayesian time series models. *Journal of Statistical Computation and Simulation*, 90(14), 2499-2523.

2. Vehtari, A., Gelman, A., & Gabry, J. (2017). Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC. *Statistics and Computing*, 27(5), 1413-1432.

3. Gneiting, T., & Raftery, A. E. (2007). Strictly proper scoring rules, prediction, and estimation. *Journal of the American Statistical Association*, 102(477), 359-378.

---

::: {.related-pages}
## Related Pages

- [Model Validation](../analysis/model-validation.qmd) - LFO results and ELPD comparisons
- [Model Specification](model-specification.qmd) - Underlying model details
- [Architecture](architecture.qmd) - Code organization
- [Diagnostics](../analysis/diagnostics.qmd) - MCMC convergence checks
:::
