---
title: ""
format:
  revealjs:
    theme: [default, ../presentations/pioneer-gng/az-reveal-theme.scss]
    width: 1600
    height: 1130
    margin: 0.03
    center: false
    controls: false
    progress: false
    slide-number: false
    embed-resources: true
    auto-stretch: false
execute:
  warning: false
  echo: false
  cache: true
---

```{r setup}
#| include: false
#| cache: false

# Set working directory to project root for relative source() calls in dependencies
knitr::opts_knit$set(root.dir = here::here())
setwd(here::here())
source(here::here("quarto", "_shared", "_setup.R"))

# Compact report theme
report_theme <- theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(size = 11, face = "bold", color = AZ_navy),
    plot.subtitle = element_text(size = 8, color = AZ_darkgrey),
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 7),
    strip.text = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.35, "cm"),
    legend.position = "bottom",
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(-5, 0, 0, 0),
    plot.margin = margin(2, 4, 2, 2, "pt"),
    panel.grid.minor = element_blank(),
    plot.caption = element_text(size = 6.5, hjust = 0, color = AZ_darkgrey)
  )

# --- Data loading (multi-DCO) ---

quant_pfs_list <- list(
  apr25 = tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_aug25),
  jan26 = jan26_quant_pfs_data
)
cond_quant_pfs_list <- list(
  apr25 = tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_aug25),
  jan26 = jan26_cond_quant_pfs_data
)
obs_km_list <- list(
  apr25 = list(
    all = tar_read(km_trial_pfs_ctdna_apr25),
    pdl1_naive = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
    pdl1 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
    naive = tar_read(km_trial_pfs_naive_ctdna_apr25)
  ),
  aug25 = list(
    all = tar_read(km_trial_pfs_ctdna_aug25),
    pdl1_naive = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
    pdl1 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
    naive = tar_read(km_trial_pfs_naive_ctdna_aug25)
  ),
  jan26 = list(
    all = jan26_obs_km_all,
    pdl1_naive = jan26_obs_km_pdl1_naive,
    pdl1 = jan26_obs_km_pdl1,
    naive = jan26_obs_km_first_line
  )
)

oos_confusion_matrix <- tar_read(oos_confusion_matrix_ctdna_jan26, store = lfo_store)

all_lfo_clean <- bind_rows(
  ctdna = tar_read(tumor_ssls_lfo_clean_lite_ctdna_aug, store = lfo_store),
  no_hist = tar_read(tumor_ssls_lfo_clean_lite_no_hist_aug, store = lfo_store_no_hist),
  .id = "model"
) |> filter(n > 3)
```

## {.smaller}

<style>
.reveal .slide-footer { display: none; }
.report-title {
  color: var(--az-navy) !important;
  font-size: 1.3em !important;
  margin: 0 !important;
  text-align: center;
  font-weight: 700;
}
.report-subtitle {
  color: var(--az-navy);
  font-size: 0.6em;
  text-align: center;
  opacity: 0.85;
  margin: 2px 0 0 0;
}
.report-authors {
  font-size: 0.42em;
  color: var(--az-darkgrey);
  text-align: center;
  margin: 2px 0 0 0;
}
.gold-rule {
  height: 3px;
  background: var(--az-gold);
  border: none;
  margin: 4px 0 6px 0;
}
.text-col {
  display: flex;
  flex-direction: column;
  justify-content: center;
}
.text-col h3, .plot-col h3 {
  color: var(--az-navy) !important;
  font-size: 1.1em !important;
  margin: 6px 0 2px 0 !important;
}
.text-col h3:first-child { margin-top: 0 !important; }
.text-col p, .plot-col p { margin: 2px 0; font-size: 0.82em; line-height: 1.4; }
.text-col ul { margin: 2px 0; }
.text-col li { margin-bottom: 2px; font-size: 0.82em; line-height: 1.4; }
.reveal .columns { gap: 0.8em; }
.plot-col .cell { margin: 0 0 4px 0; }
.plot-col img { width: 100% !important; border-radius: 2px; }
.report-footer-bar {
  position: absolute;
  bottom: 5px;
  left: 3%;
  right: 3%;
  display: flex;
  justify-content: space-between;
  font-size: 0.36em;
  color: var(--az-navy);
  border-top: 2px solid var(--az-gold);
  padding-top: 3px;
}
</style>

::: {.report-title}
PIONEER: SCLC-01 Bayesian Forecasting
:::

::: {.report-subtitle}
Early Trial Outcome Predictions via Hierarchical Tumor Dynamics Modeling
:::

::: {.report-authors}
Karim Naguib, Roger Berch&eacute;, Antonia Bevan, Haroon Hasan, Jessica Davies, Paul Metcalfe | February 2026
:::

::: {.gold-rule}
:::

:::: {.columns}

::: {.column width="25%" .text-col}

### Key Findings

- **Stable PFS forecasts** (Panel A): Model predictions closely track observed Kaplan-Meier curves across three data cutoffs; credible intervals narrow as data matures.
- **Historical data strengthens predictions** (Panel B): Historical trial data produce better out-of-sample log scores than SCLC-01 data alone, confirming the value of hierarchical borrowing.
- **High RECIST accuracy** (Panel C): Leave-future-out cross-validation achieves 70--80% out-of-sample sensitivity for RECIST response categories.
- **Consistent median PFS** (Panel D): Median PFS estimates remain stable from the earliest (Apr 2025) to the latest (Jan 2026) cutoff across all subgroups.

:::

::: {.column width="36%" .plot-col}

```{r plot-a}
#| fig-width: 7.5
#| fig-height: 5.5

plot_sclc_km_by_dco(
  estimate_type = "sample",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26)
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_naive_ctdna_jan26)
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = tar_read(all_analysis_data_ctdna_jan26)
  ),
  show_priors = FALSE,
  fill_color = AZ_platinum,
  obs_color = AZ_plum
) +
  labs(title = "A: PFS Forecast by Patient Group and Data Cutoff") +
  report_theme +
  theme(
    legend.position = "bottom",
    legend.margin = margin(-2, 0, 0, 0),
    strip.text = element_text(size = 8, face = "bold")
  )
```

```{r plot-d}
#| fig-width: 7.5
#| fig-height: 2

all_lfo_clean |>
  plot_lfo_elpd_diff(
    baseline_model = "no_hist",
    model_labels = list(
      ctdna = "Using historical trials",
      no_hist = "Without Any Historical Data"
    ),
    add_vline = TRUE
  ) +
  labs(title = "B: Impact of Historical Data on Prediction") +
  report_theme +
  theme(
    legend.position = "bottom",
    legend.margin = margin(-2, 0, 0, 0)
  )
```

### Role of Historical Data

The model's three-level hierarchy (population &rarr; trial &rarr; patient) borrows from four historical trials: HISTORICAL, ARCTIC, PEARL, and ATLANTIC. Early in SCLC-01, when patient data is sparse, historical trials anchor population-level parameters. As SCLC-01 data accrues, trial- and patient-specific effects dominate, producing the stable forecasts in Panels A and D.

:::

::: {.column width="36%" .plot-col}

```{r plot-b}
#| fig-width: 7.5
#| fig-height: 4.0

plot_oos_confusion_matrix(oos_confusion_matrix, add_caption = FALSE) +
  labs(title = "C: RECIST Prediction Accuracy") +
  report_theme +
  theme(
    legend.position = "none",
    panel.grid = element_blank()
  )
```

```{r plot-c}
#| fig-width: 7.5
#| fig-height: 3.5

plot_median_surv_by_dco(
  "sample",
  quant_pfs_list,
  cond_quant_pfs_list,
  obs_km_list,
  show_observed = FALSE
) +
  labs(title = "D: Median PFS Stability Across Data Cutoffs") +
  report_theme +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.margin = margin(0, 0, -4, 0),
    legend.box.margin = margin(0, 0, -4, 0)
  )
```

### Conclusions

Mechanistic Bayesian modeling can deliver reliable endpoint forecasts 3--6 months earlier than conventional analysis and accelerate go/no-go decisions in oncology drug development.

:::

::::

::: {.report-footer-bar}
PIONEER | SCLC-01 Forecasting | Internal

Data Science Innovation & Lung Team, CDS
:::
