---
title: "Early Prediction of PFS from Confirmed Response and Tumor Sizes"
author: "Karim Naguib"
format: html
code-fold: true
reference-location: margin
fig-cap-location: margin
toc-expand: true 
---

```{r}
#| label: setup
#| include: false

library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(targets)
library(colorspace)

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

knitr::opts_chunk$set(cache.path = file.path("/wscratch", Sys.getenv("USER"), "adc-early-predict/cache/"))

ggplot2::theme_set(theme_minimal(base_family = "Arial"))

source(here("r", "util.R"))
source(here("r", "crcr.R"))

tar_config_set(store = file.path(tmp_dir, "_targets"), script = here("_targets.R"))
```

# Data

```{r}
#| label: load-analysis-data
#| cache: true
#| cache-lazy: true

tar_load(c("treated_confirmed_resp_analysis_data", "treated_confirmed_resp_stan_data", "all_confirmed_resp_analysis_data"))
tar_load("all_confirmed_resp_stan_data")
```

```{r}
#| label: fig-confirmed-response-week
#| fig-width: 10
#| fig-cap: Timing of classification or censoring (exiting study without classification).

treated_confirmed_resp_analysis_data |> 
  ggplot(aes(confirmed_response_week)) +
  geom_bar(aes(fill = confirmed_response_censored), alpha = 0.75, position = "stack") +
  scale_x_continuous("Confirmed Response Week", breaks = (seq(0, 126, by = 12))) +
  scale_fill_discrete("Unclassified", type = AZ_palette) +
  facet_wrap(vars(trial), scales = "free_y") 
```

```{r}
#| label: fig-response-classification-time
#| fig-cap: Week patients' confirmed response is known relative to the first week of the study (not the patient).

tar_read("km_trial_confirmed_response") |>
  rowwise() |>
  transmute(
    trial,
    conf_resp_km = list( 
      set_names(conf_resp_km, c("s", "n", "c", "e")) |>
        as_tibble() |> 
        mutate(t = seq(0, n() - 1))
    )
  ) |> 
  ungroup() |> 
  unnest(conf_resp_km) |> 
  ggplot() +
  geom_step(aes(t, s, color = trial), direction = "vh") +
  scale_x_continuous("Confirmed Response Study Week", breaks = (seq(0, 126, by = 12))) +
  scale_color_discrete("", type = AZ_palette) +
  labs(title = "Study time to confirmed response classification", y = "Proportion Unclassified") +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.7)) +
  NULL

tar_read("km_trial_confirmed_response") |> 
  rowwise() |>
  transmute(
    trial,
    conf_resp_km_calendar = list( 
      set_names(conf_resp_km_calendar, c("s", "n", "c", "e")) |>
        as_tibble() |> 
        mutate(t = seq(0, n() - 1))
    )
  ) |> 
  ungroup() |> 
  unnest(conf_resp_km_calendar) |> 
  ggplot() +
  geom_step(aes(t, s, color = trial), direction = "vh") +
  scale_x_continuous("Confirmed Response Calendar Week", breaks = (seq(0, 240, by = 12))) +
  scale_color_discrete("", type = AZ_palette) +
  labs(title = "Calendar time to confirmed response classification", y = "Proportion Unclassified") +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.7)) +
  NULL
```

```{r}
#| label: fig-assessment-visits
#| fig-cap: Number of visits every week.

treated_confirmed_resp_analysis_data |> 
  select(trial, n_tumors, t_measure) |> 
  unnest(t_measure) |> 
  unnest(t_measure) |> 
  mutate(epoch = if_else(t_measure <= 0, "pre-screening", "treatment") |> as_factor()) |> 
  ggplot(aes(t_measure)) +
  geom_bar(aes(fill = epoch), alpha = 0.5) +
  scale_fill_discrete("", labels = str_to_title, type = AZ_palette) +
  scale_x_continuous("Week", breaks = (seq(0, 126, by = 12))) +
  labs(y = "Count", title = "Number of visits every week") +
  facet_wrap(vars(trial), scales = "free_y") +
  theme(legend.position = "top", axis.text.x = element_text(size = 7)) +
  coord_cartesian(xlim = c(0, 120))
```

```{r}
#| label: fig-experiment-entry-cdf
#| fig-cap: Cumulative distribution of patients in the studies over calendar time (relative to the first weeks of the studies).

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  stat_ecdf(aes(x = experiment_start_week, color = trial)) +
  # geom_vline(xintercept = c(12, 24, 48), linetype = "dotted") +
  scale_x_continuous("Experiment Entry Calendar Week") +
  labs(title = "Empirical CDF of experiment entry", y = "Proportion of Patients") +
  scale_color_discrete("", type = AZ_palette) +
  # facet_wrap(vars(trial), scales = "free") +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.3)) +
  NULL
```




```{r}
#| label: fig-response-classification
#| fig-cap: Confirmed response counts.

treated_confirmed_resp_analysis_data |> 
  mutate(
    confirmed_response = factor(confirmed_response, levels = c(TRUE, FALSE), labels = c("Response", "Non-response")) |> 
      fct_na_value_to_level("Unclassified")
  ) |>  
  ggplot(aes(confirmed_response)) +
  geom_bar(alpha = 0.25) +
  labs(title = "Confirmed response counts", x = "", y = "Count") +
  facet_wrap(vars(trial), nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-age-group
#| fig-cap: Distribution of age groups.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(ecogbl), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-hormon
#| fig-cap: Distribution of HR status.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(hormonr), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-prior-pertuzumab-treatment
#| fig-cap: Distribution of prior pertuzumab treatment. 

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(prior_pertuzumab_treatment), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-prior-cdk46-inhibit-treatment
#| fig-cap: Distribution of prior cdk46 inhibition treatment.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(prior_cdk46_inhibit_treatment), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-her2-status
#| fig-cap: Distribution of HER2 status.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(her2_status), alpha = 0.25) +
  facet_wrap(vars(trial))
```

```{r}
#| label: fig-history-visceral-disease
#| fig-cap: Distribution of a history of visceral disease.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(hist_visceral_disease), alpha = 0.25) +
  facet_wrap(vars(trial))
```

# Model

## Confirmed Response Competing Risks Model {#crcr}

* $Y^\textrm{crcr}_{rt}$ is an indicator of whether a patient become classified as $r$ at $t$.
* $T^\textrm{crcr} \in \mathcal{N}^+$ is the number of weeks a patient remains unclassified.
* $C^\textrm{crcr}$ is an indicator of whether a patient is censored (stays unclassified in the available data).
* $R$ is an indicator of whether a patient eventually will be a confirmed responder or non-responder. 
* $\boldsymbol{X}$ is a vector patient characteristics and predictors. 

First, we have a competing risks model of time-to-classification, with a separate hazard function per risk (confirmed response vs confirmed non-response).

$$
\lambda^\textrm{crcr}_r(t;\boldsymbol{x,\theta}^\textrm{crcr}_r) = \lambda^\textrm{crcr}_{0rt}\exp(\boldsymbol{x\beta}^\textrm{crcr}_r),\,r \in \{\textrm{responder, non-responder}\} 
$$
and the baseline hazard is distributions as a Gaussian process, $\boldsymbol{\lambda}^\textrm{crcr}_{0r} \sim \mathtt{GP}(K^\textrm{crcr}_r)$.

The conditional probability of classification as a responder or non-responder (exiting the state of being unclassified) is

$$
\begin{align}
\Pr[Y^\textrm{crcr}_{rt} = 1 \mid Y^\textrm{crcr}_{r,t-1} = 0, \boldsymbol{X = x, \theta}^\textrm{crcr}_r] &= 1 - \exp[-\lambda^\textrm{crcr}_r(t;\boldsymbol{x,\beta^\textrm{crcr}_r})] \\
\Pr[Y^\textrm{crcr}_{rt} = 1 \mid Y^\textrm{crcr}_{r,t-1} = 1, \boldsymbol{X = x, \theta}^\textrm{crcr}_r] &= 1 \\
\end{align}
$$

The patient level log likelihood is
$$
\log \mathcal{L}^\textrm{crcr}_i = (1-C^\textrm{crcr}_i)\log[1 - \exp(-\lambda^\textrm{crcr}_{R_i}(T^\textrm{crcr}_i + 1;\boldsymbol{X_i, \theta^\textrm{crcr}_r}))] + \sum_{r = 0}^1\sum_t^{T^\textrm{crcr}_i} - \lambda^\textrm{crcr}_{r}(t;\boldsymbol{X_i, \theta^\textrm{crcr}_r})
$$

```{r}
#| label: fig-survival-unclassified
#| fig-cap: Survival plots of time to classification.
#| cache: true

tar_load("conf_resp_hb")

bind_rows(
  prior = tar_read("prior_all_rep_confirmed_response_bindist"),
  posterior = tar_read("all_rep_confirmed_response_bindist"),
  .id = "fit_type"
) |> 
  ggplot(aes(t)) +
  stat_lineribbon(
    aes(ydist = bindist, color = fit_type, fill = fit_type, alpha = fit_type),
    linewidth = 1, step = "hv", .width = c(0.5, 0.8)
  ) +
  geom_step(
    aes(y = count, linetype = "Observed"), direction = "hv", linewidth = 0.5, 
    data = \(d) tar_read("treated_confirmed_resp_analysis_data") |> 
      semi_join(d, by = "trial") |> 
      group_by(trial) |> 
      reframe(t = conf_resp_hb[-length(conf_resp_hb)], count = sample_hist(confirmed_response_week, conf_resp_hb))
  ) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  scale_linetype_manual("", values = c(Observed = "dotted")) +
  scale_x_continuous("", breaks = seq(0, 150, 12)) +
  facet_wrap(vars(trial), scales = "free_y") +
  theme(legend.position = "bottom") +
  guides(alpha = "none") +
  NULL
```

```{r}
#| label: fig-conf-resp-hazard-ratio
#| fig-cap: Probabilistic histograms of the hazard ratio in response to patient characteristics, separated by competing risk.
#| cache: true

bind_rows(
  prior = tar_read("prior_all_crcr_hazard_ratio_bindist"),
  posterior = tar_read("all_crcr_hazard_ratio_bindist"),
  .id = "fit_type"
) |>
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = bindist, color = fit_type, fill = fit_type, alpha = fit_type),
                  linewidth = 1,
                  step = "hv",
                  .width = c(0.5, 0.8)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  facet_grid(vars(trial), vars(k), scales = "free_y") +
  guides(alpha = "none") +
  NULL
```

```{r}
#| eval: false
#| include: false

covar_param <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_covar_param, .id = "fit_type") 

covar_param |> 
  mutate(k = factor(k, levels = 1:2, labels = c("Response", "Non-response"))) |> 
  ggplot() +
  stat_halfeye(aes(xdist = crcr_covar_effect, y = factor(covar), color = fit_type, fill = fit_type), alpha = 0.25, .width = 0.8) +
  scale_y_discrete("Parameter") +
  scale_color_discrete(type = AZ_palette, aesthetics = c("color", "fill")) +
  facet_grid(vars(k), vars(trial))
```

```{r}
#| eval: false
#| include: false

tumor_param <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_tumor_param, .id = "fit_type") 

tumor_param |> 
  mutate(k = factor(k, levels = 1:2, labels = c("Response", "Non-response"))) |> 
  ggplot() +
  stat_halfeye(aes(xdist = crcr_tumor_stim_pop_coef, y = factor(m), color = fit_type, fill = fit_type), alpha = 0.25, .width = 0.8) +
  scale_y_discrete("Parameter") +
  scale_color_discrete(type = AZ_palette, aesthetics = c("color", "fill")) +
  facet_grid(vars(k), vars(trial))
```

```{r}
#| eval: false
#| include: false

conf_resp_cif <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_cif, .id = "fit_type") 

conf_resp_prob_confirmed <- conf_resp_cif |> 
  group_by(fit_type, trial, i, k) |> 
  filter(min_rank(t) == n()) |> 
  group_by(fit_type, trial, i) |> 
  mutate(cif = cif / rvar_sum(cif)) |> 
  ungroup() |> 
  filter(k == 2)

conf_resp_cif |> 
  filter(trial == "Breast04", i <= 5) |>
  mutate(k = factor(k, levels = 1:2, labels = c("Response", "Non-response"))) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = cif, color = fit_type, fill = fit_type), alpha = 0.25, linewidth = 2, .width = c(0.5, 0.8)) +
  scale_color_discrete(type = AZ_palette, aesthetics = c("color", "fill")) +
  facet_grid(vars(i), vars(k))
```

```{r}
#| eval: false
#| include: false

conf_resp_prob <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_prob, .id = "fit_type") 
```

```{r}
#| label: fig-cif
#| fig-cap: Probabilistic histograms of the Cumulative Incidence Function (CIF) for the different competing risks.
#| fig-width: 10
#| cache: true

bind_rows(
  prior = tar_read("prior_all_conf_resp_cif_bindist"),
  posterior = tar_read("all_conf_resp_cif_bindist"),
  .id = "fit_type"
) |> 
  ggplot(aes(p)) +
  stat_lineribbon(aes(ydist = cif_bindist, color = fit_type, fill = fit_type, alpha = fit_type),
                  linewidth = 1,
                  step = "vh",
                  .width = c(0.5, 0.8)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  scale_y_continuous("", breaks = NULL) +
  labs(x = "CIF") +
  facet_grid(vars(t), vars(trial, k), scales = "free_y", switch = "y", labeller = labeller(t = \(t) str_c("t = ", t))) +
  guides(alpha = "none") +
  theme(legend.position = "bottom", axis.text.x = element_blank(), strip.text.y.left = element_text(angle = 0), strip.placement = "outside") +
  NULL
```


## Recruitment Process

We model the recruitment process or the patients' entry into the experiment calendar week as

$$
W_i \sim \mathtt{NegBinomial}(\lambda^\textrm{recruit}, \phi^\textrm{recruit}),
$$
where $\lambda^\textrm{recruit}$ is the average calendar entry week and $\phi^\textrm{recruit}$ represents dispersion. At first, I tried a Poisson distribution but it was not sufficiently dispersed.

```{r}
#| label: fig-recruit-distribution-param
#| fig-cap: Parameters of the recruitment process distribution.

recruit_res_list <- lst(
  prior = tar_read("prior_recruit_fit"),
  posterior = tar_read("recruit_fit")
) 

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, lambda[trial]), .id = "fit_type") |> 
  ggplot() +
  stat_halfeye(aes(xdist = lambda, y = trial, color = fit_type, fill = fit_type, alpha = fit_type)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Expected Experiment Entry Calendar Week", y = "") +
  guides(alpha = "none") +
  coord_cartesian(xlim = c(0, 100)) +
  theme(legend.position = "bottom") +
  NULL

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, phi[trial]), .id = "fit_type") |> 
  ggplot() +
  stat_halfeye(aes(xdist = phi, y = trial, color = fit_type, fill = fit_type, alpha = fit_type)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Dispersion", y = "") +
  guides(alpha = "none") +
  coord_cartesian(xlim = c(0, 25)) +
  theme(legend.position = "bottom") +
  NULL

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, recruit_sd[trial]), .id = "fit_type") |> 
  ggplot() +
  stat_halfeye(aes(xdist = recruit_sd, y = trial, color = fit_type, fill = fit_type, alpha = fit_type)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Standard Deviation", y = "") +
  guides(alpha = "none") +
  coord_cartesian(xlim = c(0, 50)) +
  NULL
```

```{r}
#| label: fig-entry-week-cdf
#| fig-cap: Cumulative distribution of entry into the experiment over calendar weeks. 
#| cache: true

ecdf_breaks <- 1:125

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, sim_experiment_start_week[trial, i]), .id = "fit_type") |> 
  group_by(fit_type, trial) |> 
  reframe(t = ecdf_breaks, ecdf_rvar = rfun(\(w) ecdf(w)(ecdf_breaks))(sim_experiment_start_week)) |> 
  ggplot(aes(t)) +
  stat_ecdf(aes(x = experiment_start_week), linetype = "dashed", data = tar_read("treated_pfs_analysis_data")) +
  stat_lineribbon(aes(ydist = ecdf_rvar, color = fit_type, fill = fit_type, alpha = fit_type), .width = 0.8) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Calendar Week", y = "Proportion in Experiment") +
  guides(alpha = "none") +
  facet_wrap(vars(trial)) +
  theme(legend.position = "bottom")
```

Below, we predict sample sizes and (confirmed response) maturity rate: we use the recruitment process model to simulate patients entering the trial (at varying rates) as we see how many we have at various prediction calendar weeks. For the maturity rate, we confirmed response classification calendar weeks simulated from the [confirmed response model](#crcr). 

```{r}
#| label: fig-recruit-sample-size-and-maturity
#| fig-cap: Simulations to see sample sizes and data maturity with varying prediction weeks and recruitment rates.
#| cache: true

tar_load("recruit_maturity")

recruit_maturity |>
  filter(Pr(n_sample >= 10) >= 0.8) |> 
  ggplot() +
  stat_lineribbon(aes(pred_week, ydist = n_sample, fill = lambda, color = lambda, group = str_c(trial, lambda)), alpha = 0.25, .width = 0.8) +
  geom_hline(yintercept = 10, linetype = "dotted") +
  scale_x_continuous(breaks = seq(6, 48, 6)) +
  scale_color_continuous("Recruitment Rate") +
  labs(x = "Prediction Calendar Week", y = "Sample Size") +
  facet_grid(col = vars(trial), scales = "free_y") +
  NULL

recruit_maturity |> 
  filter(Pr(n_sample >= 10) >= 0.8) |> 
  ggplot() +
  stat_lineribbon(
    aes(pred_week, ydist = maturity_rate, fill = lambda, color = lambda, group = lambda), alpha = 0.25, .width = 0.8
  ) +
  scale_x_continuous(breaks = seq(6, 48, 6)) +
  scale_color_continuous("Recruitment Rate") +
  labs(x = "Prediction Calendar Week", y = "Maturity Rate") +
  facet_grid(col = vars(trial), scales = "free_y") +
  NULL
```

## PFS Model

* $Y^\textrm{pfs}_{t}$ is an indicator of whether a patient experiences disease progression at $t$.
* $T^\textrm{pfs} \in \mathcal{N}^+$ is the number of weeks a patient progression-free.
* $C^\textrm{pfs}$ is an indicator of whether a patient is censored.

First, we have a progression-free survival model,

$$
\lambda^\textrm{pfs}(t;\boldsymbol{x}, r,\boldsymbol{\theta}^\textrm{pfs}) = \lambda^\textrm{pfs}_{0t}\exp(\boldsymbol{x\beta}^\textrm{pfs} + r\rho).
$$
and the baseline hazard is distributions as a Gaussian process, $\boldsymbol{\lambda}^\textrm{pfs}_{0} \sim \mathtt{GP}(K^\textrm{pfs})$.

The conditional probability of progression is

$$
\begin{align}
\Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, R = r, \boldsymbol{\theta}^\textrm{pfs}] &= 1 - \exp[-\lambda^\textrm{pfs}(t;\boldsymbol{x},r,\boldsymbol{\beta}^\textrm{pfs},\rho)] \\
\Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 1, \boldsymbol{X = x}, R = r, \boldsymbol{\theta}^\textrm{pfs}] &= 1
\end{align}
$$
If a patient's response remains unclassified, their conditional probability is a mixture of based on the probability of a patient's eventual classification,
$$
\begin{gather}
\Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}] =\\ \gamma(\boldsymbol{x, \theta}^\textrm{crcr}) \cdot \Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, R = 1, \boldsymbol{\theta}^\textrm{pfs}] + (1 - \gamma(\boldsymbol{x, \theta}^\textrm{crcr})) \cdot \Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, R = 0, \boldsymbol{\theta}^\textrm{pfs}].
\end{gather}
$$

The probability that a patient would be classified as a confirmed responder is^[We assume that eventually everyone is classified. We substitute the limit of CIF with CIF at the last week we have in the trial, which we then normalize.]
$$
\gamma(\boldsymbol{x, \theta}^\textrm{crcr}) = 
  \begin{cases}
    R_i & \text{if } C^\textrm{crcr} = 0 \\
    \lim_{t\rightarrow\infty}\mathsf{CIF}_1(t, \boldsymbol{x, \theta}^\textrm{crcr}) & \text{if } C^\textrm{crcr} = 1,
  \end{cases}
$$
and the cumulative incidence function (CIF) for is defined as
$$
\mathsf{CIF}_r(t, \boldsymbol{x, \theta}^\textrm{crcr}) = \sum_{\tau = 0}^t S^\textrm{crcr}(t,\boldsymbol{x, \theta}^\textrm{crcr})\cdot \Pr[Y^\textrm{crcr}_{r,\tau+1} = 1 \mid Y^\textrm{crcr}_{r,\tau} = 0, \boldsymbol{x, \theta}^\textrm{crcr}_r],
$$
where $S^\textrm{crcr}(t,\boldsymbol{x, \theta}^\textrm{crcr}) = \prod_{r=0}^1\prod_{\tau = 1}^t \Pr[Y^\textrm{crcr}_{r\tau} = 0 \mid Y^\textrm{crcr}_{r,\tau-1} = 0, \boldsymbol{x, \theta}^\textrm{crcr}_r]$ is the probability of remaining unclassified up to week $t$.

The log likelihood per patient is
$$
\log \mathcal{L}^\textrm{pfs}_i = (1-C^\textrm{pfs}_i)\log \Pr[Y^\textrm{pfs}_{T^\textrm{pfs}_i} = 1 \mid Y^\textrm{pfs}_{T^\textrm{pfs}_i-1} = 0, \boldsymbol{x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}] + \log S^\textrm{pfs}(T^\textrm{pfs}_i - 1, \boldsymbol{x},\boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}) + \mathcal{L}_i^\textrm{crcr}
$$

where $S^\textrm{pfs}(t,\boldsymbol{x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}) = \prod_{\tau = 1}^t \Pr[Y^\textrm{pfs}_{\tau} = 0 \mid Y^\textrm{pfs}_{\tau-1} = 0, \boldsymbol{x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}]$ is the probability of remaining progress free up to week $t$.

```{r}
#| label: load-mpfs

pfs_cr_median_pfs <- bind_rows(
  prior = tar_read("prior_trial_conf_resp_median_pfs"), 
  posterior = tar_read("trial_conf_resp_median_pfs"),
  .id = "fit_type"
)  
```

```{r}
#| label: fig-median-pfs-conf-resp
#| fig-cap: Estimated median PFS using full trial data.

pfs_cr_median_pfs |> 
  ggplot() +
  stat_halfeye(
    aes(xdist = sim_trial_median_pfs, y = trial, color = fit_type, fill = fit_type, alpha = fit_type), 
    na.rm = TRUE, 
    .width = c(0.5, 0.8)
  ) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(y = "", x = "Median PFS [Weeks]") +
  guides(alpha = "none") +
  theme(legend.position = "bottom")
  NULL
```

```{r}
#| label: fig-pfs-conf-resp-km
#| fig-cap: Survival curves as estimated by the model and compared to the observed (KM) proportions, using full trial data.
#| cache: true

bind_rows(
  prior = tar_read("prior_trial_conf_resp_km_est"),
  posterior = tar_read("trial_conf_resp_km_est"), 
  .id = "fit_type"
) |> 
  ggplot() +
  geom_step(
    aes(x = t, y = s, group = btype, linetype = "Observed Data"), 
    data = tar_read("km_res_calculated") |> filter(treated, fct_match(btype, "lb"))
  ) +
  stat_lineribbon(aes(x = t - 1, ydist = trial_km_est, color = fit_type, fill = fit_type, alpha = fit_type), step = TRUE, .width = c(0.8)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.0625, posterior = 0.25)) +
  scale_linetype_manual("", values = "dashed") +
  labs(x = "Week", y = "Survival Probability") +
  facet_wrap(vars(trial)) +
  guides(alpha = "none") + 
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-pfs-conf-resp-hazard-ratio
#| fig-cap: Probabilistic histograms of $\exp(\boldsymbol{x\beta}^\textrm{pfs} + r\rho)$ separated by $r = 0$ (confirmed non-response) and $r = 1$ (confirmed response). 
#| fig-width: 10
#| cache: true

hb <- seq(0, 3, by = 0.1)

bind_rows(
  prior = tar_read("prior_pfs_conf_resp_hazard_ratio"),
  posterior = tar_read("pfs_conf_resp_hazard_ratio"),
  .id = "fit_type"
) |>   
  group_by(trial, k, fit_type) |> 
  reframe(t = hb[-1], bindist = rvar_sample_hist(time_invariant_hazard_ratio, hb)) |> 
  mutate(k = factor(k, levels = 1:2, labels = c("Non-response", "Response"))) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = bindist, color = fit_type, fill = fit_type, alpha = fit_type),
                  linewidth = 1,
                  step = "vh",
                  .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Hazard Ratio", y = "Density") +
  facet_grid(vars(trial), vars(k), scales = "free_y", switch = "y") +
  guides(alpha = "none") +
  theme(legend.position = "bottom", strip.text.y.left = element_text(angle = 0), strip.placement = "outside") +
  NULL
```

### Simulations 

Here we are using leave-one-trial-out evaluation of the prediction accuracy of training our models on three trials and holding the fourth out for testing.

::: {.callout-note title="Algorithm" collapse="true"}

* $\mathcal{D}_k$ is the data from trial $k \in \{1,\ldots,K\}$. $K$ is the number of trials available.
* $B$ is the number of sampling iterations.
* $\mathcal{M}$ is a set of rates of confirmed response data maturity to test. 
* $y^\textrm{pfs}$ and $y^\textrm{cr}$ are progression-free survival and confirmed response, respectively.

For every trial $k$,

* Train our models on the full data, holding out $\mathcal{D}_k$: $p_\textrm{post}(y^\textrm{cr}, y^\textrm{pfs}\mid \boldsymbol{x})$. 
* For each sampling iteration, $b \in \{1,\ldots,B\}$, 
    - Draw a random subsample, $\widetilde{\mathcal{D}}_{kb}$, of 30 patients (with replacement) from $\mathcal{D}_k$.
    - For each $m \in \mathcal{M}$,
      * Determine the set of patients who would be in the experiment such that the proportion of patients (out of 30) who have their confirmed response classified is at least $m$, $\widetilde{\mathcal{D}}_{kbm}$.
      * For each $i \in \widetilde{\mathcal{D}}_{kbm}$,
        - If $y^\textrm{pfs}_i$ is right censored,
          * If $y^\textrm{cr}_i$ is unclassified (censored), predict $(\tilde{y}^\textrm{pfs}_i, \tilde{y}^\textrm{cr}_i) \sim p_\textrm{post}(\cdot, \cdot\mid y^\textrm{pfs}_i, y^\textrm{cr}_i, \boldsymbol{x}_i)$.
          * else, let $\tilde{y}^\textrm{cr}_i = y^\textrm{cr}_i$ and predict $\tilde{y}^\textrm{pfs}_i \sim p_\textrm{post}(\cdot\mid y^\textrm{pfs}_i, y^\textrm{cr}_i, \boldsymbol{x}_i)$.
        - else, let $\tilde{y}^\textrm{pfs}_i = y^\textrm{pfs}_i$.
      * Calculate the median of $\boldsymbol{\tilde{y}}^\textrm{pfs}_{kbm}$. 
      
:::
          

```{r}
#| label: load-bootstrap-mpfs

tar_load(c("pfs_conf_resp_bootstrap_cr_median_pfs", "pfs_conf_resp_bootstrap_pfs_median_pfs"))
```

```{r}
#| label: fig-pfs-conf-resp-bootstrap-cr
#| fig-cap: Bootstrap sample median PFS prediction ribbons (50 and 80% credible intervals) over a range of confirmed response data maturity rates. Missing ribbons indicate insufficient bootstrap samples at that data maturity rate. The full data intervals shows the 80% credible intervals. 

max_all_t <- 162

pfs_conf_resp_bootstrap_cr_median_pfs |> 
  filter(bs_cr_median_pfs.upper < max_all_t) |>
  ggplot(aes(bootstrap_cr_maturity_rates)) +
  geom_lineribbon(
    aes(y = bs_cr_median_pfs, ymin = bs_cr_median_pfs.lower, ymax = bs_cr_median_pfs.upper, group = .width), 
    alpha = 0.25, size = 0.5 
  ) +
  stat_pointinterval(
    aes(ydist = sim_trial_median_pfs, color = "Full Data"),
    .width = 0.8, linewidth = 0.5, 
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |>
      filter(fct_match(fit_type, "posterior")) |>
      mutate(bootstrap_cr_maturity_rates = 0.8)
  ) +
  geom_text(
    aes(x = x, y = y, label = str_c("N = ", n), color = "Full Data"), size = 3, hjust = 1, show.legend = FALSE, 
    data = count(all_confirmed_resp_analysis_data, trial) |> mutate(x = 0.8, y = 110) 
  ) +
  scale_color_manual("", values = AZ_pink) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette) +
  labs(x = "Confirmed Response Data Maturity Rate", y = "mPFS") +
  facet_wrap(vars(trial)) +
  coord_cartesian(xlim = c(0.2, 0.8), ylim = c(10, 140))
```

```{r}
#| label: fig-pfs-conf-resp-bootstrap-pfs
#| fig-cap: Bootstrap sample median PFS prediction ribbons (50 and 80% credible intervals) over a range of confirmed response data maturity rates. Missing ribbons indicate insufficient bootstrap samples at that data maturity rate. The full data intervals shows the 80% credible intervals. 

pfs_conf_resp_bootstrap_pfs_median_pfs |> 
  filter(bs_pfs_median_pfs.upper < max_all_t) |>
  ggplot(aes(bootstrap_pfs_maturity_rates)) +
  geom_lineribbon(
    aes(y = bs_pfs_median_pfs, ymin = bs_pfs_median_pfs.lower, ymax = bs_pfs_median_pfs.upper, group = .width), 
    alpha = 0.25, linewidth = 0.5 
  ) +
  stat_pointinterval(
    aes(ydist = sim_trial_median_pfs, color = "Full Data"),
    .width = 0.8, size = 0.5,
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |>
      filter(fct_match(fit_type, "posterior")) |>
      mutate(bootstrap_pfs_maturity_rates = 0.8)
  ) +
  geom_text(
    aes(x = x, y = y, label = str_c("N = ", n), color = "Full Data"), 
    size = 3, hjust = 1, show.legend = FALSE, 
    data = count(all_confirmed_resp_analysis_data, trial) |> mutate(x = 0.8, y = 110) 
  ) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette) +
  scale_color_manual("", values = AZ_pink) +
  labs(x = "Confirmed Response Data Maturity Rate", y = "mPFS") +
  facet_wrap(vars(trial)) +
  coord_cartesian(xlim = c(0.2, 0.8), ylim = c(10, 140))
```

* What we are observing here is the probability ranges of leave-one-trial-out prediction, so those intervals combine sampling uncertainty from the bootstrap sample and the prediction model uncertainty.
  - I cannot make the decision on the acceptability of these wide intervals for decision-making.
* The models seem to be unbiased since they tend to the full data estimation intervals. Given that this isn’t a hierarchical model this means that the covariates used are doing a relatively good job.
* There is little improvement as we wait for more data maturity (this is maybe because most patients eventually reach confirmed response, so there is little information here anyway). Perhaps focusing on PFS maturity rate might be more helpful.

```{r}
#| fig-cap: Evaluating the change in precision of the predicted distribution of median PFS, in response to changes in the (i) prediction calendar week and (ii) the patient recruitment rate. The prediction model is trained on the full data while simulation uses a bootstrap subsample.
#| fig-height: 5 
#| fig-width: 6
#| eval: false
#| include: false

pfs_conf_resp_bootstrap_median_pfs |> 
  ggplot(aes(x = prediction_week)) +
  geom_lineribbon(
    aes(y = bootstrap_median_pfs, ymin = bootstrap_median_pfs.lower, ymax = bootstrap_median_pfs.upper, color = "Bootstrap", fill = "Bootstrap"),
    alpha = 0.25
  ) +
  stat_pointinterval(aes(ydist = sim_median_pfs, color = "Full Data"),
    .width = 0.8, size = 0.5,
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |> filter(fct_match(fit_type, "posterior")) |> mutate(prediction_week = 48)
  ) +
  scale_color_manual("", values = c(AZ_platinum, AZ_plum), aesthetics = c("color", "fill")) +
  scale_x_continuous("Prediction Calendar Week", breaks = seq(0, 48, 12)) +
  labs(y = "Median PFS [Weeks]") +
  facet_grid(
    col = vars(recruit_lambda), row = vars(trial), 
    scales = "free", space = "free_x"
  ) +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7), strip.text.y.right = element_text(size = 6.5)) +
  NULL
```

```{r}
#| label: fig-bootstrap-subsample-size
#| fig-cap: Bootstrap sample sizes (50 and 80% probability intervals) over a range of confirmed response data maturity rates. The dotted lines shows the maximum size of 30 patients.

pfs_conf_resp_bootstrap_median_pfs |> 
  mutate(across(starts_with("n_bs_sample"), \(n) if_else(n == 0, 30, n))) |> 
  ggplot(aes(bootstrap_cr_maturity_rates, n_bs_sample)) +
  geom_lineribbon(
    aes(ymin = n_bs_sample.lower, ymax = n_bs_sample.upper, group = .width), 
    color = AZ_platinum, fill = AZ_platinum, alpha = 0.25, 
  ) +
  geom_hline(yintercept = 30, linetype = "dotted") +
  labs(x = "Confirmed Response Data Maturity Rate", y = "Prediction Sample Size") +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(10, 30)) +
  NULL
```

```{r}
#| label: fig-bootstrap-prediction-week
#| fig-cap: Bootstrap prediction calendar weeks (50 and 80% probability intervals) over a range of confirmed response data maturity rates. The dotted line indicates the calendar week when the last (uncensored) patient was classified as a confirmed response or non-response.

last_mature_week_data <- all_confirmed_resp_analysis_data |> 
  group_by(trial) |> 
  summarize(last_mature_week = max((1 - confirmed_response_censored) * (experiment_start_week + confirmed_response_week - 1)))

pfs_conf_resp_bootstrap_median_pfs |> 
  left_join(last_mature_week_data, by = "trial") |> 
  mutate(across(starts_with("bs_prediction_calendar_week"), \(w) if_else(w > 1000, last_mature_week, w))) |> 
  ggplot(aes(bootstrap_cr_maturity_rates, bs_prediction_calendar_week)) +
  geom_lineribbon(
    aes(ymin = bs_prediction_calendar_week.lower, ymax = bs_prediction_calendar_week.upper, group = .width), 
    color = AZ_platinum, fill = AZ_platinum, alpha = 0.25, 
  ) +
  geom_hline(
    aes(yintercept = last_mature_week, linetype = "Week of last Confirmed Response classification, in the full data."),
    data = last_mature_week_data 
  ) +
  scale_linetype_manual("", values = "dotted") + 
  labs(x = "Confirmed Response Data Maturity Rate", y = "Prediction Calendar Week") +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(10, 200)) +
  theme(legend.position = "bottom") +
  NULL
```
* By the maturity rate of 40% we almost have all the 30 patients in the trial already.
* We could be making these predictions by week 50, a significant decrease from waiting till everyone (in the full trial sample) has reached confirmed response classification. That is longer than we saw with the previous in-sample evaluation, but that’s comparing apples to oranges.
 
Questions remaining:

* How much would things improve when using a hierarchical model?
* Would we gain much in updating the model from the small number of patients observed from the held out trial?
* How much would things improve if we waited for longer PFS maturity?
* How much would things improve if in included dynamic confirmed response and tumor sizes?


```{r}
#| fig-cap: Examining how the bootstrap sample size changes in response to changes in prediction calendar week and patient recruitment rate. 
#| fig-height: 5 
#| fig-width: 6
#| eval: false
#| include: false

pfs_conf_resp_bootstrap_median_pfs |> 
  ggplot(aes(x = prediction_week)) +
  geom_lineribbon(
    aes(y = n_bootstrap_sample, ymin = n_bootstrap_sample.lower, ymax = n_bootstrap_sample.upper),
    color = AZ_platinum, fill = AZ_platinum, alpha = 0.25
  ) +
  geom_hline(aes(yintercept = n), linetype = "dotted", data = count(treated_confirmed_resp_analysis_data, trial)) +
  scale_x_continuous("Prediction Week", breaks = seq(0, 48, 12)) +
  labs(y = "Subsample Size", caption = "The dotted line indicates the full sample size.") +
  facet_grid(
    col = vars(recruit_lambda), row = vars(trial), 
    scales = "free", space = "free_x"
  ) +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7), strip.text.y.right = element_text(size = 6.5)) +
  NULL
```

<!-- Below, we consider how data maturity rate changes. We define data maturity as the proportion of patients who have entered the trial by the prediction week for whose response has been classified. -->

```{r}
#| fig-cap: Examining how data maturity changes in response to changes in prediction calendar week and patient recruitment rate. 
#| fig-height: 5 
#| fig-width: 6
#| eval: false
#| include: false

pfs_conf_resp_bootstrap_median_pfs |> 
  ggplot(aes(x = prediction_week)) +
  geom_lineribbon(
    aes(y = bootstrap_maturity_rate, ymin = bootstrap_maturity_rate.lower, ymax = bootstrap_maturity_rate.upper),
    color = AZ_platinum, fill = AZ_platinum, alpha = 0.25
  ) +
  # geom_hline(aes(yintercept = m), linetype = "dotted", 
  #            data = treated_confirmed_resp_analysis_data |> group_by(trial) |> summarize(m = 1 - mean(confirmed_response_censored))) +
  scale_x_continuous("Prediction Week", breaks = seq(0, 48, 12)) +
  labs(y = "Subsample Data Maturity Rate") +
  facet_grid(
    col = vars(recruit_lambda), row = vars(trial), 
    scales = "free", space = "free_x"
  ) +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7), strip.text.y.right = element_text(size = 6.5)) +
  NULL
```