---
title: "Early Prediction of PFS from Confirmed Response and Tumor Sizes"
author: 
  - Paul Metcalfe
  - Karim Naguib
format: html
code-fold: true
reference-location: margin
fig-cap-location: bottom 
toc-expand: true 
toc-depth: 4
lightbox: true
---

```{r}
#| label: setup
#| include: false

library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(targets)
library(colorspace)
library(ggrepel)

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

knitr::opts_chunk$set(cache.path = file.path("/wscratch", Sys.getenv("USER"), "adc-early-predict/cache/"))

ggplot2::theme_set(theme_minimal(base_family = "Arial"))

source(here("r", "util.R"))
source(here("r", "crcr.R"))

tar_config_set(store = file.path(tmp_dir, "_targets"), script = here("_targets.R"))
```

# Data

```{r}
#| label: load-analysis-data

tar_load(c("treated_confirmed_resp_analysis_data", "treated_confirmed_resp_stan_data", "all_confirmed_resp_analysis_data"))
tar_load("all_confirmed_resp_stan_data")
```

```{r}
#| label: fig-confirmed-response-week
#| fig-width: 10
#| fig-cap: Timing of classification or censoring (exiting study without classification).

treated_confirmed_resp_analysis_data |> 
  ggplot(aes(confirmed_response_week)) +
  geom_bar(aes(fill = confirmed_response_censored), alpha = 0.75, position = "stack") +
  scale_x_continuous("Confirmed Response Week", breaks = (seq(0, 126, by = 12))) +
  scale_fill_discrete("Unclassified", type = AZ_palette) +
  facet_wrap(vars(trial), scales = "free_y") 
```

```{r}
#| label: fig-response-classification-time
#| fig-cap: "Week patients' confirmed response is known." 
#| fig-subcap:
#|    - Relative to the first week the patient is in the study.  
#|    - Relative to the first week of the study (not the patient).

tar_read("km_trial_confirmed_response") |>
  rowwise() |>
  transmute(
    trial,
    conf_resp_km = list( 
      set_names(conf_resp_km, c("s", "n", "c", "e")) |>
        as_tibble() |> 
        mutate(t = seq(0, n() - 1))
    )
  ) |> 
  ungroup() |> 
  unnest(conf_resp_km) |> 
  ggplot() +
  geom_step(aes(t, s, color = trial), direction = "vh") +
  scale_x_continuous("Confirmed Response Study Week", breaks = (seq(0, 126, by = 12))) +
  scale_color_discrete("", type = AZ_palette) +
  labs(title = "Study time to confirmed response classification", y = "Proportion Unclassified") +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.7)) +
  NULL

tar_read("km_trial_confirmed_response") |> 
  rowwise() |>
  transmute(
    trial,
    conf_resp_km_calendar = list( 
      set_names(conf_resp_km_calendar, c("s", "n", "c", "e")) |>
        as_tibble() |> 
        mutate(t = seq(0, n() - 1))
    )
  ) |> 
  ungroup() |> 
  unnest(conf_resp_km_calendar) |> 
  ggplot() +
  geom_step(aes(t, s, color = trial), direction = "vh") +
  scale_x_continuous("Confirmed Response Calendar Week", breaks = (seq(0, 240, by = 12))) +
  scale_color_discrete("", type = AZ_palette) +
  labs(title = "Calendar time to confirmed response classification", y = "Proportion Unclassified") +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.7)) +
  NULL
```

```{r}
#| label: fig-assessment-visits
#| fig-cap: Number of visits every week.

treated_confirmed_resp_analysis_data |> 
  select(trial, n_tumors, t_measure) |> 
  unnest(t_measure) |> 
  unnest(t_measure) |> 
  mutate(epoch = if_else(t_measure <= 0, "pre-screening", "treatment") |> as_factor()) |> 
  ggplot(aes(t_measure)) +
  geom_bar(aes(fill = epoch), alpha = 0.5) +
  scale_fill_discrete("", labels = str_to_title, type = AZ_palette) +
  scale_x_continuous("Week", breaks = (seq(0, 126, by = 12))) +
  labs(y = "Count", title = "Number of visits every week") +
  facet_wrap(vars(trial), scales = "free_y") +
  theme(legend.position = "top", axis.text.x = element_text(size = 7)) +
  coord_cartesian(xlim = c(0, 120))
```

```{r}
#| label: fig-experiment-entry-cdf
#| fig-cap: Cumulative distribution of patients in the studies over calendar time (relative to the first weeks of the studies).

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  stat_ecdf(aes(x = experiment_start_week, color = trial)) +
  # geom_vline(xintercept = c(12, 24, 48), linetype = "dotted") +
  scale_x_continuous("Experiment Entry Calendar Week") +
  labs(title = "Empirical CDF of experiment entry", y = "Proportion of Patients") +
  scale_color_discrete("", type = AZ_palette) +
  # facet_wrap(vars(trial), scales = "free") +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.3)) +
  NULL
```




```{r}
#| label: fig-response-classification
#| fig-cap: Confirmed response counts.

treated_confirmed_resp_analysis_data |> 
  mutate(
    confirmed_response = factor(confirmed_response, levels = c(TRUE, FALSE), labels = c("Response", "Non-response")) |> 
      fct_na_value_to_level("Unclassified")
  ) |>  
  ggplot(aes(confirmed_response)) +
  geom_bar(alpha = 0.25) +
  labs(title = "Confirmed response counts", x = "", y = "Count") +
  facet_wrap(vars(trial), nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-age-group
#| fig-cap: Distribution of age groups.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(ecogbl), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-hormon
#| fig-cap: Distribution of HR status.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(hormonr), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-prior-pertuzumab-treatment
#| fig-cap: Distribution of prior pertuzumab treatment. 

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(prior_pertuzumab_treatment), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-prior-cdk46-inhibit-treatment
#| fig-cap: Distribution of prior cdk46 inhibition treatment.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(prior_cdk46_inhibit_treatment), alpha = 0.25) +
  facet_wrap(vars(trial), nrow = 1)
```

```{r}
#| label: fig-her2-status
#| fig-cap: Distribution of HER2 status.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(her2_status), alpha = 0.25) +
  facet_wrap(vars(trial))
```

```{r}
#| label: fig-history-visceral-disease
#| fig-cap: Distribution of a history of visceral disease.

treated_confirmed_resp_analysis_data |> 
  ggplot() +
  geom_bar(aes(hist_visceral_disease), alpha = 0.25) +
  facet_wrap(vars(trial))
```

# Model

## Confirmed Response Competing Risks Model {#crcr}

* $Y^\textrm{crcr}_{rt}$ is an indicator of whether a patient become classified as $r$ at $t$.
* $T^\textrm{crcr} \in \mathcal{N}^+$ is the number of weeks a patient remains unclassified.
* $C^\textrm{crcr}$ is an indicator of whether a patient is censored (stays unclassified in the available data).
* $R$ is an indicator of whether a patient eventually will be a confirmed responder or non-responder. 
* $\boldsymbol{X}$ is a vector patient characteristics and predictors. 

First, we have a competing risks model of time-to-classification, with a separate hazard function per risk (confirmed response vs confirmed non-response).

$$
\lambda^\textrm{crcr}_r(t;\boldsymbol{x,\theta}^\textrm{crcr}_r) = \lambda^\textrm{crcr}_{0rt}\exp(\boldsymbol{x\beta}^\textrm{crcr}_r),\,r \in \{\textrm{responder, non-responder}\} 
$$
and the baseline hazard is distributions as a Gaussian process, $\boldsymbol{\lambda}^\textrm{crcr}_{0r} \sim \mathtt{GP}(K^\textrm{crcr}_r)$.

The conditional probability of classification as a responder or non-responder (exiting the state of being unclassified) is

$$
\begin{align}
\Pr[Y^\textrm{crcr}_{rt} = 1 \mid Y^\textrm{crcr}_{r,t-1} = 0, \boldsymbol{X = x, \theta}^\textrm{crcr}_r] &= 1 - \exp[-\lambda^\textrm{crcr}_r(t;\boldsymbol{x,\beta^\textrm{crcr}_r})] \\
\Pr[Y^\textrm{crcr}_{rt} = 1 \mid Y^\textrm{crcr}_{r,t-1} = 1, \boldsymbol{X = x, \theta}^\textrm{crcr}_r] &= 1 \\
\end{align}
$$

The patient level log likelihood is
$$
\log \mathcal{L}^\textrm{crcr}_i = (1-C^\textrm{crcr}_i)\log[1 - \exp(-\lambda^\textrm{crcr}_{R_i}(T^\textrm{crcr}_i + 1;\boldsymbol{X_i, \theta^\textrm{crcr}_r}))] + \sum_{r = 0}^1\sum_t^{T^\textrm{crcr}_i} - \lambda^\textrm{crcr}_{r}(t;\boldsymbol{X_i, \theta^\textrm{crcr}_r})
$$

```{r}
#| label: fig-survival-unclassified
#| fig-cap: Survival plots of time to classification.
#| cache: true

tar_load("conf_resp_hb")

bind_rows(
  prior = tar_read("prior_all_rep_confirmed_response_bindist"),
  posterior = tar_read("all_rep_confirmed_response_bindist"),
  .id = "fit_type"
) |> 
  ggplot(aes(t)) +
  stat_lineribbon(
    aes(ydist = bindist, color = fit_type, fill = fit_type, alpha = fit_type),
    linewidth = 1, step = "hv", .width = c(0.5, 0.8)
  ) +
  geom_step(
    aes(y = count, linetype = "Observed"), direction = "hv", linewidth = 0.5, 
    data = \(d) tar_read("treated_confirmed_resp_analysis_data") |> 
      semi_join(d, by = "trial") |> 
      group_by(trial) |> 
      reframe(t = conf_resp_hb[-length(conf_resp_hb)], count = sample_hist(confirmed_response_week, conf_resp_hb))
  ) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  scale_linetype_manual("", values = c(Observed = "dotted")) +
  scale_x_continuous("", breaks = seq(0, 150, 12)) +
  facet_wrap(vars(trial), scales = "free_y") +
  theme(legend.position = "bottom") +
  guides(alpha = "none") +
  NULL
```

```{r}
#| label: fig-conf-resp-hazard-ratio
#| fig-cap: Probabilistic histograms of the hazard ratio in response to patient characteristics, separated by competing risk.
#| cache: true

bind_rows(
  prior = tar_read("prior_all_crcr_hazard_ratio_bindist"),
  posterior = tar_read("all_crcr_hazard_ratio_bindist"),
  .id = "fit_type"
) |>
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = bindist, color = fit_type, fill = fit_type, alpha = fit_type),
                  linewidth = 1,
                  step = "hv",
                  .width = c(0.5, 0.8)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  facet_grid(vars(trial), vars(k), scales = "free_y") +
  guides(alpha = "none") +
  NULL
```

```{r}
#| eval: false
#| include: false

covar_param <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_covar_param, .id = "fit_type") 

covar_param |> 
  mutate(k = factor(k, levels = 1:2, labels = c("Response", "Non-response"))) |> 
  ggplot() +
  stat_halfeye(aes(xdist = crcr_covar_effect, y = factor(covar), color = fit_type, fill = fit_type), alpha = 0.25, .width = 0.8) +
  scale_y_discrete("Parameter") +
  scale_color_discrete(type = AZ_palette, aesthetics = c("color", "fill")) +
  facet_grid(vars(k), vars(trial))
```

```{r}
#| eval: false
#| include: false

tumor_param <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_tumor_param, .id = "fit_type") 

tumor_param |> 
  mutate(k = factor(k, levels = 1:2, labels = c("Response", "Non-response"))) |> 
  ggplot() +
  stat_halfeye(aes(xdist = crcr_tumor_stim_pop_coef, y = factor(m), color = fit_type, fill = fit_type), alpha = 0.25, .width = 0.8) +
  scale_y_discrete("Parameter") +
  scale_color_discrete(type = AZ_palette, aesthetics = c("color", "fill")) +
  facet_grid(vars(k), vars(trial))
```

```{r}
#| eval: false
#| include: false

conf_resp_cif <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_cif, .id = "fit_type") 

conf_resp_prob_confirmed <- conf_resp_cif |> 
  group_by(fit_type, trial, i, k) |> 
  filter(min_rank(t) == n()) |> 
  group_by(fit_type, trial, i) |> 
  mutate(cif = cif / rvar_sum(cif)) |> 
  ungroup() |> 
  filter(k == 2)

conf_resp_cif |> 
  filter(trial == "Breast04", i <= 5) |>
  mutate(k = factor(k, levels = 1:2, labels = c("Response", "Non-response"))) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = cif, color = fit_type, fill = fit_type), alpha = 0.25, linewidth = 2, .width = c(0.5, 0.8)) +
  scale_color_discrete(type = AZ_palette, aesthetics = c("color", "fill")) +
  facet_grid(vars(i), vars(k))
```

```{r}
#| eval: false
#| include: false

conf_resp_prob <- lst(prior = prior_confirmed_resp_comp_risk_res, post = confirmed_resp_comp_risk_res) |> 
  map_dfr(get_conf_resp_prob, .id = "fit_type") 
```

```{r}
#| label: fig-cif
#| fig-cap: Probabilistic histograms of the Cumulative Incidence Function (CIF) for the different competing risks.
#| fig-width: 10
#| cache: true

bind_rows(
  prior = tar_read("prior_all_conf_resp_cif_bindist"),
  posterior = tar_read("all_conf_resp_cif_bindist"),
  .id = "fit_type"
) |> 
  ggplot(aes(p)) +
  stat_lineribbon(aes(ydist = cif_bindist, color = fit_type, fill = fit_type, alpha = fit_type),
                  linewidth = 1,
                  step = "vh",
                  .width = c(0.5, 0.8)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  scale_y_continuous("", breaks = NULL) +
  labs(x = "CIF") +
  facet_grid(vars(t), vars(trial, k), scales = "free_y", switch = "y", labeller = labeller(t = \(t) str_c("t = ", t))) +
  guides(alpha = "none") +
  theme(legend.position = "bottom", axis.text.x = element_blank(), strip.text.y.left = element_text(angle = 0), strip.placement = "outside") +
  NULL
```


## Recruitment Process

We model the recruitment process or the patients' entry into the experiment calendar week as

$$
W_i \sim \mathtt{NegBinomial}(\lambda^\textrm{recruit}, \phi^\textrm{recruit}),
$$
where $\lambda^\textrm{recruit}$ is the average calendar entry week and $\phi^\textrm{recruit}$ represents dispersion. At first, I tried a Poisson distribution but it was not sufficiently dispersed.

```{r}
#| label: fig-recruit-distribution-param
#| fig-cap: Parameters of the recruitment process distribution.

recruit_res_list <- lst(
  prior = tar_read("prior_recruit_fit"),
  posterior = tar_read("recruit_fit")
) 

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, lambda[trial]), .id = "fit_type") |> 
  ggplot() +
  stat_halfeye(aes(xdist = lambda, y = trial, color = fit_type, fill = fit_type, alpha = fit_type)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Expected Experiment Entry Calendar Week", y = "") +
  guides(alpha = "none") +
  coord_cartesian(xlim = c(0, 100)) +
  theme(legend.position = "bottom") +
  NULL

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, phi[trial]), .id = "fit_type") |> 
  ggplot() +
  stat_halfeye(aes(xdist = phi, y = trial, color = fit_type, fill = fit_type, alpha = fit_type)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Dispersion", y = "") +
  guides(alpha = "none") +
  coord_cartesian(xlim = c(0, 25)) +
  theme(legend.position = "bottom") +
  NULL

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, recruit_sd[trial]), .id = "fit_type") |> 
  ggplot() +
  stat_halfeye(aes(xdist = recruit_sd, y = trial, color = fit_type, fill = fit_type, alpha = fit_type)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Standard Deviation", y = "") +
  guides(alpha = "none") +
  coord_cartesian(xlim = c(0, 50)) +
  NULL
```

```{r}
#| label: fig-entry-week-cdf
#| fig-cap: Cumulative distribution of entry into the experiment over calendar weeks. 
#| cache: true

ecdf_breaks <- 1:125

recruit_res_list |> 
  map_dfr(\(f) spread_rvars(f, sim_experiment_start_week[trial, i]), .id = "fit_type") |> 
  group_by(fit_type, trial) |> 
  reframe(t = ecdf_breaks, ecdf_rvar = rfun(\(w) ecdf(w)(ecdf_breaks))(sim_experiment_start_week)) |> 
  ggplot(aes(t)) +
  stat_ecdf(aes(x = experiment_start_week), linetype = "dashed", data = tar_read("treated_pfs_analysis_data")) +
  stat_lineribbon(aes(ydist = ecdf_rvar, color = fit_type, fill = fit_type, alpha = fit_type), .width = 0.8) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Calendar Week", y = "Proportion in Experiment") +
  guides(alpha = "none") +
  facet_wrap(vars(trial)) +
  theme(legend.position = "bottom")
```

<!-- Below, we predict sample sizes and (confirmed response) maturity rate: we use the recruitment process model to simulate patients entering the trial (at varying rates) as we see how many we have at various prediction calendar weeks. For the maturity rate, we confirmed response classification calendar weeks simulated from the [confirmed response model](#crcr).  -->

```{r}
#| label: fig-recruit-sample-size-and-maturity
#| fig-cap: Simulations to see sample sizes and data maturity with varying prediction weeks and recruitment rates.
#| cache: true
#| eval: false
#| include: false

tar_load("recruit_maturity")

recruit_maturity |>
  filter(Pr(n_sample >= 10) >= 0.8) |> 
  ggplot() +
  stat_lineribbon(aes(pred_week, ydist = n_sample, fill = lambda, color = lambda, group = str_c(trial, lambda)), alpha = 0.25, .width = 0.8) +
  geom_hline(yintercept = 10, linetype = "dotted") +
  scale_x_continuous(breaks = seq(6, 48, 6)) +
  scale_color_continuous("Recruitment Rate") +
  labs(x = "Prediction Calendar Week", y = "Sample Size") +
  facet_grid(col = vars(trial), scales = "free_y") +
  NULL

recruit_maturity |> 
  filter(Pr(n_sample >= 10) >= 0.8) |> 
  ggplot() +
  stat_lineribbon(
    aes(pred_week, ydist = maturity_rate, fill = lambda, color = lambda, group = lambda), alpha = 0.25, .width = 0.8
  ) +
  scale_x_continuous(breaks = seq(6, 48, 6)) +
  scale_color_continuous("Recruitment Rate") +
  labs(x = "Prediction Calendar Week", y = "Maturity Rate") +
  facet_grid(col = vars(trial), scales = "free_y") +
  NULL
```

## PFS Model

* $Y^\textrm{pfs}_{t}$ is an indicator of whether a patient experiences disease progression at $t$.
* $T^\textrm{pfs} \in \mathcal{N}^+$ is the number of weeks a patient progression-free.
* $C^\textrm{pfs}$ is an indicator of whether a patient is censored.

First, we have a progression-free survival model,

$$
\lambda^\textrm{pfs}(t;\boldsymbol{x}, r,\boldsymbol{\theta}^\textrm{pfs}) = \lambda^\textrm{pfs}_{0t}\exp(\boldsymbol{x\beta}^\textrm{pfs} + r\rho).
$$
and the baseline hazard is distributions as a Gaussian process, $\boldsymbol{\lambda}^\textrm{pfs}_{0} \sim \mathtt{GP}(K^\textrm{pfs})$.

The conditional probability of progression is

$$
\begin{align}
\Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, R = r, \boldsymbol{\theta}^\textrm{pfs}] &= 1 - \exp[-\lambda^\textrm{pfs}(t;\boldsymbol{x},r,\boldsymbol{\beta}^\textrm{pfs},\rho)] \\
\Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 1, \boldsymbol{X = x}, R = r, \boldsymbol{\theta}^\textrm{pfs}] &= 1
\end{align}
$$
If a patient's response remains unclassified, their conditional probability is a mixture of based on the probability of a patient's eventual classification,
$$
\begin{gather}
\Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}] =\\ \gamma(\boldsymbol{x, \theta}^\textrm{crcr}) \cdot \Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, R = 1, \boldsymbol{\theta}^\textrm{pfs}] + (1 - \gamma(\boldsymbol{x, \theta}^\textrm{crcr})) \cdot \Pr[Y^\textrm{pfs}_{t} = 1 \mid Y^\textrm{pfs}_{t-1} = 0, \boldsymbol{X = x}, R = 0, \boldsymbol{\theta}^\textrm{pfs}].
\end{gather}
$$

The probability that a patient would be classified as a confirmed responder is^[We assume that eventually everyone is classified. We substitute the limit of CIF with CIF at the last week we have in the trial, which we then normalize.]
$$
\gamma(\boldsymbol{x, \theta}^\textrm{crcr}) = 
  \begin{cases}
    R_i & \text{if } C^\textrm{crcr} = 0 \\
    \lim_{t\rightarrow\infty}\mathsf{CIF}_1(t, \boldsymbol{x, \theta}^\textrm{crcr}) & \text{if } C^\textrm{crcr} = 1,
  \end{cases}
$$
and the cumulative incidence function (CIF) for is defined as
$$
\mathsf{CIF}_r(t, \boldsymbol{x, \theta}^\textrm{crcr}) = \sum_{\tau = 0}^t S^\textrm{crcr}(t,\boldsymbol{x, \theta}^\textrm{crcr})\cdot \Pr[Y^\textrm{crcr}_{r,\tau+1} = 1 \mid Y^\textrm{crcr}_{r,\tau} = 0, \boldsymbol{x, \theta}^\textrm{crcr}_r],
$$
where $S^\textrm{crcr}(t,\boldsymbol{x, \theta}^\textrm{crcr}) = \prod_{r=0}^1\prod_{\tau = 1}^t \Pr[Y^\textrm{crcr}_{r\tau} = 0 \mid Y^\textrm{crcr}_{r,\tau-1} = 0, \boldsymbol{x, \theta}^\textrm{crcr}_r]$ is the probability of remaining unclassified up to week $t$.

The log likelihood per patient is
$$
\log \mathcal{L}^\textrm{pfs}_i = (1-C^\textrm{pfs}_i)\log \Pr[Y^\textrm{pfs}_{T^\textrm{pfs}_i} = 1 \mid Y^\textrm{pfs}_{T^\textrm{pfs}_i-1} = 0, \boldsymbol{x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}] + \log S^\textrm{pfs}(T^\textrm{pfs}_i - 1, \boldsymbol{x},\boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}) + \mathcal{L}_i^\textrm{crcr}
$$

where $S^\textrm{pfs}(t,\boldsymbol{x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}) = \prod_{\tau = 1}^t \Pr[Y^\textrm{pfs}_{\tau} = 0 \mid Y^\textrm{pfs}_{\tau-1} = 0, \boldsymbol{x}, \boldsymbol{\theta}^\textrm{crcr}, \boldsymbol{\theta}^\textrm{pfs}]$ is the probability of remaining progress free up to week $t$.

```{r}
#| label: load-mpfs

pfs_cr_median_pfs <- bind_rows(
  prior = tar_read("prior_trial_conf_resp_median_pfs"), 
  posterior = tar_read("trial_conf_resp_median_pfs"),
  .id = "fit_type"
)  
```

```{r}
#| label: fig-median-pfs-conf-resp
#| fig-cap: Estimated median PFS using full trial data.

pfs_cr_median_pfs |> 
  ggplot() +
  stat_halfeye(
    aes(xdist = sim_trial_median_pfs, y = trial, color = fit_type, fill = fit_type, alpha = fit_type), 
    na.rm = TRUE, 
    .width = c(0.5, 0.8)
  ) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(y = "", x = "Median PFS [Weeks]") +
  guides(alpha = "none") +
  theme(legend.position = "bottom") +
  NULL
```

```{r}
#| label: fig-pfs-conf-resp-km
#| fig-cap: Survival curves as estimated by the model and compared to the observed (KM) proportions, using full trial data.
#| cache: true

bind_rows(
  prior = tar_read("prior_trial_conf_resp_km_est"),
  posterior = tar_read("trial_conf_resp_km_est"), 
  .id = "fit_type"
) |> 
  ggplot() +
  geom_step(
    aes(x = t, y = s, group = btype, linetype = "Observed Data"), 
    data = tar_read("km_res_calculated") |> filter(treated, fct_match(btype, "lb"))
  ) +
  stat_lineribbon(aes(x = t - 1, ydist = trial_km_est, color = fit_type, fill = fit_type, alpha = fit_type), step = TRUE, .width = c(0.8)) +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.0625, posterior = 0.25)) +
  scale_linetype_manual("", values = "dashed") +
  labs(x = "Week", y = "Survival Probability") +
  facet_wrap(vars(trial)) +
  guides(alpha = "none") + 
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-pfs-conf-resp-hazard-ratio
#| fig-cap: Probabilistic histograms of $\exp(\boldsymbol{x\beta}^\textrm{pfs} + r\rho)$ separated by $r = 0$ (confirmed non-response) and $r = 1$ (confirmed response). 
#| fig-width: 10
#| cache: true

hb <- seq(0, 3, by = 0.1)

bind_rows(
  prior = tar_read("prior_pfs_conf_resp_hazard_ratio"),
  posterior = tar_read("pfs_conf_resp_hazard_ratio"),
  .id = "fit_type"
) |>   
  group_by(trial, k, fit_type) |> 
  reframe(t = hb[-1], bindist = rvar_sample_hist(time_invariant_hazard_ratio, hb)) |> 
  mutate(k = factor(k, levels = 1:2, labels = c("Non-response", "Response"))) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = bindist, color = fit_type, fill = fit_type, alpha = fit_type),
                  linewidth = 1,
                  step = "vh",
                  .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  scale_color_discrete("", label = str_to_title, type = AZ_palette, aesthetic = c("color", "fill")) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 0.25)) +
  labs(x = "Hazard Ratio", y = "Density") +
  facet_grid(vars(trial), vars(k), scales = "free_y", switch = "y") +
  guides(alpha = "none") +
  theme(legend.position = "bottom", strip.text.y.left = element_text(angle = 0), strip.placement = "outside") +
  NULL
```
# Simulations 

Here we are using *leave-one-trial-out* evaluation of the prediction accuracy of training our models on three trials and holding the fourth out for testing. We do this for each trial.^[At this point, we're are pooling all the observed trials and predicting the heldout trial: there is no hierarchical structure (yet) to the model accounting for heterogeneity in the trials not captured by covariates of the model (i.e., baseline characteristics, etc.).] We do not update the model as data from the held-out trial becomes available with increased data maturity rates.

We look at the accuracy of both the median PFS and ORR endpoints:

1. **Median PFS**: the median progression-free survival time of all the patients _currently in the trial at the point of prediction_. For all patients who are either censored or their PFS would have been observed after the point of prediction, we use the trained model to predict their survival time.
2. **Objective Response Rate (ORR)**: the proportion of patients currently in the trial at the point of prediction who would eventually be classified as confirmed responders. For those patients who are censored^[Those who exited the trial before we observed their confirmed response classification. This could be problematic if patients select themselves to exit because they don't see improvement in their condition. Ideally, our model should also model another competing risk: voluntary exit.] or their confirmed response occurs after the point of prediction (see @fig-response-classification-time), we use the trained model to predict their eventual classification. 

::: {.callout-note title="Algorithm" collapse="true"}

* $\mathcal{D}_k$ is the data from trial $k \in \{1,\ldots,K\}$. $K$ is the number of trials available.
* $B$ is the number of sampling iterations.
* $\mathcal{M}$ is a set of rates of confirmed response data maturity to test. 
* $y^\textrm{pfs}$ and $y^\textrm{cr}$ are progression-free survival and confirmed response, respectively.

For every trial $k$,

* Train our models on the full data, holding out $\mathcal{D}_k$: $p_\textrm{post}(y^\textrm{cr}, y^\textrm{pfs}\mid \boldsymbol{x})$. 
* For each sampling iteration, $b \in \{1,\ldots,B\}$, 
    - Draw a random subsample, $\widetilde{\mathcal{D}}_{kb}$, of 30 patients (with replacement) from $\mathcal{D}_k$.
    - For each $m \in \mathcal{M}$,
      * Determine the set of patients who would be in the experiment such that the proportion of patients (out of 30) who have their confirmed response classified is at least $m$, $\widetilde{\mathcal{D}}_{kbm}$.
      * For each $i \in \widetilde{\mathcal{D}}_{kbm}$,
        - If $y^\textrm{pfs}_i$ is right censored,
          * If $y^\textrm{cr}_i$ is unclassified (censored), predict $(\tilde{y}^\textrm{pfs}_i, \tilde{y}^\textrm{cr}_i) \sim p_\textrm{post}(\cdot, \cdot\mid y^\textrm{pfs}_i, y^\textrm{cr}_i, \boldsymbol{x}_i)$.
          * else, let $\tilde{y}^\textrm{cr}_i = y^\textrm{cr}_i$ and predict $\tilde{y}^\textrm{pfs}_i \sim p_\textrm{post}(\cdot\mid y^\textrm{pfs}_i, y^\textrm{cr}_i, \boldsymbol{x}_i)$.
        - else, let $\tilde{y}^\textrm{pfs}_i = y^\textrm{pfs}_i$.
      * Calculate the median of $\boldsymbol{\tilde{y}}^\textrm{pfs}_{kbm}$. 
      
:::

```{mermaid}
flowchart TD
    all-data[("Trials Data")] -.-> hold-out("Hold out one trial")
    hold-out -.-> training-data[("Training Data")] & testing-data[("Testing Data")]
    training-data -.-> model("Train models")
    model --> increment-mature("Increment maturity rate")
    increment-mature --> select-patients("Select patients")
    testing-data -.-> select-patients
    select-patients --> predict-missing("Predict incomplete data")
    hold-out --> model
    model -.-> n1{{"Prediction Model"}}
    n1 -.-> predict-missing
    predict-missing --> n2{"Any more maturity rates?"}
    n2 -- Yes --> increment-mature
    n2 -- No --> n3{"Any more trials to hold out?"}
    n3 -- Yes --> hold-out
    n3 -- No --> n4(["End"])
    n5(["Start"]) --> hold-out
     hold-out:::green
     model:::orange
     increment-mature:::green
     select-patients:::orange
     predict-missing:::orange
     n1:::pink
     n2:::blue
     n4:::purple
     n5:::purple
    classDef green fill:#B2DFDB,stroke:#00897B,stroke-width:2px
    classDef orange fill:#FFE0B2,stroke:#FB8C00,stroke-width:2px
    classDef blue fill:#BBDEFB,stroke:#1976D2,stroke-width:2px
    classDef yellow fill:#FFF9C4,stroke:#FBC02D,stroke-width:2px
    classDef pink fill:#F8BBD0,stroke:#C2185B,stroke-width:2px
    classDef purple fill:#E1BEE7,stroke:#8E24AA,stroke-width:2px
```
       
Below we look at how each of the PFS model parameters influences the hazard. These are the exponential values of the parameters: they should the multiplication effect on the hazard.
          
```{r}
#| label: fig-bs-param-est
#| fig-cap: Posterior distributions of model parameters, shown separately by the trial heldout.

tar_read("pfs_conf_resp_bootstrap_param") |> 
  ggplot(aes(y = covar)) +
  stat_slabinterval(
    aes(xdist = .exp_value, justification = 0.5, thickness = after_stat(ggdist::thickness(1)), slab_alpha = after_stat(pdf)),
    point_interval = NULL, show.legend = FALSE, data = tar_read("prior_pfs_conf_resp_param")
  ) +
  stat_pointinterval(aes(xdist = .exp_value, color = trial), point_size = 1, position = "dodge", .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  scale_color_discrete("", type = AZ_palette) +
  labs(x = "Exponential of Parameter",  y = "Parameter", caption = "Showing the posterior median, 50% CI, and 80% CI.\nGradient shows the prior.") +
  NULL
```

* While the confirmed response indicator appears to have the strongest effect on hazard, keep in mind that the majority of patients are identified as confirmed responders (see @fig-response-classification). 
* The _practically_ strongest predictor here is the baseline sum of tumor sizes. Note that the following tumor size assessment does not appear to have a lot of influence. This does not rule out the possibility of a more useful signal in the entire time-series of tumor size changes. 
* We use relatively conservative priors on these parameters. The influence of both confirmed response and baseline tumor sizes is sufficiently high to overcome this regularization. 
* Idea: look into the interaction between confirmed response and tumor sizes. Those are likely highly correlated since the former is based on the latter.
* Other observations: 
    - Highest age group has reduced hazard. This is probably a survivorship effect: those who survived that long are probably more robust.
    - Prior CDK46 inhibition treatment is probably an indicator higher risk.

```{r}
#| label: load-bootstrap

tar_load(c("bootstrap_confirmed_resp_pfs_res", "pfs_conf_resp_bootstrap_cr_median_pfs", "pfs_conf_resp_bootstrap_pfs_median_pfs", 
           "fixed_bootstrap_cr_median_pfs", "fixed_bootstrap_pfs_median_pfs", "bootstrap_cr_median_pfs_draws"))

last_mature_week_data <- all_confirmed_resp_analysis_data |> 
  group_by(trial) |> 
  summarize(last_mature_week = max((1 - confirmed_response_censored) * (experiment_start_week + confirmed_response_week - 1)))

max_all_t <- max(all_confirmed_resp_analysis_data$patient_max_t)
```

## In response to varying data maturity of confirmed response classification

In this section, we look at how prediction accuracy changes in response to the maturity of confirmed response: the proportion of patients who have entered the trial at the point of prediction whose confirmed response has been classified (is no longer _stable disease_).

```{r}
#| label: fig-posterior-mpfs-for-subsamples
#| fig-cap: Showing what the posterior of the mPFS prediction looks like for three bootstrap subsamples for Breast-04.

prediction_focus <- tibble(cr_f = c(0.1, 0.15, 0.5))

fixed_db4_subsample <- fixed_bootstrap_cr_median_pfs |>
  filter(map_lgl(fixed_bs_cr_median_pfs, \(mpfs) quantile(mpfs, probs = 0.8) < max_all_t), trial == "Breast04") |>   
  mutate(selected = f == 2)  

fixed_db4_subsample |>
  filter(selected) |> 
  ggplot(aes(bootstrap_cr_maturity_rates)) +
  stat_lineribbon(aes(ydist = fixed_bs_cr_median_pfs), alpha = 0.25, linewidth = 0.75, .width = c(0.5, 0.8)) +
  stat_pointinterval(
    aes(ydist = sim_trial_median_pfs, color = "full data"),
    .width = 0.8, linewidth = 0.5,
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |>
      filter(fct_match(fit_type, "posterior")) |>
      mutate(bootstrap_cr_maturity_rates = 0.8)
  ) +
  geom_vline(aes(xintercept = cr_f, color = factor(cr_f)), linetype = "dashed", show.legend = FALSE, data = prediction_focus) +
  scale_fill_discrete("Bootstrap sample CI", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-(1:5)]) +
  scale_color_discrete("", labels = c("full data" = "Complete trial 80% CI\n(all patients)"), type = AZ_palette) +
  labs(
    title = "Predicted median PFS for Breast-04", 
    x = "Fraction of patients with known response", y = "Median PFS [weeks]"
  ) +
  # facet_grid(vars(f), vars(trial), labeller = labeller(f = \(f) str_c("Subsample ", f))) +
  coord_cartesian(xlim = c(0.0, 0.8), ylim = c(10, 140)) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-mpfs-density
#| fig-cap: Median PFS density plot for three points in time for Breast-04
#| fig-width: 8
#| fig-height: 3

fixed_bootstrap_cr_median_pfs |>
  filter(f == 2, trial == "Breast04") |>
  semi_join(mutate(prediction_focus, cr_f2 = cr_f + 0.01), by = join_by(bootstrap_cr_maturity_rates >= cr_f, bootstrap_cr_maturity_rates < cr_f2)) |>  
  filter(map_lgl(fixed_bs_cr_median_pfs, \(mpfs) quantile(mpfs, probs = 0.8) < max_all_t)) |> 
  ggplot() +
  stat_slab(
    aes(xdist = fixed_bs_cr_median_pfs, color = factor(bootstrap_cr_maturity_rates)),
    fill = NA, linewidth = 0.5
  ) +
  stat_slab(
    aes(xdist = fixed_bs_cr_median_pfs, fill = factor(bootstrap_cr_maturity_rates)),
    color = NA,
    alpha = 0.25
  ) +
  scale_color_discrete("Fraction of patients with known response", type = AZ_palette, aesthetics = c("color", "fill")) +
  labs(x = "Median PFS [weeks]", y = "") +
  theme(legend.position = "bottom") +
  coord_cartesian(xlim = c(25, 140)) +
  NULL
  
```

```{r}
#| label: fig-db4-with-sample-draws
#| fig-cap: Showing Breast-04 prediction including a selection of five bootstrap subsamples.

pfs_conf_resp_bootstrap_cr_median_pfs |> 
  filter(bs_cr_median_pfs.upper < max_all_t) |>
  group_by(trial, bootstrap_cr_maturity_rates) |> 
  filter(n() > 1) |> 
  ungroup() |>
  filter(fct_match(trial, "Breast04")) |> 
  ggplot(aes(bootstrap_cr_maturity_rates)) +
  geom_lineribbon(
    aes(y = NULL, ymin = bs_cr_median_pfs.lower, ymax = bs_cr_median_pfs.upper, group = .width), 
    alpha = 0.25, linewidth = 0.5 
  ) +
  stat_pointinterval(
    aes(ydist = sim_trial_median_pfs, color = "Full Data"),
    .width = 0.8, linewidth = 0.5, 
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |>
      filter(fct_match(fit_type, "posterior")) |>
      mutate(bootstrap_cr_maturity_rates = 0.8)
  ) +
  # geom_line(
  #   aes(y = bs_cr_median_pfs, group = .draw), alpha = 0.125,
  #   data = \(d) semi_join(bootstrap_cr_median_pfs_draws, d, by = c("trial", "r")) |> filter(.draw <= 5) 
  # ) +
  geom_line(
    aes(y = fixed_bs_cr_median_pfs, group = f, alpha = selected), 
    linewidth = 0.75, show.legend = FALSE,
    data = fixed_db4_subsample |> point_interval(fixed_bs_cr_median_pfs) 
  ) +
  scale_color_manual("", values = AZ_palette[4], labels = c("Full Data" = "Complete trial 80% CI\n(all patients)")) +
  scale_fill_discrete("Percentage of samples", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  scale_alpha_manual("", values = c("TRUE" = 0.5, "FALSE" = 0.125)) +
  geom_vline(xintercept = 0.15, linetype = "dotted") +
  labs(
    title = "Predicted median PFS for 600 different samples from Breast-04", 
    x = "Fraction of patients with known response", y = "Median PFS [weeks]"
  ) +
  # facet_wrap(vars(trial)) +
  coord_cartesian(xlim = c(0.0, 0.8)) +
  theme(legend.position = "bottom") +
  NULL
```

`
```{r}
#| label: fig-pfs-conf-resp-bootstrap-cr
#| fig-cap: Bootstrap sample prediction ribbons (50 and 80% credible intervals) over a range of confirmed response data maturity rates. Missing ribbons indicate insufficient bootstrap samples at that data maturity rate. The full data intervals shows the 80% credible intervals. 
#| fig-subcap: 
#|   - Predicting median PFS
#|   - Predicting ORR

pfs_conf_resp_bootstrap_cr_median_pfs |> 
  filter(bs_cr_median_pfs.upper < max_all_t) |>
  group_by(trial, bootstrap_cr_maturity_rates) |> 
  filter(n() > 1) |> 
  ungroup() |>
  ggplot(aes(bootstrap_cr_maturity_rates)) +
  geom_lineribbon(
    aes(y = NULL, ymin = bs_cr_median_pfs.lower, ymax = bs_cr_median_pfs.upper, group = .width), 
    alpha = 0.25, linewidth = 0.5 
  ) +
  stat_pointinterval(
    aes(ydist = sim_trial_median_pfs, color = "Full Data"),
    .width = 0.8, linewidth = 0.5, 
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |>
      filter(fct_match(fit_type, "posterior")) |>
      mutate(bootstrap_cr_maturity_rates = 0.8)
  ) +
  # geom_text(
  #   aes(x = x, y = y, label = str_c("N = ", n), color = "Full Data"), size = 3, hjust = 1, show.legend = FALSE,
  #   data = \(d) semi_join(all_confirmed_resp_analysis_data, d, by = "trial") |> count(trial) |> mutate(x = 0.8, y = 110)
  # ) +
  scale_color_manual("", values = AZ_palette[4], labels = c("Full Data" = "Complete trial 80% CI\n(all patients)")) +
  scale_fill_discrete("Percentage of samples", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  labs(
    title = "Predicted median PFS for 600 different samples for each of the Breast trials",
    x = "Fraction of patients with known response", y = "Median PFS [weeks]"
  ) +
  facet_wrap(vars(trial)) +
  coord_cartesian(xlim = c(0.0, 0.8), ylim = c(10, 140)) +
  theme(legend.position = "bottom", title = element_text(size = 9)) +
  NULL

pfs_conf_resp_bootstrap_cr_median_pfs |> 
  filter(bs_cr_orr.lower > 0.25) |>
  group_by(trial, bootstrap_cr_maturity_rates) |> 
  filter(n() > 1) |> 
  ungroup() |> 
  ggplot(aes(bootstrap_cr_maturity_rates)) +
  geom_lineribbon(
    aes(y = NULL, ymin = bs_cr_orr.lower, ymax = bs_cr_orr.upper, group = .width), 
    alpha = 0.25, linewidth = 0.5 
  ) +
  scale_color_manual("", values = AZ_plum) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  labs(x = "Confirmed Response Data Maturity Rate", y = "ORR") +
  facet_wrap(vars(trial), scales = "free_y") +
  coord_cartesian(xlim = c(0.0, 0.8), ylim = c(0.5, 1))
```



* What we are observing here is the probability ranges of leave-one-trial-out prediction, so those intervals combine sampling uncertainty from the bootstrap sample and the prediction model uncertainty.
  - I cannot make the decision on the acceptability of these interval widths for decision-making, for that we need to better characterize the decision risk we are evaluating.
* The models seem to be unbiased since they tend to the full data estimation intervals. Given that this isn’t a hierarchical model, this means that the covariates used are doing a relatively good job at prediction.
* There is slow improvement beyond the 20% maturity rate (this is maybe because most patients eventually reach confirmed response, so there is little information here anyway). Perhaps focusing on PFS maturity rate might be more helpful.
* ORR is high (see @fig-response-classification).

Below, we take a look at how: 

(i) the sample size of patients and 
(ii) the proportion of patients that have immature (incomplete) endpoint outcomes, which needs to be predicted by the model, 

changes in response to delaying the point of prediction. 

```{r}
#| label: fig-bootstrap-subsample-size-cr
#| fig-cap: Bootstrap sample over a range of confirmed response data maturity rates. The dotted lines shows the maximum size of 30 patients.
#| fig-subcap:
#|    - The distribution of bootstrap sample sizes.
#|    - The proportion of patients who need their confirmed response or PFS predicted. 

cr_size_est_data <- pfs_conf_resp_bootstrap_cr_median_pfs |> 
  mutate(across(starts_with("n_bs_sample"), \(n) if_else(n == 0, 30, n))) |> 
  select(trial, .width, bootstrap_cr_maturity_rates, contains("predicted_prop"), matches(r"{n_bs_sample(\.(lower|upper))?$}")) |> 
  rename_with(\(v) str_replace(v, "(_cr|predicted_prop)$", "\\1.median")) |> 
  pivot_longer(
    !c(bootstrap_cr_maturity_rates, trial, .width), names_pattern = r"{(.+)\.(lower|upper|median)(?:_prop)?}", names_to = c("variable", ".value")
  )

cr_size_est_data |> 
  filter(fct_match(variable, "n_bs_sample")) |> 
  ggplot(aes(bootstrap_cr_maturity_rates, median)) +
  geom_lineribbon(aes(ymin = lower, ymax = upper, group = .width), linewidth = 0.5, alpha = 0.25) +
  geom_hline(yintercept = 30, linetype = "dotted") +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  labs(x = "Confirmed Response Data Maturity Rate", y = "Prediction Sample Size") +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(0, 30)) +
  NULL

cr_size_est_data |> 
  filter(!fct_match(variable, "n_bs_sample")) |>
  mutate(
    variable = fct_recode(variable, "confirmed response" = "n_bs_cr_conf_resp_predicted_prop", "PFS" = "n_bs_cr_pfs_predicted_prop")
  ) |> 
  ggplot(aes(bootstrap_cr_maturity_rates, median)) +
  geom_lineribbon(aes(ymin = lower, ymax = upper, group = .width), linewidth = 0.5, alpha = 0.25) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  labs(x = "Confirmed Response Data Maturity Rate", y = "Proportion of Sample Predicted") +
  facet_grid(vars(variable), vars(trial)) +
  NULL
```

* But 40% maturity we're almost always dealing with the whole (30) sample of patients.
* @fig-bootstrap-subsample-size-cr-2 shows the degree to which we are relying on observed endpoint outcomes for out sample predictions. If by the point of prediction we have a small number of observations being predicted, our model is not really adding much. From these results, that does not appear to be the case.

```{r}
#| label: fig-bootstrap-prediction-week-cr
#| fig-cap: Bootstrap prediction calendar weeks (50 and 80% probability intervals) over a range of confirmed response data maturity rates. The dotted line indicates the calendar week when the last (uncensored) patient was classified as a confirmed response or non-response. This plot has been rotated to put the time dimension on the x-axis, even though the it is the maturity rate that we manipulate in these simulations, mostly for ease of reading.

pfs_conf_resp_bootstrap_cr_median_pfs |> 
  filter(bs_cr_prediction_calendar_week.upper < 1e6) |>
  group_by(trial, bootstrap_cr_maturity_rates) |> 
  filter(n() > 1) |> 
  ungroup() |> 
  left_join(last_mature_week_data, by = "trial") |> 
  mutate(across(starts_with("bs_prediction_calendar_week"), \(w) if_else(w > 1000, last_mature_week, w))) |> 
  ggplot(aes(bootstrap_cr_maturity_rates, bs_cr_prediction_calendar_week)) +
  geom_lineribbon(
    aes(ymin = bs_cr_prediction_calendar_week.lower, ymax = bs_cr_prediction_calendar_week.upper, group = .width), 
    linewidth = 0.5, alpha = 0.25 
  ) +
  geom_hline(
    aes(yintercept = last_mature_week, linetype = "Week of last response classification, in the full data."),
    data = last_mature_week_data 
  ) +
  scale_fill_discrete("Credible intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  scale_linetype_manual("", values = "dotted") + 
  labs(x = "Fraction of patients with known response", y = "Prediction Calendar Week") +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(10, 200)) +
  theme(legend.position = "bottom") +
  NULL
```

* We could be making these predictions by week 50, a significant decrease from waiting till everyone (in the full trial sample) has reached confirmed response classification. That is longer than we saw with the previous in-sample evaluation, but that’s comparing apples to oranges.
 
Questions remaining:

* How much would things improve when using a hierarchical model?
* Would we gain much in updating the model from the small number of patients observed from the held out trial?
* How much would things improve if in included dynamic confirmed response and tumor sizes?

## In response to varying data maturity of PFS

In this section, we repeat the same exercise as above but instead we focus on the maturity of PFS rather than confirmed response. _One challenging aspect to this measure of maturity is that immaturity directly translates to better survival, which differs from confirmed response maturity, a measure of when patients' status is known._ So, the better a treatment preforms, increasing PFS, the harder it becomes to reach any kind of maturity.

```{r}
#| label: fig-pfs-conf-resp-bootstrap-pfs
#| fig-cap: Bootstrap sample prediction ribbons (50 and 80% credible intervals) over a range of PFS data maturity rates. Missing ribbons indicate insufficient bootstrap samples at that data maturity rate. The full data intervals shows the 80% credible intervals. 
#| fig-subcap: 
#|   - Predicting median PFS
#|   - Predicting ORR

pfs_conf_resp_bootstrap_pfs_median_pfs |> 
  filter(bs_pfs_median_pfs.upper < max_all_t) |>
  group_by(trial, bootstrap_pfs_maturity_rates) |> 
  filter(n() > 1) |> 
  ungroup() |> 
  ggplot(aes(bootstrap_pfs_maturity_rates)) +
  geom_lineribbon(
    aes(y = bs_pfs_median_pfs, ymin = bs_pfs_median_pfs.lower, ymax = bs_pfs_median_pfs.upper, group = .width), 
    alpha = 0.25, linewidth = 0.5 
  ) +
  stat_pointinterval(
    aes(ydist = sim_trial_median_pfs, color = "Full Data"),
    .width = 0.8, size = 0.5,
    data = \(d) semi_join(pfs_cr_median_pfs, d, by = "trial") |>
      filter(fct_match(fit_type, "posterior")) |>
      mutate(bootstrap_pfs_maturity_rates = 0.8)
  ) +
  geom_text(
    aes(x = x, y = y, label = str_c("N = ", n), color = "Full Data"), 
    size = 3, hjust = 1, show.legend = FALSE, 
    data = count(all_confirmed_resp_analysis_data, trial) |> mutate(x = 0.8, y = 110) 
  ) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  scale_color_manual("", values = AZ_pink) +
  labs(x = "PFS Data Maturity Rate", y = "mPFS") +
  facet_wrap(vars(trial)) +
  coord_cartesian(xlim = c(0.0, 0.8), ylim = c(10, 140))

pfs_conf_resp_bootstrap_pfs_median_pfs |> 
  filter(bs_pfs_orr.lower > 0.25) |>
  filter(bs_pfs_median_pfs.upper < max_all_t) |>
  group_by(trial, bootstrap_pfs_maturity_rates) |> 
  filter(n() > 1) |> 
  ggplot(aes(bootstrap_pfs_maturity_rates)) +
  geom_lineribbon(
    aes(y = bs_pfs_orr, ymin = bs_pfs_orr.lower, ymax = bs_pfs_orr.upper, group = .width), alpha = 0.25, linewidth = 0.5 
  ) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  scale_color_manual("", values = AZ_pink) +
  labs(x = "PFS Data Maturity Rate", y = "ORR") +
  facet_wrap(vars(trial)) +
  coord_cartesian(xlim = c(0.0, 0.8), ylim = c(0.5, 1))
```

* Given how long it takes to reach any maturity level, we're already dealing with a stronger signal as early as 20% maturity, hence we don't see a lot of improvement as maturity increases beyond that.

```{r}
#| label: fig-bootstrap-subsample-size-pfs
#| fig-cap: Bootstrap sample sizes (50 and 80% probability intervals) over a range of PFS data maturity rates. The dotted lines shows the maximum size of 30 patients.
#| fig-subcap:
#|    - The distribution of bootstrap sample sizes.
#|    - The proportion of patients who need their confirmed response or PFS predicted. 

pfs_size_est_data <- pfs_conf_resp_bootstrap_pfs_median_pfs |> 
  mutate(across(starts_with("n_bs_sample"), \(n) if_else(n == 0, 30, n))) |> 
  select(trial, .width, bootstrap_pfs_maturity_rates, contains("predicted_prop"), matches(r"{n_bs_sample(\.(lower|upper))?$}")) |> 
  rename_with(\(v) str_replace(v, "(_pfs|predicted_prop)$", "\\1.median")) |> 
  pivot_longer(
    !c(bootstrap_pfs_maturity_rates, trial, .width), names_pattern = r"{(.+)\.(lower|upper|median)(?:_prop)?}", names_to = c("variable", ".value")
  )

pfs_size_est_data |> 
  filter(fct_match(variable, "n_bs_sample")) |> 
  ggplot(aes(bootstrap_pfs_maturity_rates, median)) +
  geom_lineribbon(aes(ymin = lower, ymax = upper, group = .width), linewidth = 0.5, alpha = 0.25) +
  geom_hline(yintercept = 30, linetype = "dotted") +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  labs(x = "PFS Data Maturity Rate", y = "Prediction Sample Size") +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(10, 30)) +
  NULL

pfs_size_est_data |> 
  filter(!fct_match(variable, "n_bs_sample")) |>
  mutate(
    variable = fct_recode(variable, "confirmed response" = "n_bs_pfs_conf_resp_predicted_prop", "PFS" = "n_bs_pfs_pfs_predicted_prop")
  ) |> 
  ggplot(aes(bootstrap_pfs_maturity_rates, median)) +
  geom_lineribbon(aes(ymin = lower, ymax = upper, group = .width), linewidth = 0.5, alpha = 0.25) +
  scale_fill_discrete("Credible Intervals", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  labs(x = "PFS Data Maturity Rate", y = "Proportion of Sample Predicted") +
  facet_grid(vars(variable), vars(trial), scales = "free_y") +
  NULL
```

* It takes a lot longer to get to any maturity level (see @fig-bootstrap-prediction-week-pfs) so we're dealing with the full sample very quickly.
* The proportion of patients for whom prediction is needed is significantly diminished.

```{r}
#| label: fig-bootstrap-prediction-week-pfs
#| fig-cap: Bootstrap prediction calendar weeks (50 and 80% probability intervals) over a range of PFS data maturity rates. The dotted line indicates the calendar week when the last (uncensored) patient was classified as a confirmed response or non-response. This plot has been rotated to put the time dimension on the x-axis, even though the it is the maturity rate that we manipulate in these simulations, mostly for ease of reading.

pfs_conf_resp_bootstrap_pfs_median_pfs |> 
  filter(bs_pfs_prediction_calendar_week.upper < 1e6) |>
  group_by(trial, bootstrap_pfs_maturity_rates) |> 
  filter(n() > 1) |> 
  ungroup() |> 
  left_join(last_mature_week_data, by = "trial") |> 
  mutate(across(starts_with("bs_prediction_calendar_week"), \(w) if_else(w > 1000, last_mature_week, w))) |> 
  ggplot(aes(bootstrap_pfs_maturity_rates, bs_pfs_prediction_calendar_week)) +
  geom_lineribbon(
    aes(ymin = bs_pfs_prediction_calendar_week.lower, ymax = bs_pfs_prediction_calendar_week.upper, group = .width), 
    linewidth = 0.5, alpha = 0.25 
  ) +
  geom_hline(
    aes(yintercept = last_mature_week, linetype = "Week of last Confirmed Response classification, in the full data."),
    data = last_mature_week_data 
  ) +
  scale_fill_discrete("CI", labels = c("0.5" = "50%", "0.8" = "80%"), type = AZ_palette[-1]) +
  scale_linetype_manual("", values = "dotted") + 
  labs(x = "PFS Data Maturity Rate", y = "Prediction Calendar Week") +
  facet_wrap(vars(trial)) +
  coord_flip(ylim = c(10, 200)) +
  theme(legend.position = "bottom") +
  NULL
```

One thing that is immediately clear from this plot is that it takes a lot longer to reach any level of maturity: it take around a 100 weeks to get near 50% maturity.