---
title: "Early Prediction of Breast Cancer Progress Free Survival"
author: Karim Naguib
code-fold: true
code-tools: true
toc: true
toc-location: left
fig-caption-location: margin
---

```{r}
#| label: setup
#| include: false

# library(conflicted)
library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(gtsummary)
library(targets)
library(colorspace)

# conflicts_prefer(
#   dplyr::filter
# )

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

knitr::opts_chunk$set(cache.path = file.path("/wscratch", Sys.getenv("USER"), "adc-early-predict/cache/"))

options(mc.cores = 12)

ggplot2::theme_set(theme_minimal(base_family = "Arial"))

future::plan(future::multisession(workers = 12))

source(here("r", "util.R"))
```

# Exploratory Analysis Plan

## Question

Are there observable strong and deep early responses to antibody-drug conjugates (ADCs) that are predictive of long term progression free survival (PFS)? Are we able to reliably predict the heterogeneous response to ADCs using early measurements?

In this exploratory analysis we will not take up the causal question of whether these strong early responders would see similar PFS under counterfactual dosages or medications.[^1]

[^1]: This should probably revisited as the prediction within arm might not be the relevant question.

## Exploratory Analysis Context and Data

We will evaluate whether early (at baseline and at 6 and 12 weeks post treatment) measurements of tumor sizes are predictive of the PFS in response to Enhertu.

We will use data from the Breast01, Breast02, Breast03, and Breast04 trials, focusing on the treatment arms receiving the same dosage. Initially, we will be analyzing the different studies separately with no partial pooling since the inclusion criteria for the studies were different.

-   Primary outcome: Post treatment progression free survival (interval and right censored).
-   Predictive variables (at all the early intervals):
    -   Number of lesions observed.
    -   Organ of lesions.
    -   lesion sizes measured, with a maximum of ten of observed lesions and two per organ.

## Statistical Model Goals

-   Explore the heterogeneity in PFS conditional on early tumor size measures.
-   Consider more nuanced characteristics of a patient's tumors beyond the aggregated sum of largest diameters. The challenge here is how to model the nonlinear contribution of multiple influencers (tumors) on a central node's outcome (survival). A simple sum or mean ends up discarding a lot of information about the context of the disease.
-   Introduce a lesion level first-stage model. Considering the limit on the number of lesions measured, we potentially truncate some measures such as the total sum of tumor sizes. A model of tumor sizes would allow us to impute the missing data and propagate any uncertainty to the second-stage survival model.

```{r}
#| label: priors
#| cache: true

source(here("r", "priors.R"))
```

```{r}
#| label: models
#| code-fold: false   

tumor_model <- cmdstan_model(here("stan", "tumor.stan"))
pfs_model <- cmdstan_model(here("stan", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)
```

# Data

What data we have so far: 1. Breast01: phase II dose finding study (HER2-positive metastatic breast cancer). 2. Breast02: phase III study comparing to physician choice of drug (HER2-positive metastatic breast cancer). 2. Breast03: phase III study comparing to trastuzumab emtansine (HER2-positive metastatic breast cancer). 3. Breast04: This is a phase III study on patients with HER2-low metastatic breast cancer (HER2-low metastatic breast cancer).

```{mermaid}
%%| label: fig-data-structure
%%| fig-align: center

erDiagram
  Study ||--|{ Patient : enrolls
  Patient ||--|{ Tumor : has
  Tumor ||--|{ Assessment : assessed
  Organ ||--|{ Tumor : has
  Patient ||--|{ "Prior Therapy" : "previously treated"
```

```{r}
#| label: load-trials-data
#| cache: true

data_details <- tribble(
  
  ~study,          ~ idap,           ~ trial, 
  "d9673c00002",   "idap_20231204",  "Breast01",
  "d967hc00001",   "idap_20231129",  "Breast02", 
  "d9674c00001",   "idap_20231204",  "Breast03",
  "d9675c00001",   "idap_20231124",  "Breast04"
) |> 
  mutate(across(everything(), factor))

# Demographics 
sdtm_dm <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(study, idap, "deid_dm", "sdtm")))

sdtm_tu <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(study, idap, "deid_tu", "sdtm")))

# Tumor analysis_data
adam_adtr <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(study, idap, "deid_adtr", "adam"))) |> 
  unnest(data) |> 
  select(!c(tuloc, tulat)) |> 
  left_join(# I need to do this because tuloc in the adam data for breast01 is NA for all observations.
    unnest(sdtm_tu, data) |> 
      arrange(visitnum) |> 
      distinct(usubjid, tulnkid, tuloc, tulat) |> 
      group_by(usubjid, tulnkid) |>
      filter(!is.na(tuloc)) |> 
      summarize(tuloc = first(tuloc), tulat = first(tulat), .groups = "drop"), # Only the first assessment has the location of the tumor. Fix that.
    by = c("usubjid", "trlnkid" = "tulnkid")
  ) |> 
  mutate(
    tuloc = str_remove(tuloc, fixed("(EXCLUDING ORAL CAVITY)")) |>
      str_replace_all(c("SUBCARINAL" = "SUBCRANIAL", "VERTEBRA" = "SPINE", "PARAAORTIC" = "PARA-AORTIC"))
  ) |>
  nest(.by = c(idap, trial))

# Subject-level data
adam_adsl <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(study, idap, "deid_adsl", "adam")))

adam_adcm <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(study, idap, "deid_adcm", "adam")))

# Time to event
adam_adtte <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(study, idap, "deid_adtte", "adam")) |> 
           map(\(d) mutate(d, cnsr = cnsr == 1)))
```

Using SDTM and ADaM data from the trials, we prepare the analysis data.

```{r}
#| label: prepare-analysis-data
#| cache: true
#| dependson: load-trials-data
#| code-fold: false 

# This first cleaned up data set used for tumor and PFS analysis.
tumor_analysis_data <- adam_adtr |> 
  unnest(data) |> 
  filter(
    fct_match(paramcd, "LDIAM"),
    # There are a number of assessments that show up in the data but in SDTM they are marked as "NOT DONE"
    !is.na(trdy), !is.na(aval),
    tracptfl,
    aval > 0, # Drop all zero diameter assessments; these appear to be missing measurements. 
  ) %>% 
  select(studyid, idap, trial, 
         usubjid, trlnkid, trlnkgrp, tuloc, tulat, trtsdt, trtedt, adt, ends_with("fl"), trdy, trtp, trta, visit, visitnum, mmdiam = aval) |> 
  mutate(
    week = if_else(trdy > 0, # Is post-treatment day? 
                   trdy, 
                   trdy + 1) %/% 7, # Last pre-screening week is 0. 
    treated_week = adt >= trtsdt, # Was this a post-treatment week?
    usubjid = factor(usubjid)
  ) |> 
  # Turns out that some assessments are not listed in order of the days they were done. Look at subject DS8201-A-U303-10033028.
  arrange(studyid, usubjid, trlnkid, trdy) |> 
  # For each patient-tumor row, put all the tumor assessment data in a nested column
  nest(tumor_history = c(adt, trdy, week, visit, visitnum, tsplitfl, tmergefl, mmdiam, treated_week)) |> 
  mutate(# For some weeks, there are multiple (unscheduled) visits that might happen. I'm just going to use the average for tumor sizes within those
         # weeks.
    tumor_history = map(
      tumor_history, 
      \(h) nest(h, visits = c(adt, trdy, visit, visitnum, mmdiam)) |> 
        mutate(
          n_week_visits = map_int(visits, nrow),
          mmdiam = map_dbl(visits, \(v) mean(v$mmdiam))
        )
    ), 
    
    n_tumor_measures = map_int(tumor_history, nrow), # How many assessments/measurements per tumor
    
    min_t = map_int(tumor_history, \(h) min(h$week)), # First assessment week, per tumor
    max_t = map_int(tumor_history, \(h) max(h$week)), # Last assessment week, per tumor
  ) |> 
  pack(flags = ends_with("fl")) |> 
  # Each row is going to be a patient and all their tumor data are put in a nested column.
  nest(patient_tumors = !c(studyid, trial, idap, usubjid, trtsdt, trtedt, trtp, trta)) |> 
  left_join(# Adding a column `n_all_tumors` that counts all the tumors per patient (irrespective of being target tumors or not). 
    adam_adtr |> 
      unnest(data) |> 
      filter(fct_match(trlnkgrp, c("NON-TARGET", "TARGET", "NEW"))) |> 
      distinct(usubjid, trlnkid) |> 
      count(usubjid, name = "n_all_tumors"),
    by = "usubjid"
  ) |> 
  rowwise() |> # Per patient 
  mutate(
    patient_min_t = min(patient_tumors$min_t),
    patient_max_t = max(patient_tumors$max_t),
  ) |>
  ungroup() |> 
  mutate(
    patient_t_width = patient_max_t - patient_min_t + 1, # Range between first and last assessment
    n_tumors = map_int(patient_tumors, nrow), # Tumors per patient
  ) |> 
  left_join(# Get some other time-to-event variables
    adam_adtte |> 
      unnest(data) |> 
      filter(
        (fct_match(paramcd, "PFS1") & fct_match(studyid, c("DS8201-A-U301", "DS8201-A-U302", "DS8201-A-U303"))) |
          (fct_match(paramcd, "PFS") & fct_match(studyid, "DS8201-A-U201")),
        fct_match(parqual, "CENTRAL")
      ) |> 
        select(studyid, usubjid, aval, death = dthfl, dthdt, fpddt, cnsr),
    by = c("studyid", "usubjid")
  ) |> 
  mutate(
    right_censored = cnsr,
    death_week = floor(time_length(dthdt - trtsdt, unit = "weeks") + 1), # Week of death
    progress_week = floor(time_length(fpddt - trtsdt, unit = "weeks") + 1), # Week of disease progression
    progress_week = if_else(death, coalesce(progress_week, death_week), # If death, progress week is death week
                                   if_else(right_censored & is.na(progress_week), max(progress_week, na.rm = TRUE), # Get rid of NA 
                                                                                  progress_week)),
    aval_week = aval %/% 7,
    
    treated = # True is getting the same dosage 
      (fct_match(trial, "Breast01") & fct_match(trtp, "5.4 mg/kg")) | 
      (fct_match(trial, c("Breast02", "Breast03", "Breast04")) & fct_match(trtp, "T-DXd")),
    
    death_week = coalesce(death_week, 0L), # if NA, set to 0.
    
    n_measures = map(patient_tumors, \(tu) tu$n_tumor_measures), # A list of the number of tumor assessments.
    t_measure = map(patient_tumors, \(tu) map(tu$tumor_history, \(tr) pull(tr, week))), # A list of tumor assessment weeks
  ) |> 
  left_join(# Getting some demographic characteristics
    unnest(adam_adsl, data) |> 
      transmute(
        studyid, usubjid, country, age, sex, race, ecogbl, hormonr, her2ic1, nreg, 
        region = coalesce(region1, region2)
      ), 
    by = c("studyid", "usubjid")
  ) |> 
  # Get prior therapy medications per patient
  nest_join(unnest(adam_adcm, data) |> select(usubjid, cmcat, response, cmdecod), by = c("usubjid"), name = "med_data") |> 
  mutate(med_data = map(med_data, \(cm) mutate(cm, response = str_extract(response, "CR|NA|PR|PD|SD|UNK"))))
 
# For PFS focusing on the subset of patients with the needed assessments and calculating PFS. 
unfiltered_pfs_analysis_data <- tumor_analysis_data |> 
  mutate(
    patient_tumors = map(patient_tumors, \(tu) filter(tu, min_t <= 0, max_t > 0)), # Drop tumors that don't have two assessments
    
    # Recalculate these:
    n_measures = map(patient_tumors, \(tu) tu$n_tumor_measures),
    t_measure = map(patient_tumors, \(tu) map(tu$tumor_history, \(tr) pull(tr, week))),
  ) |> 
  filter(map_int(n_measures, sum) > 0) |>
  mutate(
    # After dropping tumors, these need to be recalculated 
    patient_min_t = map_int(patient_tumors, \(tu) min(tu$min_t)),
    patient_max_t = map_int(patient_tumors, \(tu) max(tu$max_t)),
    patient_t_width = patient_max_t - patient_min_t + 1,
    n_tumors = map_int(patient_tumors, nrow),
    
    # The model expects "pfs" to record the number of weeks that were **progress free**, so I need to find the last *assessment* week before
    # disease progress was detected.
    pfs = pmap_int(
      lst(progress_week, death_week, right_censored, patient_tumors), 
      function(progress_week, death_week, right_censored, patient_tumors) {
        event_week <- min(progress_week, death_week, na.rm = TRUE) # Whichever happened first, death or DP.
       
        # Get all the assessment weeks that happened before progression (if not censored). 
        pre_progress_weeks <- unnest(patient_tumors, tumor_history) |>
          distinct(week) |>
          filter(right_censored | week < event_week) |> # Either no even happened (right censored) or use weeks before event.
          pull(week)

        # There are a few patients who just have a single post treatment visit
        if (length(pre_progress_weeks) > 0) max(pre_progress_weeks) else NA_integer_
      }
    ),
  ) %>%
  mutate(
    interval_censored = with(., pfs_model$functions$identify_censoring(pfs, death_week, n_tumors, unlist(n_measures), unlist(t_measure))[[1]]) |> 
      pmax(0) # Issue https://github.com/azu-oncology-rd/ods-adc-early-predict-2023/issues/36
  ) 

pfs_analysis_data <- unfiltered_pfs_analysis_data |> 
  filter(
    !is.na(pfs), pfs >= 0, # No missing pfs and progress must be assured to have happened after treatment
    patient_min_t <= 0, # At least one screening assessment needed,
    patient_max_t > 0, # and at least one post treatment assessment.
  )

treated_pfs_analysis_data <- pfs_analysis_data |> 
  filter(treated) 

pfs_analysis_data |> 
  count(studyid, trial, treated, trtp, trta) |> 
  kableExtra::kbl() |> 
  kableExtra::kable_styling()
```

```{r}
#| label: pfs-stan-data
#| output: false
#| cache: true
#| dependson: prepare-analysis-data

# Convert analysis data into a format usable by Stan
treated_pfs_stan_data <- prepare_pfs_stan_data(treated_pfs_analysis_data, tumor_priors, pfs_priors)

# Same as above but but split by trial
treated_trial_pfs_stan_data <- treated_pfs_analysis_data |> 
  nest(.by = trial) |> 
  deframe() |> 
  imap(\(d, tr) prepare_pfs_stan_data(mutate(d, trial = tr), tumor_priors, pfs_priors))

early_tumor_pairs <- get_early_tumor_pairs(treated_pfs_stan_data)  

# Observed Kaplan-Meier
treated_obs_km <- map_dfr(treated_trial_pfs_stan_data, function(sdata) {
  with(sdata, {
    list(lb = pfs, ub = pfs + interval_censored) |> 
      map_dfr(function(s) {
        pfs_model$functions$estimate_kaplan_meier(s, right_censored, max(s)) |>
          set_names(c("s", "n", "c", "e")) |>
          as_tibble() |> 
          mutate(t = seq(0, n() - 1))
      }, .id = "btype")
    }
  )
}, .id = "trial")
```

```{r}
#| label: tbl-subject-level-summary
#| tbl-cap: Subject-level characteristics at baseline for the primary arm patients included in the meta-analysis.

over_64 <- function(x) sum(x >= 65)

subj_lvl_table1 <- pfs_analysis_data |>
  filter(treated) |> 
  mutate(
    sex = fct_recode(sex, "Female" = "F", "Male" = "M"),
    race = str_to_title(race) |> fct_recode("Missing data" = "") |> fct_relevel("Other", "Missing data", after = Inf),
    region = fct_collapse(
      region, 
      "Europe + Israel" = c("EU", "Europe", "Europe + Israel", "EUROPE"), 
      "Rest of World" = c("Rest of World", "ROW"),
      "North America" = c("North America", "USA"),
    ),
    hormonr = str_to_title(hormonr) |> fct_collapse(Unknown = c("Unknown", "Indeterminate")),
    her2ic1 = str_to_title(her2ic1) |> fct_collapse("Not Examined" = "Ne"),
    
    sum_tumor_size = map_dbl(
      patient_tumors, \(tr) map(tr$tumor_history, \(th) filter(th, week <= 0) |> filter(min_rank(week) == n()) |> pull(mmdiam)) |> 
        compact() |> 
        unlist() |> 
        sum()
    ) |> 
      divide_by(10) # cm
  ) |> 
  tbl_summary(
    by = trial,
    include = c(age, sex, race, region, ecogbl, hormonr, her2ic1, sum_tumor_size, nreg),
    label = list(
      age = "Age (years)", sex = "Sex", race = "Race", region = "Region", ecogbl = "ECOG performance-status score", hormonr = "Hormone-receptor status",
      her2ic1 = "HER2 expression", sum_tumor_size = "Sum of diameters of target lesions (cm)", nreg = "Number of previous regimens"
    ),
    type = list(age = "continuous2"),
    statistic = list(
      age = c("{median} ({min} - {max})", "{over_64}"), sum_tumor_size = "{median} ({min} - {max})",
      nreg = "{median} ({min} - {max})"
    ),
  ) |> 
  add_stat_label(label = list(age = c("Median (Range)", ">64 yr"))) 

subj_lvl_table1 |> as_gt()
```

```{r}
#| label: tbl-prior-med-summary
#| tbl-cap: Previous systemic cancer therapy for the primary arm patients included in the meta-analysis. 

prior_med_table1 <- pfs_analysis_data |>
  filter(treated) |> 
  unnest(med_data) |> 
  filter(fct_match(cmcat, "PRIOR CANCER SYSTEMIC THERAPY")) |> 
  mutate(
    med = fct_collapse(
      cmdecod, 
      "Trastuzumab" = "TRASTUZUMAB", 
      "Trastuzumab emtansine" = "TRASTUZUMAB EMTANSINE", 
      "Pertuzumab" = "PERTUZUMAB",
      "Taxane" = "TAXANES",
      other_level = "Other"
    )
  ) |> 
  tbl_summary(
    by = trial,
    include = c(med),
    label = list(
      med = "Previous systemic cancer therapy"
    ),
    statistic = list(
    ),
  )

prior_med_table1
```

In the below plots we look at the empirical Kaplan-Meier survival, the number of at risk patients, and the number of censored at each interval, for the various trials and arms. We also compare:

a)  Our calculated PFS measure defined as the observed assessment week that occured *just before* progress/death week.
b)  PFS as defined as the progress week in data.

As expected, if we use progress week, survival appears to be longer, while in fact survival is interval censored between the last progress free week and the week just before progress. In particular, the number of censored plot shows bunching at the maximum observed progress week.

```{r}
#| label: fig-survival-dynamics
#| warning: false

km_res_calculated <- pfs_analysis_data |>
  group_by(trial, treated) |>  
  group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "pfs"), .keep = TRUE) |>  
  bind_rows()  

km_res_from_data <- pfs_analysis_data |>
  group_by(trial, treated) |>  
  group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "progress_week"), .keep = TRUE) |>  
  bind_rows()  

km_res <- bind_rows(calculated = km_res_calculated, from_data = km_res_from_data, .id = "pfs_type")

km_res |> 
  filter(fct_match(btype, "lb")) |> 
  ggplot() +
  geom_step(aes(x = t, y = s, color = treated, linetype = pfs_type)) +
  scale_color_discrete("Arm") +
  labs(title = "Kaplan-Meier survival", y = "S(t)") +
  facet_wrap(vars(trial), ncol = 1) +
  NULL

km_res |> 
  filter(fct_match(btype, "lb")) |> 
  ggplot() +
  geom_step(aes(x = t, y = n, color = treated, linetype = pfs_type)) +
  scale_color_discrete("Arm") +
  labs(title = "At risk patients", y = "N") +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y") +
  NULL

km_res |>
  filter(fct_match(btype, "lb")) |> 
  group_by(pfs_type, trial, t) |> 
  summarize(c = sum(c), .groups = "drop") |> 
  ggplot() +
  geom_col(aes(x = t, y = c)) +
  scale_color_discrete("Arm") +
  labs(title = "Number of censored patients", y = "sum(C)") +
  facet_grid(vars(trial), vars(pfs_type), scales = "free_y") +
  NULL

km_res |> 
  filter(fct_match(btype, "lb")) |> 
  mutate(n = n - c - e) |> 
  pivot_longer(c(n, c, e)) |> 
  ggplot(aes(t, value)) +
  geom_col(aes(fill = name, alpha = name)) +
  scale_alpha_manual("", values = c("n" = 0.25)) +
  scale_fill_discrete("", labels = c(c = "Censored", e = "Exiting", n = "Remaining")) +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y") +
  labs(title = "Survival dynamics", y = "") +
  guides(alpha = "none") +
  NULL 
```

```{r}
#| labels: kaplan-meier-upper-and-lower-bounds

km_res |> 
  filter(fct_match(pfs_type, "calculated")) |>
  ggplot() +
  geom_step(aes(x = t, y = s, color = btype)) +
  scale_color_discrete("Survival up to", labels = c(lb = "First possible week", ub = "Last possible week")) +
  labs(title = "Kaplan-Meier survival", y = "S(t)") +
  facet_grid(vars(treated), vars(trial), scales = "free_x", labeller = labeller(treated = label_both)) +
  NULL

# treated_pfs_stan_data[c("pfs", "right_censored", "interval_censored", "patient_trial")] |> 
#   as_tibble() %>% 
#   bind_rows(lb = ., ub = mutate(., pfs = pfs + interval_censored), .id = "btype") |> 
#   rename(trial = patient_trial) |> 
#   mutate(status = !right_censored) %>% 
#   ggsurvfit::survfit2(survival::Surv(pfs, status) ~ btype + trial, data = .) |> 
#   ggsurvfit::ggsurvfit() +
#   geom_step(aes(x = t, y = s, color = btype), linewidth = 4, alpha = 0.2, data = treated_obs_km) +
#   scale_color_discrete("Survival up to", labels = c(lb = "First possible week", ub = "Last possible week")) +
#   facet_wrap(vars(trial)) +
#   labs(title = "Comparing our KM estimator to the {survival} estimation", caption = "Solid line: {survival} Kaplan-Meier curve.\nRibbon: Our estimation.", x = "t", y = "S(t)")  +
#   theme_minimal() +
#   NULL
```

```{r}
#| label: tumor-demographics-table 
#| eval: false

sdtm_dm$data |> 
  map(\(d) tbl_summary(d, include = c("age", "sex", "race", "ethnic", "dthfl"), by = "arm"))
```

```{r}
#| label: fig-tumor-history-plot 
#| fig-cap: Observed tumor sizes for a sample of patients.

treated_pfs_analysis_data |> 
  filter(patient_t_width > 1) |> 
  slice(1:20) |>
  unnest(patient_tumors) |> 
  unnest(tumor_history) |> 
  ggplot() +
  geom_step(aes(week, mmdiam, group = trlnkid), direction = "hv", alpha = 0.5) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = max_mmdiam),
    alpha = 0.25, fill = AZ_pink, 
    data = \(d) group_by(d, usubjid, trlnkid) |> 
      summarize(max_mmdiam = max(mmdiam), xmin = max(keep(week, \(w) w <= 0)), xmax = min(keep(week, \(w) w > 0)), .groups = "drop") |> 
      group_by(usubjid) |> 
      mutate(max_mmdiam = max(max_mmdiam)) |> 
      ungroup()
  ) +
  scale_y_continuous("Largest Tumor Diameter [cm]", labels = scales::label_number(scale = 1/10)) +
  facet_wrap(vars(usubjid), scale = "free") +
  labs(title = "Observed tumor sizes for a sample of patients", x = "Week", y = "Tumor Diameter [mm]") +
  theme(strip.text = element_blank()) +
  NULL
```

```{r, fig.height=7, fig.width=7}
#| label: fig-num-patient-tumors
#| fig-cap: Number of tumors per patient.

treated_pfs_analysis_data |> 
  select(trial, n_tumors, n_all_tumors) |>
  pivot_longer(n_tumors:n_all_tumors, values_to = "n", names_to = "tumor_type") |> 
  mutate(tumor_type = if_else(fct_match(tumor_type, "n_tumors"), "Target", "All")) |> 
  ggplot(aes(n, fill = tumor_type)) +
  geom_bar(position = "identity", alpha = 0.5) +
  geom_vline(xintercept = 2, linetype = "dashed") +
  scale_fill_discrete("Tumor Type", type = AZ_palette) +
  labs(x = "Number of Patient Tumors", y = "Count") +
  facet_grid(vars(trial), margins = TRUE, scales = "free_y")
```

```{r, fig.height=4, fig.width=8}
#| label: fig-num-organ-tumors-per-patient
#| fig-cap: Number of tumors per organ per patient. This plot is restricted to the organs that have at least 250 incident tumors in the data. 

sdtm_tu |> 
  unnest(data) |> 
  semi_join(treated_pfs_analysis_data, by = "usubjid") |> 
  distinct(trial, usubjid, tulnkid, tuloc, tuorres) |>
  mutate(
    tuloc = fct_lump_min(tuloc, 250) |> fct_infreq() |> fct_relabel(str_to_title) |> fct_relevel("Other", after = Inf)
  ) |> 
  filter(!fct_match(tuloc, "Other")) |> 
  group_by(trial, usubjid, tuloc) |> 
  summarize(n_target_tumors = sum(tuorres == "TARGET"), n_all_tumors = n(), .groups = "drop") |> 
  pivot_longer(c(n_all_tumors, n_target_tumors), values_to = "n", names_to = "tumor_type", names_pattern = "(all|target)") |> 
  ggplot(aes(n, fill = tumor_type)) +
  geom_bar(position = "dodge", alpha = 0.5) +
  geom_vline(xintercept = 2, linetype = "dashed") +
  scale_x_continuous("Number of Tumors", breaks = seq(0, 10)) +
  scale_fill_discrete("Tumor Type", labels = str_to_title, type = AZ_palette) +
  labs(y = "Count") +
  facet_wrap(vars(tuloc), scales = "free")
```

```{r}
#| label: fig-tumor-locations
#| fig-cap: The number of tumors in the data by location.

sdtm_tu |> 
  unnest(data) |> 
  semi_join(treated_pfs_analysis_data, by = "usubjid") |> 
  distinct(usubjid, tulnkid, tuloc, tuorres) |>
  mutate(
    tuloc = fct_lump_min(tuloc, 25) |> fct_infreq() |> fct_relabel(str_to_title) |> fct_relevel("Other", after = Inf)
  ) |>  
  ggplot() +
  geom_bar(aes(tuloc, fill = tuorres), alpha = 0.5) +
  scale_fill_discrete("Tumor Type", labels = str_to_title) +
  labs(x = "Tumor Location", y = "Tumor Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

```{r}
#| label: fig-patients-with-tumor-locations
#| fig-cap: The number of patients in the data with tumors in locations.

sdtm_tu |> 
  unnest(data) |> 
  semi_join(treated_pfs_analysis_data, by = "usubjid") |> 
  distinct(usubjid, tuloc, tuorres) |>
  mutate(
    tuloc = fct_lump_min(tuloc, 25) |> fct_infreq() |> fct_relabel(str_to_title) |> fct_relevel("Other", after = Inf)
  ) |>  
  ggplot() +
  geom_bar(aes(tuloc, fill = tuorres), alpha = 0.5) +
  scale_fill_discrete("Tumor Type", labels = str_to_title) +
  labs(x = "Tumor Location", y = "Patient Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

```{r}
#| label: fig-early-tumor-pair-count
#| fig-cap: Tumor size heatmap of counts in the data.

early_tumor_pairs |> 
  ggplot(aes(x = x0, y = x1)) +
  geom_point(aes(color = trial), alpha = 0.4) +
  # scale_color_viridis_d("") +
  scale_color_discrete("", type = AZ_palette) +
  labs(x = "Pre-screening Tumor Size [cm]", y = "Post-treatment Tumor Size [cm]") +
  coord_fixed() +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.25, 0.85)) +
  NULL
```

```{r, fig.height=2, fig.width=5}
#| label: fig-early-tumor-weeks
#| layout-ncol: 2
#| fig-cap:
#|    - First and second tumor size assessment weeks.
#|    - Time in weeks between first and second assessment.

early_tumor_pairs |>
  select(t0:t1) |> 
  pivot_longer(t0:t1, names_to = "t_index", values_to = "t") |> 
  ggplot(aes(t)) +
  geom_bar(aes(fill = t_index), position = "identity", alpha = 0.5) +
  scale_fill_discrete("Assessment", labels = c(t0 = "Pre-screening", t1 = "Post-treatment"), type = AZ_palette) +
  labs(x = "Week", y = "Count") +
  guides(fill = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.8, 0.7))

early_tumor_pairs |>
  mutate(t_diff = t1 - t0) |> 
  ggplot(aes(t_diff)) +
  geom_bar(alpha = 0.5) +
  labs(x = "Difference in Weeks", y = "Count")
```

```{r}
#| label: fig-pfs-distribution
#| code-fold: false
#| fig-cap: PFS empirical distribution. We're only showing uncensored observations.

pfs_analysis_data |> 
  filter(!right_censored, treated) |> 
  ggplot() +
  geom_bar(aes(pfs)) +
  labs(title = "PFS empirical distribution", x = "PFS", y = "Count") +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y")
```

```{r}
#| label: fig-patient-visits-pattern
#| fig-cap: Patient visits pattern.

tumor_analysis_data |> 
  select(trial, n_tumors, t_measure) |> 
  unnest(t_measure) |> 
  unnest(t_measure) |> 
  mutate(epoch = if_else(t_measure <= 0, "pre-screening", "treatment") |> as_factor()) |> 
  ggplot(aes(t_measure)) +
  geom_bar(aes(fill = epoch), alpha = 0.5) +
  scale_fill_discrete("", labels = str_to_title, type = AZ_palette) +
  labs(x = "Week", y = "Count") +
  facet_wrap(vars(trial), scales = "free_y") +
  theme(legend.position = "top")
```

```{r}
#| label: fig-treatment-start-end

tumor_analysis_data |> 
  ggplot() +
  geom_segment(aes(x = trtsdt, xend = trtedt, y = usubjid, yend = usubjid, color = country), alpha = 0.5) +
  scale_color_discrete("Country") +
  labs(title = "Patient treatment start/end timelines", x = "", y = "") +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_blank(), panel.grid.major.y = element_blank()) +
  NULL
```

```{r}
#| label: fig-interval-censoring

interval_censoring_data <- pfs_analysis_data %>% 
  mutate(
    interval_censoring = prepare_pfs_stan_data(., tumor_priors, pfs_priors) |>
      with(pfs_model$functions$identify_censoring(pfs, death_week, n_patient_tumors, n_measures, t_measure)) |> 
      pluck(1)
  )

interval_censoring_data |> 
  ggplot(aes(y = interval_censoring > 0)) +
  geom_bar(position = "dodge", alpha = 0.5) +
  scale_y_discrete("Interval Censored") +
  facet_wrap(vars(trial), ncol = 1) +
  labs(title = "Number of interval censoring", x = "") 

interval_censoring_data |> 
  filter(interval_censoring > 0) |> 
  ggplot(aes(interval_censoring)) +
  geom_histogram(alpha = 0.5, fill = AZ_platinum, position = "identity", binwidth = 5) +
  labs(
    title = "Histogram of interval censoring", 
    subtitle = "Conditional on any interval censoring", 
    x = "Interval Censoring Weeks", 
    y = "",
    caption = "Binwidth = 5"
  ) +
  facet_wrap(vars(trial), ncol = 2, scales = "free_y") +
  NULL
```

```{r}
#| label: sim-settings

n_patients <- 250
patient_measures <- 41 
max_pfs <- 45 
```

## Excluded Data

In terms of tumor data, we dropped rows from the ADTR data that 

(i) did not have a treatment start data, 
(ii) had a missing or non-positive tumor diameter, or 
(iii) did not have TRUE accepted flag.

```{r}
#| label: excluded-tumor-rows

adam_adtr |> 
  unnest(data) |> 
  filter(fct_match(paramcd, "LDIAM")) |>
  transmute(
    exclude = is.na(trdy) | is.na(aval) | !tracptfl | aval <= 0, 
    missing_trdy = is.na(trdy), 
    missing_aval = is.na(aval) | aval <= 0, 
    tracptfl 
  ) |> 
  gtsummary::tbl_summary()
```

We exclude patients 

(i) with missing PFS,
(ii) PFS could have occurred during pre-screening (before treatment),
(iii) have their first RECIST assessment after pre-screening, or
(iv) have their last RECIST assessment before treatment.

```{r}
#| label: excluded-patients

unfiltered_pfs_analysis_data |>
  transmute(
    excluded = is.na(pfs) | pfs < 0 | patient_min_t > 0 | patient_max_t <= 0,
    missing_pfs = is.na(pfs),
    pfs_in_prescreening = pfs < 0,
    first_measure_not_in_prescreening = patient_min_t > 0,
    last_measure_in_prescreening = patient_max_t <= 0
  ) |> 
  gtsummary::tbl_summary()
```

Notice that 27% of the patients are excluded because progression could have occurred before treatment started; it progression was detected on the first post-treatment visit.

```{r}
#| label: fig-early-pfs-progress-week

unfiltered_pfs_analysis_data |>
  filter(pfs < 0) |> 
  transmute(treated, progress_week = pmin(progress_week, death_week)) |> 
  ggplot(aes(progress_week)) +
  geom_bar(aes(fill = treated), position = "dodge")
```

# Tumor Size Model

I haven't yet documented much about this model or done any kind of evaluation. I'm only using it at this time to generate fake tumor size data for testing.

We model tumor size as a nonlinear function with a Gaussian process prior (with common hyperparameters),

$$
\log(X_{it}) \sim  \mathtt{GP}(\mu^\textrm{tumor}, k(\cdot;\alpha^\textrm{tumor}, \rho^\textrm{tumor})) 
$$ where $$k(p, q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right).$$

## Prior Prediction

```{r}
#| label: test-data
#| cache: true
#| dependson: priors

t_offset <- 5

test_data <- lst(
  fit_tumor_data = FALSE, # Prior prediction
  gen_tumor_sizes = TRUE,
  predict_missing_sizes = FALSE,
  multilevel_patient = FALSE,
  multilevel_tumor = FALSE,
  
  n_patients,
  n_trials = 4,
  patient_trial = sample(4, n_patients, replace = TRUE), 
  n_patient_tumors = 1 + rbinom(n_patients, 5, 0.4), # We're not yet generatively modeling the number of tumors
  n_measures = rep(patient_measures, sum(n_patient_tumors)),
  t_measure = rep(0:(patient_measures - 1), sum(n_patient_tumors)) - t_offset,
  tumor_size = rep(0, sum(n_measures)),
  tumor_location = sample(1:50, sum(n_patient_tumors), replace = TRUE), 
  n_tumor_locations = n_distinct(tumor_location),
  
  !!!tumor_priors  
)
```

```{r}
#| label: tumor-prior-predict
#| warning: false
#| cache: true
#| dependson: test-data 

tumor_prior_res <- tumor_model$sample(
  data = test_data, 
  iter_warmup = 500, iter_sampling = 500, 
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "tumor_prior", 
  refresh = 0
)

# Saving this to use in SBC runs.
tumor_prior_res$save_data_file(dir = file.path(tmp_dir, "data"), basename = "prior_tumor", timestamp = FALSE, random = FALSE)
```

```{r}
#| label: tumor-prior-predict-results
#| cache: true
#| cache-lazy: false
  
# This is the tumor data I'll be using in the PFS model
fake_tumor_data <- tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index], ndraws = 1) |> 
  prepare_tumor_size_rvars(test_data) |> 
  unnest_rvars()

# This will be used by SLURM jobs that need tumor data (e.g., PFS SBC run).
write_rds(fake_tumor_data, file.path(tmp_dir, "data", "fake_tumor.rds"))
```

```{r}
#| label: tumor-model-diagnose
#| cache: true
#| dependson: tumor-prior-predict

tumor_prior_res$cmdstan_diagnose()
```

```{r}
#| label: fig-tumor-prior-predict-plot
#| fig-cap: Prior prediction checking for tumor model.
#| cache: true
#| dependson: tumor-prior-predict

tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index]) |>
  ungroup() |> 
  prepare_tumor_size_rvars(test_data) |> 
  filter(patient_id == 1, tumor_index < 6) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = tumor_size, group = str_c(patient_id, tumor_index, .draw)), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 5)) +
  stat_lineribbon(aes(ydist = tumor_size, group = tumor_index), fill = "black", alpha = 0.25, .width = 0.8) +
  scale_x_continuous(breaks = seq(0, 40, 4)) +
  labs(title = "Prior predicted tumor sizes", y = "Tumor Size [cm]", caption = "Lines show tumors sizes for a subset of tumors of a single patient.") +
  facet_wrap(vars(tumor_index)) +
  coord_cartesian(ylim = c(0, 10)) +
  theme(strip.text = element_blank()) +
  NULL
```

## Fake Data Simulation

Getting a few fake datasets and assessing the model's ability to recover parameter values.

```{r}
#| label: tumor-fake-sim
#| output: false
#| cache: true
#| dependson: test-data, priors

tumor_fake_res <- tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index], ndraws = 24) |> # We'll simulate parameters and data for a number simulations. 
  ungroup() |> 
  prepare_tumor_size_rvars(test_data) |> 
  unnest_rvars() |> 
  nest(fake_data = !c(.draw, .iteration, .chain)) |> 
  mutate(
    fake_fit = furrr::future_map2(.draw, fake_data, function(id, d) {
    # fake_fit = map2(.draw, fake_data, function(id, d) {
      test_data |> 
        list_modify(
          fit_data = TRUE,
          gen_tumor_sizes = FALSE,
          tumor_size = d$tumor_size
        ) |> 
        tumor_model$sample(
          iter_warmup = 200, iter_sampling = 200,
          parallel_chains = 4,
          output_dir = file.path(tmp_dir, "fit"), 
          output_basename = str_c("fake_tumor_", id), 
          init = \(chain_id) lst(
            pop_tumor_gp_rho = invgamma::rinvgamma(1, 7.3, 7.5),
            tumor_mean = rnorm(1, 2.8, 0.1),
            tumor_sd = abs(rnorm(1, 0, 1.25))
          ),
          max_treedepth = 15
        )
    # })
    }, .progress = FALSE, .options = furrr::furrr_options(seed = TRUE))
  ) 
```

```{r}
#| label: fig-tumor-fake-data-params
#| fig-cap: Fake data simulation assessing model's ability to recover posterior of model parameters.
#| cache: true
#| dependson: tumor-fake-sim

tumor_fake_res |> 
  group_by(.draw) |> 
  reframe(map_dfr(fake_fit, \(f) gather_rvars(f, pop_tumor_gp_intercept, pop_tumor_gp_alpha, pop_tumor_gp_rho))) |> 
  left_join(
    tumor_prior_res |> 
      gather_rvars(pop_tumor_gp_intercept, pop_tumor_gp_alpha, pop_tumor_gp_rho) |> 
      unnest_rvars(), by = c(".draw", ".variable"), suffix = c("", "_true")) |> 
  mutate(.value = .value - .value_true) |> 
  ggplot(aes(y = .draw)) +
  stat_pointinterval(aes(xdist = .value), .width = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Comparing posterior estimate of model parameters against known values", 
    x = "Difference from True Parameter Value", 
    y = "", 
    caption = str_glue("Showing {nrow(tumor_fake_res)} simulations.")
  ) +
  facet_wrap(vars(.variable), scales = "free") +
  theme(axis.text.y = element_blank()) +
  NULL  
```

## Fit Data

Now using actual trial data.

```{r}
#| label: tumor-fit
#| output: false
#| cache: true
#| dependson: prepare-analysis-data, priors

treated_tumor_analysis_data <- pfs_analysis_data |> 
  filter(treated) 
  
tumor_stan_data <- lst(
  fit_tumor_data = TRUE,
  gen_tumor_sizes = TRUE,
  predict_missing_sizes = TRUE, 
  multilevel_patient = TRUE,
  multilevel_tumor = FALSE,
  
  !!!prepare_tumor_stan_data(treated_tumor_analysis_data),
  !!!tumor_priors
)

tumor_fit_res <- tumor_model$sample(
  tumor_stan_data, 
  iter_warmup = 200, iter_sampling = 200, 
  max_treedepth = 15,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "tumor" 
)

```

```{r}
#| label: fitted-tumor-data
#| cache: true
#| dependson: tumor-fit

n_full_measures <- with(tumor_stan_data, n_measures + pfs_model$functions$calculate_n_missing_measures(n_measures, t_measure, n_patient_tumors))

all_tumor_size <- tumor_fit_res |> 
  spread_rvars(all_tumor_size[idx]) |> 
  mutate(
    usubjid = rep(with(treated_tumor_analysis_data, rep(usubjid, n_tumors)), n_full_measures),
    trlnkid = rep(treated_tumor_analysis_data |> unnest(patient_tumors) |> pull(trlnkid), n_full_measures),
  ) |> 
  left_join(
    treated_tumor_analysis_data |> 
      unnest(patient_tumors) |> 
      select(usubjid, trlnkid, trial, matches("(patient_)?(min|max)_t")), 
    by = c("usubjid", "trlnkid")
  ) |> 
  group_by(usubjid, trlnkid) |> 
  mutate(week = first(min_t):first(patient_max_t)) |> 
  ungroup() |> 
  median_qi(all_tumor_size, .width = 0.8)
```

```{r, fig.height=5, fig.width=6}
#| label: fig-fit-tumor-history-plot
#| fig-height: 8
#| fig-cap: Showing GP tumor size dynamics with interpolation.
#| cache: true
#| dependson: fitted-tumor-data 

all_tumor_size |> 
  nest(patient_data = !c(trial, usubjid)) |> 
  group_by(trial) |> 
  sample_n(4) |> 
  ungroup() |> 
  unnest(patient_data) |> 
  ggplot(aes(week)) +
  geom_lineribbon(aes(y = all_tumor_size, ymin = .lower, ymax = .upper, group = trlnkid), fill = AZ_platinum, alpha = 0.5) +
  facet_wrap(vars(usubjid), scales = "free", ncol = 4) +
  labs(x = "Week", y = "Tumor Size [cm]") +
  theme(strip.text = element_blank())
```

# Survival Model

```{r}
#| label: pfs-common
#| include: false

hist_breaks <- seq(10, 150, by = 10)
```

We will use a piecewise-constant proportional hazard model:

-   Time is discretized into $t \in \{0,\dots,T\}$, with $t = 0$ is the baseline interval.
-   $Y_{it}$ is an indicator for disease progression at interval $t$, with $Y_{i0} = 0$.
-   $S_i$ is the number of intervals observed without disease progression, \[ \Pr[S > s] = \prod\_{r=0}\^s \Pr[Y_{i,r+1} = 0 \mid Y_{i,r} = 0]. \]
-   $X_{it}$ are--possibly time-variant--covariates.
-   $C_i$ is an indicator of whether $i$ is right censored.
-   $M_{i} \in \mathbb{Z}^{+}$ is the difference between $S_i$ and the previous interval $i$ was observed progression free ($M_i = 1$ if $C_i = 1$).

The probability of remaining progression-free at $m \geq 1$ is $$
\alpha_t(x;\theta) = \Pr[Y_t = 0\mid Y_{t-1} = 0, X_t = \bf{x}] = \exp[-\lambda(t, \bf{x}; \boldsymbol{\theta})] 
$$

and the hazard function . $$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t \kappa(\bf{x}; \boldsymbol{\beta})
$$

We assume that censoring in conditionally independent of time to progression. We will use a linear form for $\kappa(\bf{x}; \boldsymbol{\beta}) = \exp(\bf{x}\boldsymbol{\beta})$.

Aside from whatever priors we use for $\boldsymbol{\beta}$, we will use a Gaussian process prior for $\lambda_t$, $$
\log(\lambda_t) \sim \mathtt{GP}(0, k),
$$ where $k(\lambda_p, \lambda_q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right)$.

Our likelihood, accounting for interval and right censoring, is $$
\mathcal{L}_i = \begin{cases}
  \prod_{t=1}^{S_i-M_i}\alpha_t(x_i;\theta)\sum_{r=S_i-M_i+1}^{S_i + 1}\left\{\prod_{m=S_i - M_i + 1}^{r-1}\alpha_m(x_i;\theta)[1-\alpha_{r}(x_i;\theta)]\right\} & \textrm{if }C_i = 0 \\
  \prod_{t=1}^{S_i}\alpha_t(x_i;\theta) & \textrm{if }C_i = 1 
  \end{cases}
$$

## Prior Prediction

```{r}
#| label: pfs-prior-predict
#| cache: true 
#| dependson: test-data, priors

early_fake_tumors <- fake_tumor_data |> 
  group_by(patient_id) |> 
  filter(t %in% c(max(keep(t, \(x) x <= 0)), min(keep(t, \(x) x > 0)))) 

pfs_test_data <- test_data |> 
  list_modify(
    fit_data = FALSE,
    gen_pfs = TRUE,
    gen_interval_censored = FALSE,
    tumor_hazard_type = 1,
    ignore_interval_censoring = FALSE,
    add_trial_level = TRUE,
    add_tumor_location_level = FALSE,
    fit_post_2nd_meaure_only = FALSE, 
    
    pfs = rep(max_pfs, n_patients),
    death_week = rep(0, n_patients),
    right_censored = rep(FALSE, n_patients),
    tumor_size = fake_tumor_data$tumor_size,
    
    grid_tumors = array(NA, dim = 0),
    n_grid_tumors = 0,
    
    tumor_grid_range = with(early_fake_tumors, seq(min(tumor_size), max(tumor_size), 1)), 
    
    !!!pfs_priors,
    
    # !!!tumor_stan_data
    use_tumor_model = FALSE
  ) %>% 
  list_assign(n_tumor_grid_range = length(.$tumor_grid_range))

t_measure_list <- with(pfs_test_data, list_measures(t_measure, n_measures, n_patient_tumors))  
max_t <- t_measure_list |> 
  map(unlist) |> 
  map_int(max)
first_t_measure <- t_measure_list |> map(first) |> unlist()

pfs_prior_res <- pfs_model$sample(
  pfs_test_data, 
  refresh = 0, 
  iter_warmup = 500, iter_sampling = 500,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "pfs_prior" 
)

rep_pfs_rvar <- pfs_prior_res |> 
  spread_rvars(rep_pfs[patient_index], rep_right_censored[patient_index])

rep_pfs_param_rvar <- pfs_prior_res |>
  spread_rvars(
    tumor_stim_trial_intercept[trial],
    tumor_stim_trial_coef[trial, t],
    log_lambda_gp_alpha,
    log_lambda_gp_rho,
    log_lambda_gp_intercept
  ) |>
  pivot_wider(names_from = t, values_from = tumor_stim_trial_coef, names_prefix = "tumor_stim_coef_")

lambda_rvar <- pfs_prior_res |>
  spread_rvars(log_lambda[t], base_pf_cond_prob[t], base_survival[t], one_tumor_survival[t], log_trial_lambda[trial, t]) |>
  mutate(lambda = exp(log_lambda), trial_lambda = exp(log_trial_lambda)) |>
  nest(trial_data = contains("trial"))

expected_pfs_rvar <- pfs_prior_res |>
  gather_rvars(base_cond_expected_pfs, one_tumor_cond_expected_pfs, base_cond_median_pfs, one_tumor_cond_median_pfs) |>
  mutate(
    stat_type = str_extract(.variable, "expected|median"),
    tumor_size = str_extract(.variable, "base|one_tumor")
  )
```

```{r}
#| label: fig-prior-lambda
#| fig-cap: Surival prior predicted baseline hazard. Plot shows the 80% credible ribbon, and lines represent 50 draws from the prior.
#| cache: true
#| dependson: pfs-prior-predict

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(x = "m", y = latex2exp::TeX("$\\lambda_t$"))
```

```{r}
#| label: fig-prior-trial-lambda
#| fig-cap: Surival prior predicted trial-level baseline hazard. Plot shows the 80% credible ribbon, and lines represent 50 draws from the prior.
#| dependson: pfs-prior-predict

lambda_rvar |>
  unnest(trial_data) |>
  select(t, trial, trial_lambda) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = trial_lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = trial_lambda, group = .draw), alpha = 0.125, data = \(d) unnest_rvars(d) |> group_by(trial) |>  filter(.draw <= 75)) +
  labs(x = "m", y = latex2exp::TeX("$\\lambda_t$")) +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(0.0, 0.05)) +
  NULL
```

```{r}
#| label: fig-baseline-hazard-trial-variation
#| fig-cap: Prior prediction of baseline hazard variation, by trial.

lambda_rvar |>
  unnest(trial_data) |> 
  filter(trial == 1) |> # All the trials are the same here
  mutate(trial_lambda_effect = exp(log_trial_lambda - log_lambda)) |> 
  select(t, trial_lambda_effect) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = trial_lambda_effect), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = trial_lambda_effect, group = .draw), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 75)) +
  labs(y = latex2exp::TeX(r"{$\exp[\log(\lambda_{ts}) - \log(\lambda_t)]$}")) +
  coord_cartesian(ylim = c(0.0, 3)) +
  NULL
```

```{r}
#| label: fig-prior-base-pf-prob
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  gather_rvars(trial_base_pf_cond_prob[s, t]) |> 
  filter(s == 1) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = .value), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = .value, group = .draw), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 75)) +
  labs(title = "Base conditional probability of remaining progress free", x = "t", y = latex2exp::TeX(r"{$\alpha_t$}")) 
```

```{r}
#| label: pf-prob-plot
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  spread_rvars(disease_progress_prob[i]) |>
  mutate(
    patient_index = rep(seq(pfs_test_data$n_patients), each = max(max_t)), 
    trial = rep(pfs_test_data$patient_trial, each = max(max_t)), 
    t = rep(1:max(max_t), pfs_test_data$n_patients),
  ) |>
  nest(.by = c(trial, patient_index)) |> 
  group_by(trial) |> 
  sample_n(2) |> 
  ungroup() |> 
  unnest(data) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = disease_progress_prob), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = disease_progress_prob, group = str_c(patient_index, .draw)), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 20)) +
  labs(title = "Conditional probability of remaining progress free", x = "t", y = latex2exp::TeX("$\\alpha_t$")) +
  # coord_cartesian(ylim = c(0, 1)) +
  facet_wrap(vars(patient_index, trial), scales = "free_y") +
  NULL
```

```{r}
pfs_prior_res |> 
  gather_rvars(trial_base_pf_cond_prob[s, t], trial_one_tumor_pf_cond_prob[s, t]) |> 
  filter(s == 1, t <= 10) |> 
  ggplot() +
  stat_halfeye(aes(x = t, ydist = .value, fill = .variable), alpha = 0.25, .width = 0.8)
```

```{r}
pfs_prior_res |> 
  gather_rvars(tumor_stim_pop_intercept, tumor_stim_pop_coef[t], tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, t]) |>  
  mutate(
    .variable = if_else(is.na(trial), .variable, str_glue("{.variable}[{trial}]")),
    .variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]")),
  ) |> 
  ggplot() +
  stat_halfeye(aes(xdist = .value, y = .variable), color = "black", alpha = 0.5, .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  # coord_cartesian(xlim = c(0, 1)) +
  NULL
```

```{r}
#| label: base-surivival-plot
#| dependson: pfs-prior-predict

lambda_rvar |>
  select(t, base_survival, one_tumor_survival) |> 
  pivot_longer(!t, names_to = "survival_type", values_to = "survival") |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = survival, fill = survival_type, color = survival_type), step = TRUE, alpha = 0.25, .width = 0.8) +
  labs(title = "Base survival probability", x = "t", y = latex2exp::TeX("$S(t)$"))
```

```{r}
#| label: cond-expected-pfs-plot
#| dependson: pfs-prior-predict

expected_pfs_rvar |> 
  mutate(.variable = str_remove(.variable, "_cond_expected_pfs")) |> 
  ggplot(aes(y = tumor_size)) +
  stat_pointinterval(aes(xdist = .value, color = stat_type), .width = 0.8, position = "dodge") +
  labs(title = "Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "t", y = "")
```

```{r}
#| label: prior-kaplan-meier 
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  spread_rvars(km_est[m]) |> 
  ggplot(aes(m - 1)) +
  stat_lineribbon(aes(ydist = km_est), step = TRUE, fill = "black", alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", x = "t", y = latex2exp::TeX("$S(t)$")) +
  # scale_x_continuous("t", breaks = seq(0, 40, 2))
  NULL
```

```{r}
#| eval: false
#| include: false

pfs_prior_res |> 
  get_tumor_hazard_ratio_summary(pop_tumor_effect[n], pfs_test_data$tumor_grid_range) |> 
  ggplot(aes(x = x0, y = x1, z = q)) +
  stat_summary_hex(binwidth = c(1, 1)) +
  scale_fill_continuous_divergingx("Roma", name = latex2exp::TeX(r"{\exp(\beta\cdot x)}"), mid = 1) +
  labs(x = "Pre-screening Tumor Size", y = "Post-treatment Tumor Size") +
  facet_wrap(vars(p), labeller = labeller(p = label_percentile))
```

```{r}
#| eval: false
#| include: false

pfs_prior_res |> 
  get_tumor_hazard_ratio_summary(trial_tumor_effect[trial, n], pfs_test_data$tumor_grid_range, by = trial) |> 
  filter(trial == 1) |> 
  ggplot(aes(x = x0, y = x1, z = q)) +
  stat_summary_hex(binwidth = c(1, 1)) +
  scale_fill_continuous_divergingx("Roma", name = latex2exp::TeX(r"{\exp(\beta\cdot x)}"), mid = 1) +
  labs(x = "Pre-screening Tumor Size", y = "Post-treatment Tumor Size") +
  facet_wrap(vars(p), labeller = labeller(p = label_percentile))
```

```{r}
#| eval: false
#| include: false

pfs_prior_res |> 
  get_tumor_hazard_ratio_summary(location_tumor_effect[organ, n], pfs_test_data$tumor_grid_range, by = organ) |> 
  filter(organ == 1) |> 
  ggplot(aes(x = x0, y = x1, z = q)) +
  stat_summary_hex(binwidth = c(1, 1)) +
  scale_fill_continuous_divergingx("Roma", name = latex2exp::TeX(r"{\exp(\beta\cdot x)}"), mid = 1) +
  labs(x = "Pre-screening Tumor Size", y = "Post-treatment Tumor Size") +
  facet_wrap(vars(p), labeller = labeller(p = label_percentile))
```

## Fake Data Simulation

Here we're using a set of ad hoc survival model parameters and generating a number of datasets, then we compare the posterior distributions of these parameters against the known truth. We do this three times:

1.  Without any interval censoring
2.  With interval censoring but ignored by the model (observed PFS is assumed to be actual PFS)
3.  With interval censoring which is accounted for by the model

```{r}
#| label: fake-data-sim
#| output: false
#| cache: true 

n_sim <- 12  
t_to_keep <- c(-t_offset, 0, seq(1 - t_offset, patient_measures - t_offset - 1, by = 5)) |> sort()
fake_data_sim_settings <- pfs_test_data |> list_assign(add_trial_level = FALSE)

no_interval_censoring <- run_sbc_sims(pfs_model, fake_data_sim_settings, n_sim, output_name = "fake_data_sim", keep_fit = TRUE, gen_pfs = TRUE)

fake_data_sim <- bind_rows(
  no_interval_censoring = no_interval_censoring,
  
  interval_censoring = run_sbc_sims(
    pfs_model, fake_data_sim_settings, n_sim, ignore_interval_censoring = FALSE, output_name = "fake_data_sim_ic",
    keep_fit = TRUE, gen_pfs = TRUE, reuse_data = no_interval_censoring, drop_measures = t_to_keep, keep_only = TRUE
  ),
  
  interval_censoring_ignored = run_sbc_sims(
    pfs_model, fake_data_sim_settings, n_sim, ignore_interval_censoring = TRUE, output_name = "fake_data_sim_ic_ignored",
    keep_fit = TRUE, gen_pfs = TRUE, reuse_data = no_interval_censoring, drop_measures = t_to_keep, keep_only = TRUE
  ),
  
  .id = "sim_type"
) |> 
  rename(sim_id = .draw)
```

```{r}
#| eval: false
#| include: false

fake_data_sim |> 
  rowwise() |> 
  reframe(
    sim_id, sim_type,
    spread_rvars(sim_fit, km_est[t]),
    obs_km = with(sim_data, pfs_model$functions$estimate_kaplan_meier(pfs, right_censored, max(t) - 1)[[1]]),
  ) |>
  filter(sim_id <= 4) |> 
  ggplot(aes(t)) +
  geom_step(aes(y = obs_km, color = sim_type)) +
  stat_lineribbon(aes(ydist = km_est, fill = sim_type, color = sim_type), alpha = 0.25, .width = 0.8) +
  facet_wrap(vars(sim_id), scales = "free_y")
```

```{r}
#| label: fake-data-params
#| fig-cap: "Comparing fake data simulation parameter posteriors against their known true values."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  transmute(
    sim_id,
    sim_type,
    
    map2_dfc(est, select(true, all_of(names(est))), \(e, t) e - t),
  ) |> 
  # add_row(sim_id = 0, sim_type = "prior", rep_pfs_param_rvar) |>
  pivot_longer(!c(sim_id, sim_type), names_to = "param", values_to = "est") |>
  ggplot(aes(sim_id)) + 
  geom_hline(yintercept = 0, linetype = "dotted") +
  stat_pointinterval(aes(ydist = est, color = sim_type), size = 0.25, .width = 0.8, fatten_point = 0, position = "dodge") +
  scale_color_discrete("") +
  facet_wrap(vars(param), scales = "free_y") +
  labs(x = "", y = "", caption = "Black horizontal line represents the true parameter value.") +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

```{r}
#| label: fake-data-lambda
#| fig-cap: "Comparing fake data simulation base hazard posteriors against their known true values."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  rowwise() |> 
  reframe(sim_id, sim_type, spread_rvars(sim_fit, log_lambda[m])) |> 
  # mutate(sim_log_lambda_post = furrr::future_map(sim_fit, \(f) spread_rvars(f, log_lambda[m]))) |> 
  # select(!sim_fit) |>
  # unnest(sim_log_lambda_post) |> 
  # bind_rows(transmute(lambda_rvar, m = t, log_lambda, sim_id = 0, sim_type = "prior")) |> 
  mutate(lambda = exp(log_lambda)) |> 
  filter(m <= 20) |> 
  ggplot(aes(sim_id)) + 
  # geom_hline(aes(yintercept = true_val), data = \(d) distinct(d, m) |> mutate(true_val = exp(log_lambda[m]))) +
  stat_pointinterval(
    aes(ydist = lambda, color = sim_type), 
    size = 0.5, .width = 0.8, fatten_point = 0, position = "dodge"
  ) +
  labs(x = "", y = "", caption = "Black horizontal line represents the true parameter value.") +
  facet_wrap(vars(m), scales = "free_y") +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

```{r}
#| eval: false

fake_data_sim |>
  filter(sim_id <= 4) |> 
  mutate(sim_log_lambda_post = furrr::future_map(sim_fit, \(f) spread_rvars(f, log_lambda[t]))) |> 
  select(!sim_fit) |>
  unnest(sim_log_lambda_post) |>
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = log_lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = log_lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 10)) +
  geom_line(aes(y = log_lambda), color = "red", data = tibble(t = 1:max_all_t, log_lambda = log_lambda[1:max_all_t])) +
  facet_grid(vars(sim_type), vars(sim_id))
```

```{r}
#| label: fake-data-km
#| fig-cap: "Comparing the fake data Kaplan-Meier survival probability against the true probability."
#| cache: true
#| dependson: fake-data-sim
#| eval: false

fake_data_sim |> 
  mutate(km_est = furrr::future_map(sim_fit, \(f) spread_rvars(f, km_est[m]))) |> 
  select(sim_id, sim_type, km_est) |> 
  filter(sim_id <= 6) |> 
  unnest(km_est) |> 
  ggplot(aes(m - 1)) +
  # geom_line(aes(y = km_est), data = \(sim) semi_join(true_km, sim, by = "m")) +
  # This comes from a draw so not really reliable as the truth.
  stat_lineribbon(aes(ydist = km_est, color = sim_type, fill = sim_type), step = TRUE, alpha = 0.25, .width = 0.8) +
  scale_x_continuous("t", breaks = seq(0, 40, 4)) +
  labs(y = "S(t)", caption = "Black line represents the true survival probability.") +
  facet_wrap(vars(sim_id)) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fake-data-cond-expected-pfs-plot
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |>
  rowwise() |> 
  mutate(map2_dfc(est, select(true, all_of(names(est))), \(e, t) e - t)) |> 
  select(sim_id, sim_type, matches("(expected|median)_pfs")) |> 
  pivot_longer(matches("(expected|median)_pfs")) |> 
  mutate(
    stat = str_extract(name, "expected|median"),
    name = str_remove(name, "_cond_(expected|median)_pfs"),
    est_type = if_else(sim_id > 0, "post", "prior")
  ) |> 
  ggplot(aes(y = sim_id)) +
  stat_pointinterval(aes(xdist = value, color = sim_type), size = 0.5, .width = 0.8, position = "dodge") +
  geom_vline(xintercept = 0.0, linetype = "dotted") +
  scale_color_discrete("") +
  labs(title = "Fake Data Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "m", y = "") +
  # facet_wrap(vars(.variable)) +
  facet_grid(vars(stat), vars(name)) +
  NULL
```

## Simulation-Based Calibration

With SBC we use the prior to generate a number of datasets that we use the same model and prior to fit. We then calculate the number of samples that are less than the true value (from the prior-only run generating the data). If the model and algorithm are working correctly, the distribution of the number draws that are less than the true parameter values should approach a uniform distribution.

The code for this simulation is located in `sbc.R` (and `sbc.sh` to run on SLURM).

```{r}
#| label: sbc

sbc_draws <- 1000
sbc_bw <- 10

ic_labels <- c(ic = "IC Adjusted", ic_ignore = "IC Ignored", no_ic = "No IC")

sbc_data <- bind_rows( 
  no_ic = read_rds(file.path(tmp_dir, "data", "sbc4.rds")),
  ic = read_rds(file.path(tmp_dir, "data", "sbc5_ic.rds")),
  ic_ignore = read_rds(file.path(tmp_dir, "data", "sbc3_ic_ignore.rds")),
  .id = "sim_type") |> 
  select(sim_type, starts_with("r_"))

sbc_data |> 
  pivot_longer(!c(sim_type), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  filter(!str_detect(param, "pfs$")) |>
  filter(str_detect(param, "lambda_gp_")) |> 
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = sbc_bw) +
  geom_rect(aes(xmin = -10, xmax = sbc_draws, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = count(sbc_data, sim_type) |> mutate(map_dfr(n, \(n) set_names(qbinom(c(0.01, 0.5, 0.99), n, sbc_bw / sbc_draws), c("lb", "med", "ub"))))) +
  facet_grid(rows = vars(sim_type), cols = vars(param), labeller = labeller(sim_type = ic_labels)) +
  labs(y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  # labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  coord_cartesian(ylim = c(0, 20))

sbc_data |> 
  pivot_longer(!c(sim_type), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  filter(str_detect(param, "pfs$")) |>
  mutate(
    stat = str_extract(param, "expected|median"),
    tumor_state = str_extract(param, "one_tumor|base")
  ) |> 
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = sbc_bw) +
  geom_rect(aes(xmin = -10, xmax = sbc_draws, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = count(sbc_data, sim_type) |> mutate(map_dfr(n, \(n) set_names(qbinom(c(0.01, 0.5, 0.99), n, sbc_bw / sbc_draws), c("lb", "med", "ub"))))) +
  facet_grid(rows = vars(sim_type), cols = vars(stat, tumor_state), labeller = labeller(sim_type = ic_labels)) +
  labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  coord_cartesian(ylim = c(0, 30))

# sbc_data |> 
#   select(starts_with("r_")) |> 
#   pivot_longer(everything(), names_to = "param", names_prefix = "r_", values_to = "r") |> 
#   ggplot() +
#   stat_ecdf(aes(r)) +
#   facet_wrap(vars(param)) +
#   coord_cartesian(xlim = c(0, 1000))
```

It is clear that ignoring interval censoring degrades the model's ability to recover the prior parameter values; simulations show a lopsided distribution that significantly deviates from the uniform distribution.

## Fit Data

### No Tumors Model

The hazard function does not use any covariates:

$$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t
$$

```{r}
#| label: no-tumor-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors
#| output: false

no_tumor_treated_pfs_stan_data <- treated_pfs_stan_data |> 
  list_assign(tumor_hazard_type = 0, add_trial_level = TRUE)

prior_no_tumor_pfs_future <- future::future(
  no_tumor_treated_pfs_stan_data |> 
    list_modify(fit_data = FALSE) |> 
    pfs_model$sample(
      parallel_chains = 4,
      iter_warmup = 300, iter_sampling = 300, refresh = 0,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_no_tumor_pfs" 
    ),
  seed = TRUE
)

no_tumor_pfs_future <- future::future(
  pfs_model$sample(
    no_tumor_treated_pfs_stan_data, 
    parallel_chains = 4,
    iter_warmup = 300, iter_sampling = 300, refresh = 0,
    output_dir = file.path(tmp_dir, "fit"), 
    output_basename = "no_tumor_pfs" 
  ),
  seed = TRUE
)

ic_ignored_no_tumor_pfs_future <- future::future(
  no_tumor_treated_pfs_stan_data |>
    list_modify(ignore_interval_censoring = TRUE) |>
    pfs_model$sample(
      parallel_chains = 4,
      iter_warmup = 300, iter_sampling = 300, refresh = 0,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "ic_ignored_no_tumor_pfs" 
    ),
  seed = TRUE
)

prior_no_tumor_pfs_res <- future::value(prior_no_tumor_pfs_future)
no_tumor_pfs_res <- future::value(no_tumor_pfs_future)
```

```{r}
#| label: fig-no-tumor-kaplan-meier
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res, control = NULL) |> # control_no_tumor_pfs_res) |> 
  compact() |> 
  plot_km(treated_pfs_analysis_data) +
  geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) 
```

```{r}
#| label: fig-no-tumor-pfs-hist-post
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(Posterior = no_tumor_pfs_res) |> #, control = control_no_tumor_pfs_res) |> 
  compact() |> 
  plot_pfs_hist_posterior(no_tumor_treated_pfs_stan_data) +
  stat_bin(aes(pfs, color = "Observed"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data) +
  theme(legend.position = "bottom") +
  NULL
```

```{r}
#| label: fig-no-tumor-mean-median
#| fig-cap: Posterior PFS mean and median.
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res) |> #, control = control_no_tumor_pfs_res) |> 
  compact() |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = no_tumor_treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "arm") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = trial), position = "dodge", .width = 0.8) +
  scale_color_discrete("Trial") +
  labs(subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```

```{r}
#| label: fig-no-tumor-base-hazard
#| cache: true
#| dependson: no-tumor-pfs-fit
#| fig-cap: Baseline hazard for model with no tumor effects.

lst(treated = no_tumor_pfs_res, prior = prior_no_tumor_pfs_res) |> #  control = control_no_tumor_pfs_res,  
  plot_base_hazard(treated_pfs_analysis_data, ci_width = c(0.8, 0.95))
```

```{r}
#| label: fig-no-tumor-trial-lambda-variation
#| fig-cap: Baseline hazard variation by trial.
#| cache: true
#| dependson: no-tumor-pfs-fit

no_tumor_pfs_res |>  
  spread_rvars(log_lambda[t], log_trial_lambda[trial, t]) |>
  mutate(trial_lambda_effect = exp(log_trial_lambda - log_lambda)) |> 
  select(trial, t, trial_lambda_effect) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = trial_lambda_effect), fill = "black", alpha = 0.125, .width = c(0.8, 0.95)) +
  geom_line(aes(y = trial_lambda_effect, group = .draw), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 10)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "darkred") +
  labs(x = "m", y = latex2exp::TeX(r"{$\exp[\log(\lambda_{ts}) - \log(\lambda_t)]$}")) +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(0.0, 3)) +
  NULL
```

#### Looking at Interval Censoring

Running the model with and without adjustment for interval censoring.

```{r}
#| label: fit-ignore-interval-censoring
#| cache: true
#| dependson: no-tumor-pfs-fit
#| output: false

ic_ignored_no_tumor_pfs_res <- future::value(ic_ignored_no_tumor_pfs_future)
```

```{r}
#| label: fig-no-tumor-kaplan-meier-ic-compare 
#| cache: true
#| dependson: fit-ignore-interval-censoring

lst(ic = no_tumor_pfs_res, ic_ignored = ic_ignored_no_tumor_pfs_res) |> 
  plot_km(treated_pfs_analysis_data) +
  scale_fill_viridis_d(aesthetics = c("fill", "color")) 
```

```{r}
#| label: fig-no-tumor-mean-median-ic-compare 
#| cache: true
#| dependson: fit-ignore-interval-censoring

lst(ic = no_tumor_pfs_res, ic_ignored = ic_ignored_no_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = no_tumor_treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "model") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = trial), position = "dodge", .width = 0.8) +
  scale_color_discrete("Trial") +
  facet_wrap(vars(model), ncol = 1) +
  labs(title = "Posterior PFS mean and median", subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```

While SBC shows that ignoring interval censoring should be a big problem, it does not appear to bias estimates much using our data.

### Tumor Count Model

The hazard function is time invariant:

$$
\begin{align}
  \lambda(t, \bf{x}; \boldsymbol{\theta}) &= \lambda_t \cdot \exp\left[\sum_m \beta_0\right]  
  \end{align}
$$

```{r}
#| label: tumor-count-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors 
#| output: false

tumor_count_treated_pfs_stan_data <- treated_pfs_stan_data |> 
  list_assign(tumor_hazard_type = 4, add_trial_level = TRUE)

tumor_count_init <- create_pfs_initializer(tumor_count_treated_pfs_stan_data) 

prior_tumor_count_pfs_future <- future::future(
  tumor_count_treated_pfs_stan_data |> 
    list_assign(fit_data = FALSE, add_trial_level = TRUE) |> 
    pfs_model$sample(
      iter_warmup = 300, iter_sampling = 300, 
      init = tumor_count_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_tumor_count_pfs" 
    ),
  seed = TRUE
)

tumor_count_pfs_future <- future::future(
    pfs_model$sample(
      tumor_count_treated_pfs_stan_data,
      iter_warmup = 300, iter_sampling = 300, 
      init = tumor_count_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "tumor_count_pfs" 
    ),
  seed = TRUE
)

prior_tumor_count_pfs_res <- future::value(prior_tumor_count_pfs_future)
tumor_count_pfs_res <- future::value(tumor_count_pfs_future)
```

### First Two Tumor Sizes Only Model

The hazard function is time invariant:

$$
\begin{align}
  \lambda(t, \bf{x}; \boldsymbol{\theta}) &= \lambda_t \cdot \underbrace{\exp\left[\sum_m \beta_0 + \beta_1 \bar{x}_{m0} + \beta_2 \bar{x}_{m1} \right]}_{\textrm{Hazard ratio, }\kappa} \\
  \end{align}
$$ where $\bar{N}$ is the average number of tumors per patient and, for every tumor $m$, $\bar{x}_{m0}$ is the standardized last screening assessment and $\bar{x}_{m1}$ is the standardized first assessment after treatment begins.[^2] We subtract $\beta_0 \bar{N}$

[^2]: $\bar{x}_{mt} = \frac{x_{mt} - \bar{x}}{s(x)}$

```{r}
#| label: two-tumor-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors 
#| output: false

two_tumor_treated_pfs_stan_data <- treated_pfs_stan_data |> 
  list_assign(
    tumor_hazard_type = 1, 
    add_trial_level = TRUE, 
    add_tumor_location_level = FALSE,
  )

two_tumor_init <- create_pfs_initializer(two_tumor_treated_pfs_stan_data) 

prior_two_tumor_pfs_future <- future::future(
  two_tumor_treated_pfs_stan_data |> 
    list_assign(fit_data = FALSE) |> 
    pfs_model$sample(
      iter_warmup = 300, iter_sampling = 300, 
      init = two_tumor_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_two_tumor_pfs" 
    ),
  seed = TRUE
)

two_tumor_pfs_future <- future::future(
    pfs_model$sample(
      two_tumor_treated_pfs_stan_data,
      iter_warmup = 300, iter_sampling = 300, 
      init = two_tumor_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "two_tumor_pfs" 
    ),
  seed = TRUE
)

prior_two_tumor_pfs_res <- future::value(prior_two_tumor_pfs_future)
two_tumor_pfs_res <- future::value(two_tumor_pfs_future)
```

```{r}
#| label: fig-two-tumor-kaplan-meier
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res) |> 
  plot_km(treated_pfs_analysis_data) +
  geom_rug(aes(x = pfs, color = right_censored), 
           length = unit(0.05, "npc"), alpha = 0.5, 
           data = treated_pfs_analysis_data) +
  geom_step(aes(x = t, y = s, linetype = "Observed Kaplan-Meier"), 
            data = treated_obs_km |> filter(fct_match(btype, "lb"))) +
  scale_color_discrete("Observed PFS", labels = c("FALSE" = "Uncensored", "TRUE" = "Censored"), type = AZ_palette[-1]) +
  scale_linetype_manual("", values = "dotted") +
  guides(fill = "none") +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-two-tumor-pfs-hist-post
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(Posterior = two_tumor_pfs_res) |> #, control = control_no_tumor_pfs_res) |> 
  compact() |> 
  plot_pfs_hist_posterior(two_tumor_treated_pfs_stan_data) +
  stat_bin(aes(pfs, color = "Observed (All)"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data) +
  stat_bin(aes(pfs, color = "Observed (Uncensored Only)"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data |> filter(!right_censored)) +
  scale_color_discrete("", type = AZ_palette[-1]) +
  labs(y = "") +
  theme(legend.position = "bottom") +
  NULL
```

```{r}
#| label: fig-two-tumor-mean-median
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = two_tumor_treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "model") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = trial), position = "dodge", .width = 0.8) +
  scale_color_discrete("Trial") +
  facet_wrap(vars(model), ncol = 1) +
  labs(title = "Posterior PFS mean and median", subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```

```{r}
#| label: fig-two-tumor-base-hazard
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res, prior = prior_two_tumor_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  plot_base_hazard(treated_pfs_analysis_data) +
  labs(title = "") +
  guides(fill = "none")
```

```{r}
#| label: two-tumor-coef
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(prior = prior_two_tumor_pfs_res, post = two_tumor_pfs_res) |> 
  map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |> 
  map_dfr(
    \(f) gather_rvars(f, tumor_stim_pop_intercept, tumor_stim_pop_coef[t], tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, t]), 
    .id = "fit"
  ) |>
  filter(are_na(t) | t <= 2) |> 
  mutate(
    .variable = if_else(is.na(trial), .variable, str_glue("{.variable}[{trial}]")),
    .variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]")),
  ) |> 
  ggplot() +
  stat_halfeye(aes(xdist = .value, y = .variable, fill = fit, color = fit), slab_alpha = 0.5, fatten_point = 0.5, linewidth = 1, position = "dodge", .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  scale_color_discrete("", labels = c(prior = "Prior", post = "Posterior"), aesthetics = c("color", "fill")) +
  labs(x = "", y = "") +
  NULL
```

```{r}
#| label: fig-in-sample-survival-comparison
#| fig-height: 10
#| fig-width: 6
#| cache: true
#| dependson: two-tumor-pfs-fit

pfs_range_data <- treated_pfs_analysis_data |> 
  select(usubjid, trial, pfs, interval_censored) |>
  mutate(usubjid = fct_reorder(usubjid, pfs)) 

pfs_range_data |> 
  ggplot(aes(y = usubjid)) +
  stat_lineribbon(
    aes(xdist = rep_pfs, fill = "Retrodictive Credible Interval"),
    alpha = 0.125, linewidth = 0, 
    .width = c(0.5, 0.8),
    data = \(obs_data) two_tumor_pfs_res |> 
      spread_rvars(rep_pfs[patient_index]) |> 
      bind_cols(select(pfs_range_data, trial, usubjid)) |> 
      semi_join(obs_data, by = "usubjid")
  ) +
  geom_errorbar(aes(xmin = pfs, xmax = pfs + interval_censored, color = "Observed Interval Censored Range")) +
  scale_color_discrete("PFS", aesthetics = c("color", "fill")) +
  facet_wrap(vars(trial), nrow = 1, scales = "free_y") +
  labs(x = "Weeks", y = "Patients") +
  theme(axis.text.y = element_blank(), panel.grid.major.y = element_blank(), legend.position = "top")
```

```{r}
#| label: fig-two-tumor-prior-pop-tumor-effect-heatmap
#| cache: true
#| dependson: two-tumor-pfs-fit

prior_two_tumor_pfs_res |> 
  get_tumor_hazard_ratio_summary(pop_tumor_effect[n], early_tumor_pairs, two_tumor_treated_pfs_stan_data) |> 
  mutate(p = p * 100) |> 
  ggplot(aes(x = x0, y = x1, z = q)) +
  stat_summary_hex(binwidth = c(1, 1)) +
  geom_abline(slope = 1, linetype = "dashed") +
  colorspace::scale_fill_continuous_divergingx("Roma", name = "", mid = 1) +
  labs(x = "Pre-screening Tumor Size [cm]", y = "Post-treatment Tumor Size [cm]") +
  facet_wrap(vars(p), labeller = labeller(p = \(per) str_glue("Posterior {per}th percentile"))) +
  coord_fixed(xlim = c(0, 11), ylim = c(0, 11)) +
  NULL

two_tumor_pfs_res |> 
  get_tumor_hazard_ratio_summary(pop_tumor_effect[n], early_tumor_pairs, two_tumor_treated_pfs_stan_data) |> 
  mutate(p = p * 100) |> 
  filter(p == 50) |>
  ggplot(aes(x = x0, y = x1, z = q)) +
  stat_summary_hex(binwidth = c(1, 1)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_rug(sides = "tr", alpha = 0.125) +
  stat_summary_hex(aes(label = sprintf("%.2f", after_stat(value))), geom = "text", size = 3, binwidth = c(1, 1), color = "white") +
  scale_fill_gradient2("", midpoint = 1, low = AZ_darkpurple, mid = AZ_turquoise, high = AZ_gold) + 
  labs(x = "Pre-screening Tumor Size [cm]", y = "Post-treatment Tumor Size [cm]") +
  facet_wrap(
    vars(p), 
    labeller = labeller(
      p = \(per) if_else(per == 50, "Posterior median", str_glue("Posterior {per}th percentile"))
    )
  ) +
  coord_fixed(xlim = c(0, 11), ylim = c(0, 11)) +
  NULL
```

```{r}
#| label: fig-two-tumor-prior-pop-tumor-trial-effect-heatmap
#| cache: true
#| dependson: two-tumor-pfs-fit

diff_var <- FALSE 

prior_two_tumor_pfs_res |> 
  recover_types(select(treated_pfs_analysis_data, trial)) |> 
  get_tumor_hazard_ratio_summary(trial_tumor_effect[trial, n], early_tumor_pairs, two_tumor_treated_pfs_stan_data, diff_var = diff_var) |>  
  mutate(p = fct_rev(str_glue("Posterior {p * 100}th\npercentile"))) |> 
  filter(fct_match(trial, "Breast01")) |>
  ggplot(aes(x = x0, y = x1, z = q)) +
  stat_summary_hex(binwidth = c(1, 1)) +
  scale_fill_continuous_divergingx("Roma", mid = 1, name = "") + 
  labs(x = "Pre-screening Tumor Size", y = "Post-treatment Tumor Size") +
  facet_wrap(vars(p))
  # facet_grid(vars(p), vars(trial))

two_tumor_pfs_res |> 
  recover_types(select(treated_pfs_analysis_data, trial)) |> 
  get_tumor_hazard_ratio_summary(trial_tumor_effect[trial, n], early_tumor_pairs, two_tumor_treated_pfs_stan_data, diff_var = diff_var) |>  
  filter(p == 0.5) |> 
  mutate(p = fct_rev(if_else(p == 0.5, "Posterior median", str_glue("Posterior {p * 100}th\npercentile")))) |> 
  # filter(fct_match(trial, "Breast01")) |>
  ggplot(aes(x = x0, y = x1, z = q)) +
  geom_rug(sides = "tr", alpha = 0.125) +
  stat_summary_hex(binwidth = c(1, 1)) +
  scale_fill_gradient2("", midpoint = 1, low = AZ_darkpurple, mid = AZ_turquoise, high = AZ_gold) + # latex2exp::TeX(r"{\exp(\beta\cdot x)}")) +
  labs(x = "Pre-screening Tumor Size", y = "Post-treatment Tumor Size") +
  # facet_wrap(vars(trial)) +
  facet_grid(vars(p), vars(trial)) +
  coord_fixed() +
  NULL
```

```{r}

lst(prior = prior_two_tumor_pfs_res, posterior = two_tumor_pfs_res) |>
  map_dfr(\(f) gather_rvars(f, tumor_stim_pop_intercept, tumor_stim_trial_intercept[trial]), .id = "fit_type") |> 
  mutate(
    extensive_effect = exp(.value),
    .variable = if_else(are_na(trial), .variable, str_glue("exp({.variable}[{trial}])"))
  ) |> 
  ggplot() +
  stat_halfeye(
    aes(xdist = extensive_effect, y = .variable, fill = fit_type, color = fit_type), 
    position = "dodge", slab_alpha = 0.25, point_size = 0.75, .width = c(0.5, 0.8)
  ) + 
  scale_color_discrete("", labels = str_to_title, aesthetics = c("color", "fill")) +
  coord_cartesian(xlim = c(0, 2))
```

#### Variation in Tumor Predictive Power

In this section I'm trying to get a sense of how much the two first assessments of tumors contribute to the variation in the hazard and the probability of disease progression. Two plots are overlapped in these plots to: First, show the degree of variation in response to the tumor sizes, holding the model parameters fixed at the posterior median; and second, show the degree of variation using the full posterior distribution of the model parameters.

```{r}
#| label: predictor-variation
#| cache: true
#| cache-lazy: false 
#| dependson: two-tumor-pfs-fit

median_coef <- lst(post = two_tumor_pfs_res) |> 
  map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |> 
  map_dfr(\(f) spread_rvars(f, tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, k]), .id = "fit") |> 
  mutate(across(starts_with("tumor_stim"), median)) |> 
  pivot_wider(values_from = tumor_stim_trial_coef, names_prefix = "tumor_stim_trial_coef", names_from = k)

tumor_design_matrix <- with(
  treated_pfs_stan_data, {
    n_screening_t <- pfs_model$functions$calc_n_screening_t(n_patient_tumors, n_measures, t_measure)
    
    pfs_model$functions$standardize_nonzero_tumor_sizes(tumor_size) |>
      pluck(3) |> 
      pfs_model$functions$prepare_early_tumors_design_matrix(n_patient_tumors, n_measures, t_measure, n_screening_t, 2) |> 
      magrittr::extract2(1)
  }
)

tumor_design_matrix <- cbind(1, tumor_design_matrix, tumor_design_matrix[,1] * tumor_design_matrix[,2])

patient_kappa_data <- tibble(kappa = tumor_design_matrix, trial = with(treated_pfs_stan_data, rep(patient_trial, n_patient_tumors))) |>  
  left_join(median_coef, by = "trial") %>% 
  mutate(
    kappa = rowSums(kappa * select(., starts_with("tumor_stim"))),
    patient_index = rep(seq(treated_pfs_stan_data$n_patients), treated_pfs_stan_data$n_patient_tumors)
  ) |> 
  group_by(patient_index) |> 
  summarize(kappa = exp(-first(tumor_stim_trial_intercept) * mean(treated_pfs_stan_data$n_patient_tumors) + sum(kappa)))

patient_kappa_full_dist_data <- two_tumor_pfs_res |> 
  spread_rvars(total_time_invar_tumor_stim[patient_index], total_time_invar_tumor_stim_no_intercept[patient_index]) |> 
  mutate(
    trial = two_tumor_treated_pfs_stan_data$patient_trial, 
    kappa = exp(total_time_invar_tumor_stim),
    kappa_no_intercept = exp(total_time_invar_tumor_stim_no_intercept)
  )

patient_kappa_data |> 
  mutate(var = "median") |> 
  bind_rows(
    patient_kappa_full_dist_data |>
      unnest_rvars() |> 
      mutate(var = "post")
  ) |> 
  filter(fct_match(var, "post")) |> 
  ggplot(aes(kappa)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  stat_histinterval(alpha = 0.5, breaks = ggdist::breaks_fixed(width = 0.1), point_size = 0, .width = c(0.5, 0.8, 0.9)) +
  #geom_histogram(aes(y = after_stat(density)), fill = "black", position = "identity", alpha = 0.25, binwidth = 0.1) +
  # scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(x = latex2exp::TeX(r"{$\kappa$}"), y = "Density") +
  coord_cartesian(xlim = c(0, 5)) +
  NULL
```

```{r}
#| label: dp-prob-variation
#| cache: true
#| dependson: two-tumor-pfs-fit

two_tumor_pfs_res |> 
  spread_rvars(log_lambda[t]) |> 
  mutate(
    lambda = exp(log_lambda),
    median_baseline_dp_prob = 1 - exp(- median(lambda)),
    patient_dp_prob = map(lambda, \(l) 1 - exp(- median(l) * patient_kappa_data$kappa)) %>% 
      { array(unlist(.), dim = c(nrow(patient_kappa_data), length(.))) } |> 
      rvar(),
    
    patient_full_dist_dp_prob = map(lambda, \(l) draws_of(1 - exp(- median(l) * patient_kappa_full_dist_data$kappa))) %>%
      { array(unlist(.), dim = c(prod(dim(first(.))), length(.))) } |> 
      rvar()
  ) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = median_baseline_dp_prob)) +
  # Why didn't I use pivot? We can't have different sized rvars in the same column.
  stat_lineribbon(aes(ydist = patient_dp_prob, fill = "median"), alpha = 0.25, .width = 0.8) +
  stat_lineribbon(aes(ydist = patient_full_dist_dp_prob, fill = "post"), alpha = 0.25, .width = 0.8) +
  scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(title = "Variation in the probability of disease progression", y = latex2exp::TeX(r"{1 - \alpha}"))
  NULL
```

#### Partial Pooling

```{r, fig.height=4, fig.width=8}
#| label: fig-separate-two-tumor-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors 
#| output: false

two_tumor_treated_trial_pfs_stan_data <- treated_trial_pfs_stan_data |> 
  map(\(s) list_assign(s, tumor_hazard_type = 2) |> modify_at(c("patient", "patient_trial"), factor))

two_tumor_trial_pfs_res <- two_tumor_treated_trial_pfs_stan_data |> 
  furrr::future_imap(
    .progress = FALSE, .options = furrr::furrr_options(seed = TRUE),
    \(stan_data, idx) pfs_model$sample(
      stan_data,
      iter_warmup = 300, iter_sampling = 300, 
      init = create_pfs_initializer(stan_data),
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = str_glue("two_tumor_trial_pfs_{idx}")
    )
  )

bind_rows(
  separated = map_dfr(two_tumor_trial_pfs_res,\(f) f |> spread_rvars(km_est[t]), .id = "trial") |> 
    rename(trial_km_est = km_est),
  partial_pooled = two_tumor_pfs_res |> spread_rvars(trial_km_est[trial, t]),
  .id = "model_type"
) |> 
  plot_km_rvars(group_arm = TRUE, group_var = model_type) +
  scale_color_discrete("", labels = c(partial_pooled = "Partial Pooled", separated = "Fully Separated"), aesthetics = c("color", "fill"), type = AZ_palette) +
  labs(title = "") +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.9, 0.35))
```

```{r, fig.height=2, fig.width=6}
#| label: fig-separate-two-tumor-compare-loo
#| cache: true
#| dependson: fig-separate-two-tumor-pfs-fit

# This is the completely separated models LOO elpd.
two_tumor_trial_loo <- two_tumor_trial_pfs_res |> 
  map(\(f) f$loo(save_psis = TRUE, moment_match = TRUE))

# Here I'm getting LOO elpd for the partially pooled model, separately, to compare with the separated models. 
tt_ll <- two_tumor_pfs_res$draws("log_lik", format = "matrix")
split_two_tumor_loglik <- split(tt_ll, rep(two_tumor_treated_pfs_stan_data$patient_trial, each = nrow(tt_ll))) |> 
  map(\(s) matrix(s, nrow = nrow(tt_ll))) 
split_two_tumor_r_eff <- map(split_two_tumor_loglik, \(ll) relative_eff(exp(ll), rep(1:4, each = nrow(ll) / 4), cores = 4))
split_two_tumor_loo <- map2(split_two_tumor_loglik, split_two_tumor_r_eff, \(ll, re) loo(ll, r_eff = re, cores = 4, moment_match = TRUE))

map2(two_tumor_trial_loo, split_two_tumor_loo, 
     \(sep_loo, partial_pooled_loo) loo_compare(lst("Fully Separated" = sep_loo, "Partially Pooled" = partial_pooled_loo))) |> 
  map_dfr(\(l) as_tibble(l, rownames = "model"), .id = "trial") |> 
  filter(fct_match(model, "Fully Separated")) |> 
  ggplot(aes(y = trial)) +
  geom_col(aes(x = elpd_diff), fill = AZ_platinum, alpha =0.5, position = "dodge") +
  geom_errorbar(aes(xmin = elpd_diff - se_diff, xmax = elpd_diff + se_diff), width = 0.25) +
  labs(x = "Difference in log-probability score", y = "Trial") +
  NULL
```

### Tumor Size Change Model

The hazard function is time invariant:

$$
\begin{align}
  \lambda(t, \bf{x}; \boldsymbol{\theta}) &= \lambda_t \cdot \exp\left[\beta_0(N_i - \bar{N}) + \sum_m \beta_1 \frac{x_{m1} - x_{m0}}{x_{m0}}\right] \\ 
  &= \lambda_t \cdot \underbrace{\exp\left[-\beta_0 \bar{N} + \sum_m \beta_1 \frac{x_{m1} - x_{m0}}{x_{m0}}\right]}_{\kappa} 
\end{align}
$$

```{r}
#| label: tumor-change-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors 
#| output: false

# tumor_change_init <- function(chain_id) { 
#   lst(tumor_stim_coef = truncnorm::rtruncnorm(1, 0, mean = 0, sd = 0.1))
# }

tumor_change_treated_pfs_stan_data <- treated_pfs_stan_data |> 
    list_modify(tumor_hazard_type = 3, add_trial_level = TRUE)

tumor_change_init <- create_pfs_initializer(tumor_change_treated_pfs_stan_data)

prior_tumor_change_pfs_future <- future::future(
  tumor_change_treated_pfs_stan_data |> 
    list_modify(fit_data = FALSE) |> 
    pfs_model$sample(
      iter_warmup = 300, iter_sampling = 300, 
      init = tumor_change_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_tumor_change_pfs" 
    ),
  seed = TRUE
)

tumor_change_pfs_future <- future::future(
    pfs_model$sample(
      tumor_change_treated_pfs_stan_data,
      iter_warmup = 300, iter_sampling = 300, 
      init = tumor_change_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "tumor_change_pfs" 
    ),
  seed = TRUE
)

prior_tumor_change_pfs_res <- future::value(prior_tumor_change_pfs_future)
tumor_change_pfs_res <- future::value(tumor_change_pfs_future)
```

```{r}
#| label: tumor-change-kaplan-meier
#| cache: true
#| dependson: tumor-change-pfs-fit

lst(treated = tumor_change_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  plot_km(treated_pfs_analysis_data) +
  geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) 
```

### Joint Model

TODO

### Compare Models

```{r}
#| label: compare-base-hazard
#| cache: true
#| dependson: no-tumor-pfs-fit, two-tumor-pfs-fit

lst(two_tumor = two_tumor_pfs_res, no_tumor = no_tumor_pfs_res) |>  #, prior = prior_two_tumor_pfs_res) |> 
  map_dfr(
    \(f) spread_rvars(f, log_lambda[t]) |> mutate(lambda = exp(log_lambda)) |> median_qi(lambda, .width = 0.8), 
    .id = "model"
  ) |> 
  ggplot(aes(t)) +
  # stat_lineribbon(aes(ydist = lambda, color = model, fill = model), alpha = 0.25, step = TRUE, .width = 0.8) +
  geom_lineribbon(aes(y = lambda, ymin = .lower, ymax = .upper, color = model, fill = model), alpha = 0.25, step = TRUE) +
  geom_rug(aes(week), alpha = 0.125, 
           data = pfs_analysis_data |> 
             transmute(
               trtp, 
               arm = fct_collapse(trtp, treated = "T-DXd", control = c("LAPACAPE", "TRASCAPE")) |> fct_expand("prior"),
               patient_visits = map(patient_tumors, \(tu) unnest(tu, tumor_history) |> distinct(week))) |> 
             unnest(patient_visits)) +
  labs(title = "Baseline hazard", y = latex2exp::TeX(r"{$\lambda_t$}"), caption = "Ribbons shown are for the 80% CI.\nRugs below x-axis show the distribution of assessment weeks.") +
  scale_color_viridis_d(aesthetics = c("fill", "color"))
```

#### Cross Validation

Here we are calculating the LOO-PSIS cross validation approximations for the no tumors and time invariant two tumor assessments model.

```{r}
#| label: pfs-loo
#| cache: true
#| dependson: no-tumor-pfs-fit, two-tumor-pfs-fit, tumor-change-pfs-fit

no_tumor_loo <- no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
ic_ignored_no_tumor_loo <- ic_ignored_no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
tumor_count_loo <- tumor_count_pfs_res$loo(save_psis = TRUE, cores = 12)
two_tumor_loo <- two_tumor_pfs_res$loo(save_psis = TRUE, cores = 12, moment_match = TRUE) 
# tumor_change_loo <- tumor_change_pfs_res$loo(save_psis = TRUE, cores = 12)

loo_compare(lst(no_tumor_loo, tumor_count_loo, two_tumor_loo)) #, tumor_change_loo))
loo_compare(lst(no_tumor_loo, ic_ignored_no_tumor_loo))
```

```{r}
treated_pfs_analysis_data |> 
  select(trial, pfs, right_censored) |> 
  mutate(pointwise_diff = two_tumor_loo$pointwise[, "elpd_loo"] - no_tumor_loo$pointwise[, "elpd_loo"]) |> 
  ggplot(aes(pfs, pointwise_diff)) +
  geom_point(aes(color = right_censored), alpha = 0.25) +
  stat_smooth(aes(color = right_censored), se = FALSE) +
  scale_color_discrete("", labels = c("TRUE" = "Right censored", "FALSE" = "Uncensored"), type = AZ_palette) +
  # coord_cartesian(ylim = c(-1, 1)) +
  labs(x = "Actual PFS [Weeks]", y = "Difference in Pointwise Expected Log Likelihood") +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.75, 0.9))
```

```{r}
#| label: loss-functions

#linear_loss <- function(m, c, a, ac, k0 = 1, k1 = 1) if_else(c & (ac | a >= m), 0, if_else(m >= a, k0, k1) * abs(m - a))
linear_loss <- function(m, c, a, ac, k0 = 1, k1 = 1) if_else(c & (a >= m), 0, if_else(m >= a, k0, k1) * abs(m - a))
squared_error_loss <- function(m, c, a, ac) if_else(c & (ac | a >= m), 0, (m - a)^2)
```

Linear loss function, where $(m, c)$ are the PFS and right censoring status observed in the data and $(\tilde{m}, \tilde{c})$ are their inferred values.

$$
L(y, c, \tilde{y}, \tilde{c}) = \begin{cases}
0 & \textrm{if } c = 0 \land \tilde{y} \geq y \\
K_0 (y - \tilde{y}) & \textrm{if } y \geq \tilde{y} \\
K_1 (\tilde{y} - y) & \textrm{if } y < \tilde{y} 
\end{cases}
$$

```{r}
#| label: fig-linear-loss-function

tibble(delta = seq(-10, 10), sym_loss = abs(delta), asym_loss = if_else(delta < 0, 3, 1) * abs(delta)) |> 
  pivot_longer(sym_loss:asym_loss) |> 
  ggplot(aes(delta, value)) +
  geom_line(aes(y = value, color = name), linewidth = 2, alpha = 0.25) +
  scale_color_discrete("Loss Type", labels = c("sym_loss" = "Symmetric", "asym_loss" = "Asymmetric"), type = AZ_palette) +
  # labs(x = latex2exp::TeX(r"{$y - \tilde{y}$}"), y = "Loss")
  labs(x = "Prediction Error", y = "Loss") +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.7, 0.85)) +
  coord_fixed() +
  NULL
```

For each patient, $i$, we calculate the **out-of-sample** loss as $$
\int L(y_i, c_i, \tilde{y}_i, \tilde{c}_i)\cdot p(\tilde{y}_i,\tilde{c}_i \mid y_{-i}, c_{-i})\cdot p(y_i)\,\textrm{d}\tilde{y}_i\textrm{d}\tilde{c}_i\textrm{d}y_i.
$$ Where the observed PFS is modeled as uniform value within the censored interval, $Y_i \sim \mathtt{Uniform}(\underline{y}_i, \overline{y}_i)$.

```{r}
#| label: loo-predict-other-utility 
#| code-fold: false 
#| cache: true
#| dependson: pfs-loo

asym_linear_loss <- partial(linear_loss, k1 = 3)

no_tumor_asym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, asym_linear_loss)
# ic_ignored_no_tumor_asym_loss <- estimate_oos_loss(ic_ignored_no_tumor_pfs_res, ic_ignored_no_tumor_loo, no_tumor_treated_pfs_stan_data, asym_linear_loss)
two_tumor_asym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, asym_linear_loss)
tumor_count_asym_loss <- estimate_oos_loss(tumor_count_pfs_res, tumor_count_loo, tumor_count_treated_pfs_stan_data, asym_linear_loss)
no_tumor_sym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, linear_loss)
# ic_ignored_no_tumor_sym_loss <- estimate_oos_loss(ic_ignored_no_tumor_pfs_res, ic_ignored_no_tumor_loo, no_tumor_treated_pfs_stan_data, linear_loss)
two_tumor_sym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, linear_loss)
tumor_count_sym_loss <- estimate_oos_loss(tumor_count_pfs_res, tumor_count_loo, tumor_count_treated_pfs_stan_data, linear_loss)
no_tumor_sqerr_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, squared_error_loss)
two_tumor_sqerr_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, squared_error_loss)

loss_data <- bind_rows(
  lst(no_tumor_sym_loss, 
      # ic_ignored_no_tumor_asym_loss, 
      two_tumor_sym_loss, 
      no_tumor_asym_loss, 
      # ic_ignored_no_tumor_sym_loss, 
      two_tumor_asym_loss, 
      no_tumor_sqerr_loss, 
      two_tumor_sqerr_loss),
  .id = "model"
) |> 
  mutate(loss_type = str_extract(model, "(a?sym)|sqerr"), model = str_extract(model, "(ic_ignored_)?(no|two)_tumor"))
```

```{r}
#| label: fig-rmse-loss
#| fig-height: 8

loss_data |> 
  filter(!fct_match(loss_type, "sqerr")) |> #, !fct_match(model, "ic_ignored_no_tumor")) |> 
  plot_loss(binwidth = 6)

loss_data |> 
  filter(fct_match(loss_type, "sqerr")) |> #, !fct_match(model, "ic_ignored_no_tumor")) |> 
  plot_loss(binwidth = 100)
```

```{r, fig.width=10, fig.height=4}
#| label: fig-linear-loss

loss_data |> 
  filter(fct_match(model, c("no_tumor", "two_tumor")), fct_match(loss_type, c("sym", "asym"))) |> 
  group_by(trial, model, loss_type) |> 
  summarize(mean_loss = mean(value), .groups = "drop") |> 
  ggplot(aes(mean_loss)) +
  geom_col(aes(y = trial, fill = model), position = "dodge", alpha = 0.5) +
  scale_fill_discrete("Model", labels = c(no_tumor = "Baseline", two_tumor = "Linear"), type = AZ_palette) +
  labs(x = "Mean Loss", y = "") +
  facet_wrap(vars(loss_type), ncol = 1, labeller = labeller(loss_type = c(sym = "Symmetric Loss", asym = "Asymmetric Loss")))
```

```{r}
loss_data |> 
  group_by(loss_type, model) |> 
  summarize(expected_loss = mean(value), .groups = "drop") |> 
  kableExtra::kbl() |> 
  kableExtra::kable_styling()
```

```{r}
loss_data |> 
  filter(fct_match(model, c("no_tumor", "two_tumor")), !fct_match(loss_type, "sqerr")) %>%
  left_join(filter(., fct_match(model, "two_tumor")), filter(., fct_match(model, "no_tumor")), by = c("loss_type", "trial", "id")) |> 
  mutate(diff = value.x - value.y) |> 
  ggplot(aes(diff)) +
  stat_histinterval(
    point_interval = mean_qi, .width = c(0.5, 0.8), alpha = 0.5, breaks = ggdist::breaks_fixed(width = 6)
  ) +
  facet_grid(vars(trial), vars(loss_type), scales = "free", margins = "trial") +
  coord_cartesian(xlim = c(-20, 20)) +
  NULL
```

```{r}
#| label: fig-patient-hazard-ratios
#| fig-cap: Hazard ratios for the 100 patients with the greatest increase. Line intervals identify the 50% and 80% credible intervals.
#| fig-height: 4 

patient_kappa_full_dist_data |> 
  arrange(median(kappa_no_intercept)) |> 
  # head(200) |> 
  group_by(trial) |> 
  mutate(id = seq(n())) |> 
  ungroup() |>
  ggplot() +
  stat_pointinterval(aes(xdist = kappa_no_intercept, y = id), point_size = 0.25, alpha = 0.75, linewidth = 0.05, color = AZ_gold, .width = c(0.8)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_text(aes(x = x, y = n - 10, label = str_glue("N = {n}")), data = \(d) count(d, trial) |> mutate(x = 0.5)) +
  scale_x_continuous(latex2exp::TeX(r"{Hazard Ratio ($\kappa$)}"), breaks = seq(0, 100, 2)) +
  labs(y = "Patients", caption = "Ratio hazard excluding the number of tumors effect.") +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), axis.text.y = element_blank()) +
  facet_grid(vars(trial), space = "free_y", scales = "free_y") +
  coord_trans(x = "log") +
  theme(strip.text.y = element_text(angle = 0)) +
  NULL
```

<!-- Ignore everything below -->

```{r}
#| include: false
#| eval: false

loss_diff_data |> 
  slice(1:10) |> 
  select(starts_with("centered")) |> 
  mutate(id = seq(n())) |> 
  pivot_longer(starts_with("centered")) |> 
  ggplot() +
  stat_pointinterval(aes(id, ydist = value, color = name), position = "dodge") +
  #stat_lineribbon(aes(id, ydist = value, fill = name, color = name), alpha = 0.25, .width = 0.8) +
  theme(legend.position = "bottom")

```
