---
title: "DCO Comparisons"
subtitle: "Comparing results across data cutoff dates"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page compares clinical endpoint estimates across different data cutoff (DCO) dates to track how predictions evolve as more data becomes available.

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints](../analysis/clinical-endpoints.qmd) for endpoint definitions
- See [Survival Analysis](../analysis/survival-analysis.qmd) for latest KM curves
:::

## Overall Response Rate by DCO {#sec-orr-dco}

Comparing ORR estimates across data cutoffs shows how response rate predictions stabilize with additional follow-up.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-orr-dco-sample
#| fig-cap: Sample-based overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-height: 10

plot_orr_by_dco("sample")
```

### Superpopulation

```{r}
#| label: fig-orr-dco-spop
#| fig-cap: Superpopulation overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-height: 10

plot_orr_by_dco("spop")
```

:::

## Median PFS by DCO {#sec-median-pfs-dco}

Tracking median PFS estimates across data cutoffs, comparing model forecasts with observed Kaplan-Meier estimates.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-median-pfs-dco-sample
#| fig-cap: Sample-based median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10
#| fig-height: 6

plot_median_pfs_by_dco("sample")
```

### Superpopulation

```{r}
#| label: fig-median-pfs-dco-spop
#| fig-cap: Superpopulation median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10
#| fig-height: 6

plot_median_pfs_by_dco("spop")
```

:::

::: {.callout-note}
## Interpreting the Plot
- **Diamond markers**: Model forecast with 90% credible interval
- **Circle markers**: Observed KM median with confidence interval
- **Arrows**: Indicate median PFS not yet reached (right-censored)
:::

## PFS-n by DCO {#sec-pfs-n-dco}

Comparing PFS at fixed timepoints across data cutoffs.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-pfs-n-dco-sample
#| fig-cap: Sample-based progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10

plot_pfs_n_by_dco("sample")
```

### Superpopulation

```{r}
#| label: fig-pfs-n-dco-spop
#| fig-cap: Superpopulation progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10

plot_pfs_n_by_dco("spop")
```

:::

## Kaplan-Meier Curves by DCO {#sec-km-dco}

Comparing Kaplan-Meier survival curves across data cutoffs.

### By Cohort

Kaplan-Meier curves for SCLC showing All patients, First Line patients, and Part E patients across data cutoffs.

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-km-dco-sample
#| fig-cap: Sample-based Kaplan-Meier forecasts for SCLC across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_sclc_km_by_dco(
  estimate_type = "sample",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26)
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_naive_ctdna_jan26)
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = tar_read(all_analysis_data_ctdna_jan26)
  )
)
```

#### Superpopulation

```{r}
#| label: fig-km-dco-spop
#| fig-cap: Superpopulation Kaplan-Meier forecasts for SCLC across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_sclc_km_by_dco(
  estimate_type = "spop",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26)
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_naive_ctdna_jan26)
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = tar_read(all_analysis_data_ctdna_jan26)
  )
)
```

:::

### PDL1-Stratified

Kaplan-Meier curves stratified by PDL1 status (High vs Low) for All patients, First Line patients, and Part E patients across data cutoffs.

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-km-dco-sample-pdl1
#| fig-cap: Sample-based PDL1-stratified Kaplan-Meier forecasts for SCLC across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_km_pdl1_by_dco(
  cond_km_data_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26)
  ),
  obs_km_part_e_list = list(
    apr25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_jan26)
  ),
  estimate_type = "sample"
)
```

#### Superpopulation

```{r}
#| label: fig-km-dco-spop-pdl1
#| fig-cap: Superpopulation PDL1-stratified Kaplan-Meier forecasts for SCLC across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_km_pdl1_by_dco(
  cond_km_data_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26)
  ),
  obs_km_part_e_list = list(
    apr25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_jan26)
  ),
  estimate_type = "spop"
)
```

:::

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](../analysis/clinical-endpoints.qmd) - Latest endpoint estimates
- [Survival Analysis](../analysis/survival-analysis.qmd) - Kaplan-Meier curves
- [Model Validation](../analysis/model-validation.qmd) - Cross-validation results
:::
