---
title: "DCO Comparisons"
subtitle: "Comparing results across data cutoff dates"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page compares clinical endpoint estimates across different data cutoff (DCO) dates to track how predictions evolve as more data becomes available.

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints](../analysis/clinical-endpoints.qmd) for endpoint definitions
- See [Survival Analysis](../analysis/survival-analysis.qmd) for latest KM curves
:::

## Overall Response Rate by DCO {#sec-orr-dco}

Comparing ORR estimates across data cutoffs shows how response rate predictions stabilize with additional follow-up.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-orr-dco-sample
#| fig-cap: Sample-based overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-height: 5

plot_orr_by_dco("sample")
```

### Superpopulation

```{r}
#| label: fig-orr-dco-spop
#| fig-cap: Superpopulation overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-height: 5

plot_orr_by_dco("spop")
```

:::

## Median PFS by DCO {#sec-median-pfs-dco}

Tracking median PFS estimates across data cutoffs, comparing model forecasts with observed Kaplan-Meier estimates.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-median-pfs-dco-sample
#| fig-cap: Sample-based median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 6
#| fig-height: 6

plot_median_pfs_by_dco("sample")
```

### Superpopulation

```{r}
#| label: fig-median-pfs-dco-spop
#| fig-cap: Superpopulation median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 6
#| fig-height: 6

plot_median_pfs_by_dco("spop")
```

:::

::: {.callout-note}
## Interpreting the Plot
- **Diamond markers**: Model forecast with 90% credible interval
- **Circle markers**: Observed KM median with confidence interval
- **Arrows**: Indicate median PFS not yet reached (right-censored)
:::

## PFS-n by DCO {#sec-pfs-n-dco}

Comparing PFS at fixed timepoints across data cutoffs.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-pfs-n-dco-sample
#| fig-cap: Sample-based progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10

plot_pfs_n_by_dco("sample")
```

### Superpopulation

```{r}
#| label: fig-pfs-n-dco-spop
#| fig-cap: Superpopulation progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10

plot_pfs_n_by_dco("spop")
```

:::

## Kaplan-Meier Curves by DCO {#sec-km-dco}

Comparing Kaplan-Meier survival curves across data cutoffs.

### Overall SCLC

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-km-dco-sample
#| fig-cap: Sample-based Kaplan-Meier forecasts for SCLC trial across data cutoffs.
#| column: body-outset
#| fig-width: 7
#| fig-height: 3

plot_km(
  bind_rows(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25) |> filter(fct_match(trial, "sclc")),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25) |> filter(fct_match(trial, "sclc")),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26) |> filter(fct_match(trial, "sclc")),
    .id = "dco"
  ),
  km_est_var = sample_km_est,
  obs_km_data = bind_rows(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    .id = "dco"
  ) |>
    select(!quantiles) |>
    unnest(km_data),
  analysis_data = bind_rows(
    apr25 = tar_read(all_analysis_data_ctdna_apr25) |> filter(fct_match(trial, "sclc")),
    aug25 = tar_read(all_analysis_data_ctdna_aug25) |> filter(fct_match(trial, "sclc")),
    jan26 = tar_read(all_analysis_data_ctdna_jan26) |> filter(fct_match(trial, "sclc")),
    .id = "dco"
  ),
  facet_col = dco
)
```

#### Superpopulation

```{r}
#| label: fig-km-dco-spop
#| fig-cap: Superpopulation Kaplan-Meier forecasts for SCLC trial across data cutoffs.
#| column: body-outset
#| fig-width: 7
#| fig-height: 3

plot_km(
  bind_rows(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25) |> filter(fct_match(trial, "sclc")),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25) |> filter(fct_match(trial, "sclc")),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26) |> filter(fct_match(trial, "sclc")),
    .id = "dco"
  ),
  km_est_var = spop_km_est,
  obs_km_data = bind_rows(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    .id = "dco"
  ) |>
    select(!quantiles) |>
    unnest(km_data),
  analysis_data = bind_rows(
    apr25 = tar_read(all_analysis_data_ctdna_apr25) |> filter(fct_match(trial, "sclc")),
    aug25 = tar_read(all_analysis_data_ctdna_aug25) |> filter(fct_match(trial, "sclc")),
    jan26 = tar_read(all_analysis_data_ctdna_jan26) |> filter(fct_match(trial, "sclc")),
    .id = "dco"
  ),
  facet_col = dco
)
```

:::

### PDL1-Stratified

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-km-dco-sample-pdl1
#| fig-cap: Sample-based PDL1-stratified Kaplan-Meier forecasts for SCLC across data cutoffs.
#| column: body-outset
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km(
  cond_km_data = bind_rows(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26),
    .id = "dco"
  ),
  km_est_var = cond_sample_km_est,
  obs_km_data = bind_rows(
    pdl1 = bind_rows(
      apr25 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
      aug25 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
      jan26 = tar_read(km_trial_pfs_pdl1_ctdna_jan26),
      .id = "dco"
    ) |>
      filter(fct_match(trial, "sclc")) |>
      select(!quantiles) |>
      unnest(km_data),
    pdl1_naive = bind_rows(
      apr25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
      aug25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
      jan26 = tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26),
      .id = "dco"
    ) |>
      select(!quantiles) |>
      unnest(km_data),
    .id = "variable"
  )
)
```

#### Superpopulation

```{r}
#| label: fig-km-dco-spop-pdl1
#| fig-cap: Superpopulation PDL1-stratified Kaplan-Meier forecasts for SCLC across data cutoffs.
#| column: body-outset
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km(
  cond_km_data = bind_rows(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26),
    .id = "dco"
  ),
  km_est_var = cond_spop_km_est,
  obs_km_data = bind_rows(
    pdl1 = bind_rows(
      apr25 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
      aug25 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
      jan26 = tar_read(km_trial_pfs_pdl1_ctdna_jan26),
      .id = "dco"
    ) |>
      filter(fct_match(trial, "sclc")) |>
      select(!quantiles) |>
      unnest(km_data),
    pdl1_naive = bind_rows(
      apr25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
      aug25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
      jan26 = tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26),
      .id = "dco"
    ) |>
      select(!quantiles) |>
      unnest(km_data),
    .id = "variable"
  )
)
```

:::

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](../analysis/clinical-endpoints.qmd) - Latest endpoint estimates
- [Survival Analysis](../analysis/survival-analysis.qmd) - Kaplan-Meier curves
- [Model Validation](../analysis/model-validation.qmd) - Cross-validation results
:::
