---
title: "DCO Comparisons"
subtitle: "Comparing results across data cutoff dates"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")

# Load multi-DCO data for comparisons
orr_list <- list(
  apr25 = tar_read(all_tumor_ssls_orr_rvar_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_orr_rvar_ctdna_aug25),
  jan26 = jan26_orr_data
)
cond_orr_list <- list(
  apr25 = tar_read(all_tumor_ssls_cond_orr_rvar_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_cond_orr_rvar_ctdna_aug25),
  jan26 = jan26_cond_orr_data
)

quant_pfs_list <- list(
  apr25 = tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_aug25),
  jan26 = jan26_quant_pfs_data
)
cond_quant_pfs_list <- list(
  apr25 = tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_aug25),
  jan26 = jan26_cond_quant_pfs_data
)

# Observed KM data for median PFS plot
obs_km_list <- list(
  apr25 = list(
    all = tar_read(km_trial_pfs_ctdna_apr25),
    pdl1_naive = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
    pdl1 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
    naive = tar_read(km_trial_pfs_naive_ctdna_apr25)
  ),
  aug25 = list(
    all = tar_read(km_trial_pfs_ctdna_aug25),
    pdl1_naive = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
    pdl1 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
    naive = tar_read(km_trial_pfs_naive_ctdna_aug25)
  ),
  jan26 = list(
    all = jan26_obs_km_all,
    pdl1_naive = jan26_obs_km_pdl1_naive,
    pdl1 = jan26_obs_km_pdl1,
    naive = jan26_obs_km_first_line
  )
)

pfs_n_list <- list(
  apr25 = tar_read(all_tumor_ssls_forecast_target_pfs_n_rvar_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_forecast_target_pfs_n_rvar_ctdna_aug25),
  jan26 = jan26_pfs_n_data
)
cond_pfs_n_list <- list(
  apr25 = tar_read(all_tumor_ssls_cond_forecast_target_pfs_n_rvar_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_cond_forecast_target_pfs_n_rvar_ctdna_aug25),
  jan26 = jan26_cond_pfs_n_data
)
```

This page compares clinical endpoint estimates across different data cutoff (DCO) dates to track how predictions evolve as more data becomes available.

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints](../analysis/clinical-endpoints.qmd) for endpoint definitions
- See [Survival Analysis](../analysis/survival-analysis.qmd) for latest KM curves
:::

## Overall Response Rate by DCO {#sec-orr-dco}

Comparing ORR estimates across data cutoffs shows how response rate predictions stabilize with additional follow-up.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-orr-dco-sample
#| fig-cap: Sample-based overall response rate (ORR), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-height: 10

plot_orr_by_dco("sample", orr_list, cond_orr_list)
```

### Superpopulation

```{r}
#| label: fig-orr-dco-spop
#| fig-cap: Superpopulation overall response rate (ORR), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-height: 10

plot_orr_by_dco("spop", orr_list, cond_orr_list)
```

:::

## Median PFS by DCO {#sec-median-pfs-dco}

Tracking median PFS estimates across data cutoffs, comparing model forecasts with observed Kaplan-Meier estimates.

```{=html}
<style>
.toggle-switch {
  position: relative;
  display: inline-block;
  width: 50px;
  height: 24px;
  margin-right: 0.75rem;
}
.toggle-switch input {
  opacity: 0;
  width: 0;
  height: 0;
}
.toggle-slider {
  position: absolute;
  cursor: pointer;
  top: 0;
  left: 0;
  right: 0;
  bottom: 0;
  background-color: #ccc;
  transition: .4s;
  border-radius: 24px;
}
.toggle-slider:before {
  position: absolute;
  content: "";
  height: 18px;
  width: 18px;
  left: 3px;
  bottom: 3px;
  background-color: white;
  transition: .4s;
  border-radius: 50%;
}
input:checked + .toggle-slider {
  background-color: var(--az-navy);
}
input:checked + .toggle-slider:before {
  transform: translateX(26px);
}
#median-pfs-sample-with-obs, #median-pfs-spop-with-obs {
  display: none;
}
</style>

<div style="margin-bottom: 1rem;">
  <label style="display: inline-flex; align-items: center; cursor: pointer; font-weight: 500;">
    <span class="toggle-switch">
      <input type="checkbox" id="show-observed-toggle" onchange="toggleObservedKM(this.checked)">
      <span class="toggle-slider"></span>
    </span>
    Show Observed KM Estimates
  </label>
</div>
```

::: {.panel-tabset}

### Sample-Based

::: {#median-pfs-sample-with-obs}
```{r}
#| label: fig-median-pfs-dco-sample-with-obs
#| fig-cap: Sample-based median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("sample", quant_pfs_list, cond_quant_pfs_list, obs_km_list, show_observed = TRUE)
```
:::

::: {#median-pfs-sample-without-obs}
```{r}
#| label: fig-median-pfs-dco-sample-without-obs
#| fig-cap: Sample-based median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs. Model forecasts only.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("sample", quant_pfs_list, cond_quant_pfs_list, obs_km_list, show_observed = FALSE)
```
:::

### Superpopulation

::: {#median-pfs-spop-with-obs}
```{r}
#| label: fig-median-pfs-dco-spop-with-obs
#| fig-cap: Superpopulation median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("spop", quant_pfs_list, cond_quant_pfs_list, obs_km_list, show_observed = TRUE)
```
:::

::: {#median-pfs-spop-without-obs}
```{r}
#| label: fig-median-pfs-dco-spop-without-obs
#| fig-cap: Superpopulation median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs. Model forecasts only.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("spop", quant_pfs_list, cond_quant_pfs_list, obs_km_list, show_observed = FALSE)
```
:::

:::

```{=html}
<script>
function toggleObservedKM(show) {
  document.getElementById('median-pfs-sample-with-obs').style.display = show ? 'block' : 'none';
  document.getElementById('median-pfs-sample-without-obs').style.display = show ? 'none' : 'block';
  document.getElementById('median-pfs-spop-with-obs').style.display = show ? 'block' : 'none';
  document.getElementById('median-pfs-spop-without-obs').style.display = show ? 'none' : 'block';
}
</script>
```

::: {.callout-note}
## Interpreting the Plot
- **Diamond markers**: Model forecast with 90% credible interval
- **Circle markers**: Observed KM median with confidence interval (toggle above)
- **Arrows**: Indicate median PFS not yet reached (right-censored)
:::

## PFS-n by DCO {#sec-pfs-n-dco}

Comparing PFS at fixed timepoints across data cutoffs.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-pfs-n-dco-sample
#| fig-cap: Sample-based progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10

plot_surv_n_by_dco("sample", pfs_n_list, cond_pfs_n_list, pfs_timepoints)
```

### Superpopulation

```{r}
#| label: fig-pfs-n-dco-spop
#| fig-cap: Superpopulation progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10

plot_surv_n_by_dco("spop", pfs_n_list, cond_pfs_n_list, pfs_timepoints)
```

:::

## Kaplan-Meier Curves by DCO {#sec-km-dco}

Comparing Kaplan-Meier survival curves across data cutoffs.

### By Cohort

Kaplan-Meier curves for SCLC-01 showing All patients, First Line patients, and Part E patients across data cutoffs.

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-km-dco-sample
#| fig-cap: Sample-based Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_sclc_km_by_dco(
  estimate_type = "sample",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26)
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_naive_ctdna_jan26)
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = tar_read(all_analysis_data_ctdna_jan26)
  )
)
```

#### Superpopulation

```{r}
#| label: fig-km-dco-spop
#| fig-cap: Superpopulation Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_sclc_km_by_dco(
  estimate_type = "spop",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_km_rvar_ctdna_jan26)
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_naive_ctdna_jan26)
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = tar_read(all_analysis_data_ctdna_jan26)
  )
)
```

:::

### PDL1-Stratified

Kaplan-Meier curves stratified by PDL1 status (High vs Low) for All patients, First Line patients, and Part E patients across data cutoffs.

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-km-dco-sample-pdl1
#| fig-cap: Sample-based PDL1-stratified Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_km_pdl1_by_dco(
  cond_km_data_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26)
  ),
  obs_km_part_e_list = list(
    apr25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_jan26)
  ),
  estimate_type = "sample"
)
```

#### Superpopulation

```{r}
#| label: fig-km-dco-spop-pdl1
#| fig-cap: Superpopulation PDL1-stratified Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_km_pdl1_by_dco(
  cond_km_data_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug25),
    jan26 = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_ctdna_jan26)
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26)
  ),
  obs_km_part_e_list = list(
    apr25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_aug25),
    jan26 = tar_read(km_trial_pfs_part_e_pdl1_ctdna_jan26)
  ),
  estimate_type = "spop"
)
```

:::

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](../analysis/clinical-endpoints.qmd) - Latest endpoint estimates
- [Survival Analysis](../analysis/survival-analysis.qmd) - Kaplan-Meier curves
- [Model Validation](../analysis/model-validation.qmd) - Cross-validation results
:::
