---
title: "DCO Comparisons"
subtitle: "Comparing results across data cutoff dates"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page compares clinical endpoint estimates across different data cutoff (DCO) dates to track how predictions evolve as more data becomes available.

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints](../analysis/clinical-endpoints.qmd) for endpoint definitions
- See [Survival Analysis](../analysis/survival-analysis.qmd#sec-cutoff-km) for cutoff KM curves
:::

## Overall Response Rate by DCO {#sec-orr-dco}

Comparing ORR estimates across data cutoffs shows how response rate predictions stabilize with additional follow-up.

```{r}
#| label: fig-orr-dco
#| fig-cap: Overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.

bind_rows(
  aug = bind_rows(
    tar_read(all_tumor_ssls_orr_rvar_ctdna_aug) |> rename(orr = sample_target_orr),
    tar_read(all_tumor_ssls_cond_orr_rvar_ctdna_aug) |> rename(orr = cond_sample_target_orr),
  ),
  apr = bind_rows(
    tar_read(all_tumor_ssls_orr_rvar_ctdna_apr) |> rename(orr = sample_target_orr),
    tar_read(all_tumor_ssls_cond_orr_rvar_ctdna_apr) |> rename(orr = cond_sample_target_orr),
  ),
  .id = "dco"
) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(variable, c("all", "pdl1_naive")), fct_match(trial, "sclc")) |>
  ggplot() +
  stat_pointinterval(aes(xdist = orr, y = cond_group_name, color = dco, alpha = fit_type), position = "dodge", .width = c(0.9)) +
  scale_color_discrete("DCO", type = AZ_palette, label = str_to_title) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 1), label = str_to_title) +
  facet_grid(vars(variable), vars(trial), scales = "free", space = "free",
             labeller = labeller(trial = str_to_upper, variable = c("all" = "All", "pdl1_naive" = "First Line"))) +
  labs(x = "Overall Response Rate (ORR)", y = "Density") +
  NULL
```

## Median PFS by DCO {#sec-median-pfs-dco}

Tracking median PFS estimates across data cutoffs, comparing model forecasts with observed Kaplan-Meier estimates.

```{r}
#| label: fig-median-pfs-dco
#| fig-cap: Median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 6
#| fig-height: 6

obs_km <- bind_rows(
  aug = bind_rows(
    all = tar_read(km_trial_pfs_ctdna_aug),
    pdl1_naive = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug),
    pdl1 = tar_read(km_trial_pfs_pdl1_ctdna_aug),
    naive = tar_read(km_trial_pfs_naive_ctdna_aug),
    .id = "src"
  ),
  apr = bind_rows(
    all = tar_read(km_trial_pfs_ctdna_apr),
    pdl1_naive = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr),
    pdl1 = tar_read(km_trial_pfs_pdl1_ctdna_apr),
    naive = tar_read(km_trial_pfs_naive_ctdna_apr),
    .id = "src"
  ),
  .id = "dco"
  ) |>
  filter(fct_match(btype, "ub"), !fct_match(src, "naive") | first_line) |>
  unnest(quantiles) |>
  filter(quantile == 0.5) |>
  rename(median_pfs = pfs, median_pfs_lower = pfs_lower, median_pfs_upper = pfs_upper) |>
  mutate(
    variable = case_when(
      is.na(first_line) & fct_match(src, "pdl1_naive") ~ "pdl1_naive",
      fct_match(src, "naive") & first_line ~ "pdl1_naive",
      TRUE ~ "all",
    ),
    cond_group_name = case_when(
      is.na(pdl1_hi) ~ "All",
      pdl1_hi ~ "PDL1 High",
      !pdl1_hi ~ "PDL1 Low"
    )
  ) |>
  select(trial, dco, variable, cond_group_name, median_pfs, median_pfs_lower, median_pfs_upper) |>
  distinct() |>
  mutate(cond_group_name = factor(cond_group_name))

bind_rows(
  aug = bind_rows(
    tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_aug) |> rename(median_pfs = sample_quant_pfs),
    tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_aug) |> rename(median_pfs = cond_sample_quant_pfs)
  ),
  apr = bind_rows(
    tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_apr) |> rename(median_pfs = sample_quant_pfs),
    tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_apr) |> rename(median_pfs = cond_sample_quant_pfs)
  ),
  .id = "dco"
) |>
  filter(quantile == 0.5) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(variable, c("all", "pdl1_naive")), fct_match(trial, "sclc"), fct_match(fit_type, "posterior")) |>
  filter(fct_match(variable, "all") | !fct_match(cond_group_name, "PDL1 High")) |>
  mutate(cond_group_name = factor(cond_group_name)) |>
  ggplot(aes(y = cond_group_name, color = dco)) +
  stat_pointinterval(aes(xdist = median_pfs, color = dco, shape = "model"), point_size = 3, position = position_dodge(width = 0.5), .width = c(0.9)) +
  geom_errorbar(aes(x = median_pfs, xmin = median_pfs_lower, xmax = median_pfs_upper),
    width = 0.25, position = position_dodge(width = 1),
    data = \(d) semi_join(obs_km, d, by = "trial")) +
  geom_segment(aes(x = median_pfs_lower, xend = 100, yend = dodged_cond_group_name),
               position = position_dodge(width = 1),
               arrow = arrow(length = unit(0.1, "inches")),
               data = \(d) semi_join(obs_km, d, by = "trial") |>
                 filter(is.na(median_pfs_upper)) |>
                 mutate(dodged_cond_group_name = as.numeric(cond_group_name) + if_else(fct_match(dco, "aug"), 0.25, - 0.25))) +
  geom_point(aes(x = median_pfs, shape = "data"), fill = "white", size = 2, position = position_dodge(width = 1), data = \(d) semi_join(obs_km, d, by = "trial")) +
  scale_color_discrete("DCO", type = AZ_palette, label = str_to_title) +
  scale_x_continuous("Median PFS [Months]", breaks = months_to_weeks(seq(0, 60, 4)), label = label_weeks_to_months) +
  scale_y_discrete("", drop = FALSE) +
  scale_shape_manual("Method", labels = c(model = "Model Forecast", data = "Observed KM"), values = c(model = 18, data = 21)) +
  facet_grid(vars(variable), vars(trial), scales = "free", space = "free",
             labeller = labeller(trial = str_to_upper, variable = c("all" = "All", "pdl1_naive" = "First Line"))) +
  theme(legend.position = "inside", legend.position.inside = c(0.7, 0.9), legend.direction = "horizontal") +
  NULL
```

::: {.callout-note}
## Interpreting the Plot
- **Diamond markers**: Model forecast with 90% credible interval
- **Circle markers**: Observed KM median with confidence interval
- **Arrows**: Indicate median PFS not yet reached (right-censored)
:::

## PFS-n by DCO {#sec-pfs-n-dco}

Comparing PFS at fixed timepoints across data cutoffs.

```{r}
#| label: fig-pfs-n-dco
#| fig-cap: Progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial, shown for all DCOs.
#| fig-width: 10

bind_rows(
  aug = bind_rows(
    tar_read(all_tumor_ssls_forecast_target_pfs_n_rvar_ctdna_aug) |> rename(pfs_n = sample_target_pfs_n),
    tar_read(all_tumor_ssls_cond_forecast_target_pfs_n_rvar_ctdna_aug) |> rename(pfs_n = cond_sample_target_pfs_n)
  ),
  apr = bind_rows(
    tar_read(all_tumor_ssls_forecast_target_pfs_n_rvar_ctdna_apr) |> rename(pfs_n = sample_target_pfs_n),
    tar_read(all_tumor_ssls_cond_forecast_target_pfs_n_rvar_ctdna_apr) |> rename(pfs_n = cond_sample_target_pfs_n)
  ),
  .id = "dco"
) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(trial, "sclc"), fct_match(variable, c("all", "pdl1_naive"))) |>
  ggplot() +
  stat_pointinterval(aes(ydist = pfs_n, x = timepoint, group = interaction(fit_type, timepoint, dco), color = dco, alpha = fit_type),
                     position = "dodge", .width = c(0.5, 0.8)) +
  scale_x_continuous("Timepoint", breaks = tar_read(pfs_timepoints) |> pluck("timepoint")) +
  scale_color_discrete("DCO", type = AZ_palette, label = str_to_title) +
  scale_alpha_manual("", values = c(prior = 0.125, posterior = 1), label = str_to_title) +
  facet_grid(vars(variable), vars(cond_group_name), scales = "free", space = "free",
             labeller = labeller(trial = str_to_upper, variable = c("all" = "All", "pdl1_naive" = "First Line"))) +
  NULL
```

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](../analysis/clinical-endpoints.qmd) - Latest endpoint estimates
- [Survival Analysis](../analysis/survival-analysis.qmd) - Kaplan-Meier curves
- [Model Validation](../analysis/model-validation.qmd) - Cross-validation results
:::
