---
title: "Survival Analysis"
subtitle: "Kaplan-Meier forecasts and observed PFS"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
km_data     <- pub_tar_read("all_tumor_ssls_km_rvar")
obs_km_data <- pub_tar_read("km_trial_pfs") |>
  filter(fct_match(btype, "ub")) |>
  select(!quantiles) |>
  unnest(km_data)
```

## PFS Kaplan-Meier Curves {#sec-pfs-km}

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-pfs-km-sample
#| fig-cap: Sample-based PFS KM (posterior) by trial. Shaded band shows 80% credible interval.
#| fig-width: 10
#| fig-height: 5

km_data |>
  filter(fit_type == "posterior") |>
  plot_km(
    km_est_var  = sample_pfs_km_est,
    facet_col   = trial,
    obs_km_data = obs_km_data
  )
```

### Superpopulation

```{r}
#| label: fig-pfs-km-spop
#| fig-cap: Superpopulation PFS KM (posterior) by trial. Wider uncertainty than sample-based as it reflects a new patient drawn from the population.
#| fig-width: 10
#| fig-height: 5

km_data |>
  filter(fit_type == "posterior") |>
  plot_km(
    km_est_var  = spop_pfs_km_est,
    facet_col   = trial,
    obs_km_data = obs_km_data
  )
```

### Prior vs Posterior

```{r}
#| label: fig-pfs-km-prior-post
#| fig-cap: Prior vs posterior PFS KM (sample-based) by trial.
#| fig-width: 10
#| fig-height: 5

km_data |>
  plot_km(
    km_est_var  = sample_pfs_km_est,
    facet_col   = trial,
    group       = fit_type,
    alpha_group = fit_type
  )
```

:::

## Sanofi EFC5505 Model-Observed Discrepancy {#sec-sanofi-note}

::: {.callout-note}
## Why the model under-predicts PFS in the Sanofi control arm

The Sanofi EFC5505 trial (xaliproden vs placebo in mCRC) shows a substantial gap between
the observed KM (median ~16 months) and the model's posterior prediction (~10 months). This
is not a model failure — it reflects a structural PFS ascertainment bias in that trial's design.

**The mechanism:**

1. Patients discontinue oxaliplatin early (due to neuropathy — EFC5505 is a neuroprotection
   trial, so early chemo stops are driven by PSN toxicity, not progression).
2. Tumour scans continue every 8 weeks but only up to month 13.
3. After month 13, no further scans are performed. Patients without documented progression
   are censored at their last scan date.
4. Patients who die later (e.g. month 15–18) have their deaths captured for OS (survival
   follow-up continues every 3 months indefinitely after PD), but not for PFS — if no scan
   documented progression before the 13-month cap, the PFS event is never recorded.

**Why this disproportionately affects the control arm:**

The xaliproden arm was designed to keep patients on oxaliplatin longer (less neuropathy →
more treatment cycles → more scans → more chances to document progression before the cap).
The placebo arm experienced more early oxaliplatin discontinuation, fewer total scans before
the 13-month ceiling, and more patients hitting that ceiling without documented progression.
Those patients are censored for PFS even if they die shortly after.

This creates **informative censoring** in the control arm: patients sick enough to die at
months 15–18 appear as PFS-censored at month 12–13 because no scan caught progression in
the narrowing assessment window. The model, which infers PFS from tumour dynamics without
knowledge of the scan-schedule truncation, correctly predicts a shorter true PFS — the
inflated observed KM reflects ascertainment artefact rather than genuine long-term control.
:::


