---
title: "Clinical Endpoints"
subtitle: "Overall Response Rate (ORR) and Progression-Free Survival (PFS) results"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents the key clinical endpoint results from the SCLC-01 analysis.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd#sec-clinical-endpoints) for endpoint definitions
- See [Survival Analysis](survival-analysis.qmd) for Kaplan-Meier curves
:::

## Overall Response Rate (ORR) {#sec-orr}

Overall response rate represents the proportion of patients achieving complete response (CR) or partial response (PR), confirmed at two consecutive visits.

```{r}
#| label: fig-orr
#| fig-cap: Overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial.

bind_rows(
  tar_read(all_tumor_ssls_orr_rvar_ctdna_aug) |> rename(orr = sample_target_orr),
  tar_read(all_tumor_ssls_cond_orr_rvar_ctdna_aug) |> rename(orr = cond_sample_target_orr)
) |>
  filter(fct_match(dco, "aug")) |>
  plot_outcome_by_pdl1_and_trial(orr) +
  labs(x = "Overall Response Rate (ORR)", y = "Density")
```

```{r}
#| label: tbl-orr-summary
#| tbl-cap: "Tabular summary of Overall Response Rate (ORR) by PDL1 status."

orr_data_tb <- bind_rows(
  tar_read(all_tumor_ssls_orr_rvar_ctdna_aug) |> rename(orr = sample_target_orr),
  tar_read(all_tumor_ssls_cond_orr_rvar_ctdna_aug) |> rename(orr = cond_sample_target_orr)
) |>
  filter(fct_match(dco, "aug")) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(variable, c("all", "pdl1_naive")), fct_match(trial, "sclc"), fct_match(fit_type, "posterior")) |>
  point_interval(orr, .width = c(0.9))

orr_data_tb |>
  mutate(
    variable = recode(variable, all = "All Patients", pdl1_naive = "First Line Patients"),
  ) |>
  select(variable, cond_group_name, orr, .lower, .upper) |>
  group_by(variable) |>
  gt(rowname_col = NULL) |>
  tab_header(
    title = "Overall Response Rate (ORR) Posterior Summary",
  ) |>
  cols_label(
    orr = "Median ORR",
    .lower = "5%",
    .upper = "95%",
    cond_group_name = ""
  ) |>
  cols_hide(columns = variable) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()
  ) |>
  tab_spanner(
    label = "Percentiles",
    columns = c(".lower", ".upper")
  ) |>
  fmt_percent(
    columns = c(orr, .lower, .upper),
    decimals = 2
  )
```

## Median Progression-Free Survival {#sec-median-pfs}

Median PFS represents the time at which 50% of patients have experienced progression or death.

```{r}
#| label: fig-median-pfs
#| fig-cap: Median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial.

bind_rows(
  tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_aug) |> rename(median_pfs = sample_quant_pfs),
  tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_aug) |> rename(median_pfs = cond_sample_quant_pfs)
) |>
  filter(quantile == 0.5, fct_match(dco, "aug")) |>
  plot_outcome_by_pdl1_and_trial(median_pfs) +
  scale_x_continuous("Median PFS [Months]", breaks = months_to_weeks(seq(0, 48, 2)), label = label_weeks_to_months) +
  labs(y = "") +
  coord_cartesian(xlim = c(0, months_to_weeks(26))) +
  NULL
```

```{r}
#| label: tbl-median-pfs-summary
#| tbl-cap: "Tabular summary of median PFS by PDL1 status."

median_pfs_data_tb <- bind_rows(
  tar_read(all_tumor_ssls_trial_quant_pfs_ctdna_aug) |> rename(median_pfs = sample_quant_pfs),
  tar_read(all_tumor_ssls_cond_quant_pfs_ctdna_aug) |> rename(median_pfs = cond_sample_quant_pfs)
) |>
  filter(quantile == 0.5, fct_match(dco, "aug")) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(variable, c("all", "pdl1_naive")), fct_match(trial, "sclc"), fct_match(fit_type, "posterior")) |>
  point_interval(median_pfs, .width = c(0.9))

median_pfs_table <- median_pfs_data_tb |>
  mutate(
    variable = recode(variable, all = "All Patients", pdl1_naive = "First Line Patients"),
  ) |>
  select(variable, cond_group_name, median_pfs, .lower, .upper) |>
  group_by(variable) |>
  gt(rowname_col = NULL) |>
  tab_header(
    title = "Median PFS Posterior Summary",
  ) |>
  cols_label(
    median_pfs = "Median PFS (months)",
    .lower = "5%",
    .upper = "95%",
    cond_group_name = ""
  ) |>
  cols_hide(columns = variable) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()
  ) |>
  tab_spanner(
    label = "Percentiles",
    columns = c(".lower", ".upper")
  ) |>
  fmt(
    columns = c(median_pfs, .lower, .upper),
    fns = \(x) sprintf("%.2f", weeks_to_months(x))
  )

median_pfs_table
```

## PFS at Fixed Timepoints (PFS-n) {#sec-pfs-n}

PFS-n represents the probability of being progression-free at specific timepoints (e.g., 6, 9, 12, 18 months).

```{r}
#| label: fig-pfs-n
#| fig-cap: Progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial.

bind_rows(
  tar_read(all_tumor_ssls_forecast_target_pfs_n_rvar_ctdna_aug) |> rename(pfs_n = sample_pfs_n),
  tar_read(all_tumor_ssls_cond_forecast_target_pfs_n_rvar_ctdna_aug) |> rename(pfs_n = cond_sample_pfs_n)
) |>
  filter(fct_match(dco, "aug")) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(trial, "sclc"), fct_match(variable, c("all", "pdl1_naive"))) |>
  ggplot() +
  stat_pointinterval(aes(ydist = pfs_n, x = timepoint, color = fit_type, group = interaction(fit_type, timepoint)),
                     position = "dodge", .width = c(0.5, 0.8)) +
  scale_x_continuous("Timepoint", breaks = tar_read(pfs_timepoints) |> pluck("timepoint")) +
  scale_color_discrete("", type = AZ_palette, label = str_to_title) +
  facet_grid(vars(variable), vars(cond_group_name), scales = "free", space = "free",
             labeller = labeller(trial = str_to_upper, variable = c("all" = "All", "pdl1_naive" = "First Line"))) +
  NULL
```

```{r}
#| label: tbl-pfs-n-summary
#| tbl-cap: "Tabular summary of PFSn by PDL1 status."

pfs_n_data_tb <- bind_rows(
  tar_read(all_tumor_ssls_forecast_target_pfs_n_rvar_ctdna_aug) |> rename(pfs_n = sample_pfs_n),
  tar_read(all_tumor_ssls_cond_forecast_target_pfs_n_rvar_ctdna_aug) |> rename(pfs_n = cond_sample_pfs_n)
) |>
  filter(fct_match(dco, "aug")) |>
  prepare_pdl1_and_trial_info() |>
  filter(fct_match(variable, c("all", "pdl1_naive")), fct_match(trial, "sclc"), fct_match(fit_type, "posterior"), timepoint %in% c(6, 9, 12, 18)) |>
  point_interval(pfs_n, .width = c(0.9))

pfs_n_table <- pfs_n_data_tb |>
  mutate(
    variable = recode(variable, all = "All Patients", pdl1_naive = "First Line Patients"),
    timepoint_label = paste(timepoint, "Months")
  ) |>
  select(variable, timepoint_label, cond_group_name, pfs_n, .lower, .upper) |>
  group_by(timepoint_label, variable) |>
  gt() |>
  tab_header(
    title = "PFSn Posterior Summary",
  ) |>
  cols_label(
    pfs_n = "Median PFSn",
    .lower = "5%",
    .upper = "95%",
    cond_group_name = ""
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()
  ) |>
  tab_spanner(
    label = "Percentiles",
    columns = c(".lower", ".upper")
  ) |>
  fmt_percent(columns = c("pfs_n", ".lower", ".upper"), decimals = 2)

pfs_n_table
```

::: {.related-pages}
## Related Pages

- [Survival Analysis](survival-analysis.qmd) - Kaplan-Meier curves
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff comparisons
- [Model Specification](../documentation/model-specification.qmd#sec-clinical-endpoints) - Endpoint definitions
:::
