---
title: "Clinical Endpoints"
subtitle: "Overall Response Rate (ORR) and Progression-Free Survival (PFS) results"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents the key clinical endpoint results from the SCLC-01 analysis.

::: {.callout-note}
## Data Cutoff
Results on this page are based on **DCO3 (January 2026)**. For comparisons across data cutoffs, see [DCO Comparisons](../comparisons/dco-comparisons.qmd).
:::

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd#sec-clinical-endpoints) for endpoint definitions
- See [Survival Analysis](survival-analysis.qmd) for Kaplan-Meier curves
:::

## Overall Response Rate (ORR) {#sec-orr}

Overall response rate represents the proportion of patients achieving complete response (CR) or partial response (PR), confirmed at two consecutive visits.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-orr-sample
#| fig-cap: Sample-based overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial.

plot_orr_by_estimate_type("sample")
```

### Superpopulation

```{r}
#| label: fig-orr-spop
#| fig-cap: Superpopulation overall response rate (ORR), split by PDL1 status and treatment line for the SCLC trial.

plot_orr_by_estimate_type("spop")
```

:::

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: tbl-orr-summary-sample
#| tbl-cap: "Sample-based ORR summary by PDL1 status."

create_orr_table("sample")
```

### Superpopulation

```{r}
#| label: tbl-orr-summary-spop
#| tbl-cap: "Superpopulation ORR summary by PDL1 status."

create_orr_table("spop")
```

:::

## Median Progression-Free Survival {#sec-median-pfs}

Median PFS represents the time at which 50% of patients have experienced progression or death.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-median-pfs-sample
#| fig-cap: Sample-based median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial.

plot_median_pfs_by_estimate_type("sample")
```

### Superpopulation

```{r}
#| label: fig-median-pfs-spop
#| fig-cap: Superpopulation median progression-free survival (PFS), split by PDL1 status and treatment line for the SCLC trial.

plot_median_pfs_by_estimate_type("spop")
```

:::

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: tbl-median-pfs-summary-sample
#| tbl-cap: "Sample-based median PFS summary by PDL1 status."

create_median_pfs_table("sample")
```

### Superpopulation

```{r}
#| label: tbl-median-pfs-summary-spop
#| tbl-cap: "Superpopulation median PFS summary by PDL1 status."

create_median_pfs_table("spop")
```

:::

## PFS at Fixed Timepoints (PFS-n) {#sec-pfs-n}

PFS-n represents the probability of being progression-free at specific timepoints (e.g., 6, 9, 12, 18 months).

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-pfs-n-sample
#| fig-cap: Sample-based progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial.

plot_pfs_n_by_estimate_type("sample")
```

### Superpopulation

```{r}
#| label: fig-pfs-n-spop
#| fig-cap: Superpopulation progression-free survival (PFS) at different time points, split by PDL1 status and treatment line for the SCLC trial.

plot_pfs_n_by_estimate_type("spop")
```

:::

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: tbl-pfs-n-summary-sample
#| tbl-cap: "Sample-based PFS-n summary by PDL1 status."

create_pfs_n_table("sample")
```

### Superpopulation

```{r}
#| label: tbl-pfs-n-summary-spop
#| tbl-cap: "Superpopulation PFS-n summary by PDL1 status."

create_pfs_n_table("spop")
```

:::

::: {.related-pages}
## Related Pages

- [Survival Analysis](survival-analysis.qmd) - Kaplan-Meier curves
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff comparisons
- [Model Specification](../documentation/model-specification.qmd#sec-clinical-endpoints) - Endpoint definitions
:::
