---
title: "Competing Risks"
subtitle: "Cumulative incidence functions for state 0 exits: progression, on-trial death, and dropout"
---

```{r}
#| label: setup
#| include: false

library(targets)
library(posterior)
library(tidyverse)
library(ggdist)
library(patchwork)
library(cmprsk)

here::i_am("quarto/website/_setup.R")
source(here::here("r", "util.R"))
source(here::here("r", "plot_functions.R"))
source(here::here("r", "table_functions.R"))
source(here::here("r", "sclc", "plot_functions.R"))
source(here::here("r", "sclc", "table_functions.R"))

ggplot2::theme_set(
  ggplot2::theme_minimal(base_family = "Arial") +
  ggplot2::theme(
    plot.margin = ggplot2::margin(5, 10, 20, 20, "pt"),
    plot.caption = ggplot2::element_text(hjust = 0, margin = ggplot2::margin(15, 0, 0, 0, "pt"))
  )
)

# Standalone multistate model store
ms_store <- "/mnt/data/analysis-results/karim_naguib/sclc/ms-trial-intercept/_targets"
```

This page validates the **4-state multistate model's** allocation of patients across competing exit routes from state 0 (on-treatment, pre-progression). From state 0, a patient can:

- **0->1 Progression**: Tumor progression detected via RECIST
- **0->2 On-trial death**: Death without prior progression while on trial
- **0->3 Dropout**: Treatment discontinuation without progression or on-trial death

The cumulative incidence function (CIF) measures the probability of each cause by time $t$, accounting for the fact that causes compete --- a patient who progresses can no longer drop out, and vice versa.

::: {.callout-note}
## CIF vs KM
Unlike the Kaplan-Meier estimator (which estimates the marginal survival function), the CIF accounts for competing risks and always sums to at most 1 across all causes. The observed CIF is computed using the Aalen-Johansen estimator.
:::

## Standalone Multistate Model {#sec-standalone-cif}

The standalone multistate model estimates transition hazards **without SLD tumor dynamics** --- all transitions are driven purely by baseline hazards and covariates. This provides a useful diagnostic: how well can the multistate component alone reproduce the observed competing risks structure?

### Event Counts

```{r}
#| label: tbl-event-counts-standalone
#| tbl-cap: "Observed competing risks event counts from state 0 (standalone model data)."

stan_data <- tar_read(ms_standalone_stan_data_posterior_ctdna_jan26, store = ms_store)
competing_risks_event_table(stan_data)
```

### CIF Validation

```{r}
#| label: fig-cif-standalone
#| fig-cap: "Competing risks CIF: standalone multistate model (posterior predictive) vs observed Aalen-Johansen. Top panel shows per-cause CIF; bottom panel shows proportion of patients still in state 0."
#| fig-width: 10
#| fig-height: 9
#| column: body-outset

draws_cif <- tar_read(ms_standalone_draws_cif_posterior_ctdna_jan26, store = ms_store)
plot_competing_risks_cif(stan_data, draws_cif)
```

## Prior Predictive Check {#sec-prior-predictive-cif}

Following [Gabry et al. (2019)](https://doi.org/10.1111/rssa.12378), we compare the prior predictive CIF against the observed data to assess whether the priors encode a reasonable competing risks structure before seeing data.

```{r}
#| label: fig-cif-prior
#| fig-cap: "Prior predictive competing risks CIF vs observed Aalen-Johansen. The prior spreads probability roughly equally across all causes rather than concentrating on progression."
#| fig-width: 10
#| fig-height: 9
#| column: body-outset

stan_data_prior <- tar_read(ms_standalone_stan_data_prior_ctdna_jan26, store = ms_store)
draws_cif_prior <- tar_read(ms_standalone_draws_cif_prior_ctdna_jan26, store = ms_store)

plot_competing_risks_cif(stan_data_prior, draws_cif_prior)
```

The prior predictive reveals a different miscalibration pattern than the posterior:

- **0→2 on-trial death** is massively over-predicted (~25--30% vs observed ~0--5%), while 0→3 dropout is roughly in the right range.
- **0→1 progression** is under-predicted (~35--40% vs observed ~65--75%), consistent with the prior not encoding that progression is the dominant competing risk.

Comparing this to the posterior (@sec-standalone-cif), the data *corrects* the 0→2 over-prediction but redistributes that probability to 0→3 dropout rather than 0→1 progression. This confirms the miscalibration is not purely prior-driven --- the data is doing work, but the standalone model (without SLD tumor dynamics) lacks the information to route events correctly to 0→1.

::: {.related-pages}
## Related Pages

- [Overall Survival](overall-survival.qmd) - OS Kaplan-Meier curves and predictions
- [Survival Analysis](survival-analysis.qmd) - PFS Kaplan-Meier curves
- [Model Specification](../documentation/model-specification.qmd) - Multistate model details
:::
