---
title: "Data Exploration"
subtitle: "Overview of the analysis dataset and exploratory visualizations"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page provides an overview of the data used in the SCLC-01 longitudinal tumor analysis.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd) for RECIST criteria definitions
- See [Clinical Endpoints](clinical-endpoints.qmd) for ORR and PFS results
:::

## Tumor Size Dynamics {#sec-tumor-dynamics}

The following plot shows the sum of lesion diameters (SLD) over time for all patients in the analysis, grouped by trial.

```{r}
#| label: fig-sum-tumor-size
#| fig-cap: Sum of patient tumor sizes dynamics.

all_analysis_data |>
  unnest(visit_data) |>
  ggplot(aes(ady, mmsumdiam)) +
  geom_line(aes(group = usubjid), alpha = 0.0675) +
  geom_point(size = 0.5, alpha = 0.35) +
  geom_point(aes(color = "Right Censored"), size = 0.5, data = \(d) filter(d, right_censored) |> group_by(usubjid) |> filter(min_rank(week) == n())) +
  scale_color_discrete("", type = AZ_palette[-1]) +
  labs(x = "Analysis Day", y = "Sum of tumor diameters [mm]") +
  facet_wrap(vars(trial), scales = "free", labeller = labeller(trial = str_to_upper)) +
  theme(legend.position = "bottom")
```

## PFS Distribution {#sec-pfs-distribution}

The distribution of progression-free survival (PFS), split by event type (censored, death, or progressive disease).

```{r}
#| label: fig-hist-right-censored
#| fig-cap: Observed distribution histogram of PFS split by right censoring

all_analysis_data |>
  mutate(event_type = case_when(
    right_censored ~ "censored",
    progression_before_death ~ "pd",
    !progression_before_death & death ~ "death",
  )) |>
  ggplot(aes(pfs)) +
  geom_histogram(aes(y = after_stat(count), fill = event_type), binwidth = 4, alpha = 0.4, position = "identity") +
  scale_fill_discrete("Event Type", label = c(censored = "Right Censored", death = "Death", pd = "Progressive Disease"), type = AZ_palette[-c(1:3)]) +
  scale_x_continuous("PFS [Months]", breaks = months_to_weeks(seq(0, 36, 6)), label = label_weeks_to_months) +
  labs(y = "Density") +
  facet_wrap(vars(trial), labeller = labeller(trial = str_to_upper), scales = "free")
```

## RECIST Classification Comparison {#sec-recist-comparison}

The model uses RECIST 1.1 criteria to classify tumor response based on target lesion SLD measurements. This plot compares the recorded clinical RECIST classifications with those calculated from target tumor SLD alone.

::: {.callout-note}
## Why do these differ?
Clinical RECIST assessments include non-target lesions and new lesions, while our calculated response is based solely on target lesion SLD dynamics. See [Model Specification](../documentation/model-specification.qmd#sec-recist) for RECIST criteria details.
:::

```{r}
#| label: fig-data-vs-calculated-recist-confusion-matrix
#| fig-cap: Confusion matrix for the RECIST categories as observed in the data and as calculated using only the target tumor SLD.

all_analysis_data |>
  select(trial, visit_data) |>
  unnest(visit_data) |>
  filter(!is.na(response)) |>
  count(trial, response, det_response) |>
  group_by(trial, response) |>
  mutate(p = n / sum(n)) |>
  ungroup() |>
  plot_confusion_matrix(response, det_response, p, n) +
  facet_wrap(vars(trial), labeller = labeller(.default = str_to_upper)) +
  labs(x = "Recorded Response", y = "Calculated Response") +
  NULL
```

## Target vs Recorded PFS {#sec-target-vs-recorded}

Comparison between the deterministic target lesion PFS (from SLD dynamics) and the recorded clinical PFS.

```{r}
#| label: fig-compare-target-pfs-recorded-pfs
#| fig-cap: Comparing the deterministic target PFS with the recorded PFS

all_analysis_data |>
  ggplot(aes(pfs + interval_censored + 1, det_pfs)) +
  geom_abline(slope = 1, linetype = "dashed") +
  geom_point(color = AZ_platinum, alpha = 0.5) +
  labs(x = "Recorded PFS", y = "Target tumor PFS") +
  facet_wrap(vars(trial), labeller = labeller(trial = str_to_upper))
```

## Baseline Characteristics {#sec-baseline}

Distribution of baseline laboratory values across trials.

```{r}
#| label: fig-baseline-lab
#| fig-cap: Baseline laboratory values distribution by trial
#| fig-width: 10
#| message: false

all_analysis_data |>
  select(usubjid, trial, starts_with("baseline_")) |>
  pivot_longer(!c(usubjid, trial)) |>
  ggplot() +
  geom_histogram(aes(value, after_stat(density), fill = trial), position = "identity", alpha = 0.5) +
  scale_fill_discrete("", type = AZ_palette, label = str_to_upper) +
  facet_wrap(vars(name), scales = "free", labeller = labeller(name = \(n) str_remove(n, "baseline_") |> str_to_title()))
```

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and PFS results
- [Survival Analysis](survival-analysis.qmd) - Kaplan-Meier curves
- [Patient Dynamics](patient-dynamics.qmd) - Individual tumor trajectories
- [Model Specification](../documentation/model-specification.qmd) - RECIST criteria
:::
