---
title: "Sampling Diagnostics"
subtitle: "MCMC convergence and sampling quality assessment"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents MCMC sampling diagnostics to assess the quality and convergence of the Stan model fit.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd) for model details
- See [Architecture](../documentation/architecture.qmd) for computational optimization strategies
:::

## MCMC Diagnostics Summary {#sec-diagnostics}

Key diagnostics from CmdStan indicating sampling quality.

```{r}
#| label: ssls-diagnostics

tar_read(tumor_ssls_res_posterior_ctdna_aug)$diagnostic_summary()
```

::: {.callout-note}
## Interpreting Diagnostics
- **Divergences**: Should be 0 or very low; indicates problematic posterior geometry
- **Tree depth exceeded**: Maximum tree depth reached; may indicate difficult sampling
- **E-BFMI**: Energy Bayesian fraction of missing information; should be > 0.3
- **R-hat**: Convergence diagnostic; should be < 1.01 for all parameters
- **ESS (Effective Sample Size)**: Should be sufficiently high for reliable inference
:::

## Understanding Common Issues

### Divergences

Divergent transitions indicate that the sampler encountered regions of the posterior with very different curvature than expected. Common causes:

- **Strong parameter correlations**: May need non-centered parameterization
- **Heavy-tailed posteriors**: May need reparameterization
- **Model misspecification**: Prior-posterior conflict

The Sclc model uses non-centered parameterization (NCP) for hierarchical parameters to reduce divergences.

### Tree Depth

When the NUTS sampler exceeds maximum tree depth, it may be exploring regions inefficiently. The model uses computational optimizations (see [Architecture](../documentation/architecture.qmd#sec-optimization)) to improve sampling efficiency.

### R-hat

$\hat{R}$ compares between-chain and within-chain variance. Values close to 1 indicate that chains have converged to the same distribution. The Sclc model targets $\hat{R} < 1.01$ for all parameters.

## Additional Diagnostics

For detailed parameter-level diagnostics, trace plots, and pair plots, the CmdStan output files can be inspected directly:

```r
# Access the fitted model object
fit <- tar_read(tumor_ssls_res_posterior_ctdna_aug)

# View parameter summaries with diagnostics
fit$summary()

# Check specific parameters
fit$summary(variables = c("tr_intercept_pop", "frac_intercept_pop"))
```

::: {.related-pages}
## Related Pages

- [Model Validation](model-validation.qmd) - Cross-validation results
- [Parameter Inference](parameter-inference.qmd) - Posterior estimates
- [Architecture](../documentation/architecture.qmd) - Optimization strategies
:::
