---
title: "Model Validation"
subtitle: "Leave-future-out cross-validation and accuracy assessment"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents model validation results using leave-future-out (LFO) cross-validation and accuracy assessments.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd) for model details
- See [Architecture](../documentation/architecture.qmd) for computational optimization strategies
:::

## Leave-Future-Out Cross-Validation {#sec-lfo}

Leave-future-out (LFO) cross-validation is used to evaluate the model's predictive performance. Unlike leave-one-out (LOO), LFO respects the temporal structure of the data.

### LFO vs LOO Comparison

```{r}
#| label: fig-lfo-vs-loo-diagram
#| fig-cap: Comparison between LFO and LOO cross-validation showing how observations are held out in each approach.
#| fig-width: 10
#| fig-height: 4

source(here("r/plot_lfo_diagram.R"))
```

::: {.callout-note}
## Why LFO over LOO?
- **LOO** holds out individual observations regardless of time, which can lead to information leakage when predicting earlier time points using models trained on later data
- **LFO** holds out all future observations from a cutoff point, maintaining the causal structure of the prediction task
:::

### Model Comparison: Historical Data Impact

Comparison of models with different historical data configurations to understand the value of information borrowing.

```{r}
#| label: fig-lfo-elpd
#| fig-cap: Difference in leave-future-out expected log predictive density (ELPD) comparing models with different historical data configurations.
#| fig-height: 2

all_lfo_clean <- bind_rows(
    ctdna = tar_read(tumor_ssls_lfo_clean_lite_ctdna_aug, store = lfo_store2),
    no_ctdna = tar_read(tumor_ssls_lfo_clean_lite_no_ctdna_aug, store = lfo_store),
    ctdna_only = tar_read(tumor_ssls_lfo_clean_lite_ctdna_only_aug, store = lfo_store_ctdna_only),
    no_hist = tar_read(tumor_ssls_lfo_clean_lite_no_hist_aug, store = lfo_store_no_hist),
    no_pl = tar_read(tumor_ssls_lfo_clean_lite_no_pl_aug, store = lfo_store_no_pl),
  .id = "model"
) |>
  filter(n > 3)

# Define model labels for LFO plots
lfo_model_labels <- list(
  ctdna = "Covariates: All",
  no_ctdna = "Covariates: Without Baseline ctDNA",
  ctdna_only = "Covariates: Baseline ctDNA Only",
  no_pl = "Without Patient-level Hierarchy",
  all_trials = "All Historical Trials",
  no_hist = "Without Any Historical Data",
  no_covar = "Covariates: None",
  with_artic = "HISTORICAL + ARTIC",
  with_atlantic = "HISTORICAL + ATLANTIC",
  with_pearl = "HISTORICAL + PEARL"
)

all_lfo_clean |>
  filter(fct_match(model, c("ctdna", "no_pl", "no_hist"))) |>
  plot_lfo_elpd_diff(
    baseline_model = "no_hist",
    model_labels = lfo_model_labels |>
      list_assign(
        ctdna = "SCLC + HISTORICAL",
        all_trials = "SCLC + All Historical Trials",
      )
  )
```

::: {.callout-note}
## Interpreting ELPD Differences
- **Positive ELPD difference**: Better predictive performance than baseline
- **Error bars**: Standard error of the ELPD difference
- Models with historical data borrowing show improved predictive accuracy
:::

### Model Comparison: Covariate Specifications

Evaluating the impact of different covariate specifications, particularly the role of baseline ctDNA.

```{r}
#| label: fig-lfo-elpd-ctdna
#| fig-cap: Difference in leave-future-out expected log predictive density (ELPD) for models with different ctDNA covariate specifications.
#| fig-height: 2

all_lfo_clean |>
  filter(fct_match(model, c("ctdna_only", "ctdna", "no_ctdna"))) |>
  plot_lfo_elpd_diff(
    baseline_model = "ctdna_only",
    model_labels = lfo_model_labels
  )
```

## Accuracy Assessment {#sec-accuracy}

### RECIST Prediction Accuracy

Out-of-sample confusion matrix comparing observed RECIST classifications with model predictions.

```{r}
#| label: fig-data-vs-posterior-confusion-matrix
#| fig-cap: Out-of-sample (leave-future-out) confusion matrix for the RECIST categories as observed in the data and as estimated using only the target tumor SLD.
#| fig-width: 6

tar_read(oos_confusion_matrix_ctdna_aug, store = lfo_store2) |>
  mutate(mp = median(mean_pred)) |>
  plot_confusion_matrix(response, pred_response, mp) +
  labs(x = "Recorded Response", y = "Median Posterior Response", caption = "Cell proportions are the medians of the column-wise probabilities (conditional on recorded response)") +
  NULL
```

### Progressive Disease Prediction Over Time

Tracking how well the model predicts progressive disease (PD) at different points during the trial.

```{r}
#| label: fig-pd-probability-over-time
#| fig-cap: Probability of predicting PD conditional on observed RECIST being PD, over time.
#| fig-width: 6
#| fig-height: 5

tar_read(full_oos_confusion_matrix_ctdna_aug, store = lfo_store2) |>
  filter(response == "PD") |>
  group_by(n, m, cutoff_date, response) |>
  mutate(
    total = rvar_sum(oos_recist_confusion_matrix),
    prop = oos_recist_confusion_matrix / total
  ) |>
  ungroup() |>
  filter(pred_response == "PD") |>
  mutate(prop_median = median(prop)) |>
  ggplot(aes(x = cutoff_date)) +
  stat_lineribbon(aes(ydist = prop), alpha = 0.4, .width = 0.8, linewidth = 0) +
  geom_smooth(aes(y = prop_median), method = "loess", span = 0.3, se = FALSE, color = "darkblue", linewidth = 1) +
  scale_fill_manual("80% CI", values = AZ_turquoise) +
  scale_y_continuous("P(Predicted PD | Observed PD)", breaks = seq(0, 1, 0.1), labels = scales::percent_format(), limits = c(0, NA)) +
  labs(
    x = "Cutoff Date",
    caption = "Ribbon shows 80% credible interval; line shows LOESS-smoothed median"
  ) +
  theme(legend.position = "bottom")
```

::: {.related-pages}
## Related Pages

- [Survival Analysis](survival-analysis.qmd#sec-cutoff-km) - Cutoff KM comparisons
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff analyses
- [Diagnostics](diagnostics.qmd) - Sampling diagnostics
- [Model Specification](../documentation/model-specification.qmd) - Model details
:::
