---
title: "Model Validation"
subtitle: "Leave-future-out cross-validation and accuracy assessment"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")

# Load SLD data for posterior predictive checks
sld_data_historical_apr25 <- tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_apr25)
sld_data_historical_aug25 <- tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_aug25)
sld_data_historical_jan26 <- tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_jan26)
sld_data_sclc_apr25 <- tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_apr25)
sld_data_sclc_aug25 <- tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_aug25)
sld_data_sclc_jan26 <- tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_jan26)
```

This page presents model validation results using leave-future-out (LFO) cross-validation and accuracy assessments.

::: {.callout-note}
## Data Cutoff
Leave-future-out cross-validation results on this page are based on **DCO2 (August 2025)**.
:::

::: {.callout-tip}
## Related Pages
- See [Leave-Future-Out Cross-Validation](../documentation/leave-future-out-cross-validation.qmd) for LFO methodology details
- See [Model Architecture](../documentation/model-architecture.qmd) for model overview
- See [Architecture](../documentation/architecture.qmd) for computational optimization strategies
:::

## Posterior Predictive Checks - Uncensored Patients {#sec-ppc}

Posterior checking plots for patients with observed progression events, comparing model predictions with actual SLD measurements.

### HISTORICAL Trial

::: {.panel-tabset}

#### April 2025

```{r}
#| label: fig-forecast-sld-uncensored-historical-apr25
#| fig-cap: SLD posterior check for HISTORICAL uncensored patients (April 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

plot_sld_ppc_by_trial_dco("historical", sld_data_historical_apr25)
```

#### August 2025

```{r}
#| label: fig-forecast-sld-uncensored-historical-aug25
#| fig-cap: SLD posterior check for HISTORICAL uncensored patients (August 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

plot_sld_ppc_by_trial_dco("historical", sld_data_historical_aug25)
```

#### January 2026 {.active}

```{r}
#| label: fig-forecast-sld-uncensored-historical-jan26
#| fig-cap: SLD posterior check for HISTORICAL uncensored patients (January 2026 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

plot_sld_ppc_by_trial_dco("historical", sld_data_historical_jan26)
```

:::

### SCLC-01 Trial

::: {.panel-tabset}

#### April 2025

```{r}
#| label: fig-forecast-sld-uncensored-sclc-apr25
#| fig-cap: SLD posterior check for SCLC-01 uncensored patients (April 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

plot_sld_ppc_by_trial_dco("sclc", sld_data_sclc_apr25)
```

#### August 2025

```{r}
#| label: fig-forecast-sld-uncensored-sclc-aug25
#| fig-cap: SLD posterior check for SCLC-01 uncensored patients (August 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

plot_sld_ppc_by_trial_dco("sclc", sld_data_sclc_aug25)
```

#### January 2026 {.active}

```{r}
#| label: fig-forecast-sld-uncensored-sclc-jan26
#| fig-cap: SLD posterior check for SCLC-01 uncensored patients (January 2026 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

plot_sld_ppc_by_trial_dco("sclc", sld_data_sclc_jan26)
```

:::

## PFS Posterior Predictive Check {#sec-pfs-ppc}

Comparing posterior PFS predictions with observed PFS for uncensored patients to assess model calibration.

```{r}
#| label: fig-pfs-posterior-check
#| fig-cap: Posterior checking the posterior PFS predictions compared to the observed PFS, restricted to the uncensored patients.
#| column: body-outset
#| fig-width: 10

ppc_spop_pfs <- tar_read(all_tumor_ssls_pfs_rvar_ctdna_jan26) |>
  filter(fct_match(fit_type, "posterior"), !right_censored) |>
  mutate(event_type = case_when(
    !progression_before_death ~ "Death",
    !target_only_right_censored ~ "Target Lesion Progression",
    TRUE ~ "Residual Progression"
  ))

ppc_spop_pfs |>
  plot_pfs_ppc(pfs_var = spop_pfs)
```

## Leave-Future-Out Cross-Validation {#sec-lfo}

Leave-future-out (LFO) cross-validation is used to evaluate the model's predictive performance. Unlike leave-one-out (LOO), LFO respects the temporal structure of the data.

### LFO vs LOO Comparison

```{r}
#| label: fig-lfo-vs-loo-diagram
#| fig-cap: Comparison between LFO and LOO cross-validation showing how observations are held out in each approach.
#| fig-width: 10
#| fig-height: 4

source(here("r/plot_lfo_diagram.R"))
```

::: {.callout-note}
## Why LFO over LOO?
- **LOO** holds out individual observations regardless of time, which can lead to information leakage when predicting earlier time points using models trained on later data
- **LFO** holds out all future observations from a cutoff point, maintaining the causal structure of the prediction task
:::

### Model Comparison: Historical Data Impact

Comparison of models with different historical data configurations to understand the value of information borrowing.

```{r}
#| label: fig-lfo-elpd
#| fig-cap: Difference in leave-future-out expected log predictive density (ELPD) comparing models with different historical data configurations.
#| fig-height: 2

all_lfo_clean <- bind_rows(
    ctdna = tar_read(tumor_ssls_lfo_clean_lite_ctdna_aug, store = lfo_store),
    no_ctdna = tar_read(tumor_ssls_lfo_clean_lite_no_ctdna_aug, store = lfo_store),
    ctdna_only = tar_read(tumor_ssls_lfo_clean_lite_ctdna_only_aug, store = lfo_store_ctdna_only),
    no_hist = tar_read(tumor_ssls_lfo_clean_lite_no_hist_aug, store = lfo_store_no_hist),
    no_pl = tar_read(tumor_ssls_lfo_clean_lite_no_pl_aug, store = lfo_store_no_pl),
    ctdna_pn = tar_read(tumor_ssls_lfo_clean_lite_ctdna_pn_aug, store = lfo_store_ctdna_pn),
  .id = "model"
) |>
  filter(n > 3)

# Define model labels for LFO plots
lfo_model_labels <- list(
  ctdna = "Covariates: All",
  no_ctdna = "Covariates: Without Baseline ctDNA",
  ctdna_only = "Covariates: Baseline ctDNA Only",
  no_pl = "Without Patient-level Hierarchy",
  all_trials = "All Historical Trials",
  no_hist = "Without Any Historical Data",
  no_covar = "Covariates: None",
  ctdna_pn = "With Population-Level Process Noise",
  with_artic = "HISTORICAL + ARTIC",
  with_atlantic = "HISTORICAL + ATLANTIC",
  with_pearl = "HISTORICAL + PEARL"
)

all_lfo_clean |>
  filter(fct_match(model, c("ctdna", "ctdna_pn", "no_pl", "no_hist"))) |>
  plot_lfo_elpd_diff(
    baseline_model = "no_hist",
    model_labels = lfo_model_labels |>
      list_assign(
        ctdna = "SCLC-01 + HISTORICAL",
        all_trials = "SCLC-01 + All Historical Trials",
      )
  )
```

::: {.callout-note}
## Interpreting ELPD Differences
- **Positive ELPD difference**: Better predictive performance than baseline
- **Error bars**: Standard error of the ELPD difference
- Models with historical data borrowing show improved predictive accuracy
:::

### Model Comparison: Covariate Specifications

Evaluating the impact of different covariate specifications, particularly the role of baseline ctDNA.

```{r}
#| label: fig-lfo-elpd-ctdna
#| fig-cap: Difference in leave-future-out expected log predictive density (ELPD) for models with different ctDNA covariate specifications.
#| fig-height: 2

all_lfo_clean |>
  filter(fct_match(model, c("ctdna_only", "ctdna", "ctdna_pn", "no_ctdna"))) |>
  plot_lfo_elpd_diff(
    baseline_model = "ctdna_only",
    model_labels = lfo_model_labels
  )
```

## Accuracy Assessment {#sec-accuracy}

### RECIST Prediction Accuracy

Out-of-sample sensitivity analysis showing how well the model predicts each RECIST category conditional on the observed response.

::: {layout-ncol=2}

```{r}
#| label: fig-data-vs-posterior-confusion-matrix
#| fig-cap: Out-of-sample (leave-future-out) confusion matrix showing sensitivity (recall) for each RECIST category. Cell values represent P(Predicted response | Observed response).
#| fig-width: 5

oos_confusion_matrix <- tar_read(oos_confusion_matrix_ctdna_jan26, store = lfo_store)
plot_oos_confusion_matrix(oos_confusion_matrix)
```

```{r}
#| label: fig-data-vs-posterior-sankey
#| fig-cap: Sankey diagram showing the flow from recorded to predicted RECIST responses. Flow width represents prediction proportions; percentages indicate correct prediction rates.
#| fig-width: 6
#| fig-height: 5

plot_oos_confusion_sankey(oos_confusion_matrix)
```

:::

### RECIST Specificity Analysis

The specificity matrix shows the complementary view: for each category X, it displays the distribution of predictions when the true response is **NOT** X.

```{r}
#| label: fig-data-vs-posterior-specificity-matrix
#| fig-cap: Out-of-sample specificity-related matrix. Each column shows P(Predicted response | Observed ≠ category). Higher proportions in off-diagonal cells indicate better specificity (correct rejection of false positives).
#| fig-width: 6

plot_oos_specificity_matrix(oos_confusion_matrix)
```

::: {.callout-note}
## Understanding Specificity
- Each column shows predictions for cases where the true response is **NOT** that category
- Example: "NOT CR" column shows predictions for all PR, SD, and PD cases
- Higher off-diagonal proportions (e.g., P(Predicted ≠ CR | Observed ≠ CR)) indicate better specificity
- Specificity for category X = sum of all rows except X in the "NOT X" column
:::

### Summary Metrics

```{r}
#| label: tbl-accuracy-metrics
#| tbl-cap: "Out-of-sample accuracy, sensitivity, and specificity with 90% credible intervals."

compute_oos_accuracy_metrics(oos_confusion_matrix) |>
  gt() |>
  tab_header(
    title = "RECIST Prediction Performance"
  ) |>
  cols_label(
    metric = "Metric",
    class = "Class",
    median = "Median",
    q5 = "5%",
    q95 = "95%"
  ) |>
  tab_spanner(
    label = "Percentiles",
    columns = c("q5", "q95")
  ) |>
  fmt_percent(
    columns = c(median, q5, q95),
    decimals = 1
  )
```

::: {.callout-note}
## Interpreting Metrics
- **Confusion Matrix**: Shows column-wise proportions conditional on observed (true) response, i.e., sensitivity/recall for each category
- **Overall Accuracy**: Weighted average of per-class sensitivities, weighted by class prevalence
- **Sensitivity**: P(Predicted X | Observed X) - probability of correctly predicting class X given true class is X (diagonal elements in confusion matrix)
- **Specificity**: P(Predicted not X | Observed not X) - probability of correctly predicting NOT class X when true class is not X
:::

### RECIST Prediction Over Time

Tracking how well the model predicts each RECIST category at different points during the trial.

```{r}
#| label: fig-recist-probability-over-time
#| fig-cap: "Running average sensitivity for each RECIST category across LFO folds. Ribbons show ±1 SE, where SE is computed within each posterior draw across folds, then averaged. This preserves posterior uncertainty while displaying fold-to-fold variability."
#| fig-width: 10
#| fig-height: 6

tar_read(full_oos_confusion_matrix_ctdna_jan26, store = lfo_store) |>
  plot_recist_sensitivity_over_time() +
  labs(caption = "Dashed lines indicate data cutoff dates (DCO1: Apr 2025, DCO2: Aug 2025, DCO3: Jan 2026).")
```

::: {.related-pages}
## Related Pages

- [Leave-Future-Out Cross-Validation](../documentation/leave-future-out-cross-validation.qmd) - LFO methodology
- [Survival Analysis](survival-analysis.qmd#sec-cutoff-km) - Cutoff KM comparisons
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff analyses
- [Diagnostics](diagnostics.qmd) - Sampling diagnostics
- [Model Architecture](../documentation/model-architecture.qmd) - Model overview
:::
