---
title: "Overall Survival"
subtitle: "Kaplan-Meier survival curves, median OS, and OS at fixed timepoints"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents overall survival (OS) analysis results for the SCLC-01 trial using **DCO3 (January 2026)** data.

::: {.callout-note}
## OS Endpoint Definition
OS is measured from randomization to death from any cause. Patients alive at last follow-up are censored. Unlike PFS, OS does not include a progression component --- all events are deaths.
:::

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints](clinical-endpoints.qmd) for ORR and median PFS results
- See [Survival Analysis](survival-analysis.qmd) for PFS Kaplan-Meier curves
- See [Model Specification](../documentation/model-specification.qmd) for endpoint definitions
:::

## SCLC-01 OS Forecasts {#sec-sclc-os-km}

::: {.panel-tabset}

### Sample-Based

Overall survival Kaplan-Meier curves for the SCLC-01 trial, comparing model predictions with observed data across cohorts.

```{r}
#| label: fig-os-km-sclc-sample
#| fig-cap: Sample-based overall survival Kaplan-Meier forecast for SCLC-01 trial.
#| fig-width: 7

plot_sclc_km_by_estimate_type(
  "sample",
  jan26_os_km_data,
  jan26_cond_os_km_data,
  jan26_obs_os_km_all,
  jan26_obs_os_km_first_line,
  all_analysis_data,
  endpoint = "os"
)
```

### Superpopulation

Overall survival Kaplan-Meier curves for the SCLC-01 trial. Superpopulation estimates represent the expected survival for the underlying population, not just the specific patients in the trial.

```{r}
#| label: fig-os-km-sclc-spop
#| fig-cap: Superpopulation overall survival Kaplan-Meier forecast for SCLC-01 trial.
#| fig-width: 7

plot_sclc_km_by_estimate_type(
  "spop",
  jan26_os_km_data,
  jan26_cond_os_km_data,
  jan26_obs_os_km_all,
  jan26_obs_os_km_first_line,
  all_analysis_data,
  endpoint = "os"
)
```

:::

## PDL1-Stratified OS {#sec-pdl1-os-km}

::: {.panel-tabset}

### Sample-Based

Overall survival curves stratified by PDL1 status and treatment line.

```{r}
#| label: fig-os-km-pdl1-sample
#| fig-cap: Sample-based PDL1-stratified overall survival forecast.
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km_by_estimate_type(
  "sample",
  cond_km_data = jan26_cond_os_km_data,
  obs_km_all = jan26_obs_os_km_pdl1,
  obs_km_first_line = jan26_obs_os_km_pdl1_naive,
  obs_km_part_e = jan26_obs_os_km_part_e_pdl1,
  endpoint = "os"
)
```

### Superpopulation

Overall survival curves stratified by PDL1 status and treatment line.

```{r}
#| label: fig-os-km-pdl1-spop
#| fig-cap: Superpopulation PDL1-stratified overall survival forecast.
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km_by_estimate_type(
  "spop",
  cond_km_data = jan26_cond_os_km_data,
  obs_km_all = jan26_obs_os_km_pdl1,
  obs_km_first_line = jan26_obs_os_km_pdl1_naive,
  obs_km_part_e = jan26_obs_os_km_part_e_pdl1,
  endpoint = "os"
)
```

:::

## Median Overall Survival {#sec-median-os}

Median OS represents the time at which 50% of patients have died.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-median-os-sample
#| fig-cap: Sample-based median overall survival, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_median_survival_by_estimate_type("sample", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

### Superpopulation

```{r}
#| label: fig-median-os-spop
#| fig-cap: Superpopulation median overall survival, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_median_survival_by_estimate_type("spop", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

:::

::: {.panel-tabset}

### Sample-Based

#### All Patients

```{r}
#| label: tbl-median-os-summary-sample-all
#| tbl-cap: "Sample-based median OS summary for all patients."

create_median_survival_table("sample", "all", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-median-os-summary-sample-first-line
#| tbl-cap: "Sample-based median OS summary for first line patients."

create_median_survival_table("sample", "first_line", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-median-os-summary-sample-part-e
#| tbl-cap: "Sample-based median OS summary for Part E patients."

create_median_survival_table("sample", "part_e", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

### Superpopulation

#### All Patients

```{r}
#| label: tbl-median-os-summary-spop-all
#| tbl-cap: "Superpopulation median OS summary for all patients."

create_median_survival_table("spop", "all", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-median-os-summary-spop-first-line
#| tbl-cap: "Superpopulation median OS summary for first line patients."

create_median_survival_table("spop", "first_line", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-median-os-summary-spop-part-e
#| tbl-cap: "Superpopulation median OS summary for Part E patients."

create_median_survival_table("spop", "part_e", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

:::

## OS at Fixed Timepoints (OS-n) {#sec-os-n}

OS-n represents the probability of being alive at specific timepoints (e.g., 6, 9, 12, 18 months).

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-os-n-sample
#| fig-cap: Sample-based overall survival at different time points, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_survival_n_by_estimate_type("sample", jan26_os_n_data, jan26_cond_os_n_data, pfs_timepoints, endpoint = "os")
```

### Superpopulation

```{r}
#| label: fig-os-n-spop
#| fig-cap: Superpopulation overall survival at different time points, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_survival_n_by_estimate_type("spop", jan26_os_n_data, jan26_cond_os_n_data, pfs_timepoints, endpoint = "os")
```

:::

::: {.panel-tabset}

### Sample-Based

#### All Patients

```{r}
#| label: tbl-os-n-summary-sample-all
#| tbl-cap: "Sample-based OS-n summary for all patients."

create_survival_n_table("sample", "all", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-os-n-summary-sample-first-line
#| tbl-cap: "Sample-based OS-n summary for first line patients."

create_survival_n_table("sample", "first_line", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-os-n-summary-sample-part-e
#| tbl-cap: "Sample-based OS-n summary for Part E patients."

create_survival_n_table("sample", "part_e", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

### Superpopulation

#### All Patients

```{r}
#| label: tbl-os-n-summary-spop-all
#| tbl-cap: "Superpopulation OS-n summary for all patients."

create_survival_n_table("spop", "all", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-os-n-summary-spop-first-line
#| tbl-cap: "Superpopulation OS-n summary for first line patients."

create_survival_n_table("spop", "first_line", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-os-n-summary-spop-part-e
#| tbl-cap: "Superpopulation OS-n summary for Part E patients."

create_survival_n_table("spop", "part_e", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

:::

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and median PFS results
- [Survival Analysis](survival-analysis.qmd) - PFS Kaplan-Meier curves
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff comparisons
- [Model Specification](../documentation/model-specification.qmd) - Endpoint definitions
:::
