---
title: "Overall Survival"
subtitle: "Kaplan-Meier survival curves, median OS, and OS at fixed timepoints"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")

# Load multi-DCO OS data for comparisons
quant_os_list <- list(
  apr25 = tar_read(all_tumor_ssls_trial_quant_os_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_trial_quant_os_ctdna_aug25),
  jan26 = jan26_quant_os_data
)
cond_quant_os_list <- list(
  apr25 = tar_read(all_tumor_ssls_cond_quant_os_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_cond_quant_os_ctdna_aug25),
  jan26 = jan26_cond_quant_os_data
)

# Observed KM data for median OS plot
obs_os_km_list <- list(
  apr25 = list(
    all      = tar_read(km_trial_os_ctdna_apr25),
    pdl1_naive = tar_read(km_trial_os_pdl1_naive_ctdna_apr25),
    pdl1     = tar_read(km_trial_os_pdl1_ctdna_apr25),
    naive    = tar_read(km_trial_os_naive_ctdna_apr25)
  ),
  aug25 = list(
    all      = tar_read(km_trial_os_ctdna_aug25),
    pdl1_naive = tar_read(km_trial_os_pdl1_naive_ctdna_aug25),
    pdl1     = tar_read(km_trial_os_pdl1_ctdna_aug25),
    naive    = tar_read(km_trial_os_naive_ctdna_aug25)
  ),
  jan26 = list(
    all      = jan26_obs_os_km_all,
    pdl1_naive = jan26_obs_os_km_pdl1_naive,
    pdl1     = jan26_obs_os_km_pdl1,
    naive    = jan26_obs_os_km_first_line
  )
)

os_n_list <- list(
  apr25 = tar_read(all_tumor_ssls_forecast_os_n_rvar_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_forecast_os_n_rvar_ctdna_aug25),
  jan26 = jan26_os_n_data
)
cond_os_n_list <- list(
  apr25 = tar_read(all_tumor_ssls_cond_forecast_os_n_rvar_ctdna_apr25),
  aug25 = tar_read(all_tumor_ssls_cond_forecast_os_n_rvar_ctdna_aug25),
  jan26 = jan26_cond_os_n_data
)
```

This page presents overall survival (OS) analysis results for the SCLC-01 trial using **DCO3 (January 2026)** data.

::: {.callout-note}
## OS Endpoint Definition
OS is measured from randomization to death from any cause. Patients alive at last follow-up are censored. Unlike PFS, OS does not include a progression component --- all events are deaths.
:::

::: {.callout-warning}
## Known Limitation: 3-State Model
These results were generated using the **3-state illness-death model** (alive → progressed → dead), which does not account for treatment dropout (Pattern E). The model's OS predictions are **systematically more pessimistic** than the observed Kaplan-Meier --- the observed OS falls above the 80% credible interval at most timepoints for both SCLC-01 and HISTORICAL.

This is because dropout patients (who tend to have worse prognosis) are forced into the standard 0→1→2 pathway, inflating the estimated 0→2 (direct death) and 1→2 (post-progression death) hazard rates for all patients. The **4-state model** (`karim/handle-pattern-e`, [issue #31](https://github.com/azu-oncology-rd/cds-dsi-pioneer-core/issues/31)) adds a dedicated dropout state (0→3) with its own post-dropout death hazard (3→2), which is expected to correct this bias.
:::

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints](clinical-endpoints.qmd) for ORR and median PFS results
- See [Survival Analysis](survival-analysis.qmd) for PFS Kaplan-Meier curves
- See [Clinical Endpoints Specification](../documentation/clinical-endpoints-specification.qmd) for endpoint definitions
:::

## SCLC-01 OS Forecasts {#sec-sclc-os-km}

::: {.panel-tabset}

### Sample-Based

Overall survival Kaplan-Meier curves for the SCLC-01 trial, comparing model predictions with observed data across cohorts.

```{r}
#| label: fig-os-km-sclc-sample
#| fig-cap: Sample-based overall survival Kaplan-Meier forecast for SCLC-01 trial.
#| fig-width: 7

plot_sclc_km_by_estimate_type(
  "sample",
  jan26_os_km_data,
  jan26_cond_os_km_data,
  jan26_obs_os_km_all,
  jan26_obs_os_km_first_line,
  all_analysis_data,
  endpoint = "os"
)
```

### Superpopulation

Overall survival Kaplan-Meier curves for the SCLC-01 trial. Superpopulation estimates represent the expected survival for the underlying population, not just the specific patients in the trial.

```{r}
#| label: fig-os-km-sclc-spop
#| fig-cap: Superpopulation overall survival Kaplan-Meier forecast for SCLC-01 trial.
#| fig-width: 7

plot_sclc_km_by_estimate_type(
  "spop",
  jan26_os_km_data,
  jan26_cond_os_km_data,
  jan26_obs_os_km_all,
  jan26_obs_os_km_first_line,
  all_analysis_data,
  endpoint = "os"
)
```

:::

## PDL1-Stratified OS {#sec-pdl1-os-km}

::: {.panel-tabset}

### Sample-Based

Overall survival curves stratified by PDL1 status and treatment line.

```{r}
#| label: fig-os-km-pdl1-sample
#| fig-cap: Sample-based PDL1-stratified overall survival forecast.
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km_by_estimate_type(
  "sample",
  cond_km_data = jan26_cond_os_km_data,
  obs_km_all = jan26_obs_os_km_pdl1,
  obs_km_first_line = jan26_obs_os_km_pdl1_naive,
  obs_km_part_e = jan26_obs_os_km_part_e_pdl1,
  endpoint = "os"
)
```

### Superpopulation

Overall survival curves stratified by PDL1 status and treatment line.

```{r}
#| label: fig-os-km-pdl1-spop
#| fig-cap: Superpopulation PDL1-stratified overall survival forecast.
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km_by_estimate_type(
  "spop",
  cond_km_data = jan26_cond_os_km_data,
  obs_km_all = jan26_obs_os_km_pdl1,
  obs_km_first_line = jan26_obs_os_km_pdl1_naive,
  obs_km_part_e = jan26_obs_os_km_part_e_pdl1,
  endpoint = "os"
)
```

:::

## Median Overall Survival {#sec-median-os}

Median OS represents the time at which 50% of patients have died.

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-median-os-sample
#| fig-cap: Sample-based median overall survival, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_median_survival_by_estimate_type("sample", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

### Superpopulation

```{r}
#| label: fig-median-os-spop
#| fig-cap: Superpopulation median overall survival, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_median_survival_by_estimate_type("spop", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

:::

::: {.panel-tabset}

### Sample-Based

#### All Patients

```{r}
#| label: tbl-median-os-summary-sample-all
#| tbl-cap: "Sample-based median OS summary for all patients."

create_median_survival_table("sample", "all", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-median-os-summary-sample-first-line
#| tbl-cap: "Sample-based median OS summary for first line patients."

create_median_survival_table("sample", "first_line", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-median-os-summary-sample-part-e
#| tbl-cap: "Sample-based median OS summary for Part E patients."

create_median_survival_table("sample", "part_e", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

### Superpopulation

#### All Patients

```{r}
#| label: tbl-median-os-summary-spop-all
#| tbl-cap: "Superpopulation median OS summary for all patients."

create_median_survival_table("spop", "all", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-median-os-summary-spop-first-line
#| tbl-cap: "Superpopulation median OS summary for first line patients."

create_median_survival_table("spop", "first_line", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-median-os-summary-spop-part-e
#| tbl-cap: "Superpopulation median OS summary for Part E patients."

create_median_survival_table("spop", "part_e", jan26_quant_os_data, jan26_cond_quant_os_data, endpoint = "os")
```

:::

## OS at Fixed Timepoints (OS-n) {#sec-os-n}

OS-n represents the probability of being alive at specific timepoints (e.g., 6, 9, 12, 18 months).

::: {.panel-tabset}

### Sample-Based

```{r}
#| label: fig-os-n-sample
#| fig-cap: Sample-based overall survival at different time points, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_survival_n_by_estimate_type("sample", jan26_os_n_data, jan26_cond_os_n_data, pfs_timepoints, endpoint = "os")
```

### Superpopulation

```{r}
#| label: fig-os-n-spop
#| fig-cap: Superpopulation overall survival at different time points, split by PDL1 status and treatment line for the SCLC-01 trial.

plot_survival_n_by_estimate_type("spop", jan26_os_n_data, jan26_cond_os_n_data, pfs_timepoints, endpoint = "os")
```

:::

::: {.panel-tabset}

### Sample-Based

#### All Patients

```{r}
#| label: tbl-os-n-summary-sample-all
#| tbl-cap: "Sample-based OS-n summary for all patients."

create_survival_n_table("sample", "all", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-os-n-summary-sample-first-line
#| tbl-cap: "Sample-based OS-n summary for first line patients."

create_survival_n_table("sample", "first_line", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-os-n-summary-sample-part-e
#| tbl-cap: "Sample-based OS-n summary for Part E patients."

create_survival_n_table("sample", "part_e", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

### Superpopulation

#### All Patients

```{r}
#| label: tbl-os-n-summary-spop-all
#| tbl-cap: "Superpopulation OS-n summary for all patients."

create_survival_n_table("spop", "all", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### First Line Patients

```{r}
#| label: tbl-os-n-summary-spop-first-line
#| tbl-cap: "Superpopulation OS-n summary for first line patients."

create_survival_n_table("spop", "first_line", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

#### Part E

```{r}
#| label: tbl-os-n-summary-spop-part-e
#| tbl-cap: "Superpopulation OS-n summary for Part E patients."

create_survival_n_table("spop", "part_e", jan26_os_n_data, jan26_cond_os_n_data, endpoint = "os")
```

:::

## OS DCO Comparisons {#sec-os-dco}

Comparing OS predictions across data cutoff dates tracks how model forecasts evolve as more follow-up data accumulates.

### Median OS by DCO {#sec-median-os-dco}

```{=html}
<style>
.toggle-switch {
  position: relative;
  display: inline-block;
  width: 50px;
  height: 24px;
  margin-right: 0.75rem;
}
.toggle-switch input {
  opacity: 0;
  width: 0;
  height: 0;
}
.toggle-slider {
  position: absolute;
  cursor: pointer;
  top: 0;
  left: 0;
  right: 0;
  bottom: 0;
  background-color: #ccc;
  transition: .4s;
  border-radius: 24px;
}
.toggle-slider:before {
  position: absolute;
  content: "";
  height: 18px;
  width: 18px;
  left: 3px;
  bottom: 3px;
  background-color: white;
  transition: .4s;
  border-radius: 50%;
}
input:checked + .toggle-slider {
  background-color: var(--az-navy);
}
input:checked + .toggle-slider:before {
  transform: translateX(26px);
}
#median-os-sample-with-obs, #median-os-spop-with-obs {
  display: none;
}
</style>

<div style="margin-bottom: 1rem;">
  <label style="display: inline-flex; align-items: center; cursor: pointer; font-weight: 500;">
    <span class="toggle-switch">
      <input type="checkbox" id="show-os-observed-toggle" onchange="toggleOsObservedKM(this.checked)">
      <span class="toggle-slider"></span>
    </span>
    Show Observed KM Estimates
  </label>
</div>
```

::: {.panel-tabset}

#### Sample-Based

::: {#median-os-sample-with-obs}
```{r}
#| label: fig-median-os-dco-sample-with-obs
#| fig-cap: Sample-based median overall survival (OS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("sample", quant_os_list, cond_quant_os_list, obs_os_km_list, show_observed = TRUE, endpoint = "os")
```
:::

::: {#median-os-sample-without-obs}
```{r}
#| label: fig-median-os-dco-sample-without-obs
#| fig-cap: Sample-based median overall survival (OS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs. Model forecasts only.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("sample", quant_os_list, cond_quant_os_list, obs_os_km_list, show_observed = FALSE, endpoint = "os")
```
:::

#### Superpopulation

::: {#median-os-spop-with-obs}
```{r}
#| label: fig-median-os-dco-spop-with-obs
#| fig-cap: Superpopulation median overall survival (OS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("spop", quant_os_list, cond_quant_os_list, obs_os_km_list, show_observed = TRUE, endpoint = "os")
```
:::

::: {#median-os-spop-without-obs}
```{r}
#| label: fig-median-os-dco-spop-without-obs
#| fig-cap: Superpopulation median overall survival (OS), split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs. Model forecasts only.
#| fig-width: 10
#| fig-height: 6

plot_median_surv_by_dco("spop", quant_os_list, cond_quant_os_list, obs_os_km_list, show_observed = FALSE, endpoint = "os")
```
:::

:::

```{=html}
<script>
function toggleOsObservedKM(show) {
  document.getElementById('median-os-sample-with-obs').style.display = show ? 'block' : 'none';
  document.getElementById('median-os-sample-without-obs').style.display = show ? 'none' : 'block';
  document.getElementById('median-os-spop-with-obs').style.display = show ? 'block' : 'none';
  document.getElementById('median-os-spop-without-obs').style.display = show ? 'none' : 'block';
}
</script>
```

::: {.callout-note}
## Interpreting the Plot
- **Diamond markers**: Model forecast with 90% credible interval
- **Circle markers**: Observed KM median with confidence interval (toggle above)
- **Arrows**: Indicate median OS not yet reached (right-censored)
:::

### OS-n by DCO {#sec-os-n-dco}

Comparing OS at fixed timepoints across data cutoffs.

::: {.panel-tabset}

#### Sample-Based

```{r}
#| label: fig-os-n-dco-sample
#| fig-cap: Sample-based overall survival (OS) at different time points, split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10

plot_surv_n_by_dco("sample", os_n_list, cond_os_n_list, pfs_timepoints, endpoint = "os")
```

#### Superpopulation

```{r}
#| label: fig-os-n-dco-spop
#| fig-cap: Superpopulation overall survival (OS) at different time points, split by PDL1 status and treatment line for the SCLC-01 trial, shown for all DCOs.
#| fig-width: 10

plot_surv_n_by_dco("spop", os_n_list, cond_os_n_list, pfs_timepoints, endpoint = "os")
```

:::

### Kaplan-Meier Curves by DCO {#sec-os-km-dco}

#### By Cohort

Kaplan-Meier curves for SCLC-01 showing All patients, First Line patients, and Part E patients across data cutoffs.

::: {.panel-tabset}

##### Sample-Based

```{r}
#| label: fig-os-km-dco-sample
#| fig-cap: Sample-based overall survival Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_sclc_km_by_dco(
  estimate_type = "sample",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_os_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_os_rvar_ctdna_aug25),
    jan26 = jan26_os_km_data
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_aug25),
    jan26 = jan26_cond_os_km_data
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_os_ctdna_apr25),
    aug25 = tar_read(km_trial_os_ctdna_aug25),
    jan26 = jan26_obs_os_km_all
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_os_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_os_naive_ctdna_aug25),
    jan26 = jan26_obs_os_km_first_line
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = all_analysis_data
  ),
  endpoint = "os"
)
```

##### Superpopulation

```{r}
#| label: fig-os-km-dco-spop
#| fig-cap: Superpopulation overall survival Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_sclc_km_by_dco(
  estimate_type = "spop",
  all_km_list = list(
    apr25 = tar_read(all_tumor_ssls_km_os_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_km_os_rvar_ctdna_aug25),
    jan26 = jan26_os_km_data
  ),
  cond_km_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_aug25),
    jan26 = jan26_cond_os_km_data
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_os_ctdna_apr25),
    aug25 = tar_read(km_trial_os_ctdna_aug25),
    jan26 = jan26_obs_os_km_all
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_os_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_os_naive_ctdna_aug25),
    jan26 = jan26_obs_os_km_first_line
  ),
  analysis_data_list = list(
    apr25 = tar_read(all_analysis_data_ctdna_apr25),
    aug25 = tar_read(all_analysis_data_ctdna_aug25),
    jan26 = all_analysis_data
  ),
  endpoint = "os"
)
```

:::

#### PDL1-Stratified

Kaplan-Meier curves stratified by PDL1 status (High vs Low) for All patients, First Line patients, and Part E patients across data cutoffs.

::: {.panel-tabset}

##### Sample-Based

```{r}
#| label: fig-os-km-dco-sample-pdl1
#| fig-cap: Sample-based PDL1-stratified overall survival Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_km_pdl1_by_dco(
  cond_km_data_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_aug25),
    jan26 = jan26_cond_os_km_data
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_os_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_os_pdl1_ctdna_aug25),
    jan26 = jan26_obs_os_km_pdl1
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_os_pdl1_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_os_pdl1_naive_ctdna_aug25),
    jan26 = jan26_obs_os_km_pdl1_naive
  ),
  obs_km_part_e_list = list(
    apr25 = tar_read(km_trial_os_part_e_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_os_part_e_pdl1_ctdna_aug25),
    jan26 = jan26_obs_os_km_part_e_pdl1
  ),
  estimate_type = "sample",
  endpoint = "os"
)
```

##### Superpopulation

```{r}
#| label: fig-os-km-dco-spop-pdl1
#| fig-cap: Superpopulation PDL1-stratified overall survival Kaplan-Meier forecasts for SCLC-01 across cohorts and data cutoffs.
#| column: body-outset
#| fig-width: 9
#| fig-height: 7

plot_km_pdl1_by_dco(
  cond_km_data_list = list(
    apr25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_apr25),
    aug25 = tar_read(all_tumor_ssls_cond_km_os_rvar_ctdna_aug25),
    jan26 = jan26_cond_os_km_data
  ),
  obs_km_all_list = list(
    apr25 = tar_read(km_trial_os_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_os_pdl1_ctdna_aug25),
    jan26 = jan26_obs_os_km_pdl1
  ),
  obs_km_first_line_list = list(
    apr25 = tar_read(km_trial_os_pdl1_naive_ctdna_apr25),
    aug25 = tar_read(km_trial_os_pdl1_naive_ctdna_aug25),
    jan26 = jan26_obs_os_km_pdl1_naive
  ),
  obs_km_part_e_list = list(
    apr25 = tar_read(km_trial_os_part_e_pdl1_ctdna_apr25),
    aug25 = tar_read(km_trial_os_part_e_pdl1_ctdna_aug25),
    jan26 = jan26_obs_os_km_part_e_pdl1
  ),
  estimate_type = "spop",
  endpoint = "os"
)
```

:::

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and median PFS results
- [Survival Analysis](survival-analysis.qmd) - PFS Kaplan-Meier curves
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff comparisons
- [Clinical Endpoints Specification](../documentation/clinical-endpoints-specification.qmd) - Endpoint definitions
:::
