---
title: "Parameter Inference"
subtitle: "Population and patient-level parameter estimates"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents the inferred parameter estimates from the Bayesian hierarchical model, comparing prior predictions with posterior estimates.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd#sec-tumor-dynamics) for mathematical details on the tumor dynamics model
- See [Architecture](../documentation/architecture.qmd#sec-hierarchical) for the hierarchical parameter structure
:::

## Population-Level Rates {#sec-pop-rates}

The model estimates population-level tumor regression and growth rates. These represent the central tendency of rates across all trials and patients.

```{r}
#| label: fig-pop-rates
#| fig-cap: Population level log growth/regression rates comparing prior and posterior distributions.

bind_rows(
  prior = tar_read(tumor_ssls_rates_rvar_prior_ctdna_aug),
  posterior = tar_read(tumor_ssls_rates_rvar_posterior_ctdna_aug),
  .id = "stage"
) |>
  plot_prior_post_dens(.value, normalize = "groups") +
  labs(x = "") +
  facet_wrap(vars(.variable), scales = "free", labeller = labeller(.variable = \(l) str_remove(l, "log_"))) +
  NULL
```

::: {.callout-note}
## Interpretation
- **Regression rate**: Controls how quickly responding tumor cells decrease
- **Growth rate**: Controls how quickly resistant tumor cells grow
- The posterior concentrates around values consistent with observed tumor dynamics
:::

## Patient-Level Rates {#sec-patient-rates}

Individual patients have their own rates that deviate from the population mean. This plot shows the distribution of patient-level rates across all patients.

```{r}
#| label: fig-patient-rates-distogram
#| fig-cap: Distribution of patient-level posterior rate histograms showing the spread of individual patient rates.

tar_read(all_tumor_ssls_patient_rates_bpi_ctdna_aug) |>
  filter(fct_match(.variable, c("patient_log_decrease_rate_residual", "patient_log_growth_rate_residual")),  !is.na(.value)) |>
  plot_level_rates()
```

## Measurement Noise {#sec-noise}

The observation model includes measurement error representing variability in SLD measurements not explained by the underlying tumor dynamics.

```{r}
#| label: fig-noise-sd
#| fig-cap: "Measurement noise SD (observation error) comparing prior and posterior distributions."

# Note: Process noise parameters (pop_process_sd) were removed in AR(1) refactor
# When process noise is enabled, add AR(1) parameters:
# - tr_log_sd_pop_process_noise[1], tr_logit_phi_pop_process_noise[1] (patient-level)
# - tr_log_sd_pop_process_noise_pop[1], tr_logit_phi_pop_process_noise_pop[1] (pop-level)

tar_read(all_tumor_ssls_noise_sd_rvar_ctdna_aug) |>
  plot_prior_post_hist(normalize = "groups") +
  labs(x = "Measurement SD") +
  NULL
```

## Initial State {#sec-initial-state}

The initial proportion of responsive vs resistant tumor cells varies across patients and is a key determinant of tumor dynamics.

```{r}
#| label: fig-initial-state
#| fig-cap: Distribution of prior and posterior histograms of patient-level initial proportion of responsive tumors.

tar_read(all_tumor_ssls_patient_decrease_prop_bpi_ctdna_aug) |>
  filter(fct_match(.variable, "init_log_decrease_patient"), !is.na(.rs_value)) |>
  plot_level_decrease_prop() +
  coord_cartesian(xlim = c(0, 1)) +
  labs(subtitle = "Dispersion of patient-level proportion of tumors that are responsive to treatment.") +
  NULL
```

## Covariate Effects {#sec-covariates}

The model includes covariate effects on multiple model components. These coefficients show how baseline characteristics influence tumor dynamics.

```{r}
#| label: fig-pop-covar-coef
#| fig-cap: Prior predicted and posterior linear models for covariates.
#| fig-height: 6

tar_read(all_tumor_ssls_coef_ctdna_aug) |>
  filter(!fct_match(.variable, "oe_tumor_coef_pop")) |>
  left_join(all_stan_data |> pluck("covar_design_matrix") |> colnames() |> enframe(), by = c("n" = "name")) |>
  plot_ssls_coef(value, color = fit_type) +
  scale_color_discrete("", type = AZ_palette, label = str_to_title) +
  labs(x = "", y = "Covariate") +
  facet_wrap(vars(.variable),
             scales = "free_x",
             labeller = labeller(.variable = c("init_coef_qr_pop" = "Proportion of responsive tumors",
                                               "frac_coef_qr_pop" = "Proportion of rates total that is decreasing",
                                               "oe_covar_coef_qr_pop" = "Residual events hazard rate"))) +
  theme(legend.position = "bottom")
```

## Tumor Effects on Residual Events {#sec-tumor-effects}

The residual events model (non-target progression and death) includes effects of tumor state variables.

```{r}
#| label: fig-pop-tumor-coef
#| fig-cap: Prior predicted and posterior linear models for tumor size effects on the residual events model.
#| fig-height: 6

tar_read(all_tumor_ssls_coef_ctdna_aug) |>
  filter(fct_match(.variable, "oe_tumor_coef_pop")) |>
  mutate(n = factor(n, levels = 1:3, labels = c("SLD", "Decrease Rate", "Growth Rate"))) |>
  plot_ssls_coef(color = fit_type) +
  scale_y_discrete() +
  scale_color_discrete("", type = AZ_palette, label = str_to_title) +
  labs(x = "", y = "Tumor Covariate") +
  theme(legend.position = "bottom")
```

::: {.related-pages}
## Related Pages

- [Model Specification](../documentation/model-specification.qmd) - Mathematical details
- [Architecture](../documentation/architecture.qmd#sec-hierarchical) - Hierarchical structure
- [Patient Dynamics](patient-dynamics.qmd) - Individual patient trajectories
- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and PFS results
:::
