---
title: "Patient Dynamics"
subtitle: "Individual tumor trajectories and RECIST predictions"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents individual patient tumor dynamics, including SLD forecasts and RECIST response predictions.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd#sec-tumor-dynamics) for the bi-exponential tumor model
- See [Data Exploration](data-exploration.qmd) for observed tumor dynamics
:::

## SLD Forecasts - Censored Patients {#sec-sld-forecast-censored}

Sum of lesion diameters (SLD) forecasts for SCLC-01 patients who are right censored (progression-free at data cutoff).

```{r}
#| label: fig-forecast-sld
#| fig-cap: Forecast measured SLD, for SCLC-01 patients that are right censored.
#| column: body-outset
#| fig-width: 7
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_sld_rvar_right_censored_posterior_ctdna_aug) |>
  filter(fct_match(trial, "sclc")) |>
  mutate(patient_sld = patient_sld * 10) |>
  plot_dynamics(patient_sld) +
  scale_x_continuous("Months", breaks = months_to_weeks(seq(0, 48, 12)), label = label_weeks_to_months) +
  labs(y = "SLD [mm]") +
  theme(legend.position = "bottom", strip.text = element_blank()) +
  NULL
```

## Posterior Predictive Checks - Uncensored Patients {#sec-ppc}

Posterior checking plots for patients with observed progression events, comparing model predictions with actual SLD measurements.

### HISTORICAL Trial

```{r}
#| label: fig-forecast-sld-uncensored-historical
#| fig-cap: Forecast measured SLD, for HISTORICAL patients that are uncensored (posterior checking).
#| column: body-outset
#| fig-width: 7
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_aug) |>
  filter(fct_match(trial, "historical"), fct_match(stage, "obs")) |>
  mutate(patient_sld = patient_sld * 10) |>
  plot_dynamics(patient_sld) +
  scale_fill_manual("Stage",
                    values = c("obs" = AZ_palette[2]),
                    labels = c("obs" = "Observed")) +
  scale_x_continuous("Months", breaks = months_to_weeks(seq(0, 48, 2)), label = label_weeks_to_months) +
  theme(legend.position = "bottom") +
  NULL
```

### SCLC Trial

```{r}
#| label: fig-forecast-sld-uncensored-sclc
#| fig-cap: Forecast measured SLD, for SCLC patients that are uncensored (posterior checking).
#| column: body-outset
#| fig-width: 7
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_sld_rvar_uncensored_posterior_ctdna_aug) |>
  filter(fct_match(trial, "sclc"), fct_match(stage, "obs")) |>
  mutate(patient_sld = patient_sld * 10) |>
  plot_dynamics(patient_sld) +
  scale_fill_manual("Stage",
                    values = c("obs" = AZ_palette[2]),
                    labels = c("obs" = "Observed")) +
  scale_x_continuous("Months", breaks = months_to_weeks(seq(0, 48, 2)), label = label_weeks_to_months) +
  theme(legend.position = "bottom") +
  NULL
```

## RECIST Predictions {#sec-recist}

Predicted RECIST categories over time for censored patients, showing the model's forecast of future tumor response classifications.

```{r}
#| label: fig-predict-recist
#| fig-cap: Predicted and forecast RECIST categories at every week, for SCLC-01 patients who are right censored.
#| column: body-outset
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_recist_rvar_right_censored_posterior_ctdna_aug) |>
  filter(fct_match(trial, "sclc")) |>
  plot_recist_predictions(x_breaks = months_to_weeks(seq(0, 60, 12)))
```

::: {.callout-note}
## RECIST Categories
- **CR** (Complete Response): Disappearance of all target lesions
- **PR** (Partial Response): ≥30% decrease in SLD from baseline
- **SD** (Stable Disease): Neither PR nor PD criteria met
- **PD** (Progressive Disease): ≥20% increase in SLD from nadir plus ≥5mm absolute increase
:::

## PFS Posterior Predictive Check {#sec-pfs-ppc}

Comparing posterior PFS predictions with observed PFS for uncensored patients to assess model calibration.

```{r}
#| label: fig-pfs-posterior-check
#| fig-cap: Posterior checking the posterior PFS predictions compared to the observed PFS, restricted to the uncensored patients.
#| column: body-outset
#| fig-width: 10

ppc_spop_pfs <- tar_read(all_tumor_ssls_pfs_rvar_ctdna_aug) |>
  filter(fct_match(fit_type, "posterior"), !right_censored) |>
  mutate(event_type = case_when(
    !progression_before_death ~ "Death",
    !target_only_right_censored ~ "Target Lesion Progression",
    TRUE ~ "Residual Progression"
  ))

ppc_spop_pfs |>
  plot_pfs_ppc(pfs_var = spop_pfs)
```

::: {.related-pages}
## Related Pages

- [Data Exploration](data-exploration.qmd) - Observed tumor dynamics
- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and PFS results
- [Survival Analysis](survival-analysis.qmd) - Kaplan-Meier curves
- [Model Specification](../documentation/model-specification.qmd#sec-recist) - RECIST criteria
:::
