---
title: "Patient Dynamics"
subtitle: "Individual tumor trajectories and RECIST predictions"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents individual patient tumor dynamics, including SLD forecasts and RECIST response predictions.

::: {.callout-note}
## Data Cutoffs
This page shows results across all data cutoffs: **April 2025**, **August 2025**, and **January 2026 (DCO3)**. Use the tabs to switch between different data cutoffs.
:::

::: {.callout-tip}
## Related Pages
- See [Tumor Dynamics Specification](../documentation/tumor-dynamics-specification.qmd#sec-tumor-dynamics) for the bi-exponential tumor model
- See [Data Exploration](data-exploration.qmd) for observed tumor dynamics
:::

## SLD Forecasts - Censored Patients {#sec-sld-forecast-censored}

Sum of lesion diameters (SLD) forecasts for SCLC-01 patients who are right censored (progression-free at data cutoff).

::: {.panel-tabset}

### April 2025

```{r}
#| label: fig-forecast-sld-apr25
#| fig-cap: Forecast measured SLD for SCLC-01 right censored patients (April 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_sld_rvar_right_censored_posterior_ctdna_apr25) |>
  filter(fct_match(trial, "sclc")) |>
  plot_sld_forecast_censored()
```

### August 2025

```{r}
#| label: fig-forecast-sld-aug25
#| fig-cap: Forecast measured SLD for SCLC-01 right censored patients (August 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_sld_rvar_right_censored_posterior_ctdna_aug25) |>
  filter(fct_match(trial, "sclc")) |>
  plot_sld_forecast_censored()
```

### January 2026

```{r}
#| label: fig-forecast-sld-jan26
#| fig-cap: Forecast measured SLD for SCLC-01 right censored patients (January 2026 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_sld_rvar_right_censored_posterior_ctdna_jan26) |>
  filter(fct_match(trial, "sclc")) |>
  plot_sld_forecast_censored()
```

:::

## RECIST Predictions {#sec-recist}

Predicted RECIST categories over time for censored patients, showing the model's forecast of future tumor response classifications.

::: {.panel-tabset}

### April 2025

```{r}
#| label: fig-predict-recist-apr25
#| fig-cap: RECIST predictions for SCLC-01 right censored patients (April 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_recist_rvar_right_censored_posterior_ctdna_apr25) |>
  filter(fct_match(trial, "sclc")) |>
  plot_recist_censored()
```

### August 2025

```{r}
#| label: fig-predict-recist-aug25
#| fig-cap: RECIST predictions for SCLC-01 right censored patients (August 2025 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_recist_rvar_right_censored_posterior_ctdna_aug25) |>
  filter(fct_match(trial, "sclc")) |>
  plot_recist_censored()
```

### January 2026

```{r}
#| label: fig-predict-recist-jan26
#| fig-cap: RECIST predictions for SCLC-01 right censored patients (January 2026 DCO).
#| fig-width: 10
#| fig-height: 12
#| warning: false

tar_read(tumor_ssls_staged_recist_rvar_right_censored_posterior_ctdna_jan26) |>
  filter(fct_match(trial, "sclc")) |>
  plot_recist_censored()
```

:::

::: {.callout-note}
## RECIST Categories
- **CR** (Complete Response): Disappearance of all target lesions
- **PR** (Partial Response): ≥30% decrease in SLD from baseline
- **SD** (Stable Disease): Neither PR nor PD criteria met
- **PD** (Progressive Disease): ≥20% increase in SLD from nadir plus ≥5mm absolute increase
:::

::: {.related-pages}
## Related Pages

- [Data Exploration](data-exploration.qmd) - Observed tumor dynamics
- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and PFS results
- [Survival Analysis](survival-analysis.qmd) - Kaplan-Meier curves
- [Tumor Dynamics Specification](../documentation/tumor-dynamics-specification.qmd#sec-recist) - RECIST criteria
:::
