---
title: "Survival Analysis"
subtitle: "Kaplan-Meier survival curves and PFS forecasts"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")

# PFS KM data
all_analysis_data    <- tar_read(all_analysis_data_ctdna_jan26)
jan26_km_data        <- tar_read(all_tumor_ssls_km_rvar_ctdna_jan26)
jan26_cond_km_data   <- tar_read(all_tumor_ssls_cond_km_rvar_ctdna_jan26)
jan26_obs_km_all     <- tar_read(km_trial_pfs_ctdna_jan26)
jan26_obs_km_first_line   <- tar_read(km_trial_pfs_naive_ctdna_jan26)
jan26_obs_km_pdl1         <- tar_read(km_trial_pfs_pdl1_ctdna_jan26)
jan26_obs_km_pdl1_naive   <- tar_read(km_trial_pfs_pdl1_naive_ctdna_jan26)
jan26_obs_km_part_e_pdl1  <- tar_read(km_trial_pfs_part_e_pdl1_ctdna_jan26)
```

This page presents Kaplan-Meier survival analysis results including historical trial comparisons and SCLC forecasts using **DCO3 (January 2026)** data.

::: {.callout-tip}
## DCO Comparisons
To see how results evolve across data cutoff dates, visit [DCO Comparisons](../comparisons/dco-comparisons.qmd).
:::

::: {.callout-tip}
## Related Pages
- See [Clinical Endpoints Specification](../documentation/clinical-endpoints-specification.qmd#sec-kaplan-meier) for KM estimation details
- See [Clinical Endpoints](clinical-endpoints.qmd) for median PFS estimates
:::

## Historical Trials {#sec-historical-km}

Kaplan-Meier curves for the historical trials (excluding SCLC-01) used to inform the hierarchical model.

```{r}
#| label: fig-kaplan-meier-historical
#| fig-cap: Kaplan-Meier plots for the historical trials. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.
#| column: body-outset
#| fig-width: 7

plot_km(
  tar_read(all_tumor_ssls_km_rvar_ctdna_jan26) |> filter(!fct_match(trial, "sclc")),
  km_est_var = sample_km_est,
  obs_km_data = tar_read(km_trial_pfs_ctdna_jan26) |>
    filter(fct_match(btype, "ub"), !fct_match(trial, "sclc")) |>
    select(!quantiles) |>
    unnest(km_data),
  analysis_data = all_analysis_data |> filter(!fct_match(trial, "sclc")),
  facet_col = trial
)
```

## SCLC-01 Forecasts {#sec-sclc-km}

::: {.panel-tabset}

### Sample-Based

Kaplan-Meier curves for the SCLC-01 trial, comparing model predictions with observed data. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.

```{r}
#| label: fig-kaplan-meier-sclc-forecast
#| fig-cap: Sample-based Kaplan-Meier forecast for SCLC-01 trial.
#| fig-width: 7

plot_sclc_km_by_estimate_type(
  "sample",
  jan26_km_data,
  jan26_cond_km_data,
  jan26_obs_km_all,
  jan26_obs_km_first_line,
  all_analysis_data
)
```

### Superpopulation

Kaplan-Meier curves for the SCLC-01 trial, comparing model predictions with observed data. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.

```{r}
#| label: fig-kaplan-meier-sclc-superpopulation
#| fig-cap: Superpopulation Kaplan-Meier forecast for SCLC-01 trial. Superpopulation KM estimates represent the expected survival for the underlying population, not just the specific patients in the trial.
#| fig-width: 7

plot_sclc_km_by_estimate_type(
  "spop",
  jan26_km_data,
  jan26_cond_km_data,
  jan26_obs_km_all,
  jan26_obs_km_first_line,
  all_analysis_data
)
```

:::

## PDL1-Stratified Analysis {#sec-pdl1-km}

::: {.panel-tabset}

### Sample-Based

Kaplan-Meier curves stratified by PDL1 status and treatment line. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.

```{r}
#| label: fig-kaplan-meier-pdl1-forecast
#| fig-cap: Sample-based PDL1-stratified Kaplan-Meier forecast.
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km_by_estimate_type(
  "sample",
  cond_km_data = jan26_cond_km_data,
  obs_km_all = jan26_obs_km_pdl1,
  obs_km_first_line = jan26_obs_km_pdl1_naive,
  obs_km_part_e = jan26_obs_km_part_e_pdl1
)
```

### Superpopulation

Kaplan-Meier curves stratified by PDL1 status and treatment line. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.

```{r}
#| label: fig-kaplan-meier-pdl1-superpopulation
#| fig-cap: Superpopulation PDL1-stratified Kaplan-Meier forecast.
#| fig-width: 7
#| fig-height: 5

plot_pdl1_km_by_estimate_type(
  "spop",
  cond_km_data = jan26_cond_km_data,
  obs_km_all = jan26_obs_km_pdl1,
  obs_km_first_line = jan26_obs_km_pdl1_naive,
  obs_km_part_e = jan26_obs_km_part_e_pdl1
)
```

:::

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and median PFS results
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff comparisons
- [Model Validation](model-validation.qmd) - Cross-validation results
- [Clinical Endpoints Specification](../documentation/clinical-endpoints-specification.qmd#sec-kaplan-meier) - KM estimation
:::
