---
title: "Survival Analysis"
subtitle: "Kaplan-Meier survival curves and PFS forecasts"
---

```{r}
#| label: setup
#| include: false

source("_setup.R")
```

This page presents Kaplan-Meier survival analysis results including historical trial comparisons and SCLC forecasts.

::: {.callout-tip}
## Related Pages
- See [Model Specification](../documentation/model-specification.qmd#sec-kaplan-meier) for KM estimation details
- See [Clinical Endpoints](clinical-endpoints.qmd) for median PFS estimates
:::

## Historical Trials {#sec-historical-km}

Kaplan-Meier curves for the historical trials (excluding SCLC) used to inform the hierarchical model.

```{r}
#| label: fig-kaplan-meier-historical
#| fig-cap: Kaplan-Meier plots for the historical trials. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.
#| column: body-outset
#| fig-width: 10

plot_km(
  tar_read(all_tumor_ssls_km_rvar_ctdna_aug) |> filter(!fct_match(trial, "sclc")),
  km_est_var = sample_km_est,
  obs_km_data = tar_read(km_trial_pfs_ctdna_aug) |>
    filter(fct_match(btype, "ub"), !fct_match(trial, "sclc")) |>
    select(!quantiles) |>
    unnest(km_data),
  analysis_data = all_analysis_data |> filter(!fct_match(trial, "sclc")),
  facet_col = trial
)
```

## SCLC Forecasts {#sec-sclc-km}

### Sample-Based Forecast

Kaplan-Meier curves for the SCLC trial, comparing model predictions with observed data.

```{r}
#| label: fig-kaplan-meier-sclc-forecast
#| fig-cap: Forecast Kaplan-Meier plots for SCLC trial. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.
#| column: body-outset
#| fig-width: 10
#| fig-height: 6

plot_sclc_km <- partial(
  plot_km,
  bind_rows(
    apr = tar_read(all_tumor_ssls_km_rvar_ctdna_apr) |> filter(fct_match(trial, "sclc")),
    aug = tar_read(all_tumor_ssls_km_rvar_ctdna_aug) |> filter(fct_match(trial, "sclc")),
    .id = "dco"
  ),
  obs_km_data = bind_rows(
    apr = tar_read(km_trial_pfs_ctdna_apr) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    aug = tar_read(km_trial_pfs_ctdna_aug) |> filter(fct_match(btype, "ub"), fct_match(trial, "sclc")),
    .id = "dco"
  ) |>
    select(!quantiles) |>
    unnest(km_data),
  analysis_data = bind_rows(
    apr = tar_read(all_analysis_data_ctdna_apr) |> filter(fct_match(trial, "sclc")),
    aug = all_analysis_data |> filter(fct_match(trial, "sclc")),
    .id = "dco"
  ),
  facet_col = dco
)

plot_sclc_km(km_est_var = sample_km_est)
```

### Superpopulation Forecast

Superpopulation KM estimates represent the expected survival for the underlying population, not just the specific patients in the trial.

```{r}
#| label: fig-kaplan-meier-sclc-superpopulation
#| fig-cap: Superpopulation Kaplan-Meier plots for SCLC trial. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.
#| column: body-outset
#| fig-width: 10
#| fig-height: 6

plot_sclc_km(km_est_var = spop_km_est)
```

## PDL1-Stratified Analysis {#sec-pdl1-km}

### Sample-Based Forecast by PDL1 Status

```{r}
#| label: fig-kaplan-meier-pdl1-forecast
#| fig-cap: Forecast Kaplan-Meier plots. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.
#| column: body-outset
#| fig-width: 10
#| fig-height: 7

plot_sclc_pdl1_km <- partial(
  plot_pdl1_km,
  cond_km_data = bind_rows(
    apr = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_apr),
    aug = tar_read(all_tumor_ssls_cond_km_rvar_ctdna_aug),
    .id = "dco"
  ),
  obs_km_data = bind_rows(
    pdl1 = bind_rows(
      apr = tar_read(km_trial_pfs_pdl1_ctdna_apr),
      aug = tar_read(km_trial_pfs_pdl1_ctdna_aug),
      .id = "dco"
    ) |>
      filter(fct_match(trial, "sclc")) |>
      select(!quantiles) |>
      unnest(km_data),
    pdl1_naive = bind_rows(
      apr = tar_read(km_trial_pfs_pdl1_naive_ctdna_apr),
      aug = tar_read(km_trial_pfs_pdl1_naive_ctdna_aug),
      .id = "dco"
    ) |>
      select(!quantiles) |>
      unnest(km_data),
    .id = "variable"
  )
)

plot_sclc_pdl1_km(km_est_var = cond_sample_km_est)
```

### Superpopulation Forecast by PDL1 Status

```{r}
#| label: fig-kaplan-meier-pdl1-superpopulation
#| fig-cap: Superpopulation Kaplan-Meier plots. The full forecasting row predicts PFS beyond the observed right censoring while the observed censoring row censors predictions that exceed the observed right censoring.
#| column: body-outset
#| fig-width: 10
#| fig-height: 7

plot_sclc_pdl1_km(km_est_var = cond_spop_km_est)
```

## Cutoff Comparison {#sec-cutoff-km}

Comparison of observed and model-predicted Kaplan-Meier curves at different data cutoff dates.

```{r}
#| label: fig-cutoff-km-est
#| fig-cap: Comparing observed and model Kaplan-Meier curves for different data cutoffs
#| eval: false

cutoff_plot_dates <- c("2023-04-17", "2024-02-11", "2024-10-08", "2025-06-05")

# Create enrollment timeline plot
timeline_plot <- tar_read(all_analysis_data_ctdna_aug, store = lfo_store2) |>
  filter(fct_match(trial, "sclc")) |>
  distinct(usubjid, trial, trtsdt) |>
  arrange(trtsdt) |>
  mutate(patient_number = row_number()) |>
  ggplot(aes(x = trtsdt)) +
  geom_vline(aes(xintercept = as_date(value)), color = AZ_pink, data = enframe(cutoff_plot_dates)) +
  geom_histogram(binwidth = 30, fill = AZ_navy, alpha = 0.75, color = "white") +
  geom_label(
    aes(x = as_date(value), label = name),
    y = 12,
    data = enframe(cutoff_plot_dates),
    size = 4,
    color = "white",
    fill = AZ_pink,
    fontface = "bold",
    label.padding = unit(0.35, "lines"),
    label.r = unit(0.5, "lines")
  ) +
  geom_label(
    aes(x = as_date(value), label = format(as_date(value), "%Y-%m-%d")),
    y = 10.5,
    data = enframe(cutoff_plot_dates),
    angle = 90,
    hjust = 1,
    size = 3,
    color = AZ_pink,
    fontface = "bold"
  ) +
  scale_x_date("") +
  scale_y_continuous("Number of Patients Enrolled", breaks = seq(0, 12, 2)) +
  theme_minimal() +
  theme(legend.position = "top", panel.grid.minor = element_blank())

predicted_data <- tar_read(tumor_ssls_cutoff_km_est_ctdna_aug, store = lfo_store2) |>
  mutate(trial = factor(trial, levels = 1:2, labels = c("historical", "sclc"))) |>
  filter(fct_match(trial, "sclc"), cutoff_date %in% cutoff_plot_dates) |>
  mutate(
    trial = factor("sclc"),
    cutoff_number = base::match(cutoff_date, cutoff_plot_dates),
    facet_group = cutoff_date
  )

observed_data <- tar_read(cutoff_km_trial_pfs_ctdna_aug, store = lfo_store2) |>
  filter(fct_match(trial, "sclc"), fct_match(btype, "ub")) |>
  select(!c(quantiles, cutoff_date)) |>
  unnest(km_data) |>
  mutate(trial = factor("sclc")) |>
  right_join(select(predicted_data, t, cutoff_date, cutoff_calendar_day, trial, cutoff_number), by = c("trial", "t", "cutoff_calendar_day")) |>
  mutate(facet_group = cutoff_date)

observed_data <- observed_data |>
  nest(data = !facet_group) |>
  transmute(
    data = map2(data, facet_group, \(d, fg) bind_rows(d, filter(observed_data, cutoff_date < first(d$cutoff_date))) |> mutate(facet_group = fg))) |>
  unnest(data) |>
  mutate(primary = facet_group == cutoff_date)

predicted_data <- predicted_data |>
  nest(data = !facet_group) |>
  transmute(
    data = map2(data, facet_group, \(d, fg) bind_rows(d, filter(predicted_data, cutoff_date < first(d$cutoff_date))) |> mutate(facet_group = fg))) |>
  unnest(data) |>
  mutate(primary = facet_group == cutoff_date)

label_data <- tibble(
  facet_group = cutoff_plot_dates,
  cutoff_number = seq_along(cutoff_plot_dates)
)

pred_vs_obs_km_plot <- predicted_data |>
  ggplot(aes(x = t - 1)) +
  stat_lineribbon(
    aes(ydist = sample_km_est, group = factor(cutoff_date), fill = "Predicted"),
    alpha = 0.25, linewidth = 0.0, .width = 0.8) +
  geom_step(aes(y = s, group = factor(cutoff_date), color = "Observed", alpha = primary, size = primary), data = observed_data) +
  geom_label(
    aes(label = cutoff_number),
    x = months_to_weeks(13),
    y = 1,
    data = label_data,
    size = 4,
    color = "white",
    fill = AZ_pink,
    fontface = "bold",
    label.padding = unit(0.35, "lines"),
    label.r = unit(0.5, "lines")
  ) +
  scale_x_continuous("Months", breaks = months_to_weeks(seq(0, 60, 5)), label = label_weeks_to_months) +
  scale_y_continuous("Progression-Free Survival", breaks = seq(0, 1, 0.1)) +
  scale_color_manual(
    "", values = c(Observed = AZ_navy, Predicted = AZ_pink), labels = c(Predicted = "Forecast 80% CI"), aesthetics = c("color", "fill")) +
  scale_alpha_manual("", values = c(`TRUE` = 1, `FALSE` = 0.25), guide = "none") +
  scale_size_manual("", values = c(`TRUE` = 1.0, `FALSE` = 0.5), guide = "none") +
  facet_wrap(vars(facet_group), ncol = 4) +
  coord_cartesian(xlim = c(0, months_to_weeks(30)), ylim = c(0, 1)) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    strip.text = element_blank()
  ) +
  NULL

pred_vs_obs_km_plot / timeline_plot +
  plot_layout(heights = c(0.7, 0.3))
```

::: {.related-pages}
## Related Pages

- [Clinical Endpoints](clinical-endpoints.qmd) - ORR and median PFS results
- [DCO Comparisons](../comparisons/dco-comparisons.qmd) - Data cutoff comparisons
- [Model Validation](model-validation.qmd) - Cross-validation results
- [Model Specification](../documentation/model-specification.qmd#sec-kaplan-meier) - KM estimation
:::
