---
title: "Early Prediction of Breast Cancer Progress Free Survival"
author: Karim Naguib
code-fold: true
code-tools: true
table-of-contents: true
---

```{r}
#| label: setup
#| include: false

# library(conflicted)
library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(gtsummary)

# conflicts_prefer(
#   dplyr::filter
# )

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

options(mc.cores = 12)

ggplot2::theme_set(theme_minimal())

future::plan(future::multisession(workers = 12))

source(here("src", "util.R"))
```

# Exploratory Analysis Plan

## Question

Are there observable strong and deep early responses to antibody-drug conjugates (ADCs) that are predictive of long term progression free survival (PFS)? Are we able to reliably predict the heterogeneous response to ADCs using early measurements?

In this exploratory analysis we will not take up the causal question of whether these strong early responders would see similar PFS under counterfactual dosages or medications.[^1]

[^1]: This should probably revisited as the prediction within arm might not be the relevant question.

## Exploratory Analysis Context and Data

We will evaluate whether early (at baseline and at 6 and 12 weeks post treatment) measurements of tumor sizes are predictive of the PFS in response to Enhertu.

We will use data from the Breast01, Breast02, Breast03, and Breast04 trials, focusing on the treatment arms receiving the same dosage. Initially, we will be analyzing the different studies separately with no partial pooling since the inclusion criteria for the studies were different.

* Primary outcome: Post treatment progression free survival (interval and right censored).
* Predictive variables (at all the early intervals): 
    - Number of lesions observed. 
    - Organ of lesions. 
    - lesion sizes measured, with a maximum of ten of observed lesions and two per organ.

## Statistical Model Goals

-   Explore the heterogeneity in PFS conditional on early tumor size measures.
-   Consider more nuanced characteristics of a patient's tumors beyond the aggregated sum of largest diameters. The challenge here is how to model the nonlinear contribution of multiple influencers (tumors) on a central node's outcome (survival). A simple sum or mean ends up discarding a lot of information about the context of the disease.
-   Introduce a lesion level first-stage model. Considering the limit on the number of lesions measured, we potentially truncate some measures such as the total sum of tumor sizes. A model of tumor sizes would allow us to impute the missing data and propagate any uncertainty to the second-stage survival model.

# Data

```{r}
#| label: load-trials-data

# Trial arms
sdtm_ta <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_ta.parquet") |> 
  rename_with(str_to_lower)

# Protocol deviations
sdtm_dv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dv.parquet") |> 
  rename_with(str_to_lower)

# Inclusion/exclusion
sdtm_ie <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_ie.parquet") |> 
  rename_with(str_to_lower)

# Demographics 
sdtm_dm <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dm.parquet") |> 
  rename_with(str_to_lower)

# Death details 
sdtm_dd <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dd.parquet") |> 
  rename_with(str_to_lower)

# Tumor identification
sdtm_tu <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tu.parquet") |> 
  rename_with(str_to_lower)

# Trial visits 
sdtm_tv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tv.parquet") |> 
  rename_with(str_to_lower)

# Subject visits
sdtm_sv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_sv.parquet") |> 
  rename_with(str_to_lower)

# Tumor measurements/assessments
sdtm_tr <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tr.parquet") |>
# sdtm_tr <- read_rds("/scratch/ewfteams/dpo0083/sdtm/prod/data/deid_tr.rds") |> 
  rename_with(str_to_lower) |> 
  mutate(tracptfl = fct_match(tracptfl, "Y"))

# Disease response 
sdtm_rs <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_rs.parquet") |> 
  rename_with(str_to_lower)

# Tumor analysis_data
adam_adtr <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/adam/deid_adtr.parquet") |>
  rename_with(str_to_lower) |> 
  mutate(across(ends_with("fl"), \(fl) fct_match(fl, "Y")))

# Tumor analysis_data
adam_adef <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/adam/deid_adef.parquet") |>
  rename_with(str_to_lower) |> 
  mutate(across(ends_with("fl"), \(fl) fct_match(fl, "Y")))

```

Using SDTM and ADaM data from the trials, we prepare the analysis data.

```{r}
#| label: prepare-analysis-data
#| code-fold: false 

tumor_analysis_data <- adam_adtr |> 
  filter(
    paramcd == "DIAMETER",
    # There are a number of assessments that show up in the data but in SDTM they are marked as "NOT DONE"
    !is.na(trdy), !is.na(aval),
    tracptfl,
    adt <= trtedt, # Drop assessments occurring after treatment ended.
    aval > 0, # Drop all zero diameter assessments; these appear to be missing measurements. 
  ) %>% 
  select(studyid, usubjid, trlnkid, tuloc, tulat, trtsdt, trtedt, adt, ends_with("fl"), trdy, trtp, trta, visit, visitnum, mmdiam = aval) |> 
  mutate(
    week = if_else(trdy > 0, trdy - 1, trdy) %/% 7, 
    treated = adt >= trtsdt
  ) |> 
  arrange(usubjid, trlnkid, visitnum) |> 
  group_by(usubjid, trlnkid) |> 
  mutate(tuloc = first(tuloc), tulat = first(tulat) |> na_if("")) |> # Only the first assessment has the location of the tumor. Fix that.
  ungroup() |> 
  nest(tumor_history = c(adt, trdy, week, visit, visitnum, tsplitfl, tmergefl, mmdiam, treated)) |> 
  mutate(n_measures = map_int(tumor_history, nrow)) |> 
  pack(flags = ends_with("fl")) |> 
  nest(patient_tumors = !c(studyid, usubjid, trtsdt, trtedt, trtp, trta)) |> 
  rowwise() |> 
  mutate(
    patient_min_t = min(map_int(patient_tumors$tumor_history, \(h) min(h$week))),
    patient_max_t = max(map_int(patient_tumors$tumor_history, \(h) max(h$week))),
    patient_last_dt = max(do.call(c, map(patient_tumors$tumor_history, \(h) max(h$adt)))),
  ) |>
  ungroup() |> 
  mutate(
    patient_t_width = patient_max_t - patient_min_t + 1,
    n_tumors = map_int(patient_tumors, nrow),
  ) |> 
  # Let's get survival outcomes
  left_join(select(adam_adef, usubjid, bicrdt, bicrpfs), by = "usubjid") |> 
  mutate(
    death = fct_match(bicrpfs, "DEATH"),
    progress_week = time_length(bicrdt - trtsdt, unit = "days") %/% 7,
    death_week = if_else(death & bicrdt <= trtedt, progress_week, 0L), # Deaths after treatment ended will be treated as right censored
    
    # The model expects "pfs" to record the number of weeks that were progress free, so I need to find the last assessment week before
    # disease progress was detected.
    pfs = map2_int(progress_week, patient_tumors, function(progress_week, patient_tumors) {
      pre_progress_weeks <- unnest(patient_tumors, tumor_history) |> 
        distinct(week) |> 
        filter(week < progress_week) |> 
        pull(week) 
     
      # There are a few patients who just have a single post treatment visit 
      if (length(pre_progress_weeks) > 0) max(pre_progress_weeks) else NA_integer_
    }),
  )
```


```{r}
#| label: tumor-demographics-table 

sdtm_dm |> 
  tbl_summary(
    include = c("age", "sex", "race", "ethnic", "dthfl"), by = "arm"
  )
```

```{r}
#| label: tumor-history-plot 

tumor_analysis_data |>
  filter(patient_t_width > 1) |> 
  slice(1:20) |> 
  unnest(patient_tumors) |> 
  unnest(tumor_history) |> 
  ggplot() +
  geom_step(aes(week, mmdiam, group = trlnkid), direction = "hv") +
  facet_wrap(vars(usubjid), scale = "free") +
  labs(title = "Observed tumor sizes for a sample of patients") +
  theme(strip.text = element_blank()) +
  NULL
```

```{r}
#| label: sim-settings

n_patients <- 250
patient_measures <- 41 
max_pfs <- 50 
```

# Tumor Size Model

I haven't yet documented much about this model or done any kind of evaluation. I'm only using it at this time to generate fake tumor size data for testing.

We model tumor size as a nonlinear function with a Gaussian process prior (with common hyperparameters),

$$
X_{it} \sim  \mathtt{GP}(\mu^\textrm{tumor}, k(\cdot;\alpha^\textrm{tumor}, \rho^\textrm{tumor})) 
$$ 
where 
$$k(p, q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right).$$
```{r}
#| label: tumor-priors

tumor_priors <- lst(
  pop_tumor_gp_rho_alpha = 7.3,
  pop_tumor_gp_rho_beta = 7.5, 
)
```

## Prior Prediction

```{r}
#| label: tumor-prior-predict
#| warning: false

tumor_model <- cmdstan_model(here("src", "tumor.stan"))

t_offset <- 5

test_data <- lst(
  fit_tumor_data = FALSE, # Prior prediction
  gen_tumor_sizes = TRUE,
  predict_missing_sizes = FALSE,
  multilevel_patient = FALSE,
  multilevel_tumor = FALSE,
  
  n_patients,
  n_patient_tumors = 1 + rbinom(n_patients, 5, 0.4), # We're not yet generatively modeling the number of tumors
  n_measures = rep(patient_measures, sum(n_patient_tumors)),
  t_measure = rep(0:(patient_measures - 1), sum(n_patient_tumors)) - t_offset,
  tumor_size = rep(0, sum(n_measures)),

  !!!tumor_priors  
)

tumor_prior_res <- tumor_model$sample(data = test_data, iter_warmup = 500, iter_sampling = 500, refresh = 0)

tumor_prior_res$save_data_file(dir = file.path(tmp_dir, "data"), basename = "prior_tumor", timestamp = FALSE, random = FALSE)
  
res_rvar <- tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index]) |> 
  ungroup() |> 
  mutate( # A bunch of acrobatics to get the IDs, indices, and intervals right.
    patient_id = rep(1:n_patients, with(test_data, n_patient_tumors * patient_measures)),
    tumor_index = ((tumor_index - 1) %/% patient_measures) + 1,
    t = rep(0:(patient_measures - 1), sum(test_data$n_patient_tumors)) - t_offset
  ) |> 
  mutate(tumor_index = tumor_index - min(tumor_index) + 1, .by = patient_id) |> 
  rename(tumor_size = rep_tumor_size)

tumor_prior_param_rvar <- tumor_prior_res |> 
  gather_rvars(pop_tumor_gp_intercept, pop_tumor_gp_alpha, pop_tumor_gp_rho)

# This is the tumor data I'll be using in the PFS model
fake_tumor_data <- res_rvar |>
  unnest_rvars() |>
  ungroup() |> 
  filter(.draw == 1) 

# This will be used by SLURM jobs that need tumor data (e.g., PFS SBC run).
write_rds(fake_tumor_data, file.path(tmp_dir, "data", "fake_tumor.rds"))
```

```{r}
#| label: tumor-model-diagnose

tumor_prior_res$cmdstan_diagnose()
```

```{r}
#| label: tumor-prior-predict-plot
#| fig-cap: "Prior prediction checking for tumor model."

res_rvar |> 
  filter(patient_id == 1, tumor_index < 6) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = tumor_size, group = str_c(patient_id, tumor_index, .draw)), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 5)) +
  stat_lineribbon(aes(ydist = tumor_size, group = tumor_index), fill = "black", alpha = 0.25, .width = 0.8) +
  scale_x_continuous(breaks = seq(0, 40, 4)) +
  labs(title = "Prior predicted tumor sizes", y = "Tumor Size", caption = "Lines show tumors sizes for a subset of tumors of a single patient.") +
  facet_wrap(vars(tumor_index)) +
  coord_cartesian(ylim = c(0, 10)) +
  theme(strip.text = element_blank()) +
  NULL
```

## Fake Data Simulation

Getting a few fake datasets and assessing the model's ability to recover parameter values.

```{r}
#| label: tumor-fake-sim
#| output: false
#| cache: true

tumor_fake_res <- res_rvar |> 
  unnest_rvars() |> 
  filter(.draw <= 24) |> # We'll simulate parameters and data for a number simulations. 
  nest(fake_data = !c(.draw, .iteration, .chain)) |> 
  mutate(
    fake_fit = furrr::future_map2(.draw, fake_data, function(id, d) {
    # fake_fit = map2(.draw, fake_data, function(id, d) {
      test_data |> 
        list_modify(
          fit_data = TRUE,
          gen_tumor_sizes = FALSE,
          tumor_size = d$tumor_size
        ) |> 
        tumor_model$sample(
          iter_warmup = 200, iter_sampling = 200,
          parallel_chains = 4,
          output_dir = file.path(tmp_dir, "fit"), 
          output_basename = str_c("fake_tumor_", id), 
          init = \(chain_id) lst(
            pop_tumor_gp_rho = invgamma::rinvgamma(1, 7.3, 7.5),
            tumor_mean = rnorm(1, 2.8, 0.1),
            tumor_sd = abs(rnorm(1, 0, 1.25))
          ),
          max_treedepth = 15
        )
    # })
    }, .progress = FALSE, .options = furrr::furrr_options(seed = TRUE))
  ) 
```

```{r}
#| eval: false
#| include: false

f <- tumor_fake_res$fake_fit[[1]]
f_lp <- log_posterior(f)
f_np <- nuts_params(f)

mcmc_trace(f$draws(inc_warmup = TRUE), pars = "tumor_mean", n_warmup = 500) #, np = f_np)
```


```{r}
#| label: tumor-fake-data-params
#| fig-cap: "Fake data simulation assessing model's ability to recover posterior of model parameters."
#| cache: true
#| dependson: tumor-fake-sim

tumor_fake_res |> 
  group_by(.draw) |> 
  reframe(map_dfr(fake_fit, \(f) gather_rvars(f, pop_tumor_gp_intercept, pop_tumor_gp_alpha, pop_tumor_gp_rho))) |> 
  left_join(tumor_prior_param_rvar |> unnest_rvars(), by = c(".draw", ".variable"), suffix = c("", "_true")) |> 
  mutate(.value = .value - .value_true) |> 
  ggplot(aes(y = .draw)) +
  stat_pointinterval(aes(xdist = .value), .width = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Comparing posterior estimate of model parameters against known values", 
    x = "Difference from True Parameter Value", 
    y = "", 
    caption = str_glue("Showing {nrow(tumor_fake_res)} simulations.")
  ) +
  facet_wrap(vars(.variable), scales = "free") +
  theme(axis.text.y = element_blank()) +
  NULL  
```


## Fit Data

Now using actual trial data.

```{r}
#| label: tumor-fit
#| output: false
#| cache: true

prepare_tumor_stan_data <- function(analysis_data) {
  lst(
    n_patients = nrow(analysis_data),
    n_patient_tumors = analysis_data$n_tumors,
    n_measures = map(analysis_data$patient_tumors, \(tu) tu$n_measures) |> unlist(),
    t_measure = unnest(analysis_data, patient_tumors) |> pull(tumor_history) |> map(\(h) h$week) |> unlist(),
    tumor_size = unnest(analysis_data, patient_tumors) |> pull(tumor_history) |> map(\(h) h$mmdiam / 10) |> unlist(),
  )
}

treated_tumor_analysis_data <- tumor_analysis_data |> 
  filter(
    fct_match(trtp, "T-DXd"), 
    map_lgl(patient_tumors, \(tu) all(tu$n_measures >= 2)) # Drop patients that don't have two assessments per tumor
  )
  
tumor_stan_data <- lst(
  fit_tumor_data = TRUE,
  gen_tumor_sizes = TRUE,
  predict_missing_sizes = TRUE, 
  multilevel_patient = TRUE,
  multilevel_tumor = FALSE,
  
  !!!prepare_tumor_stan_data(treated_tumor_analysis_data),
  !!!tumor_priors
)

tumor_fit_res <- tumor_model$sample(
  tumor_stan_data, 
  iter_warmup = 500, iter_sampling = 500, 
  max_treedepth = 15,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "tumor" 
)
```

```{r}
#| label: fitted-tumor-data

all_tumor_size <- tumor_fit_res |> 
  spread_rvars(all_tumor_size[idx]) |> 
  mutate(
    usubjid = with(treated_tumor_analysis_data, rep(usubjid, patient_t_width * n_tumors)),
    trlnkid = rep(
      treated_tumor_analysis_data |> unnest(patient_tumors) |> pull(trlnkid), 
      with(treated_tumor_analysis_data, rep(patient_t_width, n_tumors)) 
    )
  ) |> 
  nest(tumor_history = c(idx, all_tumor_size)) |> 
  left_join(select(treated_tumor_analysis_data, usubjid, matches("patient_(min|max)_t"), patient_t_width), by = "usubjid") |> 
  mutate(
    tumor_history = pmap(lst(h = tumor_history, min_t = patient_min_t, max_t = patient_max_t), 
                         \(h, min_t, max_t) mutate(h, week = min_t:max_t))
  ) |> 
  nest(patient_tumors = c(trlnkid, tumor_history))
```


```{r}
#| label: fit-tumor-history-plot

all_tumor_size |> 
  slice(1:9) |> 
  unnest(patient_tumors) |> 
  unnest(tumor_history) |> 
  ggplot(aes(week)) +
  stat_lineribbon(aes(ydist = all_tumor_size, group = trlnkid), fill = "black", alpha = 0.25, .width = 0.8) +
  facet_wrap(vars(usubjid), scales = "free") +
  labs(title = "Showing GP tumor size dynamics with interpolation", x = "Week", y = "") +
  theme(strip.text = element_blank())
```


# Survival Model

We will use a piecewise-constant proportional hazard model:

-   Time is discretized into $t \in \{0,\dots,T\}$, with $t = 0$ is the baseline interval.
-   $Y_{it}$ is an indicator for disease progression at interval $t$, with $Y_{i0} = 0$.
-   $S_i$ is the number of intervals observed without disease progression, \[ \Pr[S > s] = \prod\_{r=0}\^s \Pr[Y_{i,r+1} = 0 \mid Y_{i,r} = 0]. \]
-   $X_{it}$ are--possibly time-variant--covariates.
-   $C_i$ is an indicator of whether $i$ is right censored.
-   $M_{i} \in \mathbb{Z}^{+}$ is the difference between $S_i$ and the previous interval $i$ was observed progression free ($M_i = 1$ if $C_i = 1$).

The probability of remaining progression-free at $m \geq 1$ is 
$$
\alpha_t(x;\theta) = \Pr[Y_t = 0\mid Y_{t-1} = 0, X_t = \bf{x}] = \exp[-\lambda(t, \bf{x}; \boldsymbol{\theta})] 
$$

and the hazard function . 
$$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t \kappa(\bf{x}; \boldsymbol{\beta})
$$

We assume that censoring in conditionally independent of time to progression. We will use a linear form for $\kappa(\bf{x}; \boldsymbol{\beta}) = \exp(\bf{x}\boldsymbol{\beta})$.

Aside from whatever priors we use for $\boldsymbol{\beta}$, we will use a Gaussian process prior for $\lambda_t$, 
$$
\lambda_t \sim \mathtt{GP}(0, k),
$$ 
where $k(\lambda_p, \lambda_q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right)$.

Our likelihood, accounting for interval and right censoring, is 
$$
\mathcal{L}_i = \begin{cases}
  \prod_{t=1}^{S_i-M_i}\alpha_t(x_i;\theta)\sum_{r=S_i-M_i+1}^{S_i}\alpha_r(x_i;\theta)[1-\alpha_{r+1}(x_i;\theta)] & \textrm{if }C_i = 0 \\
  \prod_{t=1}^{S_i}\alpha_t(x_i;\theta) & \textrm{if }C_i = 1 
  \end{cases}
$$
```{r}
#| label: pfs-priors

pfs_priors <- lst(
  log_lambda_gp_intercept_mean = -3,
  log_lambda_gp_intercept_sd = 0.5,
  log_lambda_gp_alpha_sd = 0.5,
  log_lambda_gp_rho_alpha = 7.3,
  log_lambda_gp_rho_beta = 7.5, 
  tumor_stim_intercept_sd = 0.5,
  tumor_stim_coef_sd = c(0.25, 0.25, 0.125),
)
```


```{r}
#| label: pfs-model

pfs_model <- cmdstan_model(here("src", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)

pfs_test_data <- test_data |> 
  list_modify(
    fit_data = FALSE,
    gen_pfs = TRUE,
    gen_interval_censored = FALSE,
    early_tumors_only = TRUE,
    add_interaction_term = FALSE,
    ignore_interval_censoring = FALSE,
    no_tumor_stim = FALSE,
    
    pfs = rep(max_pfs, n_patients),
    death_week = rep(0, n_patients),
    tumor_size = fake_tumor_data$tumor_size,
    
    !!!pfs_priors,
    
    # !!!tumor_stan_data
    use_tumor_model = FALSE
  )  

t_measure_list <- with(pfs_test_data, list_measures(t_measure, n_measures, n_patient_tumors))  
max_t <- t_measure_list |> 
  map(unlist) |> 
  map_int(max)
first_t_measure <- t_measure_list |> map(first) |> unlist()
```

## Prior Prediction

```{r}
#| label: pfs-prior-predict
#| cache: true

pfs_prior_res <- pfs_model$sample(
  pfs_test_data, 
  refresh = 0, 
  iter_warmup = 500, iter_sampling = 500,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "pfs_prior" 
)

rep_pfs_rvar <- pfs_prior_res |> 
  spread_rvars(rep_pfs[patient_index], rep_right_censored[patient_index])

rep_pfs_param_rvar <- pfs_prior_res |> 
  spread_rvars(
    tumor_stim_intercept,
    tumor_stim_coef[t],
    log_lambda_gp_alpha, 
    log_lambda_gp_rho, 
    log_lambda_gp_intercept
  ) |> 
  pivot_wider(names_from = t, values_from = tumor_stim_coef, names_prefix = "tumor_stim_coef_") 

lambda_rvar <- pfs_prior_res |> 
  spread_rvars(log_lambda[t], base_pf_cond_prob[t], base_survival[t], one_tumor_survival[t]) |> 
  mutate(lambda = exp(log_lambda))

expected_pfs_rvar <- pfs_prior_res |> 
  gather_rvars(base_cond_expected_pfs, one_tumor_cond_expected_pfs, base_cond_median_pfs, one_tumor_cond_median_pfs) |> 
  mutate(
    stat_type = str_extract(.variable, "expected|median"),
    tumor_size = str_extract(.variable, "base|one_tumor")
  )
```

```{r}
#| label: lambda-plot
#| dependson: pfs-prior-predict

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(title = "Base hazard", x = "m", y = latex2exp::TeX("$\\lambda_t$"))
```

```{r}
#| label: base-pf-prob-plot
#| dependson: pfs-prior-predict

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = base_pf_cond_prob), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = base_pf_cond_prob, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(title = "Base conditional probability of remaining progress free", x = "t", y = "")
```

```{r}
#| label: pf-prob-plot
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  spread_rvars(disease_progress_prob[i]) |>
  mutate(
    patient_index = rep(seq(pfs_test_data$n_patients), each = max(max_t)), 
    t = rep(1:max(max_t), pfs_test_data$n_patients) 
  ) |> 
  filter(patient_index <= 6) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = disease_progress_prob), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = disease_progress_prob, group = str_c(patient_index, .draw)), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 20)) +
  labs(title = "Conditional probability of remaining progress free", x = "t", y = latex2exp::TeX("$\\alpha_t$")) +
  coord_cartesian(ylim = c(0, 1)) +
  facet_wrap(vars(patient_index)) +
  NULL
```

```{r}
#| label: base-surivival-plot
#| dependson: pfs-prior-predict

lambda_rvar |>
  select(t, base_survival, one_tumor_survival) |> 
  pivot_longer(!t, names_to = "survival_type", values_to = "survival") |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = survival, fill = survival_type, color = survival_type), step = TRUE, alpha = 0.25, .width = 0.8) +
  labs(title = "Base survival probability", x = "t", y = latex2exp::TeX("$S(t)$"))
```

```{r}
#| label: cond-expected-pfs-plot
#| dependson: pfs-prior-predict

expected_pfs_rvar |> 
  mutate(.variable = str_remove(.variable, "_cond_expected_pfs")) |> 
  ggplot(aes(y = tumor_size)) +
  stat_pointinterval(aes(xdist = .value, color = stat_type), .width = 0.8, position = "dodge") +
  labs(title = "Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "t", y = "")
```

```{r}
#| label: prior-kaplan-meier 
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  spread_rvars(km_est[m]) |> 
  ggplot(aes(m - 1)) +
  stat_lineribbon(aes(ydist = km_est), step = TRUE, fill = "black", alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", x = "t", y = latex2exp::TeX("$S(t)$")) +
  # scale_x_continuous("t", breaks = seq(0, 40, 2))
  NULL
```

## Fake Data Simulation

Here we're using a set of ad hoc survival model parameters and generating a number of datasets, then we compare the posterior distributions of these parameters against the known truth. We do this three times:

1.  Without any interval censoring
2.  With interval censoring but ignored by the model (observed PFS is assumed to be actual PFS)
3.  With interval censoring which is accounted for by the model

```{r}
#| label: fake-data-model
#| include: false

log_lambda <- pfs_model$functions$calc_gp_pred(seq(max_pfs) / 12, intercept = -3.0, alpha = 0.3, rho = 1.0, delta = 1e-9, eta = rnorm(max_pfs))

true_tumor_intercept <- 0.2
true_tumor_coef <- c(0.15, 0.05)

fake_data_sim_settings <- pfs_test_data |>
  list_modify(
    tumor_stim_intercept_sd = 0.25,
    tumor_stim_coef_sd = c(0.125, 0.125, 0.0625),
    
    fit_tumor_data = FALSE,
    gen_tumor_sizes = FALSE,
    predict_missing_sizes = FALSE, 
    multilevel_patient = FALSE,
    multilevel_tumor = FALSE,
  )

patient_interval_data <- gen_patient_interval_properties(
  fake_tumor_data, log_lambda, true_tumor_intercept, true_tumor_coef, fake_data_sim_settings
)

true_base_pf_cond_prob <- 1 - pfs_model$functions$calculate_progress_linear_prob(log_lambda, true_tumor_intercept, true_tumor_coef, t(c(0, 0)))
true_one_tumor_pf_cond_prob <- 1 - pfs_model$functions$calculate_progress_linear_prob(log_lambda, true_tumor_intercept, true_tumor_coef, t(c(1, 1)))

max_all_t <- max(max_t) #patient_measures - t_offset - 1

true_base_marginal_prob_res = pfs_model$functions$calculate_marginal_dp_prob(true_base_pf_cond_prob[1:max_all_t], max_all_t)  
true_one_tumor_marginal_prob_res = pfs_model$functions$calculate_marginal_dp_prob(true_one_tumor_pf_cond_prob[1:max_all_t], max_all_t)  

base_dp_prob_not_censored <- true_base_marginal_prob_res[[1]] / (1 - true_base_marginal_prob_res[[2]][max_all_t])
one_tumor_dp_prob_not_censored <- true_one_tumor_marginal_prob_res[[1]] / (1 - true_one_tumor_marginal_prob_res[[2]][max_all_t])

true_base_cond_expected_pfs = c(base_dp_prob_not_censored %*% seq(0, length.out = max_all_t))
true_one_tumor_cond_expected_pfs = c(one_tumor_dp_prob_not_censored %*% seq(0, length.out = max_all_t))
true_base_cond_median_pfs = pfs_model$functions$pfs_quantiles_from_prob(base_dp_prob_not_censored, 0.5) 
true_one_tumor_cond_median_pfs = pfs_model$functions$pfs_quantiles_from_prob(one_tumor_dp_prob_not_censored, 0.5)  

true_km <- patient_interval_data |> 
  group_by(patient_id) %>%
  # This really isn't the _true_ km; I'm taking random draws here give true parameters.
  summarize(pfs = pfs_model$functions$pfs_rng(progress_prob, 1:n()) |> list()) %>% 
  mutate(!!!set_names(list_transpose(.$pfs), c("interval_censored", "right_censored", "pfs", "actual_pfs"))) |>   
  reframe(km_est = pfs_model$functions$estimate_kaplan_meier(pfs, right_censored, max_all_t - 1)) |> 
  mutate(m = seq(n()))
```

```{r}
#| label: fake-data-sim
#| output: false
#| cache: true 

n_sim <- 12  
fake_data_sim_settings_with_ic <- drop_missing_measures(
  fake_data_sim_settings, c(-t_offset, seq(1 - t_offset, patient_measures - t_offset - 1, by = 4)), keep_only = TRUE
)

fake_data_sim <- bind_rows(
  interval_censoring = fit_simulations(
    n_sim, patient_interval_data, fake_data_sim_settings_with_ic, fit_basename = "fake_data_sim_ic", tmp_dir = tmp_dir
  ),

  interval_censoring_ignored = fit_simulations(
    n_sim, patient_interval_data, fake_data_sim_settings_with_ic, ignore_interval_censoring = TRUE, fit_basename = "fake_data_sim_ic_ignored", tmp_dir = tmp_dir
  ),

  no_interval_censoring = fit_simulations(
    n_sim, patient_interval_data, fake_data_sim_settings, fit_basename = "fake_data_sim", tmp_dir = tmp_dir
  ),
  
  .id = "sim_type"
)
```

```{r}
#| label: fake-data-params
#| fig-cap: "Comparing fake data simulation parameter posteriors against their known true values."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  transmute(
    sim_id,
    sim_type,
    
    sim_param_post = furrr::future_map(
      sim_fit,
      \(f) spread_rvars(f, log_lambda_gp_alpha, log_lambda_gp_rho, log_lambda_gp_intercept, tumor_stim_intercept, tumor_stim_coef[t]) |>
        pivot_wider(names_from = t, values_from = tumor_stim_coef, names_prefix = "tumor_stim_coef_")
    ),
  ) |> 
  unnest(sim_param_post) |> 
  # add_row(sim_id = 0, sim_type = "prior", rep_pfs_param_rvar) |>
  pivot_longer(!c(sim_id, sim_type), names_to = "param", values_to = "est") |>
  ggplot(aes(sim_id)) + 
  geom_hline(aes(yintercept = true_val), data = tribble(~ param,                   ~ true_val,
                                                        "log_lambda_gp_intercept", -3.0,
                                                        "log_lambda_gp_alpha",     0.3,
                                                        "log_lambda_gp_rho",       1.0,
                                                        "tumor_stim_intercept",    0.2,
                                                        "tumor_stim_coef_1",       0.15,
                                                        "tumor_stim_coef_2",       0.05)) +
  stat_pointinterval(aes(ydist = est, color = sim_type), size = 0.25, .width = 0.8, fatten_point = 0, position = "dodge") +
  scale_color_discrete("") +
  facet_wrap(vars(param), scales = "free_y") +
  labs(x = "", y = "", caption = "Black horizontal line represents the true parameter value.") +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

```{r}
#| label: fake-data-lambda
#| fig-cap: "Comparing fake data simulation base hazard posteriors against their known true values."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  mutate(sim_log_lambda_post = furrr::future_map(sim_fit, \(f) spread_rvars(f, log_lambda[m]))) |> 
  select(!sim_fit) |>
  unnest(sim_log_lambda_post) |> 
  # bind_rows(transmute(lambda_rvar, m = t, log_lambda, sim_id = 0, sim_type = "prior")) |> 
  mutate(lambda = exp(log_lambda)) |> 
  filter(m <= 20) |> 
  ggplot(aes(sim_id)) + 
  geom_hline(aes(yintercept = true_val), data = \(d) distinct(d, m) |> mutate(true_val = exp(log_lambda[m]))) +
  stat_pointinterval(
    aes(ydist = lambda, color = sim_type), 
    size = 0.5, .width = 0.8, fatten_point = 0, position = "dodge"
  ) +
  labs(x = "", y = "", caption = "Black horizontal line represents the true parameter value.") +
  facet_wrap(vars(m)) +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

```{r}
#| eval: false

fake_data_sim |>
  filter(sim_id <= 4) |> 
  mutate(sim_log_lambda_post = furrr::future_map(sim_fit, \(f) spread_rvars(f, log_lambda[t]))) |> 
  select(!sim_fit) |>
  unnest(sim_log_lambda_post) |>
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = log_lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = log_lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 10)) +
  geom_line(aes(y = log_lambda), color = "red", data = tibble(t = 1:max_all_t, log_lambda = log_lambda[1:max_all_t])) +
  facet_grid(vars(sim_type), vars(sim_id))
```


```{r}
#| label: fake-data-km
#| fig-cap: "Comparing the fake data Kaplan-Meier survival probability against the true probability."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  mutate(km_est = furrr::future_map(sim_fit, \(f) spread_rvars(f, km_est[m]))) |> 
  select(sim_id, sim_type, km_est) |> 
  filter(sim_id <= 6) |> 
  unnest(km_est) |> 
  ggplot(aes(m - 1)) +
  # geom_line(aes(y = km_est), data = \(sim) semi_join(true_km, sim, by = "m")) +
  # This comes from a draw so not really reliable as the truth.
  stat_lineribbon(aes(ydist = km_est, color = sim_type, fill = sim_type), step = TRUE, alpha = 0.25, .width = 0.8) +
  scale_x_continuous("t", breaks = seq(0, 40, 4)) +
  labs(y = "S(t)", caption = "Black line represents the true survival probability.") +
  facet_wrap(vars(sim_id)) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fake-data-cond-expected-pfs-plot
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |>
  mutate(
    cond_stat = furrr::future_map(
      sim_fit, \(f) gather_rvars(f, base_cond_expected_pfs, one_tumor_cond_expected_pfs, base_cond_median_pfs, one_tumor_cond_median_pfs)
    )
  ) |> 
  unnest(cond_stat) |> 
  # add_row(sim_id = 0, sim_type = "prior", expected_pfs_rvar) |> 
  mutate(
    stat = str_extract(.variable, "expected|median"),
    .variable = str_remove(.variable, "_cond_(expected|median)_pfs"),
    est_type = if_else(sim_id > 0, "post", "prior")
  ) |> 
  ggplot(aes(y = sim_id)) +
  stat_pointinterval(aes(xdist = .value, color = sim_type), size = 0.5, .width = 0.8, position = "dodge") +
  geom_vline(
    aes(xintercept = true_e), 
    data = tibble(
      .variable = rep(c("base", "one_tumor"), 2), 
      stat = rep(c("expected", "median"), each = 2),
      true_e = c(true_base_cond_expected_pfs, true_one_tumor_cond_expected_pfs, true_base_cond_median_pfs, true_one_tumor_cond_median_pfs)
    )
  ) +
  scale_color_discrete("") +
  labs(title = "Fake Data Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "m", y = "") +
  # facet_wrap(vars(.variable)) +
  facet_grid(vars(stat), vars(.variable)) +
  NULL
```

## Simulation-Based Calibration

With SBC we use the prior to generate a number of datasets that we use the same model and prior to fit. We then calculate the number of samples that are less than the true value (from the prior-only run generating the data). If the model and algorithm are working correctly, the distribution of the number draws that are less than the true parameter values should approach a uniform distribution.

The code for this simulation is located in `sbc.R` (and `sbc.sh` to run on SLURM).

```{r}
#| label: sbc

sbc_draws <- 1000
sbc_bw <- 10

sbc_data <- bind_rows( 
  no_ic = read_rds(file.path(tmp_dir, "data", "sbc.rds")),
  ic = read_rds(file.path(tmp_dir, "data", "sbc_ic.rds")),
  ic_ignore = read_rds(file.path(tmp_dir, "data", "sbc_ic_ignore.rds")),
  .id = "sim_type") |> 
  select(sim_type, starts_with("r_"))

sbc_data |> 
  pivot_longer(!c(sim_type), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  filter(!str_detect(param, "pfs$")) |>
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = sbc_bw) +
  geom_rect(aes(xmin = -10, xmax = sbc_draws, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = count(sbc_data, sim_type) |> mutate(map_dfr(n, \(n) set_names(qbinom(c(0.01, 0.5, 0.99), n, sbc_bw / sbc_draws), c("lb", "med", "ub"))))) +
  facet_grid(rows = vars(sim_type), cols = vars(param)) +
  labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  coord_cartesian(ylim = c(0, 20))

sbc_data |> 
  pivot_longer(!c(sim_type), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  filter(str_detect(param, "pfs$")) |>
  mutate(
    stat = str_extract(param, "expected|median"),
    tumor_state = str_extract(param, "one_tumor|base")
  ) |> 
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = sbc_bw) +
  geom_rect(aes(xmin = -10, xmax = sbc_draws, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = count(sbc_data, sim_type) |> mutate(map_dfr(n, \(n) set_names(qbinom(c(0.01, 0.5, 0.99), n, sbc_bw / sbc_draws), c("lb", "med", "ub"))))) +
  facet_grid(rows = vars(sim_type), cols = vars(stat, tumor_state)) +
  labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  coord_cartesian(ylim = c(0, 30))

# sbc_data |> 
#   select(starts_with("r_")) |> 
#   pivot_longer(everything(), names_to = "param", names_prefix = "r_", values_to = "r") |> 
#   ggplot() +
#   stat_ecdf(aes(r)) +
#   facet_wrap(vars(param)) +
#   coord_cartesian(xlim = c(0, 1000))
```

It is clear that ignoring interval censoring degrades the model's ability to recover the prior parameter values; simulations show a lopsided distribution that significantly deviates from the uniform distribution.

## Fit Data

```{r}
#| label: fit-pfs-util

prepare_pfs_stan_data <- function(analysis_data, ...) {
    tumor_stan_data <- prepare_tumor_stan_data(analysis_data)
    pfs_data <- select(analysis_data, pfs, death_week)
  
    lst(
    fit_data = TRUE,
    use_tumor_model = FALSE,
    gen_pfs = TRUE,
    gen_interval_censored = FALSE,
    early_tumors_only = TRUE,
    add_interaction_term = TRUE,
    ignore_interval_censoring = FALSE,
    
    fit_tumor_data = FALSE,
    gen_tumor_sizes = FALSE,
    predict_missing_sizes = FALSE, 
    multilevel_patient = FALSE,
    multilevel_tumor = FALSE,
    
    !!!tumor_stan_data,
    !!!pfs_data,
    
    !!!pfs_priors,
    !!!tumor_priors,
    
    ...
  ) |> 
    list_modify(fit_tumor_data = FALSE)
}

pfs_analysis_data <- tumor_analysis_data |> 
  filter(
    !is.na(pfs), pfs > 0, # No missing pfs and progress must be assured to have happened after treatment
    patient_min_t <= 0, # At least one screening assessment needed
  )

treated_pfs_analysis_data <- pfs_analysis_data |> 
  filter(fct_match(trtp, "T-DXd"), map_lgl(patient_tumors, \(tu) all(tu$n_measures >= 2)))

obs_km <- with(
  prepare_pfs_stan_data(treated_pfs_analysis_data), 
  tibble(
    t = seq(0, max(t_measure)),
    s = pfs_model$functions$estimate_kaplan_meier(
      pfs,
      pfs_model$functions$identify_censoring(pfs, death_week, n_patient_tumors, n_measures, t_measure)[[2]],
      max(t_measure)
    )
  )
)
```

### No Tumors

The hazard function does not use any covariates: 

$$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t
$$

```{r}
#| label: no-tumor-pfs-fit
#| cache: true
#| output: false

no_tumor_pfs_stan_data <- treated_pfs_analysis_data |> 
  prepare_pfs_stan_data(no_tumor_stim = TRUE)

no_tumor_pfs_res <- pfs_model$sample(
  no_tumor_pfs_stan_data, 
  iter_warmup = 500, iter_sampling = 500,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "no_tumor_pfs" 
)

control_no_tumor_pfs_res <- pfs_model$sample(
  data = prepare_pfs_stan_data(filter(pfs_analysis_data, !fct_match(trtp, "T-DXd")), no_tumor_stim = TRUE), 
  iter_warmup = 500, iter_sampling = 500,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "control_no_tumor_pfs" 
)
```

```{r}
#| label: no-tumor_kaplan-meier
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res, control = control_no_tumor_pfs_res) |> 
  map_dfr(\(r) spread_rvars(r, km_est[m]), .id = "arm") |> 
  # filter(fct_match(arm, "treated")) |>  
  ggplot() +
  geom_step(aes(x = t, y = s), data = obs_km) +
  stat_lineribbon(aes(x = m - 1, ydist = km_est, fill = arm, color = arm), step = TRUE, alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", x = "Week", y = latex2exp::TeX("$S(t)$")) +
  NULL
```

```{r}
#| label: no-tumor-mean-median
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res, control = control_no_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), .id = "arm") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = arm), position = "dodge", .width = 0.8)
```

#### Looking at Interval Censoring

Running the model with and without adjustment for interval censoring.

```{r}
#| label: fit-ignore-interval-censoring
#| cache: true
#| dependson: no-tumor-pfs-fit
#| output: false

ic_ignored_no_tumor_pfs_res <- no_tumor_pfs_stan_data |>
  list_modify(ignore_interval_censoring = TRUE) |>
  pfs_model$sample(
    iter_warmup = 500, iter_sampling = 500,
    output_dir = file.path(tmp_dir, "fit"), 
    output_basename = "ic_ignored_no_tumor_pfs" 
  )
```

```{r}
#| label: no-tumor-kaplan-meier-ic-compare 
#| cache: true
#| dependson: fit-ignore-interval-censoring

lst(ic = no_tumor_pfs_res, ic_ignored = ic_ignored_no_tumor_pfs_res) |> 
  map_dfr(\(r) spread_rvars(r, km_est[m]), .id = "model") |> 
  ggplot(aes(m - 1)) +
  stat_lineribbon(aes(ydist = km_est, fill = model, color = model), step = TRUE, alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", y = latex2exp::TeX("$S(t)$")) +
  scale_fill_viridis_d(aesthetics = c("fill", "color")) +
  NULL
```

```{r}
#| label: no-tumor-mean-median-ic-compare 
#| cache: true
#| dependson: fit-ignore-interval-censoring

lst(ic = no_tumor_pfs_res, ic_ignored = ic_ignored_no_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), .id = "model") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = model), position = "dodge", .width = 0.8)
```

While SBC shows that ignoring interval censoring should be a big problem, it does not appear to bias estimates much using our data.

### First Two Tumor Sizes Only

The hazard function is time invariant:

$$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t \exp[\beta_0 + \beta_1 x_0 + \beta_2 x_1 + \beta_3 (x_0 \times x_1)]
$$
where $x_0$ is the last screening assessment and $x_1$ is the first assessment after treatment begins.

```{r}
#| label: two-tumor-pfs-fit
#| cache: true
#| output: false

two_tumor_init <- function(chain_id) { 
  lst(
    tumor_stim_coef = c(truncnorm::rtruncnorm(1, 0, mean = 0, sd = 0.1),
                        truncnorm::rtruncnorm(1, 0, mean = 0, sd = 0.1),
                        truncnorm::rtruncnorm(1, 0, mean = 0, sd = 0.05)))
}

two_tumor_pfs_future <- future::future(
  treated_pfs_analysis_data |> 
    prepare_pfs_stan_data(no_tumor_stim = FALSE) |>  
    list_modify(add_interaction_term = TRUE) |> 
    pfs_model$sample(
      iter_warmup = 500, iter_sampling = 500, 
      init = two_tumor_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "two_tumor_pfs" 
    ),
  seed = TRUE
)

control_two_tumor_pfs_future <- future::future(
  pfs_analysis_data |> 
    filter(!fct_match(trtp, "T-DXd"), map_lgl(patient_tumors, \(tu) all(tu$n_measures >= 2))) |> 
    prepare_pfs_stan_data(no_tumor_stim = FALSE) |>  
    list_modify(add_interaction_term = TRUE) |> 
    pfs_model$sample(
      iter_warmup = 500, iter_sampling = 500,
      init = two_tumor_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "control_two_tumor_pfs" 
    ),
  seed = TRUE
)

two_tumor_pfs_res <- future::value(two_tumor_pfs_future)
control_two_tumor_pfs_res <- future::value(control_two_tumor_pfs_future)
```

```{r}
#| label: two-tumor-kaplan-meier
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res, control = control_two_tumor_pfs_res) |> 
  map_dfr(\(r) spread_rvars(r, km_est[m]), .id = "arm") |> 
  ggplot(aes(m - 1)) +
  geom_step(aes(x = t, y = s), data = obs_km) +
  stat_lineribbon(aes(ydist = km_est, fill = arm, color = arm), step = TRUE, alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", y = latex2exp::TeX("$S(t)$")) +
  NULL
```

```{r}
#| label: two-tumor-mean-median
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res, control = control_two_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), .id = "arm") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = arm), position = "dodge", .width = 0.8)
```


```{r}
#| label: two-tumor-coef
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(prior = pfs_prior_res, post = two_tumor_pfs_res) |> 
  map_dfr(\(f) gather_rvars(f, tumor_stim_intercept, tumor_stim_coef[t]), .id = "fit") |> 
  mutate(.variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]"))) |> 
  ggplot() +
  stat_halfeye(aes(xdist = .value, y = .variable, fill = fit, color = fit), alpha = 0.25, .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  coord_cartesian(xlim = c(0, 1))
```

### Joint Model

### Cross Validation

Here we are calculating the LOO-PSIS cross validation approximations for the no tumors and time invariant two tumor assessments model.

```{r}
#| label: pfs-loo

no_tumor_loo <- no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
two_tumor_loo <- two_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)

loo_compare(lst(no_tumor_loo, two_tumor_loo))
```

We see that in terms of elpd the no tumors model is inferior but given the higher SE this difference does not appear to be very significant. 

Below we do the same thing but using a linear loss function and RMSE. Again, there is little difference between them. 

```{r}
#| label: loo-predict-other-utility 

linear_loss <- function(m, a, k0 = 1, k1 = 1) if_else(m >= a, k0, k1) * abs(m - a)

two_tumor_log_lik <- two_tumor_pfs_res$draws("log_lik", format = "matrix")
no_tumor_log_lik <- no_tumor_pfs_res$draws("log_lik", format = "matrix")
 
two_tumor_loo_pfs <- E_loo(
    two_tumor_pfs_res$draws("rep_pfs", format = "matrix"), two_tumor_loo$psis_object, type = "mean", log_ratios = -two_tumor_log_lik
  ) |>  
  pluck("value")

no_tumor_loo_pfs <- E_loo(
    no_tumor_pfs_res$draws("rep_pfs", format = "matrix"), no_tumor_loo$psis_object, type = "mean", log_ratios = -no_tumor_log_lik
  ) |>  
  pluck("value")

linear_loss(two_tumor_loo_pfs, treated_pfs_analysis_data$pfs, 2, 1) |> 
  mean()

linear_loss(no_tumor_loo_pfs, treated_pfs_analysis_data$pfs, 2, 1) |> 
  mean()

subtract(two_tumor_loo_pfs, treated_pfs_analysis_data$pfs) |> 
  magrittr::raise_to_power(2) |> 
  mean() |> 
  sqrt()

subtract(no_tumor_loo_pfs, treated_pfs_analysis_data$pfs) |> 
  magrittr::raise_to_power(2) |> 
  mean() |> 
  sqrt()
```

