---
title: "Early Prediction of Breast Cancer Survival"
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(rlang)
library(cmdstanr)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)

options(mc.cores = 12)

ggplot2::theme_set(theme_minimal())

future::plan(future::multisession(workers = 3))

source(here("src", "util.R"))
```

```{r}
#| label: sim-settings

n_patients <- 1000
patient_measures <- 10
max_pfs <- 20
```

# Tumor Size Model

```{r}
#| label: tumor

tumor_model <- cmdstan_model(here("src", "tumor.stan"))

test_data <- lst(
  fit_data = FALSE,
  gen_tumor_sizes = TRUE,
  n_patients,
  n_measures = rep(patient_measures, n_patients),
  t_measure = rep(1:patient_measures, n_patients),
  n_patient_tumors = 1 + rbinom(n_patients, 5, 0.4),
  tumor_size = array(dim = 0)
)

if (is_empty(dir(here("temp", "fit"), pattern = "prior_tumor", full.names = TRUE))) {
  res <- tumor_model$sample(data = test_data)
  
  res$save_output_files(
    here("temp", "fit"), basename = "prior_tumor", timestamp = FALSE, random = FALSE
  )
  
  res$save_data_file(
    here("temp", "data"), basename = "prior_tumor", timestamp = FALSE, random = FALSE
  )
} else {
  res <- dir(here("temp", "fit"), pattern = "prior_tumor", full.names = TRUE) |>
    as_cmdstan_fit()
  
  test_data <- rjson::fromJSON(file = here("temp", "data", "prior_tumor.json"))
}

res_rvar <- res |> 
  spread_rvars(rep_tumor_size[tumor_index]) |> 
  ungroup() |> 
  mutate(
    patient_id = rep(1:n_patients, with(test_data, n_patient_tumors * patient_measures)),
    tumor_index = ((tumor_index - 1) %/% 10) + 1,
    t = rep(1:patient_measures, sum(test_data$n_patient_tumors))
  ) |> 
  mutate(tumor_index = tumor_index - min(tumor_index) + 1, .by = patient_id) |> 
  rename(tumor_size = rep_tumor_size)

res_rvar |> 
  filter(patient_id == 1, tumor_index < 6) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = tumor_size, group = str_c(patient_id, tumor_index, .draw)), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  stat_lineribbon(aes(ydist = tumor_size, group = tumor_index), alpha = 0.25, .width = 0.8) +
  coord_cartesian(ylim = c(0, 10)) +
  facet_wrap(vars(tumor_index)) +
  NULL

fake_tumor_data <- res_rvar |>
  unnest_rvars() |>
  ungroup() |> 
  filter(.draw == 1) 

write_rds(fake_tumor_data, here("temp", "data", "fake_tumor.rds"))
```

# Survival Model

We will be using a piecewise-constant proportional hazard model:

-   Time is discretized into $m \in \{0,\dots,M\}$, with $m = 0$ is the baseline timepoint.
-   $Y_m$ is an indicator for disease progression at timepoint $m$, with $Y_0 = 0$.
-   $X_m$ are--possibly time-variant--covariates.

The probability of remaining progression-free at $m \geq 1$ is 
$$
\alpha_m(x;\theta) = \Pr[Y_m = 0\mid Y_{m-1} = 0, X_m = \bf{x}] = \exp[-\lambda(m, \bf{x}; \boldsymbol{\theta})] 
$$

and the hazard function

$$
  \lambda(m, \bf{x}; \boldsymbol{\theta}) = \lambda_m \kappa(\bf{x}; \boldsymbol{\beta})
$$

We assume that censoring in conditionally independent of time to progression. We will use a linear for for $\kappa(\bf{x}; \boldsymbol{\beta}) = \exp(\bf{x}\boldsymbol{\beta})$.

Aside from whatever priors we use for $\boldsymbol{\beta}$, we will use a Gaussian process prior for $\lambda_m$, $$
\lambda_m \sim \mathtt{GP}(0, k),
$$ where $k(\lambda_p, \lambda_q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right)$.

```{r}
#| label: pfs-model

pfs_model <- cmdstan_model(here("src", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)

pfs_test_data <- test_data |> 
  list_modify(
    gen_pfs = TRUE,
    early_tumors_only = TRUE,
    tumor_size = fake_tumor_data$tumor_size,
    pfs = rep(max_pfs, test_data$n_patients),
    right_censored = rep(1, test_data$n_patients),
    interval_censored = rep(0, test_data$n_patients),
    
    log_lambda_gp_intercept_mean = -3,
    log_lambda_gp_intercept_sd = 0.5,
    tumor_stim_intercept_sd = 0.5,
    tumor_stim_coef_sd = c(0.25, 0.25),
  )
```

## Prior Prediction

```{r}
#| label: pfs-prior-predict

pfs_res <- pfs_model$sample(data = pfs_test_data, refresh = 0)

rep_pfs_rvar <- pfs_res |> 
  spread_rvars(rep_pfs[patient_index], rep_right_censored[patient_index])

rep_pfs_param_rvar <- pfs_res |> 
  spread_rvars(log_lambda_gp_alpha, log_lambda_gp_rho, log_lambda_gp_intercept, tumor_stim_intercept, tumor_stim_coef[t]) |> 
  pivot_wider(names_from = t, values_from = tumor_stim_coef, names_prefix = "tumor_stim_coef_") 

lambda_rvar <- pfs_res |> 
  spread_rvars(log_lambda[t], base_pf_cond_prob[t], base_survival[t], one_tumor_survival[t]) |> 
  mutate(lambda = exp(log_lambda))

expected_pfs_rvar <- pfs_res |> 
  gather_rvars(base_cond_expected_pfs, one_tumor_cond_expected_pfs)  
```

```{r}
#| label: lambda-plot

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = lambda), fill = "black", alpha = 0.5, .width = 0.8) +
  geom_line(aes(y = lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(x = "m", y = latex2exp::TeX("$\\lambda_m$"))
```

```{r}
#| label: base-pf-prob-plot

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = base_pf_cond_prob), fill = "black", alpha = 0.5, .width = 0.8) +
  geom_line(aes(y = base_pf_cond_prob, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(title = "Base conditional probability of remaining progress free", x = "m", y = "")
```

```{r}
#| label: pf-prob-plot

pfs_res |> 
  spread_rvars(disease_progress_prob[i]) |> 
  mutate(patient_index = rep(seq(test_data$n_patients), each = max_pfs), t = rep(seq(max_pfs), n_patients)) |> 
  filter(patient_index <= 6) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = disease_progress_prob), fill = "black", alpha = 0.5, .width = 0.8) +
  geom_line(aes(y = disease_progress_prob, group = str_c(patient_index, .draw)), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 20)) +
  labs(title = "Conditional probability of remaining progress free", x = "m", y = latex2exp::TeX("$\\alpha_m$")) +
  coord_cartesian(ylim = c(0, 1)) +
  facet_wrap(vars(patient_index)) +
  NULL
```

```{r}
#| label: base-surivival-plot

lambda_rvar |>
  select(t, base_survival, one_tumor_survival) |> 
  pivot_longer(!t, names_to = "survival_type", values_to = "survival") |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = survival, fill = survival_type, color = survival_type), alpha = 0.25, .width = 0.8) +
  labs(title = "Base survival probability", x = "m", y = latex2exp::TeX("$S(m)$"))
```

```{r}
#| label: cond-expected-pfs-plot

expected_pfs_rvar |> 
  mutate(.variable = str_remove(.variable, "_cond_expected_pfs")) |> 
  ggplot(aes(y = .variable)) +
  stat_pointinterval(aes(xdist = .value), .width = 0.8) +
  labs(title = "Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "m", y = "")
```

```{r}
pfs_res |> 
  spread_rvars(km_est[m]) |> 
  ggplot(aes(m)) +
  stat_lineribbon(aes(ydist = km_est), fill = "black", alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", y = latex2exp::TeX("$S(m)$"))
```

## Fake Data Simulation

```{r}
#| label: fake-pfs-data
#| include: false
#| cache: true

# Using a fixed set of log lambda values, generated using:
#   pfs_model$functions$calc_gp_pred(seq(max_pfs), intercept = -2.0, alpha = 0.3, rho = 1.5, delta = 1e-9, eta = rnorm(max_pfs))

log_lambda <- c(-2.48777760709026, -2.57152726838213, -2.23380584287276, -1.90739893145911, -1.86259266390287, -1.92891255821976, -1.93801857664684, -1.78872336415524, -1.54522270753866, -1.48718504020436, -1.70706597143648, -1.82018710107277, -1.67764635563667, -1.69933003110133, -1.86937604603516, -2.03108481232034, -2.15728306810556, -2.11459406967666, -2.02311574428507, -2.08871876979315) 

fake_data_sim_settings <- pfs_test_data |>
  list_modify(
    log_lambda_gp_intercept_mean = -2,
    log_lambda_gp_intercept_sd = 0.5,
    tumor_stim_intercept_sd = 0.25,
    tumor_stim_coef_sd = c(0.125, 0.125),
  )

patient_interval_data <- gen_patient_interval_properties(fake_tumor_data, log_lambda, fake_data_sim_settings)

fake_data_sim <- tibble(sim_id = 1:12) |> 
  rowwise() |> 
  mutate(sim_data = list(gen_fake_pfs_data(patient_interval_data))) |> 
  ungroup() |> 
  mutate(
    sim_fit = furrr::future_map(.progress = TRUE, .options = furrr::furrr_options(seed = TRUE),
    # sim_fit = map(
      sim_data, \(sdata) fit_sim_data(fake_data_sim_settings, sdata)
    ), 
  ) 
```

```{r}
fake_data_sim |> 
  transmute(
    sim_id,
    sim_param_post = map(
      sim_fit,
      \(f) spread_rvars(f, log_lambda_gp_alpha, log_lambda_gp_rho, log_lambda_gp_intercept, tumor_stim_intercept, tumor_stim_coef[t]) |>
        pivot_wider(names_from = t, values_from = tumor_stim_coef, names_prefix = "tumor_stim_coef_")
    ),
  ) |> 
  unnest(sim_param_post) |> 
  add_row(sim_id = 0, rep_pfs_param_rvar) |>
  pivot_longer(!sim_id, names_to = "param", values_to = "est") |>
  mutate(est_type = if_else(sim_id > 0, "post", "prior")) |> 
  ggplot(aes(sim_id)) + 
  geom_hline(aes(yintercept = true_val), data = tribble(~ param,                   ~ true_val,
                                                        "log_lambda_gp_intercept", -2.0,
                                                        "log_lambda_gp_alpha",     0.3,
                                                        "log_lambda_gp_rho",       1.5,
                                                        "tumor_stim_intercept",    0.2,
                                                        "tumor_stim_coef_1",       0.15,
                                                        "tumor_stim_coef_2",       0.05)) +
  stat_pointinterval(aes(ydist = est, color = est_type), size = 0.25, .width = 0.8, fatten_point = 0) +
  facet_wrap(vars(param), scales = "free_y") +
  theme(axis.text.x = element_blank(), legend.position = "none")
```

```{r}
fake_data_sim |> 
  mutate(sim_log_lambda_post = map(sim_fit, \(f) spread_rvars(f, log_lambda[m]))) |> 
  select(!c(sim_data, sim_fit)) |>
  unnest(sim_log_lambda_post) |> 
  bind_rows(transmute(lambda_rvar, m = t, log_lambda, sim_id = 0)) |> 
  mutate(est_type = if_else(sim_id > 0, "post", "prior")) |> 
  ggplot(aes(sim_id)) + 
  geom_hline(aes(yintercept = true_val), data = tibble(m = 1:20, true_val = log_lambda)) +
  stat_pointinterval(aes(ydist = log_lambda, color = est_type), size = 0.5, .width = 0.8, fatten_point = 0) +
  facet_wrap(vars(m)) +
  theme(axis.text.x = element_blank(), legend.position = "none")
```

```{r}
fake_data_sim |> 
  mutate(km_est = map2(sim_fit, sim_data, \(f, simd) spread_rvars(f, km_est[m]) |> mutate(emp_s = pfs_model$functions$estimate_kaplan_meier(simd$pfs, max(simd$pfs))))) |> 
  select(sim_id, km_est) |> 
  filter(sim_id <= 6) |> 
  unnest(km_est) |> 
  ggplot(aes(m)) +
  stat_lineribbon(aes(ydist = km_est), alpha = 0.25, fill = "black", .width = 0.8) +
  geom_line(aes(y = emp_s), color = "red") +
  facet_wrap(vars(sim_id))
```

## Simulation-Based Calibration

```{r}
#| label: sbc

sbc_data <- rep_pfs_rvar |> 
  unnest_rvars() |> 
  ungroup() |> 
  filter(.draw <= 3) |> 
  select(.draw, pfs = rep_pfs, right_censored = rep_right_censored) |> 
  mutate(interval_censored = 0) |> 
  nest(sim_data = !.draw) |> 
  left_join(
    rep_pfs_param_rvar |> 
      unnest_rvars() |> 
      ungroup() |> 
      select(!c(.iteration, .chain)),
    by = ".draw"
  ) |> 
  mutate(
    furrr::future_map_dfr(.progress = TRUE, .options = furrr::furrr_options(seed = TRUE),
      sim_data,
      # We need thinning when doing SBC using MCMC to break the correlation between samples.
      \(d) fit_sim_data(d, thin = 4) |> 
        spread_rvars(tumor_stim_intercept, tumor_stim_coef, ndraws = 1000) |> 
        rename_with(\(n) str_c("est_", n))
    ),
    
    r_tumor_stim_intercept = sum(est_tumor_stim_intercept < tumor_stim_intercept),
    r_tumor_stim_coef_1 = c(sum(est_tumor_stim_coef[, 1] < tumor_stim_coef_1)),
    r_tumor_stim_coef_2 = c(sum(est_tumor_stim_coef[, 2] < tumor_stim_coef_2)),
  )
```
