---
title: "Early Prediction of Breast Cancer Survival"
---

```{r}
#| label: setup
#| include: false

library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)

options(mc.cores = 12)

ggplot2::theme_set(theme_minimal())

future::plan(future::multisession(workers = 12))

source(here("src", "util.R"))
```

```{r}
#| label: sim-settings

n_patients <- 1000
patient_measures <- 11
max_pfs <- 20
```

# Tumor Size Model

I haven't yet documented much about this model or done any kind of evaluation. I'm only using it at this time to generate fake tumor size data for testing.

```{r}
#| label: tumor

tumor_model <- cmdstan_model(here("src", "tumor.stan"))

if (!file.exists(here("temp", "fit", "prior_tumor.rds"))) {
  test_data <- lst(
    fit_data = FALSE,
    gen_tumor_sizes = TRUE,
    n_patients,
    n_measures = rep(patient_measures, n_patients),
    t_measure = rep(1:(patient_measures - 1), n_patients),
    n_patient_tumors = 1 + rbinom(n_patients, 5, 0.4),
    tumor_size = array(dim = 0)
  )
  
  res <- tumor_model$sample(data = test_data, output_dir = here("temp", "fit"))
  
  res$save_object(here("temp", "fit", "prior_tumor.rds"))
  
  res$save_data_file(
    here("temp", "data"), basename = "prior_tumor", timestamp = FALSE, random = FALSE
  )
} else {
  res <- read_rds(here("temp", "fit", "prior_tumor.rds"))
  test_data <- rjson::fromJSON(file = here("temp", "data", "prior_tumor.json"))
}

res_rvar <- res |> 
  spread_rvars(rep_tumor_size[tumor_index]) |> 
  ungroup() |> 
  mutate(
    patient_id = rep(1:n_patients, with(test_data, n_patient_tumors * patient_measures)),
    tumor_index = ((tumor_index - 1) %/% patient_measures) + 1,
    t = rep(0:(patient_measures - 1), sum(test_data$n_patient_tumors))
  ) |> 
  mutate(tumor_index = tumor_index - min(tumor_index) + 1, .by = patient_id) |> 
  rename(tumor_size = rep_tumor_size)

res_rvar |> 
  filter(patient_id == 1, tumor_index < 6) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = tumor_size, group = str_c(patient_id, tumor_index, .draw)), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  stat_lineribbon(aes(ydist = tumor_size, group = tumor_index), alpha = 0.25, .width = 0.8) +
  coord_cartesian(ylim = c(0, patient_measures)) +
  facet_wrap(vars(tumor_index)) +
  NULL

fake_tumor_data <- res_rvar |>
  unnest_rvars() |>
  ungroup() |> 
  filter(.draw == 1) 

write_rds(fake_tumor_data, here("temp", "data", "fake_tumor.rds"))
```

```{r}
#| label: tumor-model-diagnose
#| eval: false

res$cmdstan_diagnose()
```

## Fake Data Simulation

```{r}
#| include: false
#| eval: false

test_data |> 
  list_modify(
    fit_data = TRUE,
    gen_tumor_sizes = FALSE,
    tumor_size = fake_tumor_data$tumor_size
  ) |> 
  tumor_model$sample(output_dir = here("temp", "fit"), output_basename = "tumor")

```

# Survival Model

We will be using a piecewise-constant proportional hazard model:

-   Time is discretized into $m \in \{0,\dots,M\}$, with $m = 0$ is the baseline timepoint.
-   $Y_m$ is an indicator for disease progression at timepoint $m$, with $Y_0 = 0$.
-   $X_m$ are--possibly time-variant--covariates.

The probability of remaining progression-free at $m \geq 1$ is 
$$
\alpha_m(x;\theta) = \Pr[Y_m = 0\mid Y_{m-1} = 0, X_m = \bf{x}] = \exp[-\lambda(m, \bf{x}; \boldsymbol{\theta})] 
$$

and the hazard function

$$
  \lambda(m, \bf{x}; \boldsymbol{\theta}) = \lambda_m \kappa(\bf{x}; \boldsymbol{\beta})
$$

We assume that censoring in conditionally independent of time to progression. We will use a linear for for $\kappa(\bf{x}; \boldsymbol{\beta}) = \exp(\bf{x}\boldsymbol{\beta})$.

Aside from whatever priors we use for $\boldsymbol{\beta}$, we will use a Gaussian process prior for $\lambda_m$, 
$$
\lambda_m \sim \mathtt{GP}(0, k),
$$ 
where $k(\lambda_p, \lambda_q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right)$.

```{r}
#| label: pfs-model

pfs_model <- cmdstan_model(here("src", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)

pfs_test_data <- test_data |> 
  list_modify(
    gen_pfs = TRUE,
    early_tumors_only = TRUE,
    ignore_interval_censoring = FALSE,
    tumor_size = fake_tumor_data$tumor_size,
    pfs = rep(max_pfs, test_data$n_patients),
    right_censored = rep(1, test_data$n_patients),
    
    log_lambda_gp_intercept_mean = -3,
    log_lambda_gp_intercept_sd = 0.5,
    tumor_stim_intercept_sd = 0.5,
    tumor_stim_coef_sd = c(0.25, 0.25),
  )
```

## Prior Prediction

```{r}
#| label: pfs-prior-predict

pfs_res <- pfs_model$sample(data = pfs_test_data, refresh = 0)

rep_pfs_rvar <- pfs_res |> 
  spread_rvars(rep_pfs[patient_index], rep_right_censored[patient_index])

rep_pfs_param_rvar <- pfs_res |> 
  spread_rvars(log_lambda_gp_alpha, log_lambda_gp_rho, log_lambda_gp_intercept, tumor_stim_intercept, tumor_stim_coef[t]) |> 
  pivot_wider(names_from = t, values_from = tumor_stim_coef, names_prefix = "tumor_stim_coef_") 

lambda_rvar <- pfs_res |> 
  spread_rvars(log_lambda[t], base_pf_cond_prob[t], base_survival[t], one_tumor_survival[t]) |> 
  mutate(lambda = exp(log_lambda))

expected_pfs_rvar <- pfs_res |> 
  gather_rvars(base_cond_expected_pfs, one_tumor_cond_expected_pfs)  
```

```{r}
#| label: lambda-plot

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = lambda), fill = "black", alpha = 0.5, .width = 0.8) +
  geom_line(aes(y = lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(title = "Base hazard", x = "m", y = latex2exp::TeX("$\\lambda_m$"))
```

```{r}
#| label: base-pf-prob-plot

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = base_pf_cond_prob), fill = "black", alpha = 0.5, .width = 0.8) +
  geom_line(aes(y = base_pf_cond_prob, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(title = "Base conditional probability of remaining progress free", x = "m", y = "")
```

```{r}
#| label: pf-prob-plot

max_t <- with(pfs_test_data, list_measures(t_measure, n_measures - 1)) |> 
  map_int(max)

pfs_res |> 
  spread_rvars(disease_progress_prob[i]) |> 
  mutate(patient_index = rep(seq(test_data$n_patients), max_t), t = pfs_test_data$t_measure) |> 
  filter(patient_index <= 6) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = disease_progress_prob), fill = "black", alpha = 0.5, .width = 0.8) +
  geom_line(aes(y = disease_progress_prob, group = str_c(patient_index, .draw)), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 20)) +
  labs(title = "Conditional probability of remaining progress free", x = "m", y = latex2exp::TeX("$\\alpha_m$")) +
  coord_cartesian(ylim = c(0, 1)) +
  facet_wrap(vars(patient_index)) +
  NULL
```

```{r}
#| label: base-surivival-plot

lambda_rvar |>
  select(t, base_survival, one_tumor_survival) |> 
  pivot_longer(!t, names_to = "survival_type", values_to = "survival") |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = survival, fill = survival_type, color = survival_type), alpha = 0.25, .width = 0.8) +
  labs(title = "Base survival probability", x = "m", y = latex2exp::TeX("$S(m)$"))
```

```{r}
#| label: cond-expected-pfs-plot

expected_pfs_rvar |> 
  mutate(.variable = str_remove(.variable, "_cond_expected_pfs")) |> 
  ggplot(aes(y = .variable)) +
  stat_pointinterval(aes(xdist = .value), .width = 0.8) +
  labs(title = "Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "m", y = "")
```

```{r}
#| label: prior-kaplan-meier 

pfs_res |> 
  spread_rvars(km_est[m]) |> 
  ggplot(aes(m - 1)) +
  stat_lineribbon(aes(ydist = km_est), fill = "black", alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", y = latex2exp::TeX("$S(m)$")) +
  scale_x_continuous("m", breaks = seq(0, 20, 2))
```

## Fake Data Simulation

```{r}
#| label: fake-data-model
#| include: false
#| cache: true

# Using a fixed set of log lambda values, generated using:
#   pfs_model$functions$calc_gp_pred(seq(max_pfs), intercept = -2.0, alpha = 0.3, rho = 1.5, delta = 1e-9, eta = rnorm(max_pfs))

log_lambda <- c(-2.48777760709026, -2.57152726838213, -2.23380584287276, -1.90739893145911, -1.86259266390287, -1.92891255821976, -1.93801857664684, -1.78872336415524, -1.54522270753866, -1.48718504020436, -1.70706597143648, -1.82018710107277, -1.67764635563667, -1.69933003110133, -1.86937604603516, -2.03108481232034, -2.15728306810556, -2.11459406967666, -2.02311574428507, -2.08871876979315) 

true_tumor_intercept <- 0.2
true_tumor_coef <- c(0.15, 0.05)

patient_interval_data <- gen_patient_interval_properties(
  fake_tumor_data, log_lambda, true_tumor_intercept, true_tumor_coef, fake_data_sim_settings
)

fake_data_sim_settings <- pfs_test_data |>
  list_modify(
    log_lambda_gp_intercept_mean = -2,
    log_lambda_gp_intercept_sd = 0.5,
    tumor_stim_intercept_sd = 0.25,
    tumor_stim_coef_sd = c(0.125, 0.125),
  )

true_base_pf_cond_prob = 
  1 - pfs_model$functions$calculate_progress_linear_prob(
    log_lambda, true_tumor_intercept, true_tumor_coef, t(c(0, 0)) 
  )

true_one_tumor_pf_cond_prob = 
  1 - pfs_model$functions$calculate_progress_linear_prob(
    log_lambda, true_tumor_intercept, true_tumor_coef, t(c(1, 1)) 
  )

true_base_marginal_prob_res = pfs_model$functions$calculate_marginal_dp_prob(true_base_pf_cond_prob, 10)  
true_one_tumor_marginal_prob_res = pfs_model$functions$calculate_marginal_dp_prob(true_one_tumor_pf_cond_prob, 10)  
true_base_cond_expected_pfs = 
  c((true_base_marginal_prob_res[[1]] / (1 - true_base_marginal_prob_res[[2]][10])) %*% 1:10)
true_one_tumor_cond_expected_pfs = 
  c((true_one_tumor_marginal_prob_res[[1]] / (1 - true_one_tumor_marginal_prob_res[[2]][10])) %*% 1:10)

true_km <- patient_interval_data |> 
  group_by(patient_id) %>% 
  summarize(pfs = pfs_model$functions$pfs_rng(progress_prob, 1:n()) |>   
              list()) %>% 
  mutate(!!!set_names(list_transpose(.$pfs), c("interval_censored", "right_censored", "pfs", "actual_pfs"))) |>   
  reframe(km_est = pfs_model$functions$estimate_kaplan_meier(pfs, right_censored, max_pfs - 1)) |> 
  mutate(m = seq(n()))
```

```{r}
#| label: fake-data-sim 

n_sim <- 3 
fake_data_sim_settings_with_ic <- drop_missing_measures(fake_data_sim_settings, c(3, 4, 6, 9))

fake_data_sim <- bind_rows(
  interval_censoring = fit_simulations(n_sim, patient_interval_data, fake_data_sim_settings_with_ic, fit_basename = "fake_data_sim_ic"),
  interval_censoring_ignored = fit_simulations(n_sim, patient_interval_data, fake_data_sim_settings_with_ic, ignore_interval_censoring = TRUE, fit_basename = "fake_data_sim_ic_ignored"),
  no_interval_censoring = fit_simulations(n_sim, patient_interval_data, fake_data_sim_settings, fit_basename = "fake_data_sim"),
  .id = "sim_type"
)
```

```{r}
#| label: fake-data-params

fake_data_sim |> 
  transmute(
    sim_id,
    sim_type,
    
    sim_param_post = furrr::future_map(
    # sim_param_post = map(
      sim_fit,
      \(f) spread_rvars(f, log_lambda_gp_alpha, log_lambda_gp_rho, log_lambda_gp_intercept, tumor_stim_intercept, tumor_stim_coef[t]) |>
        pivot_wider(names_from = t, values_from = tumor_stim_coef, names_prefix = "tumor_stim_coef_")
    ),
  ) |> 
  unnest(sim_param_post) |> 
  add_row(sim_id = 0, sim_type = "prior", rep_pfs_param_rvar) |>
  pivot_longer(!c(sim_id, sim_type), names_to = "param", values_to = "est") |>
  ggplot(aes(sim_id)) + 
  geom_hline(aes(yintercept = true_val), data = tribble(~ param,                   ~ true_val,
                                                        "log_lambda_gp_intercept", -2.0,
                                                        "log_lambda_gp_alpha",     0.3,
                                                        "log_lambda_gp_rho",       1.5,
                                                        "tumor_stim_intercept",    0.2,
                                                        "tumor_stim_coef_1",       0.15,
                                                        "tumor_stim_coef_2",       0.05)) +
  stat_pointinterval(aes(ydist = est, color = sim_type), size = 0.25, .width = 0.8, fatten_point = 0, position = "dodge") +
  facet_wrap(vars(param), scales = "free_y") +
  theme(axis.text.x = element_blank())
```

```{r}
#| label: fake-data-lambda

fake_data_sim |> 
  mutate(
    # sim_log_lambda_post = furrr::future_map(
    sim_log_lambda_post = map(
      sim_fit, \(f) spread_rvars(f, log_lambda[m]))
  ) |> 
  select(!c(sim_data, sim_fit)) |>
  unnest(sim_log_lambda_post) |> 
  bind_rows(transmute(lambda_rvar, m = t, log_lambda, sim_id = 0, sim_type = "prior")) |> 
  ggplot(aes(sim_id)) + 
  geom_hline(aes(yintercept = true_val), data = \(d) distinct(d, m) |> mutate(true_val = log_lambda[m])) +
  stat_pointinterval(
    aes(ydist = log_lambda, color = sim_type), 
    size = 0.5, .width = 0.8, fatten_point = 0, position = "dodge"
  ) +
  facet_wrap(vars(m)) +
  theme(axis.text.x = element_blank())
```

```{r}
#| label: fake-data-km

fake_data_sim |> 
  # mutate(km_est = map2(
  mutate(km_est = furrr::future_map2(
    sim_fit, sim_data, 
    \(f, simd) spread_rvars(f, km_est[m])  
  )) |> 
  select(sim_id, sim_type, km_est) |> 
  filter(sim_id <= 9) |> 
  unnest(km_est) |> 
  ggplot(aes(m - 1)) +
  geom_line(aes(y = km_est), data = \(sim) semi_join(true_km, sim, by = "m")) +
  stat_lineribbon(aes(ydist = km_est, color = sim_type, fill = sim_type), alpha = 0.25, .width = 0.8) +
  scale_x_continuous("m", breaks = seq(0, 10, 2)) +
  facet_wrap(vars(sim_id)) +
  theme(legend.position = "top")
```

```{r}
#| label: fake-data-cond-expected-pfs-plot

fake_data_sim |>
  group_by(sim_id, sim_type) |> # So reframe doesn't throw away sim_id 
  reframe(furrr::future_map_dfr(sim_fit, \(f) gather_rvars(f, base_cond_expected_pfs, one_tumor_cond_expected_pfs))) |> 
  add_row(sim_id = 0, sim_type = "prior", expected_pfs_rvar) |> 
  mutate(
    .variable = str_remove(.variable, "_cond_expected_pfs"),
    est_type = if_else(sim_id > 0, "post", "prior")
  ) |> 
  ggplot(aes(y = sim_id)) +
  stat_pointinterval(aes(xdist = .value, color = sim_type), size = 0.5, .width = 0.8, position = "dodge") +
  geom_vline(aes(xintercept = true_e), 
             data = tibble(.variable = c("base", "one_tumor"), true_e = c(true_base_cond_expected_pfs, true_one_tumor_cond_expected_pfs))) +
  labs(title = "Fake Data Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "m", y = "") +
  facet_wrap(vars(.variable)) +
  NULL
```

## Simulation-Based Calibration

The code for this simulation is located in `sbc.R` (and `sbc.sh` to run on SLURM).

```{r}
#| label: sbc

sbc_data <- read_rds(here("temp", "data", "sbc.rds"))

sbc_data |> 
  select(starts_with("r_")) |> 
  pivot_longer(everything(), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = 10) +
  geom_rect(aes(xmin = -10, xmax = 1010, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = \(d) tibble(!!!set_names(qbinom(c(0.01, 0.5, 0.99), nrow(sbc_data), 1/100), c("lb", "med", "ub")))) +
  facet_wrap(vars(param)) +
  labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom(1090, 0.01).\n{nrow(sbc_data)} simulation used."))

# sbc_data |> 
#   select(starts_with("r_")) |> 
#   pivot_longer(everything(), names_to = "param", names_prefix = "r_", values_to = "r") |> 
#   ggplot() +
#   stat_ecdf(aes(r)) +
#   facet_wrap(vars(param)) +
#   coord_cartesian(xlim = c(0, 1000))
```
