---
title: "Early Prediction of Breast Cancer Progress Free Survival"
author: Karim Naguib
code-fold: true
code-tools: true
toc: true
toc-location: left
fig-caption-location: margin
---

```{r}
#| label: setup
#| include: false

# library(conflicted)
library(magrittr)
library(tidyverse)
library(rlang)
library(cmdstanr)
library(loo)
library(posterior)
library(tidybayes)
library(bayesplot)
library(here)
library(gtsummary)
library(targets)

# conflicts_prefer(
#   dplyr::filter
# )

# Using different temp directories on SCP and locally.
tmp_dir <- if (nchar(Sys.getenv("TMP")) > 0) file.path(Sys.getenv("TMPDIR"), "adc-early-predict") else here("temp")

knitr::opts_chunk$set(cache.path = file.path("/wscratch", Sys.getenv("USER"), "adc-early-predict/cache/"))

options(mc.cores = 4)

ggplot2::theme_set(theme_minimal())

future::plan(future::multisession(workers = 12))

source(here("src", "util.R"))
```

# Exploratory Analysis Plan

## Question

Are there observable strong and deep early responses to antibody-drug conjugates (ADCs) that are predictive of long term progression free survival (PFS)? Are we able to reliably predict the heterogeneous response to ADCs using early measurements?

In this exploratory analysis we will not take up the causal question of whether these strong early responders would see similar PFS under counterfactual dosages or medications.[^1]

[^1]: This should probably revisited as the prediction within arm might not be the relevant question.

## Exploratory Analysis Context and Data

We will evaluate whether early (at baseline and at 6 and 12 weeks post treatment) measurements of tumor sizes are predictive of the PFS in response to Enhertu.

We will use data from the Breast01, Breast02, Breast03, and Breast04 trials, focusing on the treatment arms receiving the same dosage. Initially, we will be analyzing the different studies separately with no partial pooling since the inclusion criteria for the studies were different.

* Primary outcome: Post treatment progression free survival (interval and right censored).
* Predictive variables (at all the early intervals): 
    - Number of lesions observed. 
    - Organ of lesions. 
    - lesion sizes measured, with a maximum of ten of observed lesions and two per organ.

## Statistical Model Goals

-   Explore the heterogeneity in PFS conditional on early tumor size measures.
-   Consider more nuanced characteristics of a patient's tumors beyond the aggregated sum of largest diameters. The challenge here is how to model the nonlinear contribution of multiple influencers (tumors) on a central node's outcome (survival). A simple sum or mean ends up discarding a lot of information about the context of the disease.
-   Introduce a lesion level first-stage model. Considering the limit on the number of lesions measured, we potentially truncate some measures such as the total sum of tumor sizes. A model of tumor sizes would allow us to impute the missing data and propagate any uncertainty to the second-stage survival model.

```{r}
#| label: priors
#| cache: true

tumor_priors <- lst(
  pop_tumor_gp_rho_alpha = 7.3,
  pop_tumor_gp_rho_beta = 7.5, 
)

pfs_priors <- lst(
  log_lambda_gp_intercept_mean = -1,
  log_lambda_gp_intercept_sd = 0.25,
  log_lambda_gp_alpha_sd = 0.25,
  log_lambda_gp_rho_alpha = 7.3,
  log_lambda_gp_rho_beta = 7.5, 
  tumor_stim_intercept_sd = 0.125,
  log_lambda_gp_trial_alpha_sd = 0.0625,
  log_lambda_gp_trial_intercept_sd_sd = 0.25,
  tumor_stim_coef_sd = c(0.125, 0.125, 0.0625, 0.0625, 0.0625) / 2,
  tumor_stim_trial_coef_sd_sd = 0.0625
)
```

```{r}
#| label: models
#| code-fold: false   

tumor_model <- cmdstan_model(here("src", "tumor.stan"))
pfs_model <- cmdstan_model(here("src", "pfs.stan"), compile_standalone = TRUE, force_recompile = TRUE)
```


# Data

What data we have so far:
1. Breast01: phase II dose finding study (HER2-positive metastatic breast cancer).
2. Breast02: phase III study comparing to physician choice of drug (HER2-positive metastatic breast cancer). 
2. Breast03^[Pending]: phase III study comparing to trastuzumab emtansine (HER2-positive metastatic breast cancer).
3. Breast04: This is a phase III study on patients with HER2-low metastatic breast cancer (HER2-low metastatic breast cancer).

```{mermaid}
%%| label: fig-data-structure
%%| fig-align: center

erDiagram
  Study ||--|{ Patient : enrolls
  Patient ||--|{ Tumor : has
  Tumor ||--|{ Assessment : assessed 
```

```{r}
#| label: load-trials-data
#| cache: true

data_details <- tribble(
  ~ idap,          ~ trial, 
  "idap_20231204", "Breast01",
  "idap_20231129", "Breast02",  
  "idap_20231124", "Breast04"
) |> 
  mutate(across(everything(), factor))

# Trial arms
sdtm_ta <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_ta", "sdtm")))

# Trial elements 
sdtm_te <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_te", "sdtm")))

# Protocol deviations
# sdtm_dv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dv.parquet") |> 
#   rename_with(str_to_lower)

# Inclusion/exclusion
# sdtm_ie <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_ie.parquet") |> 
#   rename_with(str_to_lower)

# Demographics 
sdtm_dm <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_dm", "sdtm")))

# Death details 
# sdtm_dd <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_dd.parquet") |> 
#   rename_with(str_to_lower)

# Exposure 
sdtm_ex <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_ex", "sdtm")))

# Tumor identification
# sdtm_tu <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tu.parquet") |> 
#   rename_with(str_to_lower)

# Trial visits 
# sdtm_tv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tv.parquet") |> 
#   rename_with(str_to_lower)

# Subject visits
# sdtm_sv <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_sv.parquet") |> 
#   rename_with(str_to_lower)

# Tumor measurements/assessments
# sdtm_tr <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_tr.parquet") |>
# sdtm_tr <- read_rds("/scratch/ewfteams/dpo0083/sdtm/prod/data/deid_tr.rds") |> 
  # rename_with(str_to_lower) |> 
  # mutate(tracptfl = fct_match(tracptfl, "Y"))

# Disease response 
# sdtm_rs <- arrow::read_parquet("/scratch/ewfteams/dpo0083/parquet/sdtm/deid_rs.parquet") |> 
#   rename_with(str_to_lower)

# Tumor analysis_data
adam_adtr <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adtr", "adam")))
  
# Tumor analysis_data
# adam_adef <- map_dfr(idaps, \(idap) read_rds(file.path(team_dir, idap, "adam/prod/data/deid_adef.rds")), .id = "idap") |> 
#   rename_with(str_to_lower) |> 
#   mutate(across(ends_with("fl"), \(fl) fct_match(fl, "Y")))

# Time to event
adam_adtte <- data_details |> 
  rowwise() |> 
  mutate(data = list(read_entimice_data(idap, "deid_adtte", "adam")) |> 
           map(\(d) mutate(d, cnsr = cnsr == 1)))
```

Using SDTM and ADaM data from the trials, we prepare the analysis data.

```{r}
#| label: prepare-analysis-data
#| cache: true
#| dependson: load-trials-data
#| code-fold: false 

tumor_analysis_data <- adam_adtr |> 
  unnest(data) |> 
  filter(
    # fct_match(paramcd, c("DIAMETER", "DIAM")),
    fct_match(paramcd, "LDIAM"),
    # There are a number of assessments that show up in the data but in SDTM they are marked as "NOT DONE"
    !is.na(trdy), !is.na(aval),
    tracptfl,
    # adt <= trtedt, # Drop assessments occurring after treatment ended.
    aval > 0, # Drop all zero diameter assessments; these appear to be missing measurements. 
  ) %>% 
  select(studyid, idap, trial, 
         usubjid, trlnkid, tuloc, tulat, trtsdt, trtedt, adt, ends_with("fl"), trdy, trtp, trta, visit, visitnum, mmdiam = aval) |> 
  mutate(
    week = if_else(trdy > 0, trdy, trdy + 1) %/% 7, 
    treated_week = adt >= trtsdt,
  ) |> 
  arrange(usubjid, trlnkid, visitnum) |> 
  group_by(usubjid, trlnkid) |> 
  mutate(tuloc = first(tuloc), tulat = first(tulat) |> na_if("")) |> # Only the first assessment has the location of the tumor. Fix that.
  ungroup() |>
  # Turns out that some assessments are not listed in order of the days they were done. Look at subject DS8201-A-U303-10033028.
  arrange(studyid, usubjid, trlnkid, trdy) |> 
  nest(tumor_history = c(adt, trdy, week, visit, visitnum, tsplitfl, tmergefl, aocc01fl, mmdiam, treated_week)) |> 
  mutate(# For some weeks, there are multiple (unscheduled) visits that might happen. I'm just going to use the average for tumor sizes within those
         # weeks.
    tumor_history = map(
      tumor_history, 
      \(h) nest(h, visits = c(adt, trdy, visit, visitnum, mmdiam)) |> 
        mutate(
          n_week_visits = map_int(visits, nrow),
          mmdiam = map_dbl(visits, \(v) mean(v$mmdiam))
        )
    ),
    n_tumor_measures = map_int(tumor_history, nrow),
    
    min_t = map_int(tumor_history, \(h) min(h$week)),
    max_t = map_int(tumor_history, \(h) max(h$week)),
  ) |> 
  pack(flags = ends_with("fl")) |> 
  nest(patient_tumors = !c(studyid, trial, idap, usubjid, trtsdt, trtedt, trtp, trta)) |> 
  rowwise() |> 
  mutate(
    patient_min_t = min(patient_tumors$min_t),
    patient_max_t = max(patient_tumors$max_t),
  ) |>
  ungroup() |> 
  mutate(
    patient_t_width = patient_max_t - patient_min_t + 1,
    n_tumors = map_int(patient_tumors, nrow),
  ) |> 
  left_join(
    adam_adtte |> 
      unnest(data) |> 
      filter(
        (fct_match(paramcd, "PFS1") & fct_match(studyid, c("DS8201-A-U301", "DS8201-A-U303"))) | 
          (fct_match(paramcd, "PFS") & fct_match(studyid, "DS8201-A-U201")), 
        fct_match(parqual, "CENTRAL")
      ) |> 
        select(studyid, usubjid, aval, death = dthfl, dthdt, fpddt, cnsr),
    by = c("studyid", "usubjid")
  ) |> 
  mutate(
    right_censored = cnsr,
    death_week = floor(time_length(dthdt - trtsdt, unit = "weeks")),
    progress_week = floor(time_length(fpddt - trtsdt, unit = "weeks")),
    progress_week = if_else(death, coalesce(progress_week, death_week), 
                                   if_else(right_censored & is.na(progress_week), max(progress_week, na.rm = TRUE), progress_week)),
    aval_week = aval %/% 7,
    
    treated = 
      (fct_match(trial, "Breast01") & fct_match(trtp, "5.4 mg/kg")) | 
      (fct_match(trial, c("Breast02", "Breast04")) & fct_match(trtp, "T-DXd")),
    
    # The model expects "pfs" to record the number of weeks that were progress free, so I need to find the last assessment week before
    # disease progress was detected.
    pfs = pmap_int(
      lst(progress_week, death_week, right_censored, patient_tumors), function(progress_week, death_week, right_censored, patient_tumors) {
        pre_progress_weeks <- unnest(patient_tumors, tumor_history) |>
          distinct(week) |>
          filter(right_censored | week < min(progress_week, death_week, na.rm = TRUE)) |>
          pull(week)

        # There are a few patients who just have a single post treatment visit
        if (length(pre_progress_weeks) > 0) max(pre_progress_weeks) else NA_integer_
      }
    ),

    death_week = coalesce(death_week, 0L),
    
    n_measures = map(patient_tumors, \(tu) tu$n_tumor_measures),
    t_measure = map(patient_tumors, \(tu) map(tu$tumor_history, \(tr) pull(tr, week))),
  ) |> 
  left_join(select(unnest(sdtm_dm, data), studyid, usubjid, country), by = c("studyid", "usubjid")) %>%
  mutate(
    interval_censored = with(., pfs_model$functions$identify_censoring(pfs, death_week, n_tumors, unlist(n_measures), unlist(t_measure))[[1]]) |> 
      pmax(0) # Issue https://github.com/azu-oncology-rd/ods-adc-early-predict-2023/issues/36
  )
  
pfs_analysis_data <- tumor_analysis_data |> 
  filter(
    !is.na(pfs), pfs >= 0, # No missing pfs and progress must be assured to have happened after treatment
    patient_min_t <= 0, # At least one screening assessment needed,
    patient_max_t > 0, # and at least one post treatment assessment.
  )

treated_pfs_analysis_data <- pfs_analysis_data |> 
  filter(treated) 

treated_pfs_stan_data <- prepare_pfs_stan_data(treated_pfs_analysis_data, tumor_priors, pfs_priors)

treated_trial_pfs_stan_data <- treated_pfs_analysis_data |> 
  nest(.by = trial) |> 
  deframe() |> 
  imap(\(d, tr) prepare_pfs_stan_data(mutate(d, trial = tr), tumor_priors, pfs_priors))

treated_obs_km <- map_dfr(treated_trial_pfs_stan_data, function(sdata) {
  with(sdata, {
    list(lb = pfs, ub = pfs + interval_censored) |> 
      map_dfr(function(s) {
        pfs_model$functions$estimate_kaplan_meier(s, right_censored, max(s)) |>
          set_names(c("s", "n", "c", "e")) |>
          as_tibble() |> 
          mutate(t = seq(0, n() - 1))
      }, .id = "btype")
    }
  )
}, .id = "trial")

pfs_analysis_data |> 
  count(studyid, trial, treated, trtp, trta) |> 
  kableExtra::kbl() |> 
  kableExtra::kable_styling()
```

In the below plots we look at the empirical Kaplan-Meier survival, the number of at risk patients, and the number of censored at each interval, for the various trials and arms. We also compare: 

a) Our calculated PFS measure defined as the observed assessment week that occured _just before_ progress/death week.
b) PFS as defined as the progress week in data.

As expected, if we use progress week, survival appears to be longer, while in fact survival is interval censored between the last progress free week and the week just before progress. In particular, the number of censored plot shows bunching at the maximum observed progress week.

```{r}
#| label: fig-survival-dynamics
#| warning: false

km_res_calculated <- pfs_analysis_data |>
  group_by(trial, treated) |>  
  group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "pfs"), .keep = TRUE) |>  
  bind_rows()  

km_res_from_data <- pfs_analysis_data |>
  group_by(trial, treated) |>  
  group_map(\(trt_data, key) km_to_tibble(trt_data, key, pfs_var = "progress_week"), .keep = TRUE) |>  
  bind_rows()  

km_res <- bind_rows(calculated = km_res_calculated, from_data = km_res_from_data, .id = "pfs_type")

km_res |> 
  filter(fct_match(btype, "lb")) |> 
  ggplot() +
  geom_step(aes(x = t, y = s, color = treated, linetype = pfs_type)) +
  scale_color_discrete("Arm") +
  labs(title = "Kaplan-Meier survival", y = "S(t)") +
  facet_wrap(vars(trial), ncol = 1) +
  NULL

km_res |> 
  filter(fct_match(btype, "lb")) |> 
  ggplot() +
  geom_step(aes(x = t, y = n, color = treated, linetype = pfs_type)) +
  scale_color_discrete("Arm") +
  labs(title = "At risk patients", y = "N") +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y") +
  NULL

km_res |>
  filter(fct_match(btype, "lb")) |> 
  group_by(pfs_type, trial, t) |> 
  summarize(c = sum(c), .groups = "drop") |> 
  ggplot() +
  geom_col(aes(x = t, y = c)) +
  scale_color_discrete("Arm") +
  labs(title = "Number of censored patients", y = "sum(C)") +
  facet_grid(vars(trial), vars(pfs_type), scales = "free_y") +
  NULL

km_res |> 
  filter(fct_match(btype, "lb")) |> 
  mutate(n = n - c - e) |> 
  pivot_longer(c(n, c, e)) |> 
  ggplot(aes(t, value)) +
  geom_col(aes(fill = name, alpha = name)) +
  scale_alpha_manual("", values = c("n" = 0.25)) +
  scale_fill_discrete("", labels = c(c = "Censored", e = "Exiting", n = "Remaining")) +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y") +
  labs(title = "Survival dynamics", y = "") +
  guides(alpha = "none") +
  NULL 
```


```{r}
#| labels: kaplan-meier-upper-and-lower-bounds

km_res |> 
  filter(fct_match(pfs_type, "calculated")) |>
  ggplot() +
  geom_step(aes(x = t, y = s, color = btype)) +
  scale_color_discrete("Survival up to", labels = c(lb = "First possible week", ub = "Last possible week")) +
  labs(title = "Kaplan-Meier survival", y = "S(t)") +
  facet_grid(vars(treated), vars(trial), scales = "free_x", labeller = labeller(treated = label_both)) +
  NULL

# treated_pfs_stan_data[c("pfs", "right_censored", "interval_censored", "patient_trial")] |> 
#   as_tibble() %>% 
#   bind_rows(lb = ., ub = mutate(., pfs = pfs + interval_censored), .id = "btype") |> 
#   rename(trial = patient_trial) |> 
#   mutate(status = !right_censored) %>% 
#   ggsurvfit::survfit2(survival::Surv(pfs, status) ~ btype + trial, data = .) |> 
#   ggsurvfit::ggsurvfit() +
#   geom_step(aes(x = t, y = s, color = btype), linewidth = 4, alpha = 0.2, data = treated_obs_km) +
#   scale_color_discrete("Survival up to", labels = c(lb = "First possible week", ub = "Last possible week")) +
#   facet_wrap(vars(trial)) +
#   labs(title = "Comparing our KM estimator to the {survival} estimation", caption = "Solid line: {survival} Kaplan-Meier curve.\nRibbon: Our estimation.", x = "t", y = "S(t)")  +
#   theme_minimal() +
#   NULL
```


```{r}
#| label: tumor-demographics-table 
#| eval: false

sdtm_dm$data |> 
  map(\(d) tbl_summary(d, include = c("age", "sex", "race", "ethnic", "dthfl"), by = "arm"))
```

```{r}
#| label: tumor-history-plot 

tumor_analysis_data |>
  filter(patient_t_width > 1) |> 
  slice(1:20) |>
  unnest(patient_tumors) |> 
  unnest(tumor_history) |> 
  ggplot() +
  geom_step(aes(week, mmdiam, group = trlnkid), direction = "hv") +
  facet_wrap(vars(usubjid), scale = "free") +
  labs(title = "Observed tumor sizes for a sample of patients", x = "Week", y = "Tumor Diameter [mm]") +
  theme(strip.text = element_blank()) +
  NULL
```

```{r}
#| label: fig-pfs-distribution
#| code-fold: false
#| fig-cap: PFS empirical distribution. We're only showing uncensored observations.

pfs_analysis_data |> 
  filter(!right_censored, treated) |> 
  ggplot() +
  geom_bar(aes(pfs)) +
  labs(title = "PFS empirical distribution", x = "PFS", y = "Count") +
  facet_wrap(vars(trial), ncol = 1)
```

```{r}
#| label: patient-visits-pattern

tumor_analysis_data |> 
  select(trial, n_tumors, t_measure) |> 
  unnest(t_measure) |> 
  unnest(t_measure) |> 
  mutate(epoch = if_else(t_measure <= 0, "pre-screening", "treatment") |> as_factor()) |> 
  ggplot(aes(t_measure), alpha = 0.25) +
  geom_bar(aes(fill = epoch)) +
  labs(title = "Patient visit patterns", x = "Week", y = "Count") +
  facet_wrap(vars(trial), scales = "free_y") +
  theme(legend.position = "top")
```

```{r}
#| label: fig-treatment-start-end

tumor_analysis_data |> 
  ggplot() +
  geom_segment(aes(x = trtsdt, xend = trtedt, y = usubjid, yend = usubjid, color = country), alpha = 0.5) +
  scale_color_discrete("Country") +
  labs(title = "Patient treatment start/end timelines", x = "", y = "") +
  facet_wrap(vars(trial), ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_blank(), panel.grid.major.y = element_blank()) +
  NULL
```

```{r}
#| label: fig-interval-censoring

interval_censoring_data <- pfs_analysis_data %>% 
  mutate(
    interval_censoring = prepare_pfs_stan_data(., tumor_priors, pfs_priors) |>
      with(pfs_model$functions$identify_censoring(pfs, death_week, n_patient_tumors, n_measures, t_measure)) |> 
      pluck(1)
  )

interval_censoring_data |> 
  ggplot(aes(y = interval_censoring > 0)) +
  geom_bar(position = "dodge", alpha = 0.5) +
  scale_y_discrete("Interval Censored") +
  facet_wrap(vars(trial), ncol = 1) +
  labs(title = "Proportion of interval censoring", x = "") 

interval_censoring_data |> 
  filter(interval_censoring > 0) |> 
  ggplot(aes(interval_censoring)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth = 5) +
  labs(
    title = "Histogram of interval censoring", 
    subtitle = "Conditional on any interval censoring", 
    x = "Interval Censoring Weeks", 
    y = "",
    caption = "Binwidth = 5"
  ) +
  facet_wrap(vars(trial), ncol = 2, scales = "free_y") +
  NULL
```

```{r}
#| label: sim-settings

n_patients <- 250
patient_measures <- 41 
max_pfs <- 50 
```

# Tumor Size Model

I haven't yet documented much about this model or done any kind of evaluation. I'm only using it at this time to generate fake tumor size data for testing.

We model tumor size as a nonlinear function with a Gaussian process prior (with common hyperparameters),

$$
\log(X_{it}) \sim  \mathtt{GP}(\mu^\textrm{tumor}, k(\cdot;\alpha^\textrm{tumor}, \rho^\textrm{tumor})) 
$$ 
where 
$$k(p, q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right).$$

## Prior Prediction


```{r}
#| label: test-data
#| cache: true
#| dependson: priors

t_offset <- 5

test_data <- lst(
  fit_tumor_data = FALSE, # Prior prediction
  gen_tumor_sizes = TRUE,
  predict_missing_sizes = FALSE,
  multilevel_patient = FALSE,
  multilevel_tumor = FALSE,
  
  n_patients,
  n_trials = 4,
  patient_trial = sample(4, n_patients, replace = TRUE), 
  n_patient_tumors = 1 + rbinom(n_patients, 5, 0.4), # We're not yet generatively modeling the number of tumors
  n_measures = rep(patient_measures, sum(n_patient_tumors)),
  t_measure = rep(0:(patient_measures - 1), sum(n_patient_tumors)) - t_offset,
  tumor_size = rep(0, sum(n_measures)),

  !!!tumor_priors  
)
```


```{r}
#| label: tumor-prior-predict
#| warning: false
#| cache: true
#| dependson: test-data 

tumor_prior_res <- tumor_model$sample(
  data = test_data, 
  iter_warmup = 500, iter_sampling = 500, 
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "tumor_prior", 
  refresh = 0
)

# Saving this to use in SBC runs.
tumor_prior_res$save_data_file(dir = file.path(tmp_dir, "data"), basename = "prior_tumor", timestamp = FALSE, random = FALSE)
```

```{r}
#| label: tumor-prior-predict-results
#| cache: true
#| cache-lazy: false
  
# This is the tumor data I'll be using in the PFS model
fake_tumor_data <- tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index], ndraws = 1) |> 
  prepare_tumor_size_rvars(test_data) |> 
  unnest_rvars()

# This will be used by SLURM jobs that need tumor data (e.g., PFS SBC run).
write_rds(fake_tumor_data, file.path(tmp_dir, "data", "fake_tumor.rds"))
```


```{r}
#| label: tumor-model-diagnose
#| cache: true
#| dependson: tumor-prior-predict

tumor_prior_res$cmdstan_diagnose()
```

```{r}
#| label: tumor-prior-predict-plot
#| fig-cap: "Prior prediction checking for tumor model."
#| cache: true
#| dependson: tumor-prior-predict

tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index]) |>
  ungroup() |> 
  prepare_tumor_size_rvars(test_data) |> 
  filter(patient_id == 1, tumor_index < 6) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = tumor_size, group = str_c(patient_id, tumor_index, .draw)), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 5)) +
  stat_lineribbon(aes(ydist = tumor_size, group = tumor_index), fill = "black", alpha = 0.25, .width = 0.8) +
  scale_x_continuous(breaks = seq(0, 40, 4)) +
  labs(title = "Prior predicted tumor sizes", y = "Tumor Size [cm]", caption = "Lines show tumors sizes for a subset of tumors of a single patient.") +
  facet_wrap(vars(tumor_index)) +
  coord_cartesian(ylim = c(0, 10)) +
  theme(strip.text = element_blank()) +
  NULL
```

## Fake Data Simulation

Getting a few fake datasets and assessing the model's ability to recover parameter values.

```{r}
#| label: tumor-fake-sim
#| output: false
#| cache: true
#| dependson: test-data, priors

tumor_fake_res <- tumor_prior_res |> 
  spread_rvars(rep_tumor_size[tumor_index], ndraws = 24) |> # We'll simulate parameters and data for a number simulations. 
  ungroup() |> 
  prepare_tumor_size_rvars(test_data) |> 
  unnest_rvars() |> 
  nest(fake_data = !c(.draw, .iteration, .chain)) |> 
  mutate(
    fake_fit = furrr::future_map2(.draw, fake_data, function(id, d) {
    # fake_fit = map2(.draw, fake_data, function(id, d) {
      test_data |> 
        list_modify(
          fit_data = TRUE,
          gen_tumor_sizes = FALSE,
          tumor_size = d$tumor_size
        ) |> 
        tumor_model$sample(
          iter_warmup = 200, iter_sampling = 200,
          parallel_chains = 4,
          output_dir = file.path(tmp_dir, "fit"), 
          output_basename = str_c("fake_tumor_", id), 
          init = \(chain_id) lst(
            pop_tumor_gp_rho = invgamma::rinvgamma(1, 7.3, 7.5),
            tumor_mean = rnorm(1, 2.8, 0.1),
            tumor_sd = abs(rnorm(1, 0, 1.25))
          ),
          max_treedepth = 15
        )
    # })
    }, .progress = FALSE, .options = furrr::furrr_options(seed = TRUE))
  ) 
```

```{r}
#| eval: false
#| include: false

f <- tumor_fake_res$fake_fit[[1]]
f_lp <- log_posterior(f)
f_np <- nuts_params(f)

mcmc_trace(f$draws(inc_warmup = TRUE), pars = "tumor_mean", n_warmup = 500) #, np = f_np)
```


```{r}
#| label: tumor-fake-data-params
#| fig-cap: "Fake data simulation assessing model's ability to recover posterior of model parameters."
#| cache: true
#| dependson: tumor-fake-sim

tumor_fake_res |> 
  group_by(.draw) |> 
  reframe(map_dfr(fake_fit, \(f) gather_rvars(f, pop_tumor_gp_intercept, pop_tumor_gp_alpha, pop_tumor_gp_rho))) |> 
  left_join(
    tumor_prior_res |> 
      gather_rvars(pop_tumor_gp_intercept, pop_tumor_gp_alpha, pop_tumor_gp_rho) |> 
      unnest_rvars(), by = c(".draw", ".variable"), suffix = c("", "_true")) |> 
  mutate(.value = .value - .value_true) |> 
  ggplot(aes(y = .draw)) +
  stat_pointinterval(aes(xdist = .value), .width = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Comparing posterior estimate of model parameters against known values", 
    x = "Difference from True Parameter Value", 
    y = "", 
    caption = str_glue("Showing {nrow(tumor_fake_res)} simulations.")
  ) +
  facet_wrap(vars(.variable), scales = "free") +
  theme(axis.text.y = element_blank()) +
  NULL  
```


## Fit Data

Now using actual trial data.

```{r}
#| label: tumor-fit
#| output: false
#| cache: true
#| dependson: prepare-analysis-data, priors

treated_tumor_analysis_data <- pfs_analysis_data |> 
  filter(treated) 
  
tumor_stan_data <- lst(
  fit_tumor_data = TRUE,
  gen_tumor_sizes = TRUE,
  predict_missing_sizes = TRUE, 
  multilevel_patient = TRUE,
  multilevel_tumor = FALSE,
  
  !!!prepare_tumor_stan_data(treated_tumor_analysis_data),
  !!!tumor_priors
)

tumor_fit_res <- tumor_model$sample(
  tumor_stan_data, 
  iter_warmup = 200, iter_sampling = 200, 
  max_treedepth = 15,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "tumor" 
)

```

```{r}
#| label: fitted-tumor-data
#| cache: true
#| dependson: tumor-fit

n_full_measures <- with(tumor_stan_data, n_measures + pfs_model$functions$calculate_n_missing_measures(n_measures, t_measure, n_patient_tumors))

all_tumor_size <- tumor_fit_res |> 
  spread_rvars(all_tumor_size[idx]) |> 
  mutate(
    usubjid = rep(with(treated_tumor_analysis_data, rep(usubjid, n_tumors)), n_full_measures),
    trlnkid = rep(treated_tumor_analysis_data |> unnest(patient_tumors) |> pull(trlnkid), n_full_measures),
  ) |> 
  left_join(
    treated_tumor_analysis_data |> 
      unnest(patient_tumors) |> 
      select(usubjid, trlnkid, trial, matches("(patient_)?(min|max)_t")), 
    by = c("usubjid", "trlnkid")
  ) |> 
  group_by(usubjid, trlnkid) |> 
  mutate(week = first(min_t):first(patient_max_t)) |> 
  ungroup() |> 
  median_qi(all_tumor_size, .width = 0.8)
```

```{r}
#| label: fig-fit-tumor-history-plot
#| fig-height: 8
#| fig-cap: Showing GP tumor size dynamics with interpolation.
#| cache: true
#| dependson: fitted-tumor-data 

all_tumor_size |> 
  nest(patient_data = !c(trial, usubjid)) |> 
  group_by(trial) |> 
  sample_n(3) |> 
  ungroup() |> 
  unnest(patient_data) |> 
  ggplot(aes(week)) +
  geom_lineribbon(aes(y = all_tumor_size, ymin = .lower, ymax = .upper, group = trlnkid), fill = "black", alpha = 0.25) +
  facet_wrap(vars(usubjid), scales = "free", ncol = 3) +
  labs(title = "Showing GP tumor size dynamics with interpolation", x = "Week", y = "Tumor Size [cm]") +
  theme(strip.text = element_blank())
```


# Survival Model

We will use a piecewise-constant proportional hazard model:

-   Time is discretized into $t \in \{0,\dots,T\}$, with $t = 0$ is the baseline interval.
-   $Y_{it}$ is an indicator for disease progression at interval $t$, with $Y_{i0} = 0$.
-   $S_i$ is the number of intervals observed without disease progression, \[ \Pr[S > s] = \prod\_{r=0}\^s \Pr[Y_{i,r+1} = 0 \mid Y_{i,r} = 0]. \]
-   $X_{it}$ are--possibly time-variant--covariates.
-   $C_i$ is an indicator of whether $i$ is right censored.
-   $M_{i} \in \mathbb{Z}^{+}$ is the difference between $S_i$ and the previous interval $i$ was observed progression free ($M_i = 1$ if $C_i = 1$).

The probability of remaining progression-free at $m \geq 1$ is 
$$
\alpha_t(x;\theta) = \Pr[Y_t = 0\mid Y_{t-1} = 0, X_t = \bf{x}] = \exp[-\lambda(t, \bf{x}; \boldsymbol{\theta})] 
$$

and the hazard function . 
$$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t \kappa(\bf{x}; \boldsymbol{\beta})
$$

We assume that censoring in conditionally independent of time to progression. We will use a linear form for $\kappa(\bf{x}; \boldsymbol{\beta}) = \exp(\bf{x}\boldsymbol{\beta})$.

Aside from whatever priors we use for $\boldsymbol{\beta}$, we will use a Gaussian process prior for $\lambda_t$, 
$$
\log(\lambda_t) \sim \mathtt{GP}(0, k),
$$ 
where $k(\lambda_p, \lambda_q; \alpha, \rho) = \alpha^2\exp\left(\frac{|p - q|^2}{2\rho^2}\right)$.

Our likelihood, accounting for interval and right censoring, is 
$$
\mathcal{L}_i = \begin{cases}
  \prod_{t=1}^{S_i-M_i}\alpha_t(x_i;\theta)\sum_{r=S_i-M_i+1}^{S_i + 1}\left\{\prod_{m=S_i - M_i + 1}^{r-1}\alpha_m(x_i;\theta)[1-\alpha_{r}(x_i;\theta)]\right\} & \textrm{if }C_i = 0 \\
  \prod_{t=1}^{S_i}\alpha_t(x_i;\theta) & \textrm{if }C_i = 1 
  \end{cases}
$$

## Prior Prediction

```{r}
#| label: pfs-prior-predict
#| cache: true
#| dependson: test-data, priors

pfs_test_data <- test_data |> 
  list_modify(
    fit_data = FALSE,
    gen_pfs = TRUE,
    gen_interval_censored = FALSE,
    tumor_hazard_type = 1,
    ignore_interval_censoring = FALSE,
    add_trial_level = TRUE,
    
    pfs = rep(max_pfs, n_patients),
    death_week = rep(0, n_patients),
    right_censored = rep(FALSE, n_patients),
    tumor_size = fake_tumor_data$tumor_size,
    
    !!!pfs_priors,
    
    # !!!tumor_stan_data
    use_tumor_model = FALSE
  )

t_measure_list <- with(pfs_test_data, list_measures(t_measure, n_measures, n_patient_tumors))  
max_t <- t_measure_list |> 
  map(unlist) |> 
  map_int(max)
first_t_measure <- t_measure_list |> map(first) |> unlist()

pfs_prior_res <- pfs_model$sample(
  pfs_test_data, 
  refresh = 0, 
  iter_warmup = 500, iter_sampling = 500,
  output_dir = file.path(tmp_dir, "fit"), 
  output_basename = "pfs_prior" 
)

rep_pfs_rvar <- pfs_prior_res |> 
  spread_rvars(rep_pfs[patient_index], rep_right_censored[patient_index])

rep_pfs_param_rvar <- pfs_prior_res |> 
  spread_rvars(
    tumor_stim_trial_intercept[trial],
    tumor_stim_trial_coef[trial, t],
    log_lambda_gp_alpha, 
    log_lambda_gp_rho, 
    log_lambda_gp_intercept
  ) |> 
  pivot_wider(names_from = t, values_from = tumor_stim_trial_coef, names_prefix = "tumor_stim_coef_") 

lambda_rvar <- pfs_prior_res |> 
  spread_rvars(log_lambda[t], base_pf_cond_prob[t], base_survival[t], one_tumor_survival[t], log_trial_lambda[trial, t]) |> 
  mutate(lambda = exp(log_lambda), trial_lambda = exp(log_trial_lambda)) |> 
  nest(trial_data = contains("trial"))

expected_pfs_rvar <- pfs_prior_res |> 
  gather_rvars(base_cond_expected_pfs, one_tumor_cond_expected_pfs, base_cond_median_pfs, one_tumor_cond_median_pfs) |> 
  mutate(
    stat_type = str_extract(.variable, "expected|median"),
    tumor_size = str_extract(.variable, "base|one_tumor")
  )
```

```{r}
#| label: fig-prior-lambda
#| cache: true
#| dependson: pfs-prior-predict

lambda_rvar |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 50)) +
  labs(title = "Base hazard", x = "m", y = latex2exp::TeX("$\\lambda_t$"))
```

```{r}
#| label: fig-prior-trial-lambda
#| dependson: pfs-prior-predict

lambda_rvar |>
  unnest(trial_data) |>
  select(t, trial, trial_lambda) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = trial_lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = trial_lambda, group = .draw), alpha = 0.125, data = \(d) unnest_rvars(d) |> group_by(trial) |>  filter(.draw <= 75)) +
  labs(title = "Base hazard", x = "m", y = latex2exp::TeX("$\\lambda_t$")) +
  facet_wrap(vars(trial)) +
  coord_cartesian(ylim = c(0.0, 3)) +
  NULL
```

```{r}
#| label: fig-prior-base-pf-prob
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  gather_rvars(trial_base_pf_cond_prob[s, t]) |> 
  filter(s == 1) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = .value), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = .value, group = .draw), alpha = 0.125, data = \(d) unnest_rvars(d) |> filter(.draw <= 75)) +
  labs(title = "Base conditional probability of remaining progress free", x = "t", y = latex2exp::TeX(r"{$\alpha_t$}")) 
```

```{r}
#| label: pf-prob-plot
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  spread_rvars(disease_progress_prob[i]) |>
  mutate(
    patient_index = rep(seq(pfs_test_data$n_patients), each = max(max_t)), 
    trial = rep(pfs_test_data$patient_trial, each = max(max_t)), 
    t = rep(1:max(max_t), pfs_test_data$n_patients),
  ) |>
  nest(.by = c(trial, patient_index)) |> 
  group_by(trial) |> 
  sample_n(2) |> 
  ungroup() |> 
  unnest(data) |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = disease_progress_prob), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = disease_progress_prob, group = str_c(patient_index, .draw)), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 20)) +
  labs(title = "Conditional probability of remaining progress free", x = "t", y = latex2exp::TeX("$\\alpha_t$")) +
  # coord_cartesian(ylim = c(0, 1)) +
  facet_wrap(vars(patient_index, trial), scales = "free_y") +
  NULL
```

```{r}
pfs_prior_res |> 
  gather_rvars(trial_base_pf_cond_prob[s, t], trial_one_tumor_pf_cond_prob[s, t]) |> 
  filter(s == 1, t <= 10) |> 
  ggplot() +
  stat_halfeye(aes(x = t, ydist = .value, fill = .variable), alpha = 0.25, .width = 0.8)
```
```{r}
pfs_prior_res |> 
  gather_rvars(tumor_stim_intercept, tumor_stim_coef[t], tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, t]) |>  
  mutate(
    .variable = if_else(is.na(trial), .variable, str_glue("{.variable}[{trial}]")),
    .variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]")),
  ) |> 
  ggplot() +
  stat_halfeye(aes(xdist = .value, y = .variable), color = "black", alpha = 0.5, .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  # coord_cartesian(xlim = c(0, 1)) +
  NULL
```


```{r}
#| label: base-surivival-plot
#| dependson: pfs-prior-predict

lambda_rvar |>
  select(t, base_survival, one_tumor_survival) |> 
  pivot_longer(!t, names_to = "survival_type", values_to = "survival") |> 
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = survival, fill = survival_type, color = survival_type), step = TRUE, alpha = 0.25, .width = 0.8) +
  labs(title = "Base survival probability", x = "t", y = latex2exp::TeX("$S(t)$"))
```

```{r}
#| label: cond-expected-pfs-plot
#| dependson: pfs-prior-predict

expected_pfs_rvar |> 
  mutate(.variable = str_remove(.variable, "_cond_expected_pfs")) |> 
  ggplot(aes(y = tumor_size)) +
  stat_pointinterval(aes(xdist = .value, color = stat_type), .width = 0.8, position = "dodge") +
  labs(title = "Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "t", y = "")
```

```{r}
#| label: prior-kaplan-meier 
#| dependson: pfs-prior-predict

pfs_prior_res |> 
  spread_rvars(km_est[m]) |> 
  ggplot(aes(m - 1)) +
  stat_lineribbon(aes(ydist = km_est), step = TRUE, fill = "black", alpha = 0.25, .width = 0.8) +
  labs(title = "Kaplan-Meier estimate", x = "t", y = latex2exp::TeX("$S(t)$")) +
  # scale_x_continuous("t", breaks = seq(0, 40, 2))
  NULL
```

## Fake Data Simulation

Here we're using a set of ad hoc survival model parameters and generating a number of datasets, then we compare the posterior distributions of these parameters against the known truth. We do this three times:

1.  Without any interval censoring
2.  With interval censoring but ignored by the model (observed PFS is assumed to be actual PFS)
3.  With interval censoring which is accounted for by the model

```{r}
#| label: fake-data-sim
#| output: false
#| cache: true 

n_sim <- 12  
t_to_keep <- c(-t_offset, 0, seq(1 - t_offset, patient_measures - t_offset - 1, by = 5)) |> sort()
fake_data_sim_settings <- pfs_test_data # |> list_assign(add_trial_level = FALSE)

no_interval_censoring <- run_sbc_sims(pfs_model, fake_data_sim_settings, n_sim, output_name = "fake_data_sim", keep_fit = TRUE, gen_pfs = TRUE)

fake_data_sim <- bind_rows(
  no_interval_censoring = no_interval_censoring,
  
  interval_censoring = run_sbc_sims(
    pfs_model, fake_data_sim_settings, n_sim, ignore_interval_censoring = FALSE, output_name = "fake_data_sim_ic",
    keep_fit = TRUE, gen_pfs = TRUE, reuse_data = no_interval_censoring, drop_measures = t_to_keep, keep_only = TRUE
  ),
  
  interval_censoring_ignored = run_sbc_sims(
    pfs_model, fake_data_sim_settings, n_sim, ignore_interval_censoring = TRUE, output_name = "fake_data_sim_ic_ignored",
    keep_fit = TRUE, gen_pfs = TRUE, reuse_data = no_interval_censoring, drop_measures = t_to_keep, keep_only = TRUE
  ),
  
  .id = "sim_type"
) |> 
  rename(sim_id = .draw)
```

```{r}
#| eval: false
#| include: false

fake_data_sim |> 
  rowwise() |> 
  reframe(
    sim_id, sim_type,
    spread_rvars(sim_fit, km_est[t]),
    obs_km = with(sim_data, pfs_model$functions$estimate_kaplan_meier(pfs, right_censored, max(t) - 1)[[1]]),
  ) |>
  filter(sim_id <= 4) |> 
  ggplot(aes(t)) +
  geom_step(aes(y = obs_km, color = sim_type)) +
  stat_lineribbon(aes(ydist = km_est, fill = sim_type, color = sim_type), alpha = 0.25, .width = 0.8) +
  facet_wrap(vars(sim_id), scales = "free_y")
```

```{r}
#| label: fake-data-params
#| fig-cap: "Comparing fake data simulation parameter posteriors against their known true values."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  transmute(
    sim_id,
    sim_type,
    
    map2_dfc(est, select(true, all_of(names(est))), \(e, t) e - t),
  ) |> 
  # add_row(sim_id = 0, sim_type = "prior", rep_pfs_param_rvar) |>
  pivot_longer(!c(sim_id, sim_type), names_to = "param", values_to = "est") |>
  ggplot(aes(sim_id)) + 
  geom_hline(yintercept = 0, linetype = "dotted") +
  stat_pointinterval(aes(ydist = est, color = sim_type), size = 0.25, .width = 0.8, fatten_point = 0, position = "dodge") +
  scale_color_discrete("") +
  facet_wrap(vars(param), scales = "free_y") +
  labs(x = "", y = "", caption = "Black horizontal line represents the true parameter value.") +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

```{r}
#| label: fake-data-lambda
#| fig-cap: "Comparing fake data simulation base hazard posteriors against their known true values."
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |> 
  rowwise() |> 
  reframe(sim_id, sim_type, spread_rvars(sim_fit, log_lambda[m])) |> 
  # mutate(sim_log_lambda_post = furrr::future_map(sim_fit, \(f) spread_rvars(f, log_lambda[m]))) |> 
  # select(!sim_fit) |>
  # unnest(sim_log_lambda_post) |> 
  # bind_rows(transmute(lambda_rvar, m = t, log_lambda, sim_id = 0, sim_type = "prior")) |> 
  mutate(lambda = exp(log_lambda)) |> 
  filter(m <= 20) |> 
  ggplot(aes(sim_id)) + 
  # geom_hline(aes(yintercept = true_val), data = \(d) distinct(d, m) |> mutate(true_val = exp(log_lambda[m]))) +
  stat_pointinterval(
    aes(ydist = lambda, color = sim_type), 
    size = 0.5, .width = 0.8, fatten_point = 0, position = "dodge"
  ) +
  labs(x = "", y = "", caption = "Black horizontal line represents the true parameter value.") +
  facet_wrap(vars(m), scales = "free_y") +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

```{r}
#| eval: false

fake_data_sim |>
  filter(sim_id <= 4) |> 
  mutate(sim_log_lambda_post = furrr::future_map(sim_fit, \(f) spread_rvars(f, log_lambda[t]))) |> 
  select(!sim_fit) |>
  unnest(sim_log_lambda_post) |>
  ggplot(aes(t)) +
  stat_lineribbon(aes(ydist = log_lambda), fill = "black", alpha = 0.25, .width = 0.8) +
  geom_line(aes(y = log_lambda, group = .draw), alpha = 0.25, data = \(d) unnest_rvars(d) |> filter(.draw <= 10)) +
  geom_line(aes(y = log_lambda), color = "red", data = tibble(t = 1:max_all_t, log_lambda = log_lambda[1:max_all_t])) +
  facet_grid(vars(sim_type), vars(sim_id))
```


```{r}
#| label: fake-data-km
#| fig-cap: "Comparing the fake data Kaplan-Meier survival probability against the true probability."
#| cache: true
#| dependson: fake-data-sim
#| eval: false

fake_data_sim |> 
  mutate(km_est = furrr::future_map(sim_fit, \(f) spread_rvars(f, km_est[m]))) |> 
  select(sim_id, sim_type, km_est) |> 
  filter(sim_id <= 6) |> 
  unnest(km_est) |> 
  ggplot(aes(m - 1)) +
  # geom_line(aes(y = km_est), data = \(sim) semi_join(true_km, sim, by = "m")) +
  # This comes from a draw so not really reliable as the truth.
  stat_lineribbon(aes(ydist = km_est, color = sim_type, fill = sim_type), step = TRUE, alpha = 0.25, .width = 0.8) +
  scale_x_continuous("t", breaks = seq(0, 40, 4)) +
  labs(y = "S(t)", caption = "Black line represents the true survival probability.") +
  facet_wrap(vars(sim_id)) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fake-data-cond-expected-pfs-plot
#| cache: true
#| dependson: fake-data-sim

fake_data_sim |>
  rowwise() |> 
  mutate(map2_dfc(est, select(true, all_of(names(est))), \(e, t) e - t)) |> 
  select(sim_id, sim_type, matches("(expected|median)_pfs")) |> 
  pivot_longer(matches("(expected|median)_pfs")) |> 
  # reframe(sim_id, sim_type, gather_rvars(sim_fit, base_cond_expected_pfs, one_tumor_cond_expected_pfs, base_cond_median_pfs, one_tumor_cond_median_pfs)) |> 
  # mutate(
  #   cond_stat = furrr::future_map(
  #     sim_fit, \(f) gather_rvars(f, base_cond_expected_pfs, one_tumor_cond_expected_pfs, base_cond_median_pfs, one_tumor_cond_median_pfs)
  #   )
  # ) |> 
  # unnest(cond_stat) |> 
  # add_row(sim_id = 0, sim_type = "prior", expected_pfs_rvar) |> 
  mutate(
    stat = str_extract(name, "expected|median"),
    name = str_remove(name, "_cond_(expected|median)_pfs"),
    est_type = if_else(sim_id > 0, "post", "prior")
  ) |> 
  ggplot(aes(y = sim_id)) +
  stat_pointinterval(aes(xdist = value, color = sim_type), size = 0.5, .width = 0.8, position = "dodge") +
  # geom_vline(
  #   aes(xintercept = true_e), 
  #   data = tibble(
  #     .variable = rep(c("base", "one_tumor"), 2), 
  #     stat = rep(c("expected", "median"), each = 2),
  #     true_e = c(true_base_cond_expected_pfs, true_one_tumor_cond_expected_pfs, true_base_cond_median_pfs, true_one_tumor_cond_median_pfs)
  #   )
  # ) +
  geom_vline(xintercept = 0.0, linetype = "dotted") +
  scale_color_discrete("") +
  labs(title = "Fake Data Expected PFS", subtitle = "Conditional on progress within maximum observed PFS", x = "m", y = "") +
  # facet_wrap(vars(.variable)) +
  facet_grid(vars(stat), vars(name)) +
  NULL
```

## Simulation-Based Calibration

With SBC we use the prior to generate a number of datasets that we use the same model and prior to fit. We then calculate the number of samples that are less than the true value (from the prior-only run generating the data). If the model and algorithm are working correctly, the distribution of the number draws that are less than the true parameter values should approach a uniform distribution.

The code for this simulation is located in `sbc.R` (and `sbc.sh` to run on SLURM).

```{r}
#| label: sbc

sbc_draws <- 1000
sbc_bw <- 10

ic_labels <- c(ic = "IC Adjusted", ic_ignore = "IC Ignored", no_ic = "No IC")

sbc_data <- bind_rows( 
  no_ic = read_rds(file.path(tmp_dir, "data", "sbc3.rds")),
  ic = read_rds(file.path(tmp_dir, "data", "sbc3_ic.rds")),
  ic_ignore = read_rds(file.path(tmp_dir, "data", "sbc3_ic_ignore.rds")),
  .id = "sim_type") |> 
  select(sim_type, starts_with("r_"))

sbc_data |> 
  pivot_longer(!c(sim_type), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  filter(!str_detect(param, "pfs$")) |>
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = sbc_bw) +
  geom_rect(aes(xmin = -10, xmax = sbc_draws, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = count(sbc_data, sim_type) |> mutate(map_dfr(n, \(n) set_names(qbinom(c(0.01, 0.5, 0.99), n, sbc_bw / sbc_draws), c("lb", "med", "ub"))))) +
  facet_grid(rows = vars(sim_type), cols = vars(param), labeller = labeller(sim_type = ic_labels)) +
  labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  coord_cartesian(ylim = c(0, 20))

sbc_data |> 
  pivot_longer(!c(sim_type), names_to = "param", names_prefix = "r_", values_to = "r") |> 
  filter(str_detect(param, "pfs$")) |>
  mutate(
    stat = str_extract(param, "expected|median"),
    tumor_state = str_extract(param, "one_tumor|base")
  ) |> 
  ggplot() +
  geom_histogram(aes(r), alpha = 0.5, binwidth = sbc_bw) +
  geom_rect(aes(xmin = -10, xmax = sbc_draws, ymin = lb, ymax = ub), 
            alpha = 0.25, fill = "pink",
            data = count(sbc_data, sim_type) |> mutate(map_dfr(n, \(n) set_names(qbinom(c(0.01, 0.5, 0.99), n, sbc_bw / sbc_draws), c("lb", "med", "ub"))))) +
  facet_grid(rows = vars(sim_type), cols = vars(stat, tumor_state), labeller = labeller(sim_type = ic_labels)) +
  labs(title = "Histogram of rank statistics", y = "", caption = str_glue("The histogram uses a bin width of 10.\nThe pink area represents the 1%-99% region of a Binom({sbc_draws}, {sbc_bw / sbc_draws}).\n{sbc_draws} simulation used per model.")) +
  coord_cartesian(ylim = c(0, 30))

# sbc_data |> 
#   select(starts_with("r_")) |> 
#   pivot_longer(everything(), names_to = "param", names_prefix = "r_", values_to = "r") |> 
#   ggplot() +
#   stat_ecdf(aes(r)) +
#   facet_wrap(vars(param)) +
#   coord_cartesian(xlim = c(0, 1000))
```

It is clear that ignoring interval censoring degrades the model's ability to recover the prior parameter values; simulations show a lopsided distribution that significantly deviates from the uniform distribution.

## Fit Data

```{r}
hist_breaks <- seq(10, 150, by = 10)
```

### No Tumors Model

The hazard function does not use any covariates: 

$$
  \lambda(t, \bf{x}; \boldsymbol{\theta}) = \lambda_t
$$

```{r}
#| label: no-tumor-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors
#| output: false

no_tumor_treated_pfs_stan_data <- treated_pfs_stan_data |> 
  list_assign(tumor_hazard_type = 0, add_trial_level = TRUE)

prior_no_tumor_pfs_future <- future::future(
  no_tumor_treated_pfs_stan_data |> 
    list_modify(fit_data = FALSE) |> 
    pfs_model$sample(
      parallel_chains = 4,
      iter_warmup = 300, iter_sampling = 300, refresh = 0,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_no_tumor_pfs" 
    ),
  seed = TRUE
)

no_tumor_pfs_future <- future::future(
  pfs_model$sample(
    no_tumor_treated_pfs_stan_data, 
    parallel_chains = 4,
    iter_warmup = 300, iter_sampling = 300, refresh = 0,
    output_dir = file.path(tmp_dir, "fit"), 
    output_basename = "no_tumor_pfs" 
  ),
  seed = TRUE
)

ic_ignored_no_tumor_pfs_future <- future::future(
  no_tumor_treated_pfs_stan_data |>
    list_modify(ignore_interval_censoring = TRUE) |>
    pfs_model$sample(
      parallel_chains = 4,
      iter_warmup = 300, iter_sampling = 300, refresh = 0,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "ic_ignored_no_tumor_pfs" 
    ),
  seed = TRUE
)

prior_no_tumor_pfs_res <- future::value(prior_no_tumor_pfs_future)
no_tumor_pfs_res <- future::value(no_tumor_pfs_future)
```

```{r}
#| label: fig-no-tumor-kaplan-meier
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res, control = NULL) |> # control_no_tumor_pfs_res) |> 
  compact() |> 
  plot_km(treated_pfs_analysis_data) +
  geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) 
```

```{r}
#| label: fig-no-tumor-pfs-hist-post
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res) |> #, control = control_no_tumor_pfs_res) |> 
  compact() |> 
  plot_pfs_hist_posterior(no_tumor_treated_pfs_stan_data) +
  stat_bin(aes(pfs, color = "Observed"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data) +
  theme(legend.position = "bottom") +
  NULL
```

```{r}
#| label: fig-no-tumor-mean-median
#| cache: true
#| dependson: no-tumor-pfs-fit

lst(treated = no_tumor_pfs_res) |> #, control = control_no_tumor_pfs_res) |> 
  compact() |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = no_tumor_treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "arm") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = trial), position = "dodge", .width = 0.8) +
  scale_color_discrete("Trial") +
  labs(title = "Posterior PFS mean and median", subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```

```{r}
#| label: fig-no-tumor-base-hazard
#| cache: true
#| dependson: no-tumor-pfs-fit
#| fig-cap: Baseline hazard for model with no tumor effects.

lst(treated = no_tumor_pfs_res, prior = prior_no_tumor_pfs_res) |> #  control = control_no_tumor_pfs_res,  
  plot_base_hazard(treated_pfs_analysis_data)
```


#### Looking at Interval Censoring

Running the model with and without adjustment for interval censoring.

```{r}
#| label: fit-ignore-interval-censoring
#| cache: true
#| dependson: no-tumor-pfs-fit
#| output: false

ic_ignored_no_tumor_pfs_res <- future::value(ic_ignored_no_tumor_pfs_future)
```

```{r}
#| label: fig-no-tumor-kaplan-meier-ic-compare 
#| cache: true
#| dependson: fit-ignore-interval-censoring

lst(ic = no_tumor_pfs_res, ic_ignored = ic_ignored_no_tumor_pfs_res) |> 
  plot_km(treated_pfs_analysis_data) +
  scale_fill_viridis_d(aesthetics = c("fill", "color")) 
```

```{r}
#| label: fig-no-tumor-mean-median-ic-compare 
#| cache: true
#| dependson: fit-ignore-interval-censoring

lst(ic = no_tumor_pfs_res, ic_ignored = ic_ignored_no_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = no_tumor_treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "model") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = trial), position = "dodge", .width = 0.8) +
  scale_color_discrete("Trial") +
  facet_wrap(vars(model), ncol = 1) +
  labs(title = "Posterior PFS mean and median", subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```


While SBC shows that ignoring interval censoring should be a big problem, it does not appear to bias estimates much using our data.

### Tumor Count Model

The hazard function is time invariant:

$$
\begin{align}
  \lambda(t, \bf{x}; \boldsymbol{\theta}) &= \lambda_t \cdot \exp\left[\beta_0(N_i - \bar{N})\right] \\ 
  \end{align}
$$

### First Two Tumor Sizes Only Model

The hazard function is time invariant:

$$
\begin{align}
  \lambda(t, \bf{x}; \boldsymbol{\theta}) &= \lambda_t \cdot \exp\left[\beta_0(N_i - \bar{N}) + \sum_m \beta_1 \bar{x}_{m0} + \beta_2 \bar{x}_{m1} + \beta_3 (\bar{x}_{m0} \times \bar{x}_{m1})\right] \\ 
  &= \lambda_t \cdot \underbrace{\exp\left[-\beta_0 \bar{N} + \sum_m \beta_0 + \beta_1 \bar{x}_{m0} + \beta_2 \bar{x}_{m1} + \beta_3 (\bar{x}_{m0} \times \bar{x}_{m1})\right]}_{\kappa} \\
  \end{align}
$$
where $\bar{N}$ is the average number of tumors per patient and, for every tumor $m$, $\bar{x}_{m0}$ is the standardized last screening assessment and $\bar{x}_{m1}$ is the standardized first assessment after treatment begins.^[$\bar{x}_{mt} = \frac{x_{mt} - \bar{x}}{s(x)}$] We subtract $\beta_0 \bar{N}$ 

```{r}
#| label: two-tumor-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors 
#| output: false

two_tumor_treated_pfs_stan_data <- treated_pfs_stan_data |> 
  list_assign(tumor_hazard_type = 2, add_trial_level = TRUE)

two_tumor_init <- create_pfs_initializer(two_tumor_treated_pfs_stan_data) 

prior_two_tumor_pfs_future <- future::future(
  two_tumor_treated_pfs_stan_data |> 
    list_assign(fit_data = FALSE, add_trial_level = TRUE) |> 
    pfs_model$sample(
      iter_warmup = 300, iter_sampling = 300, 
      init = two_tumor_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_two_tumor_pfs" 
    ),
  seed = TRUE
)

two_tumor_pfs_future <- future::future(
    pfs_model$sample(
      two_tumor_treated_pfs_stan_data,
      iter_warmup = 300, iter_sampling = 300, 
      init = two_tumor_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "two_tumor_pfs" 
    ),
  seed = TRUE
)

# control_two_tumor_pfs_future <- future::future(
#   pfs_analysis_data |> 
#     filter(!treated) |> 
#     prepare_pfs_stan_data(tumor_priors, pfs_priors, tumor_hazard_type = 2) |>  
#     pfs_model$sample(
#       iter_warmup = 500, iter_sampling = 500,
#       init = two_tumor_init,
#       parallel_chains = 4,
#       output_dir = file.path(tmp_dir, "fit"), 
#       output_basename = "control_two_tumor_pfs" 
#     ),
#   seed = TRUE
# )

prior_two_tumor_pfs_res <- future::value(prior_two_tumor_pfs_future)
two_tumor_pfs_res <- future::value(two_tumor_pfs_future)
control_two_tumor_pfs_res <- NULL # future::value(control_two_tumor_pfs_future)

```

```{r}
#| label: fig-two-tumor-kaplan-meier
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  plot_km(treated_pfs_analysis_data) +
  geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) 
```

```{r}
#| label: fig-two-tumor-pfs-hist-post
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res) |> #, control = control_no_tumor_pfs_res) |> 
  compact() |> 
  plot_pfs_hist_posterior(two_tumor_treated_pfs_stan_data) +
  stat_bin(aes(pfs, color = "Observed"), geom = "step", linewidth = 0.5, breaks = c(0, hist_breaks), data = treated_pfs_analysis_data) +
  theme(legend.position = "bottom") +
  NULL
```

```{r}
#| label: two-tumor-mean-median
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, rep_pfs[i]) |> 
            mutate(trial = two_tumor_treated_pfs_stan_data$patient_trial) |> 
            group_by(trial) |> 
            summarize(rep_mean_pfs = rvar_mean(rep_pfs), rep_median_pfs = rvar_median(rep_pfs)), 
          .id = "model") |> 
  pivot_longer(rep_mean_pfs:rep_median_pfs, names_to = "stat", names_pattern = "(mean|median)") |> 
  ggplot() +
  stat_pointinterval(aes(xdist = value, y = stat, color = trial), position = "dodge", .width = 0.8) +
  scale_color_discrete("Trial") +
  facet_wrap(vars(model), ncol = 1) +
  labs(title = "Posterior PFS mean and median", subtitle = "Treated arm", x = "PFS [weeks]", y = "")
```

```{r}
#| label: two-tumor-base-hazard
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(treated = two_tumor_pfs_res, prior = prior_two_tumor_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  plot_base_hazard(treated_pfs_analysis_data)
```

```{r}
#| label: two-tumor-coef
#| cache: true
#| dependson: two-tumor-pfs-fit

lst(prior = prior_two_tumor_pfs_res, post = two_tumor_pfs_res) |> 
  map(\(f) recover_types(f, select(treated_pfs_analysis_data, trial))) |> 
  map_dfr(
    \(f) gather_rvars(f, tumor_stim_intercept, tumor_stim_coef[t], tumor_stim_trial_intercept[trial], tumor_stim_trial_coef[trial, t]), 
    .id = "fit"
  ) |> 
  mutate(
    .variable = if_else(is.na(trial), .variable, str_glue("{.variable}[{trial}]")),
    .variable = if_else(is.na(t), .variable, str_glue("{.variable}[{t}]")),
  ) |> 
  ggplot() +
  stat_halfeye(aes(xdist = .value, y = .variable, fill = fit, color = fit), alpha = 0.25, .width = c(0.5, 0.8)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  coord_cartesian(xlim = c(0, 1))
```

```{r}
#| label: fig-in-sample-survival-comparison
#| fig-height: 10
#| fig-width: 6
#| cache: true
#| dependson: two-tumor-pfs-fit

pfs_range_data <- treated_pfs_analysis_data |> 
  select(usubjid, trial, pfs, interval_censored) |>
  mutate(usubjid = fct_reorder(usubjid, pfs)) 

pfs_range_data |> 
  ggplot(aes(y = usubjid)) +
  stat_lineribbon(
    aes(xdist = rep_pfs, fill = "Retrodictive Credible Interval"),
    alpha = 0.125, linewidth = 0, 
    .width = c(0.5, 0.8),
    data = \(obs_data) two_tumor_pfs_res |> 
      spread_rvars(rep_pfs[patient_index]) |> 
      bind_cols(select(pfs_range_data, trial, usubjid)) |> 
      semi_join(obs_data, by = "usubjid")
  ) +
  geom_errorbar(aes(xmin = pfs, xmax = pfs + interval_censored, color = "Observed Interval Censored Range")) +
  scale_color_discrete("PFS", aesthetics = c("color", "fill")) +
  facet_wrap(vars(trial), nrow = 1, scales = "free_y") +
  labs(x = "Weeks", y = "Patients") +
  theme(axis.text.y = element_blank(), panel.grid.major.y = element_blank(), legend.position = "top")
```

#### Variation in Tumor Predictive Power

In this section I'm trying to get a sense of how much the two first assessments of tumors contribute to the variation in the hazard and the probability of disease progression. Two plots are overlapped in these plots to: First, show the degree of variation in response to the tumor sizes, holding the model parameters fixed at the posterior median; and second, show the degree of variation using the full posterior distribution of the model parameters.

```{r}
#| label: predictor-variation
#| cache: true
#| cache-lazy: false 
#| dependson: two-tumor-pfs-fit

median_coef <- lst(post = two_tumor_pfs_res) |> 
  map_dfr(\(f) spread_rvars(f, tumor_stim_intercept, tumor_stim_coef), .id = "fit") |> 
  mutate(across(starts_with("tumor_stim"), median)) 

tumor_design_matrix <- with(
  treated_pfs_stan_data, {
    n_screening_t <- pfs_model$functions$calc_n_screening_t(n_patient_tumors, n_measures, t_measure)
    
    pfs_model$functions$standardize_nonzero_tumor_sizes(tumor_size) |>
      pluck(3) |> 
      pfs_model$functions$prepare_early_tumors_design_matrix(n_patient_tumors, n_measures, n_screening_t)
  }
)

tumor_design_matrix <- cbind(1, tumor_design_matrix, tumor_design_matrix[,1] * tumor_design_matrix[,2])

patient_kappa_data <- tibble(kappa = tumor_design_matrix %*% t(median_coef[, -1])) |> 
  mutate(patient_index = rep(seq(treated_pfs_stan_data$n_patients), treated_pfs_stan_data$n_patient_tumors)) |> 
  group_by(patient_index) |> 
  summarize(kappa = exp(-median_coef$tumor_stim_intercept[1] * mean(treated_pfs_stan_data$n_patient_tumors) + sum(kappa)))

patient_kappa_full_dist_data <- two_tumor_pfs_res |> 
  spread_rvars(total_time_invar_tumor_stim[patient_index]) |> 
  mutate(kappa = exp(total_time_invar_tumor_stim))

patient_kappa_data |> 
  mutate(var = "median") |> 
  bind_rows(
    patient_kappa_full_dist_data |>
      unnest_rvars() |> 
      mutate(var = "post")
  ) |> 
  ggplot(aes(kappa)) +
  geom_vline(xintercept = 1, linetype = "dotted") +
  geom_histogram(aes(y = after_stat(density), fill = var), position = "identity", alpha = 0.25, binwidth = 0.1) +
  scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(title = latex2exp::TeX(r"{Histogram of hazard $\kappa$ function}"), x = latex2exp::TeX(r"{$\kappa$}"), y = "Density") +
  coord_cartesian(xlim = c(0, 5)) +
  NULL
```

```{r}
#| label: dp-prob-variation
#| cache: true
#| dependson: two-tumor-pfs-fit

two_tumor_pfs_res |> 
  spread_rvars(log_lambda[t]) |> 
  mutate(
    lambda = exp(log_lambda),
    median_baseline_dp_prob = 1 - exp(- median(lambda)),
    patient_dp_prob = map(lambda, \(l) 1 - exp(- median(l) * patient_kappa_data$kappa)) %>% 
      { array(unlist(.), dim = c(nrow(patient_kappa_data), length(.))) } |> 
      rvar(),
    
    patient_full_dist_dp_prob = map(lambda, \(l) draws_of(1 - exp(- median(l) * patient_kappa_full_dist_data$kappa))) %>%
      { array(unlist(.), dim = c(prod(dim(first(.))), length(.))) } |> 
      rvar()
  ) |> 
  ggplot(aes(t)) +
  geom_line(aes(y = median_baseline_dp_prob)) +
  # Why didn't I use pivot? We can't have different sized rvars in the same column.
  stat_lineribbon(aes(ydist = patient_dp_prob, fill = "median"), alpha = 0.25, .width = 0.8) +
  stat_lineribbon(aes(ydist = patient_full_dist_dp_prob, fill = "post"), alpha = 0.25, .width = 0.8) +
  scale_fill_discrete("Variation", labels = c(median = "Posterior Median Parameters", post = "Full Posterior")) +
  labs(title = "Variation in the probability of disease progression", y = latex2exp::TeX(r"{1 - \alpha}"))
  NULL
```

### Tumor Size Change Model

The hazard function is time invariant:

$$
\begin{align}
  \lambda(t, \bf{x}; \boldsymbol{\theta}) &= \lambda_t \cdot \exp\left[\beta_0(N_i - \bar{N}) + \sum_m \beta_1 \frac{x_{m1} - x_{m0}}{x_{m0}}\right] \\ 
  &= \lambda_t \cdot \underbrace{\exp\left[-\beta_0 \bar{N} + \sum_m \beta_1 \frac{x_{m1} - x_{m0}}{x_{m0}}\right]}_{\kappa} 
\end{align}
$$

```{r}
#| label: tumor-change-pfs-fit
#| cache: true
#| dependson: fit-pfs-util, priors 
#| output: false

# tumor_change_init <- function(chain_id) { 
#   lst(tumor_stim_coef = truncnorm::rtruncnorm(1, 0, mean = 0, sd = 0.1))
# }

tumor_change_treated_pfs_stan_data <- treated_pfs_stan_data |> 
    list_modify(tumor_hazard_type = 3, add_trial_level = TRUE)

tumor_change_init <- create_pfs_initializer(tumor_change_treated_pfs_stan_data)

prior_tumor_change_pfs_future <- future::future(
  tumor_change_treated_pfs_stan_data |> 
    list_modify(fit_data = FALSE) |> 
    pfs_model$sample(
      iter_warmup = 300, iter_sampling = 300, 
      init = tumor_change_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "prior_tumor_change_pfs" 
    ),
  seed = TRUE
)

tumor_change_pfs_future <- future::future(
    pfs_model$sample(
      tumor_change_treated_pfs_stan_data,
      iter_warmup = 300, iter_sampling = 300, 
      init = tumor_change_init,
      parallel_chains = 4,
      output_dir = file.path(tmp_dir, "fit"), 
      output_basename = "tumor_change_pfs" 
    ),
  seed = TRUE
)

# control_tumor_change_pfs_future <- future::future(
#   pfs_analysis_data |> 
#     filter(!treated) |> 
#     prepare_pfs_stan_data(tumor_priors, pfs_priors, tumor_hazard_type = 3) |>  
#     pfs_model$sample(
#       iter_warmup = 500, iter_sampling = 500,
#       init = tumor_change_init,
#       parallel_chains = 4,
#       output_dir = file.path(tmp_dir, "fit"), 
#       output_basename = "control_tumor_change_pfs" 
#     ),
#   seed = TRUE
# )

prior_tumor_change_pfs_res <- future::value(prior_tumor_change_pfs_future)
tumor_change_pfs_res <- future::value(tumor_change_pfs_future)
# control_tumor_change_pfs_res <- future::value(control_tumor_change_pfs_future)
```

```{r}
#| label: tumor-change-kaplan-meier
#| cache: true
#| dependson: tumor-change-pfs-fit

lst(treated = tumor_change_pfs_res) |> #, control = control_two_tumor_pfs_res) |> 
  plot_km(treated_pfs_analysis_data) +
  geom_step(aes(x = t, y = s, group = btype), data = treated_obs_km) 
```

### Joint Model

TODO

### Compare Models

```{r}
#| label: compare-base-hazard
#| cache: true
#| dependson: no-tumor-pfs-fit, two-tumor-pfs-fit

lst(two_tumor = two_tumor_pfs_res, no_tumor = no_tumor_pfs_res) |>  #, prior = prior_two_tumor_pfs_res) |> 
  map_dfr(
    \(f) spread_rvars(f, log_lambda[t]) |> mutate(lambda = exp(log_lambda)) |> median_qi(lambda, .width = 0.8), 
    .id = "model"
  ) |> 
  ggplot(aes(t)) +
  # stat_lineribbon(aes(ydist = lambda, color = model, fill = model), alpha = 0.25, step = TRUE, .width = 0.8) +
  geom_lineribbon(aes(y = lambda, ymin = .lower, ymax = .upper, color = model, fill = model), alpha = 0.25, step = TRUE) +
  geom_rug(aes(week), alpha = 0.125, 
           data = pfs_analysis_data |> 
             transmute(
               trtp, 
               arm = fct_collapse(trtp, treated = "T-DXd", control = c("LAPACAPE", "TRASCAPE")) |> fct_expand("prior"),
               patient_visits = map(patient_tumors, \(tu) unnest(tu, tumor_history) |> distinct(week))) |> 
             unnest(patient_visits)) +
  labs(title = "Baseline hazard", y = latex2exp::TeX(r"{$\lambda_t$}"), caption = "Ribbons shown are for the 80% CI.\nRugs below x-axis show the distribution of assessment weeks.") +
  scale_color_viridis_d(aesthetics = c("fill", "color"))
```

#### Cross Validation

Here we are calculating the LOO-PSIS cross validation approximations for the no tumors and time invariant two tumor assessments model.

```{r}
#| label: pfs-loo
#| cache: true
#| dependson: no-tumor-pfs-fit, two-tumor-pfs-fit, tumor-change-pfs-fit

no_tumor_loo <- no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
ic_ignored_no_tumor_loo <- ic_ignored_no_tumor_pfs_res$loo(save_psis = TRUE, cores = 12)
two_tumor_loo <- two_tumor_pfs_res$loo(save_psis = TRUE, cores = 12) # This doesn't change anything -> , moment_match = TRUE)
tumor_change_loo <- tumor_change_pfs_res$loo(save_psis = TRUE, cores = 12)

loo_compare(lst(no_tumor_loo, ic_ignored_no_tumor_loo, two_tumor_loo, tumor_change_loo))
```

```{r}
#| label: loss-functions

linear_loss <- function(m, c, a, ac, k0 = 1, k1 = 1) if_else(c & (ac | a >= m), 0, if_else(m >= a, k0, k1) * abs(m - a))
squared_error_loss <- function(m, c, a, ac) if_else(c & (ac | a >= m), 0, (m - a)^2)
```

Linear loss function, where $(m, c)$ are the PFS and right censoring status observed in the data and $(\tilde{m}, \tilde{c})$ are their inferred values.

$$
L(y, c, \tilde{y}, \tilde{c}) = \begin{cases}
0 & \textrm{if } c = 0 \land (\tilde{c} \lor \tilde{y} \geq y) \\
K_0 (y - \tilde{y}) & \textrm{if } y \geq \tilde{y} \\
K_1 (\tilde{y} - y) & \textrm{if } y < \tilde{y} 
\end{cases}
$$

```{r}
#| label: fig-linear-loss

tibble(delta = seq(-10, 10), sym_loss = abs(delta), asym_loss = if_else(delta < 0, 3, 1) * abs(delta)) |> 
  pivot_longer(sym_loss:asym_loss) |> 
  ggplot(aes(delta, value)) +
  geom_line(aes(y = value, color = name), linewidth = 2, alpha = 0.25) +
  scale_color_discrete("Loss Type", labels = c("sym_loss" = "Symmetric", "asym_loss" = "Asymmetric")) +
  labs(x = latex2exp::TeX(r"{$y - \tilde{y}$}"), y = "Loss")
```

For each patient, $i$, we calculate the **out-of-sample** loss as
$$
\int L(y_i, c_i, \tilde{y}_i, \tilde{c}_i)\cdot p(\tilde{y}_i,\tilde{c}_i \mid y_{-i}, c_{-i})\cdot p(y_i)\,\textrm{d}\tilde{y}_i\textrm{d}\tilde{c}_i\textrm{d}y_i.
$$
Where the observed PFS is modeled as uniform value within the censored interval, $Y_i \sim \mathtt{Uniform}(\underline{y}_i, \overline{y}_i)$.

```{r}
#| label: loo-predict-other-utility 
#| code-fold: false 
#| cache: true
#| dependson: pfs-loo

asym_linear_loss <- partial(linear_loss, k1 = 3)
no_tumor_asym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, asym_linear_loss)
ic_ignored_no_tumor_asym_loss <- estimate_oos_loss(ic_ignored_no_tumor_pfs_res, ic_ignored_no_tumor_loo, no_tumor_treated_pfs_stan_data, asym_linear_loss)
two_tumor_asym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, asym_linear_loss)
no_tumor_sym_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, linear_loss)
ic_ignored_no_tumor_sym_loss <- estimate_oos_loss(ic_ignored_no_tumor_pfs_res, ic_ignored_no_tumor_loo, no_tumor_treated_pfs_stan_data, linear_loss)
two_tumor_sym_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, linear_loss)
no_tumor_sqerr_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, squared_error_loss)
two_tumor_sqerr_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, squared_error_loss)

loss_data <- bind_rows(
  lst(no_tumor_sym_loss, 
      ic_ignored_no_tumor_asym_loss, 
      two_tumor_sym_loss, 
      no_tumor_asym_loss, 
      ic_ignored_no_tumor_sym_loss, 
      two_tumor_asym_loss, 
      no_tumor_sqerr_loss, 
      two_tumor_sqerr_loss),
  .id = "model"
) |> 
  mutate(loss_type = str_extract(model, "(a?sym)|sqerr"), model = str_extract(model, "(ic_ignored_)?(no|two)_tumor"))
```

```{r}
#| label: fig-loss
#| fig-height: 8

loss_data |> 
  filter(!fct_match(loss_type, "sqerr"), !fct_match(model, "ic_ignored_no_tumor")) |> 
  plot_loss(binwidth = 6)

loss_data |> 
  filter(fct_match(loss_type, "sqerr"), !fct_match(model, "ic_ignored_no_tumor")) |> 
  plot_loss(binwidth = 100)
```

```{r}
loss_data |> 
  group_by(loss_type, model) |> 
  summarize(expected_loss = mean(value), .groups = "drop") |> 
  kableExtra::kbl() |> 
  kableExtra::kable_styling()
```
```{r}
loss_data |> 
  filter(fct_match(model, c("no_tumor", "two_tumor")), !fct_match(loss_type, "sqerr")) %>%
  left_join(filter(., fct_match(model, "two_tumor")), filter(., fct_match(model, "no_tumor")), by = c("loss_type", "trial", "id")) |> 
  mutate(diff = value.x - value.y) |> 
  ggplot(aes(diff)) +
  stat_histinterval(
    point_interval = mean_qi, .width = c(0.5, 0.8), alpha = 0.5, breaks = ggdist::breaks_fixed(width = 6)
  ) +
  facet_grid(vars(trial), vars(loss_type), scales = "free", margins = "trial") +
  coord_cartesian(xlim = c(-20, 20)) +
  NULL
```


```{r} 
#| label: in-sample-loss
#| eval: false
#| include: false

no_tumor_asym_is_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, asym_linear_loss, insample = TRUE)
two_tumor_asym_is_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, asym_linear_loss, insample = TRUE)
no_tumor_sym_is_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, linear_loss, insample = TRUE)
two_tumor_sym_is_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, linear_loss, insample = TRUE)
no_tumor_sqerr_is_loss <- estimate_oos_loss(no_tumor_pfs_res, no_tumor_loo, no_tumor_treated_pfs_stan_data, squared_error_loss, insample = TRUE)
two_tumor_sqerr_is_loss <- estimate_oos_loss(two_tumor_pfs_res, two_tumor_loo, two_tumor_treated_pfs_stan_data, squared_error_loss, insample = TRUE)

is_loss_data <- bind_rows(
  lst(no_tumor_sym_is_loss, two_tumor_sym_is_loss, no_tumor_asym_is_loss, two_tumor_asym_is_loss, no_tumor_sqerr_is_loss, two_tumor_sqerr_is_loss),
  .id = "model"
) |> 
  mutate(loss_type = str_extract(model, "(a?sym)|sqerr"), model = str_extract(model, "(no|two)_tumor"))

is_loss_data |> 
  group_by(loss_type, model) |> 
  summarize(expected_loss = mean(value), .groups = "drop") |> 
  kableExtra::kbl() |> 
  kableExtra::kable_styling()
```

